---
title: "Grand Canyon QAQC for FMH Data"
author: "Alexandra Lalor"
date: "2025-01-22"
output:
  html_document:
    theme: readable
    highlight: 
    toc: yes
    toc_depth: 3
    toc_float:
      smooth_scroll: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Install packages (if needed)
# install.packages("tidyverse")
# install.packages("Rtools")
# install.packages("janitor")
# install.packages("condformat")
```

# BEFORE STARTING

### Load Packages

```{r}
# Load packages
library(tidyverse)
library(writexl)
library(dplyr)
library(janitor)
#detach(package:plyr)
library(knitr)
# EnvStats is needed for the rosnerTest() function
library(EnvStats)
# condformat is used to format excel output files
library(condformat)
```

### Adjust File Paths

Make sure to update file paths to be specific for your data.

```{r}
# Identify working directory (specifically user name)
getwd()

# Load in data. Change file path based on user name!
path_data <- "X:/FFI Data Management/Exports from FFI/GRCA_FMH/2025-01-23/"
path_errors <- "C:/Users/alalor.NPS/OneDrive - DOI/Desktop/R/GRCA/output/errors/"
path_errors_name <- "2024 GRCA FMH QAQC Errors"
```

### Create Function

```{r}
# QAQC function
qaqc <- function(data, query, query_message, values_check) {
  
  # Check if there is data to QAQC
  if(nrow(data) == 0) {
  # If there is no data, default to "No Error" data frame
  errors <- errors_blank_NoError %>%     
      mutate(SavedQuery = query)
  # If there is data, perform error check
} else {
  
  # Use relevant dataset to look for errors
  errors <- data %>%  
    # Add columns relevant to error checking
           # Populate the "SavedQuery" column with the query name
    mutate("SavedQuery" = query, 
           # Populate the "Error" column with the query message and relevant data
           "Error" = paste(query_message, "=", values_data),
           # Create a blank "Fixed" column
           "Fixed" = "",
           # Create a blank "Explanation" column
           "Explanation" = "",
           # Create a blank "Queryers" column
           "Queryers" = "") %>%   
    # Filter for data which is "false" (data does not match valid conditions)
    # NA values are treated as "false" (missing data is not valid)
    filter(!(values_check %>% replace_na(FALSE))) %>%   
    # Select relevant columns to view errors
    select("SavedQuery","MacroPlot.Name","Date","Error", "Fixed", "Explanation", "Queryers")
  
  # Next, check if there are duplicate errors
  errors_temp <- errors %>% 
    get_dupes(SavedQuery,MacroPlot.Name, Date, Error)
  # Merge duplicate errors
  errors <- unique(merge(errors, errors_temp, all = T))
  # If duplicate errors exist, add number of duplicates to error message
  errors <- errors %>% 
    mutate(Error = ifelse(is.na(dupe_count), Error, paste0("(x", dupe_count, ") ", Error))) %>% 
    select(!"dupe_count")
}
  
  # Check if there are no errors
  if (nrow(errors) == 0) {  
    # If there are no errors, default to "No Error" data frame
    errors <- errors_blank_NoError %>%     
      mutate(SavedQuery = query) 
    # If there are errors, keep error log you just created
  } else {   
    errors <- errors
  }
}
```

```{r}
# Blank data frame for no errors
errors_blank_NoError <- data.frame("SavedQuery" = "","MacroPlot.Name" = "","Date" = "","Error" = "No Error", "Fixed" = "", "Explanation" = "", "Queryers" = "")
# Blank data frame
errors_blank <- data.frame("SavedQuery" = "","MacroPlot.Name" = "","Date" = "","Error" = "", "Fixed" = "", "Explanation" = "", "Queryers" = "")
```

### Load Data

```{r}
# Load in data
Fuels1000_all <- read.csv(paste0(path_data, "Surface Fuels - 1000Hr_XPT.csv"), quote = "")
FuelsDuffLitt_all <- read.csv(paste0(path_data, "Surface Fuels - Duff_Litter_XPT.csv"), quote = "")
FuelsFine_all <- read.csv(paste0(path_data, "Surface Fuels - Fine_XPT.csv"), quote = "")
# HerbsPoints_all <- read.csv(paste0(path_data, "Cover - Points (metric)_XPT.csv"), quote = "")
# HerbsObs_all <- read.csv(paste0(path_data, "Cover - Species Composition (metric)_XPT.csv"), quote = "")
# Shrubs_all <- read.csv(paste0(path_data, "Density - Belts (metric)_XPT.csv"), quote = "")
# Seedlings_all <- read.csv(paste0(path_data, "Density - Quadrats (metric)_XPT.csv"), quote = "")
Trees_all <- read.csv(paste0(path_data, "Trees - Individuals (metric)_XPT.csv"), quote = "")
PostBurn_all <- read.csv(paste0(path_data, "Post Burn Severity_XPT.csv"), quote = "")
```

### Clean & Filter Data

Fuels1000

```{r}
# Clean up data frame
Fuels1000_all <- Fuels1000_all %>%
  # Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>%
  select(!c("Month", "Day", "Time"))

# Ensure blanks in Visited column are NA
Fuels1000_all$Visited[Fuels1000_all$Visited==""] <- NA

# Filter for only data
Fuels1000_data <- Fuels1000_all %>%
  filter(is.na(Visited), Year %in% c(2024))

# Filter for only headers
Fuels1000_header <- Fuels1000_all %>%
  filter(!is.na(Visited), Year %in% c(2024))
```

FuelsDuffLitt

```{r}
# Clean up data frame 
FuelsDuffLitt_all <- FuelsDuffLitt_all %>%  
  # Separate Date column into Date and Time columns   
  separate(Date, sep = " ", into = c("Date", "Time")) %>% 
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>%
  select(!c("Month", "Day", "Time"))

# Ensure blanks in Visited column are NA 
FuelsDuffLitt_all$Visited[FuelsDuffLitt_all$Visited==""] <- NA 

# Filter for only data 
FuelsDuffLitt_data <- FuelsDuffLitt_all %>%   
  filter(is.na(Visited), Year %in% c(2024))  

# Filter for only headers 
FuelsDuffLitt_header <- FuelsDuffLitt_all %>%   
  filter(!is.na(Visited), Year %in% c(2024))
```

FuelsFine

```{r}
# Clean up data frame 
FuelsFine_all <- FuelsFine_all %>%
  # Separate Date column into Date and Time columns   
  separate(Date, sep = " ", into = c("Date", "Time")) %>%   
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>%
  select(!c("Month", "Day", "Time"))

# Ensure blanks in Visited column are NA 
FuelsFine_all$Visited[FuelsFine_all$Visited==""] <- NA  

# Filter for only data 
FuelsFine_data <- FuelsFine_all %>%   
  filter(is.na(Visited), Year %in% c(2024))  

# Filter for only headers 
FuelsFine_header <- FuelsFine_all %>%   
  filter(!is.na(Visited), Year %in% c(2024))
```

HerbsPoints

```{r}
# # Clean up data frame 
# HerbsPoints_all <- HerbsPoints_all %>%  
#   # Separate Date column into Date and Time columns   
#   separate(Date, sep = " ", into = c("Date", "Time")) %>%   
#   separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>%
#   select(!c("Month", "Day", "Time"))
# 
# # Ensure blanks in Visited column are NA 
# HerbsPoints_all$Visited[HerbsPoints_all$Visited==""] <- NA  
# 
# # Filter for only data 
# HerbsPoints_data <- HerbsPoints_all %>%   
#   filter(is.na(Visited), Year %in% c(2024))  
# 
# # Filter for only headers 
# HerbsPoints_header <- HerbsPoints_all %>%   
#   filter(!is.na(Visited), Year %in% c(2024))
```

HerbsObs

```{r}
# # Clean up data frame 
# HerbsObs_all <- HerbsObs_all %>%  
#   # Separate Date column into Date and Time columns   
#   separate(Date, sep = " ", into = c("Date", "Time")) %>%   
#   separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>%
#   select(!c("Month", "Day", "Time"))
# 
# # Ensure blanks in Visited column are NA 
# HerbsObs_all$Visited[HerbsObs_all$Visited==""] <- NA  
# 
# # Filter for only data 
# HerbsObs_data <- HerbsObs_all %>%   
#   filter(is.na(Visited), Year %in% c(2024))  
# 
# # Filter for only headers 
# HerbsObs_header <- HerbsObs_all %>%   
#   filter(!is.na(Visited), Year %in% c(2024))
```

Shrubs

```{r}
# # Clean up data frame 
# Shrubs_all <- Shrubs_all %>%
#   # Separate Date column into Date and Time columns   
#   separate(Date, sep = " ", into = c("Date", "Time")) %>%   
#   separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>%
#   select(!c("Month", "Day", "Time"))
# 
# # Ensure blanks in Visited column are NA 
# Shrubs_all$Visited[Shrubs_all$Visited==""] <- NA 
# 
# # Filter for only data 
# Shrubs_data <- Shrubs_all %>%   
#   filter(is.na(Visited), Year %in% c(2024))  
# 
# # Filter for only headers 
# Shrubs_header <- Shrubs_all %>%   
#   filter(!is.na(Visited), Year %in% c(2024))
```

Seedlings

```{r}
# # Clean up data frame 
# Seedlings_all <- Seedlings_all %>%
#   # Separate Date column into Date and Time columns   
#   separate(Date, sep = " ", into = c("Date", "Time")) %>%   
#   separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>%
#   select(!c("Month", "Day", "Time"))
# 
# # Ensure blanks in Visited column are NA 
# Seedlings_all$Visited[Seedlings_all$Visited==""] <- NA 
# 
# # Filter for only data 
# Seedlings_data <- Seedlings_all %>%   
#   filter(is.na(Visited), Year %in% c(2024))  
# 
# # Filter for only headers 
# Seedlings_header <- Seedlings_all %>%   
#   filter(!is.na(Visited), Year %in% c(2024))
```

Trees

```{r}
# Clean up data frame 
Trees_all <- Trees_all %>%
  # Separate Date column into Date and Time columns   
  separate(Date, sep = " ", into = c("Date", "Time")) %>%   
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>%
  select(!c("Month", "Day", "Time"))

# Ensure blanks in Visited column are NA 
Trees_all$Visited[Trees_all$Visited==""] <- NA  

# Filter for only data 
Trees_data <- Trees_all %>%   
  filter(is.na(Visited), Year %in% c(2024))  

# Filter for only headers 
Trees_header <- Trees_all %>%   
  filter(!is.na(Visited), Year %in% c(2024))
```

PostBurn

```{r}
# Clean up data frame
PostBurn_all <- PostBurn_all %>%
  # Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>%
  select(!c("Month", "Day", "Time"))

# Ensure blanks in Visited column are NA
PostBurn_all$Visited[PostBurn_all$Visited==""] <- NA

# Filter for only data
PostBurn_data <- PostBurn_all %>%
  filter(is.na(Visited), Year %in% c(2024))

# Filter for only headers
PostBurn_header <- PostBurn_all %>%
  filter(!is.na(Visited), Year %in% c(2024))
```

# PROTOCOL - SURFACE FUELS

## 1000HR (CWD)

This code conducts quality control checks on coarse woody debris (CWD) surface fuels data within the "Surface Fuels - 1000Hr" data set.

It checks for errors in: headers, decay class values, fuel constants, duplicates, transect numbers, slope, and diameter.

### Fuel 1000 Header

[Problem:]{.underline} Incorrect header information entered for FMH plots.

[Procedure:]{.underline}

-   Check that NumTran = 4

```{r}
# Set parameters 
data <- Fuels1000_header
query <- "Fuel 1000 Header NumTran"
query_message <- "Number of Transects" 
values_data <- data$NumTran
values_valid <- 4
values_check <- values_data == values_valid

# Identify errors
errors_Fuels1000_NumTran <- qaqc(data, query, query_message, values_check)
```

-   Check that North Rim TranLen = 50 (or very close). North Rim plots include PIAB, PIEN, and PIPN.

```{r}
# Set parameters 
data <- Fuels1000_header %>% 
  filter(MacroPlot.Purpose %in% c("PIAB", "PIEN", "PIPN"))
query <- "Fuel 1000 Header NRim TranLen" 
query_message <- "Transect Length" 
values_data <- data$TranLen 
values_valid <- 50  
values_check <-values_data == values_valid

# Identify errors
errors_Fuels1000_TranLen_N <- qaqc(data, query, query_message, values_check)
```

-   Check that S Rim TranLen = 100 (or very close). South Rim plots include PIED and PIPO.

```{r}
# Set parameters 
data <- Fuels1000_header %>% 
  filter(MacroPlot.Purpose %in% c("PIED", "PIPO"))
query <- "Fuel 1000 Header SRim TranLen" 
query_message <- "Transect Length" 
values_data <- data$TranLen 
values_valid <- 100 
values_check <- values_data == values_valid

# Identify errors
errors_Fuels1000_TranLen_S <- qaqc(data, query, query_message, values_check)
```

#### Combine Header Errors

```{r}
# Combine
errors_Fuels1000_Header <- unique(rbind(errors_Fuels1000_NumTran, errors_Fuels1000_TranLen_N, errors_Fuels1000_TranLen_S))

remove(errors_Fuels1000_NumTran, errors_Fuels1000_TranLen_N, errors_Fuels1000_TranLen_S)
```

### Fuel 1000 DecayClass

[Problem:]{.underline} Logs have an entry for decay class that we don’t use.

[Procedure:]{.underline}

-   Check that DecayCl = 3 or 4

```{r}
# Set parameters
data <- Fuels1000_data
query <- "Fuel 1000 DecayClass"
query_message <- paste0("Log ", data$LogNum, ". ", "Decay Class")
values_data <- data$DecayCl
values_valid <- c(3,4)
values_check <- values_data %in% values_valid

# Identify errors
errors_Fuels1000_DecayCl <- qaqc(data, query, query_message, values_check)
```

### Fuel 1000 FuelConst

[Problem:]{.underline} Incorrect coarse woody debris fuel constant entered for PIAB or PIED or PIEN or PIPN or PIPO. Fuel constant name should correspond to the MacroPlot Purpose.

[Procedure:]{.underline}

-   Check that CWDFuConSt = MacroPlot.Purpose

```{r}
# Set parameters
data <- Fuels1000_data
query <- "Fuel 1000 FuelConst"
query_message <- "Fuel Constant"
values_data <- data$CWDFuConSt
values_valid <- data$MacroPlot.Purpose
values_check <- values_data == values_valid

# Identify errors
errors_Fuels1000_FuelConst <- qaqc(data, query, query_message, values_check)
```

### Fuel 1000 LogNo

[Problem:]{.underline} "Duplicate" data exists due to multiple 1000 hr fuels with the same transect / diameter / decay class. Log Numbers are missing (which would differentiate 1000 hr fuels).

[Procedure:]{.underline}

-   Check for duplicate logs across Date, MacroPlot.Name, Transect, LogNum, Dia [and]{.underline} DecayCl.

```{r}
# Check for log duplication and store results in new column
data_temp <- Fuels1000_data %>%
  get_dupes(Date, MacroPlot.Name, Transect, LogNum, Dia, DecayCl)

# Set parameters 
data <- merge(Fuels1000_data, data_temp, all.x = T)
data$dupe_count[is.na(data$dupe_count)] <- 0
query <- "Fuel 1000 LogNo" 
query_message <- paste0("Log ", data$LogNum, ". ", "Transect ", data$Transect, ". ",  "Dia ", data$Dia, ". ",  "DecayCl ", data$DecayCl, ". ", "Duplicates")
values_data <- data$dupe_count
values_valid <- 0
values_check <- values_data == values_valid

# Identify errors
errors_Fuels1000_LogNo <- qaqc(data, query, query_message, values_check)
```

### Fuel 1000 Values

[Problem:]{.underline}  Values for transect, slope, and diameter are missing or unreasonable.

[Procedure:]{.underline}

-   Check that Transect = 1, 2, 3, or 4

```{r}
# Set parameters
data <- Fuels1000_data
query <- "Fuel 1000 Values TransNum"
query_message <- "Transect Number"
values_data <- data$Transect
values_valid <- c(1, 2, 3, 4)
values_check <- values_data %in% values_valid

# Identify errors
errors_Fuels1000_TransNum <- qaqc(data, query, query_message, values_check)
```

-   Check that Slope values are reasonable.

```{r}
# Set parameters
data <- Fuels1000_data
query <- "Fuel 1000 Values Slope"
query_message <- paste0("Transect = ", data$Transect, ". ", "Slope")
values_data <- data$Slope
values_valid <- seq(0, 40, by = 1)
values_check <- values_data %in% values_valid

# Identify errors
errors_Fuels1000_Slope <- qaqc(data, query, query_message, values_check)
```

-   Check that diameter values are reasonable, are measured at 0.5 increments, and ≥ 3.0.

```{r}
# Perform Rosner's Test for outliers using all values
data_temp <- Fuels1000_all %>%
  filter(!is.na(Dia))
data_outlier <- rosnerTest(data_temp$Dia)
outliers <- data_outlier$all.stats$Value[data_outlier$all.stats$Outlier]
max_Dia <- ifelse(length(outliers) == 0, max(data_temp$Dia), min(outliers) - 1)

# Set parameters
data <- Fuels1000_data
query <- "Fuel 1000 Values Dia"
query_message <- paste0("Log ", data$LogNum, ". ", "Diameter")
values_data <- data$Dia
values_valid <- seq(3, max_Dia, by = 0.5)
values_check <- values_data %in% values_valid

# Identify errors
errors_Fuels1000_Dia <- qaqc(data, query, query_message, values_check)
```

#### Combine Values Errors

```{r}
# Combine
errors_Fuels1000_Values <- unique(rbind(errors_Fuels1000_TransNum, errors_Fuels1000_Slope, errors_Fuels1000_Dia))

remove(errors_Fuels1000_TransNum, errors_Fuels1000_Slope, errors_Fuels1000_Dia)
```

### Fuel 1000 Errors

```{r}
# Save to master error list
# errors_Fuels1000 <- unique(rbind(errors_Fuels1000_DecayCl, errors_Fuels1000_FuelConst, errors_Fuels1000_Header, errors_Fuels1000_LogNo, errors_Fuels1000_Values))

errors_Fuels1000 <- rbind(errors_Fuels1000_DecayCl, errors_blank, errors_Fuels1000_FuelConst, errors_blank, errors_Fuels1000_Header, errors_blank, errors_Fuels1000_LogNo, errors_blank, errors_Fuels1000_Values)

remove(errors_Fuels1000_DecayCl, errors_Fuels1000_FuelConst, errors_Fuels1000_Header, errors_Fuels1000_LogNo, errors_Fuels1000_Values)
```

```{r}
# Table of results for quick check
#kable(errors_Fuels1000, "pipe")

# Save as CSV or XLSX
path_errors
#write.csv(errors_Fuels1000, paste0(path_errors, "errors_GRCA_FMH_Fuels1000.csv"), quote=FALSE, row.names = FALSE, na = "")
#write_xlsx(errors_Fuels1000, paste0(path_errors, "errors_GRCA_FMH_Fuels1000.xlsx"))
```

## DUFF LITTER (DL)

This code conducts quality control checks on duff litter (DL) surface fuels data within the "Surface Fuels - Duff_Litter" data set.

It checks for errors in: headers, fuel constants, number of points sampled, litter values, and duff values.

### Fuel DL Header

[Problem:]{.underline} Incorrect header information entered for FMH plots.

[Procedure:]{.underline}

-   Check that NumTran = 4

```{r}
# Set parameters  
data <- FuelsDuffLitt_header
query <- "Fuel DL Header NumTran" 
query_message <- "Number of Transects"  
values_data <- data$NumTran  
values_valid <- 4    
values_check <- values_data == values_valid

# Identify errors
errors_FuelsDuffLitt_Header <- qaqc(data, query, query_message, values_check)
```

### Fuel DL FuelConst

[Problem:]{.underline} Incorrect duff & litter fuel constant entered for PIAB or PIED or PIEN or PIPN or PIPO. Fuel constant name should correspond to the MacroPlot Purpose.

[Procedure:]{.underline}

-   Check that DLFuConSt = MacroPlot.Purpose

```{r}
# Set parameters 
data <- FuelsDuffLitt_data
query <- "Fuel DL FuelConst" 
query_message <- "Fuel Constant" 
values_data <- data$DLFuConSt 
values_valid <- data$MacroPlot.Purpose 
values_check <- values_data == values_valid

# Identify errors
errors_FuelsDuffLitt_FuelConst <- qaqc(data, query, query_message, values_check)
```

### Fuel DL Hits

[Problem:]{.underline} Extra or missing data points in FMH plots. There should be 4 transects with 10 sample points per transect. 

[Procedure:]{.underline}

-   Group by Macroplot. Calculate number of hits per plot. Check that sum_hits = 40

```{r}
# Filter NA values. Add number of hits and store in new column
data_temp1 <- FuelsDuffLitt_data %>%
  filter(!is.na(LittDep)) %>%
  filter(!is.na(DuffDep)) %>% 
  group_by(MacroPlot.Name, Date) %>% 
  dplyr::summarise("sum_hits" = n())

data_temp2 <- FuelsDuffLitt_header %>% 
  group_by(MacroPlot.Name, Date) %>% 
  dplyr::summarise()

data_temp <- merge(data_temp1, data_temp2, by = c("MacroPlot.Name","Date"), all = TRUE)
data_temp$sum_hits[is.na(data_temp$sum_hits)] <- 0
remove(data_temp1, data_temp2)

# Set parameters 
data <- merge(FuelsDuffLitt_data, data_temp, by = c("MacroPlot.Name", "Date"), all = T)
query <- "Fuel DL Hits" 
query_message <- "Complete Sample Locations"
values_data <- data$sum_hits
values_valid <- 40
values_check <- values_data == values_valid

# Identify errors
errors_FuelsDuffLitt_Hits <- qaqc(data, query, query_message, values_check)
```

### Fuel DL Values

[Problem:]{.underline}  Values for transect, litter, and duff depth are missing or unreasonable.

[Procedure:]{.underline}

-   Check that Transect = 1, 2, 3, or 4

```{r}
# Set parameters
data <- FuelsDuffLitt_data
query <- "Fuel DL Values TransNum"
query_message <- "Transect Number"
values_data <- data$Transect
values_valid <- c(1, 2, 3, 4)
values_check <- values_data %in% values_valid

# Identify errors
errors_FuelsDuffLitt_TransNum <- qaqc(data, query, query_message, values_check)
```

-   Sort by LittDep. Check that litter depth values are reasonable.

```{r}
# Perform Rosner's Test for outliers using all values
data_temp <- FuelsDuffLitt_all %>%
  filter(!is.na(LittDep))
data_outlier <- rosnerTest(data_temp$LittDep)
outliers <- data_outlier$all.stats$Value[data_outlier$all.stats$Outlier]
max_Litt <- ifelse(length(outliers) == 0, max(data_temp$LittDep), min(outliers) - 0.1)

# Set parameters 
data <- FuelsDuffLitt_data
query <- "Fuel DL Values LittDep" 
query_message <- paste0("Transect = ", data$Transect, ". ", "Sample Location = ", data$SampLoc, ". ", "Litter Depth")
values_data <- data$LittDep
values_valid <- max_Litt
values_check <- values_data <= values_valid

# Identify errors
errors_FuelsDuffLitt_LittDep <- qaqc(data, query, query_message, values_check)
```

-   Sort by DuffDep. Check that duff depth values are reasonable.

```{r}
# Perform Rosner's Test for outliers using all values
data_temp <- FuelsDuffLitt_all %>%
  filter(!is.na(DuffDep))
data_outlier <- rosnerTest(data_temp$DuffDep)
outliers <- data_outlier$all.stats$Value[data_outlier$all.stats$Outlier]
max_Duff <- ifelse(length(outliers) == 0, max(data_temp$DuffDep), min(outliers) - 0.1)

# Set parameters 
data <- FuelsDuffLitt_data
query <- "Fuel DL Values DuffDep" 
query_message <- paste0("Transect = ", data$Transect, ". ", "Sample Location = ", data$SampLoc, ". ", "Duff Depth")
values_data <- data$DuffDep
values_valid <- max_Duff
values_check <- values_data <= values_valid

# Identify errors
errors_FuelsDuffLitt_DuffDep <- qaqc(data, query, query_message, values_check)
```

#### Combine Values Errors

```{r}
# Combine
errors_FuelsDuffLitt_Values <- unique(rbind(errors_FuelsDuffLitt_TransNum, errors_FuelsDuffLitt_LittDep, errors_FuelsDuffLitt_DuffDep))

remove(errors_FuelsDuffLitt_TransNum, errors_FuelsDuffLitt_LittDep, errors_FuelsDuffLitt_DuffDep)
```

### Fuel DL Errors

```{r}
# Save to master error list 
# errors_FuelsDuffLitt <- unique(rbind(errors_FuelsDuffLitt_FuelConst, errors_FuelsDuffLitt_Header, errors_FuelsDuffLitt_Hits, errors_FuelsDuffLitt_Values))

errors_FuelsDuffLitt <- rbind(errors_FuelsDuffLitt_FuelConst, errors_blank,  errors_FuelsDuffLitt_Header, errors_blank, errors_FuelsDuffLitt_Hits, errors_blank, errors_FuelsDuffLitt_Values)

remove(errors_FuelsDuffLitt_FuelConst, errors_FuelsDuffLitt_Header, errors_FuelsDuffLitt_Hits, errors_FuelsDuffLitt_Values)
```

```{r}
# Table of results for quick check
#kable(errors_FuelsDuffLitt, "pipe")

# Save as CSV or XLSX
path_errors
#write.csv(errors_FuelsDuffLitt, paste0(path_errors, "errors_GRCA_FMH_FuelsDuffLitt.csv"), quote=FALSE, row.names = FALSE, na = "")
#write_xlsx(errors_FuelsDuffLitt, paste0(path_errors, "errors_GRCA_FMH_FuelsDuffLitt.xlsx"))
```

## FINE FUELS (FWD)

This code conducts quality control checks on fine woody debris (FWD) surface fuels data within the "Surface Fuels - Fine" data set.

It checks for errors in: headers, fuel constants, transect numbers, azimuth, slope, and One/Ten/Hundred Hour count values.

### Fuel Fine Header

[Problem:]{.underline} Incorrect header information entered for FMH plots.

[Procedure:]{.underline}

-   Check that NumTran = 4

```{r}
# Set parameters  
data <- FuelsFine_header
query <- "Fuel Fine Header NumTran" 
query_message <- "Number of Transects"  
values_data <- data$NumTran  
values_valid <- 4    
values_check <- values_data == values_valid

# Identify errors
errors_FuelsFine_NumTran <- qaqc(data, query, query_message, values_check)
```

-   Check that OneHrTranLen = 6

```{r}
# Set parameters  
data <- FuelsFine_header
query <- "Fuel Fine Header OneHrTranLen" 
query_message <- "One Hour Transect Length"  
values_data <- data$OneHrTranLen
values_valid <- 6
values_check <- values_data == values_valid

# Identify errors
errors_FuelsFine_OneHrTranLen <- qaqc(data, query, query_message, values_check)
```

-   Check that TenHrTranLen = 6

```{r}
# Set parameters  
data <- FuelsFine_header
query <- "Fuel Fine Header TenHrTranLen" 
query_message <- "Ten Hour Transect Length"  
values_data <- data$TenHrTranLen 
values_valid <- 6
values_check <- values_data == values_valid

# Identify errors
errors_FuelsFine_TenHrTranLen <- qaqc(data, query, query_message, values_check)
```

-   Check that HunHrTranLen = 12

```{r}
# Set parameters  
data <- FuelsFine_header
query <- "Fuel Fine Header HunHrTranLen" 
query_message <- "Hundrend Hour Transect Length"  
values_data <- data$HunHrTranLen 
values_valid <- 12
values_check <- values_data == values_valid

# Identify errors
errors_FuelsFine_HunHrTranLen <- qaqc(data, query, query_message, values_check)
```

#### Combine Header Errors

```{r}
# Combine
errors_FuelsFine_Header <- unique(rbind(errors_FuelsFine_NumTran, errors_FuelsFine_OneHrTranLen, errors_FuelsFine_TenHrTranLen, errors_FuelsFine_HunHrTranLen))

remove(errors_FuelsFine_NumTran, errors_FuelsFine_OneHrTranLen, errors_FuelsFine_TenHrTranLen, errors_FuelsFine_HunHrTranLen)
```

### Fuel Fine FuelConst

[Problem:]{.underline} Incorrect fine woody debris fuel constant entered for PIAB or PIED or PIEN or PIPN or PIPO. Fuel constant name should correspond to the MacroPlot Purpose.

[Procedure:]{.underline}

-   Check that FWDFuConSt = MacroPlot.Purpose

```{r}
# Set parameters 
data <- FuelsFine_data
query <- "Fuel Fine FuelConst" 
query_message <- "Fuel Constant" 
values_data <- data$FWDFuConSt 
values_valid <- data$MacroPlot.Purpose  
values_check <- values_data == values_valid

# Identify errors
errors_FuelsFine_FuelConst <- qaqc(data, query, query_message, values_check)
```

### Fuel Fine Hits

[Problem:]{.underline} Extra or missing data points in FMH plots. There should be 4 transects with 1 sample point per transect. 

[Procedure:]{.underline}

-   Group by Macroplot. Calculate number of hits per plot. Check that sum_hits = 4

```{r}
# Filter NA values. Add number of hits and store in new column
data_temp1 <- FuelsFine_data %>%
  group_by(MacroPlot.Name, Date) %>% 
  dplyr::summarize("sum_hits" = n())

data_temp2 <- FuelsFine_header %>% 
  group_by(MacroPlot.Name, Date) %>% 
  dplyr::summarise()

data_temp <- merge(data_temp1, data_temp2, by = c("MacroPlot.Name","Date"), all = TRUE)
data_temp$sum_hits[is.na(data_temp$sum_hits)] <- 0
remove(data_temp1, data_temp2)

# Set parameters 
data <- merge(FuelsFine_data, data_temp, by = c("MacroPlot.Name", "Date"), all = T)
query <- "Fuel Fine Hits" 
query_message <- "Complete Transects" 
values_data <- data$sum_hits
values_valid <- 4
values_check <- values_data == values_valid

# Identify errors
errors_FuelsFine_Hits <- qaqc(data, query, query_message, values_check)
```

### Fuel Fine Values

[Problem:]{.underline}  Values for transect, azimuth, slope, and counts for One, Ten, Hundred hour fuels are missing or unreasonable.

[Procedure:]{.underline}

-   Check that Transect = 1, 2, 3, or 4

```{r}
# Set parameters
data <- FuelsFine_data
query <- "Fuel Fine Values TransNum"
query_message <- "Transect Number"
values_data <- data$Transect
values_valid <- c(1, 2, 3, 4)
values_check <- values_data %in% values_valid

# Identify errors
errors_FuelsFine_TransNum <- qaqc(data, query, query_message, values_check)
```

-   Check that Azimuth values are reasonable.

```{r}
# Set parameters
data <- FuelsFine_data
query <- "Fuel Fine Values Azimuth"
query_message <- paste0("Transect = ", data$Transect, ". ", "Azimuth")
values_data <- data$Azimuth
values_valid <- seq(0, 359, by = 1)
values_check <- values_data %in% values_valid

# Identify errors
errors_FuelsFine_Azimuth <- qaqc(data, query, query_message, values_check)
```

-   Check that Slope values are reasonable.

```{r}
# Set parameters
data <- FuelsFine_data
query <- "Fuel Fine Values Slope"
query_message <- paste0("Transect = ", data$Transect, ". ", "Slope")
values_data <- data$Slope
values_valid <- seq(0, 40, by = 1)
values_check <- values_data %in% values_valid

# Identify errors
errors_FuelsFine_Slope <- qaqc(data, query, query_message, values_check)
```

-   Check that One Hour values are reasonable.

```{r}
# Perform Rosner's Test for outliers using all values
data_temp <- FuelsFine_all %>%
  filter(!is.na(OneHr))
data_outlier <- rosnerTest(data_temp$OneHr)
outliers <- data_outlier$all.stats$Value[data_outlier$all.stats$Outlier]
max_One <- ifelse(length(outliers) == 0, max(data_temp$OneHr), min(outliers) - 1)

# Set parameters 
data <- FuelsFine_data
query <- "Fuel Fine Values OneHr" 
query_message <- paste0("Transect = ", data$Transect, ". ", "One Hour Count")
values_data <- data$OneHr
values_valid <- seq(0, max_One, by = 1)
values_check <- values_data %in% values_valid

# Identify errors
errors_FuelsFine_OneHr <- qaqc(data, query, query_message, values_check)
```

-   Check that Ten Hour values are reasonable.

```{r}
# Perform Rosner's Test for outliers using all values
data_temp <- FuelsFine_all %>%
  filter(!is.na(TenHr))
data_outlier <- rosnerTest(data_temp$TenHr)
outliers <- data_outlier$all.stats$Value[data_outlier$all.stats$Outlier]
max_Ten <- ifelse(length(outliers) == 0, max(data_temp$TenHr), min(outliers) - 1)

# Set parameters 
data <- FuelsFine_data
query <- "Fuel Fine Values TenHr" 
query_message <- paste0("Transect = ", data$Transect, ". ", "Ten Hour Count")
values_data <- data$TenHr
values_valid <- seq(0, max_Ten, by = 1)
values_check <- values_data %in% values_valid

# Identify errors
errors_FuelsFine_TenHr <- qaqc(data, query, query_message, values_check)
```

-   Check that Hundred Hour values are reasonable.

```{r}
# Perform Rosner's Test for outliers using all values
data_temp <- FuelsFine_all %>%
  filter(!is.na(HunHr))
data_outlier <- rosnerTest(data_temp$HunHr)
outliers <- data_outlier$all.stats$Value[data_outlier$all.stats$Outlier]
max_Hun <- ifelse(length(outliers) == 0, max(data_temp$HunHr), min(outliers) - 1)

# Set parameters 
data <- FuelsFine_data
query <- "Fuel Fine Values HunHr" 
query_message <- paste0("Transect = ", data$Transect, ". ", "Hundred Hour Count") 
values_data <- data$HunHr
values_valid <- seq(0, max_Hun, by = 1)
values_check <- values_data %in% values_valid

# Identify errors
errors_FuelsFine_HunHr <- qaqc(data, query, query_message, values_check)
```

#### Combine Values Errors

```{r}
# Combine
errors_FuelsFine_Values <- unique(rbind(errors_FuelsFine_TransNum, errors_FuelsFine_Azimuth, errors_FuelsFine_Slope, errors_FuelsFine_OneHr, errors_FuelsFine_TenHr, errors_FuelsFine_HunHr))

remove(errors_FuelsFine_TransNum, errors_FuelsFine_Azimuth, errors_FuelsFine_Slope, errors_FuelsFine_OneHr, errors_FuelsFine_TenHr, errors_FuelsFine_HunHr)
```

### Fuel Fine Errors

```{r}
# Save to master error list 
# errors_FuelsFine <- unique(rbind(errors_FuelsFine_FuelConst, errors_FuelsFine_Header, errors_FuelsFine_Hits, errors_FuelsFine_Values))

errors_FuelsFine <- rbind(errors_FuelsFine_FuelConst, errors_blank,  errors_FuelsFine_Header, errors_blank, errors_FuelsFine_Hits, errors_blank,  errors_FuelsFine_Values)

remove(errors_FuelsFine_FuelConst, errors_FuelsFine_Header, errors_FuelsFine_Hits, errors_FuelsFine_Values)
```

```{r}
# Table of results for quick check
#kable(errors_FuelsFine, "pipe")

# Save as CSV or XLSX 
path_errors
#write.csv(errors_FuelsFine, paste0(path_errors, "errors_GRCA_FMH_FuelsFine.csv"), quote=FALSE, row.names = FALSE, na = "") 
#write_xlsx(errors_FuelsFine, paste0(path_errors, "errors_GRCA_FMH_FuelsFine.xlsx"))
```

# PROTOCOL - POST BURN

This code conducts quality control checks on post burn data within the "Post Burn Severity" data set.

It checks for errors in: headers, substrate, and vegetation values.

### PBSev Header

[Problem:]{.underline} Incorrect header information entered for FMH plots.

[Procedure:]{.underline}

-   Check that NumTran = 4

```{r}
# Set parameters
data <- PostBurn_header
query <- "PBSev Header NumTran"
query_message <- "Number of Transects"
values_data <- data$NumTran
values_valid <- 4
values_check <- values_data == values_valid

# Identify errors
errors_PostBurn_NumTran <- qaqc(data, query, query_message, values_check)
```

-   Check that TranLen = 50

```{r}
# Set parameters
data <- PostBurn_header
query <- "PBSev Header TranLen"
query_message <- "Transect Length"
values_data <- data$TranLen
values_valid <- 50
values_check <- values_data == values_valid

# Identify errors
errors_PostBurn_TranLen <- qaqc(data, query, query_message, values_check)
```

-   Check that NumPtsTran = 10

```{r}
# Set parameters
data <- PostBurn_header
query <- "PBSev Header NumPtsTran"
query_message <- "Number of Points"
values_data <- data$NumPtsTran
values_valid <- 10
values_check <- values_data == values_valid

# Identify errors
errors_PostBurn_NumPtsTran <- qaqc(data, query, query_message, values_check)
```

#### Combine Header Errors

```{r}
# Combine
errors_PostBurn_Header <- unique(rbind(errors_PostBurn_NumTran, errors_PostBurn_TranLen, errors_PostBurn_NumPtsTran))

remove(errors_PostBurn_NumTran, errors_PostBurn_TranLen, errors_PostBurn_NumPtsTran)
```

### PBSev Hits

[Problem:]{.underline} Extra or missing data points in FMH plots. There should be 4 transects with 10 sample points per transect.

[Procedure:]{.underline}

-   Group by Macroplot. Calculate number of hits per plot. Check that sum_hits = 40

```{r}
# Filter NA values. Add number of hits and store in new column
data_temp1 <- PostBurn_data %>%
  filter(!is.na(Veg)) %>%
  filter(!is.na(Sub)) %>% 
  group_by(MacroPlot.Name, Date) %>% 
  dplyr::summarise("sum_hits" = n())

data_temp2 <- PostBurn_header %>% 
  group_by(MacroPlot.Name, Date) %>% 
  dplyr::summarise()

data_temp <- merge(data_temp1, data_temp2, by = c("MacroPlot.Name","Date"), all = TRUE)
data_temp$sum_hits[is.na(data_temp$sum_hits)] <- 0
remove(data_temp1, data_temp2)

# Set parameters 
data <- merge(PostBurn_data, data_temp, by = c("MacroPlot.Name", "Date"), all = T)
query <- "PBSev Hits" 
query_message <- "Complete Sample Locations"
values_data <- data$sum_hits
values_valid <- 40
values_check <- values_data == values_valid

# Identify errors
errors_PostBurn_Hits <- qaqc(data, query, query_message, values_check)
```

### PBSev Values

[Problem:]{.underline} Incorrect values entered for transect, substrate, and vegetation.

[Procedure:]{.underline}

-   Check that Transect = 1, 2, 3, or 4

```{r}
# Set parameters
data <- PostBurn_data
query <- "PBSev Values TransNum"
query_message <- "Transect Number"
values_data <- data$Transect
values_valid <- c(1, 2, 3, 4)
values_check <- values_data %in% values_valid

# Identify errors
errors_PostBurn_TransNum <- qaqc(data, query, query_message, values_check)
```

-   Check that Sub = 0, 1, 2, 3, 4, or 5

```{r}
# Set parameters
data <- PostBurn_data
query <- "PBSev Values Substrate"
query_message <- paste0("Transect = ", data$Transect, ". ", "Point = ", data$Point, ". ", "Substrate Severity")
values_data <- data$Sub
values_valid <- c(0, 1, 2, 3, 4, 5)
values_check <- values_data %in% values_valid

# Identify errors
errors_PostBurn_Sub <- qaqc(data, query, query_message, values_check)
```

-   Check that Veg = 0, 1, 2, 3, 4, or 5

```{r}
# Set parameters
data <- PostBurn_data
query <- "PBSev Values Veg"
query_message <- paste0("Transect = ", data$Transect, ". ", "Point = ", data$Point, ". ", "Vegetation Severity")
values_data <- data$Veg
values_valid <- c(0, 1, 2, 3, 4, 5)
values_check <- values_data %in% values_valid

# Identify errors
errors_PostBurn_Veg <- qaqc(data, query, query_message, values_check)
```

#### Combine Values Errors

```{r}
# Combine
errors_PostBurn_Values <- unique(rbind(errors_PostBurn_TransNum, errors_PostBurn_Sub, errors_PostBurn_Veg))

remove(errors_PostBurn_TransNum, errors_PostBurn_Sub, errors_PostBurn_Veg)
```

### PBSev Errors

```{r}
# Save to master error list
# errors_PostBurn <- unique(rbind(errors_PostBurn_Header, errors_PostBurn_Hits, errors_PostBurn_Values))

errors_PostBurn <- rbind(errors_PostBurn_Header, errors_blank, errors_PostBurn_Hits, errors_blank, errors_PostBurn_Values)

remove(errors_PostBurn_Header, errors_PostBurn_Hits, errors_PostBurn_Values)
```

```{r}
# Table of results for quick check
#kable(errors_PostBurn, "pipe")

# Save as CSV or XLSX
path_errors
#write.csv(errors_PostBurn, paste0(path_errors, "errors_GRCA_FMH_PostBurn.csv"), quote=FALSE, row.names = FALSE, na = "")
#write_xlsx(errors_PostBurn, paste0(path_errors, "errors_GRCA_FMH_PostBurn.xlsx"))
```

# PROTOCOL - TREES

This code conducts quality control checks on tree data within the "Trees - Individuals (metric)" data set.

It checks for errors in: headers, crown base height, crown class, damage codes, DBH, dead and down, duplicates, subplot fractions, height, life form, non-post severity, post severity, status, and user variables.

### Trees Header

[Problem:]{.underline} Incorrect header information entered for FMH plots.

[Procedure:]{.underline}

-   Check that MacroPlotSize = 0.1

```{r}
# Set parameters  
data <- Trees_header
query <- "Trees Header MacroPlotSize" 
query_message <- "Macro Plot Size"  
values_data <- data$MacroPlotSize  
values_valid <- 0.1    
values_check <- values_data == values_valid

# Identify errors
errors_Trees_MacroPlotSize <- qaqc(data, query, query_message, values_check)
```

-   Check that SnagPlotSize = 0.1

```{r}
# Set parameters  
data <- Trees_header
query <- "Trees Header SnagPlotSize" 
query_message <- "Snag Plot Size"  
values_data <- data$SnagPlotSize
values_valid <- 0.1    
values_check <- values_data == values_valid

# Identify errors
errors_Trees_SnagPlotSize <- qaqc(data, query, query_message, values_check)
```

-   Check that BrkPntDia = 15

```{r}
# Set parameters  
data <- Trees_header
query <- "Trees Header BrkPntDia" 
query_message <- "Break Point Diameter"  
values_data <- data$BrkPntDia
values_valid <- 15
values_check <- values_data == values_valid

# Identify errors
errors_Trees_BrkPntDia <- qaqc(data, query, query_message, values_check)
```

#### Combine Header Errors

```{r}
# Save to master error list 
errors_Trees_Header <- unique(rbind(errors_Trees_MacroPlotSize, errors_Trees_SnagPlotSize, errors_Trees_BrkPntDia))

remove(errors_Trees_MacroPlotSize, errors_Trees_SnagPlotSize, errors_Trees_BrkPntDia)
```

### Trees CBH

#### Trees CBH Live

[Problem]{.underline}: Live overstory trees do not have live crown base height entered.

[Procedure:]{.underline}

-   Sort by Status. If Status = L, then check that LiCrBHt is [not]{.underline} blank

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Status == "L", DBH > 15.0)
data$LiCrBHt[is.na(data$LiCrBHt)] <- ""
query <- "Trees CBH Live" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Status = L. DBH = overstory. Live Crown Base Height")
values_data <- data$LiCrBHt
values_valid <- ""
values_check <- values_data != values_valid

# Identify errors
errors_Trees_CBH_Live <- qaqc(data, query, query_message, values_check)
```

#### Trees CBH Dead

[Problem]{.underline}: Dead trees have live crown base height entered.

[Procedure:]{.underline}

-   Sort by Status. If Status = D, then check that LiCrBHt = blank

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Status == "D")
data$LiCrBHt[is.na(data$LiCrBHt)] <- ""
query <- "Trees CBH Dead" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Status = D. Live Crown Base Height")
values_data <- data$LiCrBHt
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_CBH_Dead <- qaqc(data, query, query_message, values_check)
```

#### Trees CBH vs Ht

[Problem]{.underline}: Check that Live Crown Base Height is shorter than total tree Height

[Procedure:]{.underline}

-   Check that LiCrBHt \< Ht

```{r}
# Set parameters
data <- Trees_data %>% 
  filter(!is.na(LiCrBHt), Status == "L")
query <- "Trees CBH vs Ht" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Height shorter than CBH. CBH")
values_data <- data$LiCrBHt
values_valid <- data$Ht
values_check <- values_data < values_valid

# Identify errors
errors_Trees_CBH_Ht <- qaqc(data, query, query_message, values_check)
```

#### Combine CBH Errors

```{r}
# Combine
errors_Trees_CBH <- unique(rbind(errors_Trees_CBH_Live, errors_Trees_CBH_Dead, errors_Trees_CBH_Ht))

remove(errors_Trees_CBH_Live, errors_Trees_CBH_Dead, errors_Trees_CBH_Ht)
```

### Trees Crown Class

[Problem]{.underline}: Incompatible values entered for tree crown class.

[Procedure:]{.underline}

-   Filter by Status = D and DBH = blank. Check that CrwnCl = BBD, CUS, or DD

```{r}
# Set parameters  
data <- Trees_data %>% 
  filter(is.na(DBH), Status == "D")
query <- "Trees Crown Class Dead" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Status = D. DBH = blank. Crown Class")  
values_data <- data$CrwnCl
values_valid <- c("DD", "BBD", "CUS")
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_CrwnCl_1 <- qaqc(data, query, query_message, values_check)
```

-   Filter by Status = L and DBH = blank. Check that CrwnCl = blank

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(is.na(DBH), Status == "L")
query <- "Trees Crown Class Live" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Status = L. DBH = blank. Crown Class")  
values_data <- data$CrwnCl
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_CrwnCl_2 <- qaqc(data, query, query_message, values_check)
```

-   Filter by Status = D and DBH = Overstory. Check that CrwnCl = BAD, CS, LBS, or RS or blank

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(DBH > 15.0, Status == "D")
query <- "Trees Crown Class Overstory Dead" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Status = D. DBH = overstory. Crown Class")
values_data <- data$CrwnCl
values_valid <- c("BAD", "CS", "LBS", "RS")
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_CrwnCl_3 <- qaqc(data, query, query_message, values_check)
```

-   Filter by Status = L and DBH = Overstory. Check that CrwnCl = C, D, I, O, or SC or blank

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(DBH > 15.0, Status == "L")
query <- "Trees Crown Class Overstory Live" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Status = L. DBH = overstory. Crown Class") 
values_data <- data$CrwnCl
values_valid <- c("C", "D", "I", "O", "SC")
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_CrwnCl_4 <- qaqc(data, query, query_message, values_check)
```

-   Filter by DBH = Pole. Check that CrwnCl = X or blank

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(DBH <= 15.0)
query <- "Trees Crown Class Pole" 
query_message <- paste0("Tag ", data$TagNo, ". ", "DBH = pole. Crown Class") 
values_data <- data$CrwnCl
values_valid <- c("X", "")
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_CrwnCl_5 <- qaqc(data, query, query_message, values_check)
```

#### Combine Crown Class Errors

```{r}
# Save to master error list 
errors_Trees_CrwnCl <- unique(rbind(errors_Trees_CrwnCl_1, errors_Trees_CrwnCl_2, errors_Trees_CrwnCl_3, errors_Trees_CrwnCl_4, errors_Trees_CrwnCl_5))

remove(errors_Trees_CrwnCl_1, errors_Trees_CrwnCl_2, errors_Trees_CrwnCl_3, errors_Trees_CrwnCl_4, errors_Trees_CrwnCl_5)
```

### Trees Damage

#### Trees Damage Live

[Problem]{.underline}: Tree damage codes are entered incorrectly.

[Procedure:]{.underline}

-   Sort by DamCd 1. Check that DamCd 1 corresponds to one of the choices on the FMH-1 overstory data sheets.

```{r}
# Make sure damage codes are blank vs NA
Trees_data$DamCd1[is.na(Trees_data$DamCd1)] <- ""
Trees_data$DamCd2[is.na(Trees_data$DamCd2)] <- ""
Trees_data$DamCd3[is.na(Trees_data$DamCd3)] <- ""
Trees_data$DamCd4[is.na(Trees_data$DamCd4)] <- ""
Trees_data$DamCd5[is.na(Trees_data$DamCd5)] <- ""
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Status == "L")
query <- "Trees Damage Live" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Status = L. Damage Code 1")  
values_data <- data$DamCd1
values_valid <- c("", "ABGR", "BIRD", "BLIG", "BROK", "BROM", "BURL", "CONK", "CROK", "DTOP", "EPIC", "EPIP", "FIRE", "FORK", "FRST", "GALL", "HOLW", "INSE", "LEAN", "LICH", "LIGT", "MAMM", "MISL", "MOSS", "OZON", "ROOT", "ROTT", "SPAR", "SPRT", "TWIN", "UMAN", "WOND")
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_L_1 <- qaqc(data, query, query_message, values_check)
```

-   Sort by DamCd 2. Check that DamCd 2 corresponds to one of the choices on the FMH-1 overstory data sheets.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Status == "L")
query <- "Trees Damage Live" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Status = L. Damage Code 2")  
values_data <- data$DamCd2
values_valid <- c("", "ABGR", "BIRD", "BLIG", "BROK", "BROM", "BURL", "CONK", "CROK", "DTOP", "EPIC", "EPIP", "FIRE", "FORK", "FRST", "GALL", "HOLW", "INSE", "LEAN", "LICH", "LIGT", "MAMM", "MISL", "MOSS", "OZON", "ROOT", "ROTT", "SPAR", "SPRT", "TWIN", "UMAN", "WOND")
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_L_2 <- qaqc(data, query, query_message, values_check)
```

-   Sort by DamCd 3. Check that DamCd 3 corresponds to one of the choices on the FMH-1 overstory data sheets.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Status == "L")
query <- "Trees Damage Live" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Status = L. Damage Code 3") 
values_data <- data$DamCd3
values_valid <- c("", "ABGR", "BIRD", "BLIG", "BROK", "BROM", "BURL", "CONK", "CROK", "DTOP", "EPIC", "EPIP", "FIRE", "FORK", "FRST", "GALL", "HOLW", "INSE", "LEAN", "LICH", "LIGT", "MAMM", "MISL", "MOSS", "OZON", "ROOT", "ROTT", "SPAR", "SPRT", "TWIN", "UMAN", "WOND")
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_L_3 <- qaqc(data, query, query_message, values_check)
```

-   Sort by DamCd 4. Check that DamCd 4 corresponds to one of the choices on the FMH-1 overstory data sheets.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Status == "L")
query <- "Trees Damage Live" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Status = L. Damage Code 4")  
values_data <- data$DamCd4
values_valid <- c("", "ABGR", "BIRD", "BLIG", "BROK", "BROM", "BURL", "CONK", "CROK", "DTOP", "EPIC", "EPIP", "FIRE", "FORK", "FRST", "GALL", "HOLW", "INSE", "LEAN", "LICH", "LIGT", "MAMM", "MISL", "MOSS", "OZON", "ROOT", "ROTT", "SPAR", "SPRT", "TWIN", "UMAN", "WOND")
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_L_4 <- qaqc(data, query, query_message, values_check)
```

-   Sort by DamCd 5. Check that DamCd 5 corresponds to one of the choices on the FMH-1 overstory data sheets.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Status == "L")
query <- "Trees Damage Live" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Status = L. Damage Code 5")  
values_data <- data$DamCd5
values_valid <- c("", "ABGR", "BIRD", "BLIG", "BROK", "BROM", "BURL", "CONK", "CROK", "DTOP", "EPIC", "EPIP", "FIRE", "FORK", "FRST", "GALL", "HOLW", "INSE", "LEAN", "LICH", "LIGT", "MAMM", "MISL", "MOSS", "OZON", "ROOT", "ROTT", "SPAR", "SPRT", "TWIN", "UMAN", "WOND")
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_L_5 <- qaqc(data, query, query_message, values_check)
```

#### Trees Damage Dead

[Problem]{.underline}: Damage codes entered for trees that are dead.

[Procedure:]{.underline}

-   Sort by Status. If Status = D, check that DamCd1, DamCd2, DamCd3, DamCd4, [and]{.underline} DamCd5 are blank.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Status == "D")
query <- "Trees Damage Dead" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Status = D. Damage Code 1")
values_data <- data$DamCd1
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_D_1 <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Status == "D")
query <- "Trees Damage Dead" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Status = D. Damage Code 2")  
values_data <- data$DamCd2
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_D_2 <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Status == "D")
query <- "Trees Damage Dead" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Status = D. Damage Code 3")
values_data <- data$DamCd3
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_D_3 <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Status == "D")
query <- "Trees Damage Dead" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Status = D. Damage Code 4")
values_data <- data$DamCd4
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_D_4 <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Status == "D")
query <- "Trees Damage Dead" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Status = D. Damage Code 5")
values_data <- data$DamCd5
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_D_5 <- qaqc(data, query, query_message, values_check)
```

#### Trees Damage Blank

[Problem]{.underline}: Tree damage codes are entered incorrectly. Blank damages exist before entered damages

[Procedure:]{.underline}

-   Filter for DamCd1 = blank. Check that DamCd2, DamCd3, DamCd4, [and]{.underline} DamCd5 are blank.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(DamCd1 == "")
query <- "Trees Damage 1 Blank" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Damage Code 1 = Blank, Damage Code 2")
values_data <- data$DamCd2
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_1_Blank_2 <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(DamCd1 == "")
query <- "Trees Damage 1 Blank" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Damage Code 1 = Blank, Damage Code 3")
values_data <- data$DamCd3
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_1_Blank_3 <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(DamCd1 == "")
query <- "Trees Damage 1 Blank" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Damage Code 1 = Blank, Damage Code 4")
values_data <- data$DamCd4
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_1_Blank_4 <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data  %>% 
  filter(DamCd1 == "")
query <- "Trees Damage 1 Blank" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Damage Code 1 = Blank, Damage Code 5")
values_data <- data$DamCd5
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_1_Blank_5 <- qaqc(data, query, query_message, values_check)
```

-   Filter for DamCd2 = blank. Check that DamCd3, DamCd4, [and]{.underline} DamCd5 are blank.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(DamCd2 == "")
query <- "Trees Damage 2 Blank" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Damage Code 2 = Blank, Damage Code 3")
values_data <- data$DamCd3
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_2_Blank_3 <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(DamCd2 == "")
query <- "Trees Damage 2 Blank" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Damage Code 2 = Blank, Damage Code 4")
values_data <- data$DamCd4
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_2_Blank_4 <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(DamCd2 == "")
query <- "Trees Damage 2 Blank" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Damage Code 2 = Blank, Damage Code 5")
values_data <- data$DamCd5
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_2_Blank_5 <- qaqc(data, query, query_message, values_check)
```

-   Filter for DamCd3 = blank. Check that DamCd4 [and]{.underline} DamCd5 are blank.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(DamCd3 == "")
query <- "Trees Damage 3 Blank" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Damage Code 3 = Blank, Damage Code 4")
values_data <- data$DamCd4
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_3_Blank_4 <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(DamCd3 == "")
query <- "Trees Damage 3 Blank" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Damage Code 3 = Blank, Damage Code 5")
values_data <- data$DamCd5
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_3_Blank_5 <- qaqc(data, query, query_message, values_check)
```

-   Filter for DamCd4 = blank. Check that DamCd5 is blank.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(DamCd4 == "")
query <- "Trees Damage 4 Blank" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Damage Code 4 = Blank, Damage Code 5")
values_data <- data$DamCd5
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_4_Blank_5 <- qaqc(data, query, query_message, values_check)
```

#### Combine Damage Errors

```{r}
# Damage Errors Live Trees
errors_Trees_DamCd_L <- unique(rbind(errors_Trees_DamCd_L_1, errors_Trees_DamCd_L_2, errors_Trees_DamCd_L_3, errors_Trees_DamCd_L_4, errors_Trees_DamCd_L_5))

remove(errors_Trees_DamCd_L_1, errors_Trees_DamCd_L_2, errors_Trees_DamCd_L_3, errors_Trees_DamCd_L_4, errors_Trees_DamCd_L_5)

# Damage Errors Dead Trees
errors_Trees_DamCd_D <- unique(rbind(errors_Trees_DamCd_D_1, errors_Trees_DamCd_D_2, errors_Trees_DamCd_D_3, errors_Trees_DamCd_D_4, errors_Trees_DamCd_D_5))

remove(errors_Trees_DamCd_D_1, errors_Trees_DamCd_D_2, errors_Trees_DamCd_D_3, errors_Trees_DamCd_D_4, errors_Trees_DamCd_D_5)

# Damage Errors Blank Entries
errors_Trees_DamCd_Blank <- unique(rbind(errors_Trees_DamCd_1_Blank_2, errors_Trees_DamCd_1_Blank_3, errors_Trees_DamCd_1_Blank_4, errors_Trees_DamCd_1_Blank_5, errors_Trees_DamCd_2_Blank_3, errors_Trees_DamCd_2_Blank_4, errors_Trees_DamCd_2_Blank_5, errors_Trees_DamCd_3_Blank_4, errors_Trees_DamCd_3_Blank_5, errors_Trees_DamCd_4_Blank_5))

remove(errors_Trees_DamCd_1_Blank_2, errors_Trees_DamCd_1_Blank_3, errors_Trees_DamCd_1_Blank_4, errors_Trees_DamCd_1_Blank_5, errors_Trees_DamCd_2_Blank_3, errors_Trees_DamCd_2_Blank_4, errors_Trees_DamCd_2_Blank_5, errors_Trees_DamCd_3_Blank_4, errors_Trees_DamCd_3_Blank_5, errors_Trees_DamCd_4_Blank_5)

# Damage Errors Combined
errors_Trees_DamCd <- unique(rbind(errors_Trees_DamCd_L, errors_Trees_DamCd_D, errors_Trees_DamCd_Blank))

remove(errors_Trees_DamCd_L, errors_Trees_DamCd_D, errors_Trees_DamCd_Blank)
                                  
```

### Trees DBH

[Problem]{.underline}: DBH values smaller than sampling parameters or excessively large. DBH values for standing trees do not exist.

[Procedure:]{.underline}

-   Filter to exclude DD trees. Check that DBH ≥ 2.5.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(!CrwnCl %in% c("DD", "BBD", "CUS"))
query <- "Trees DBH Min" 
query_message <- paste0("Tag ", data$TagNo, ". ", "DBH")
values_data <- data$DBH
values_valid <- min(2.5)
values_check <- values_data >= values_valid

# Identify errors
errors_Trees_DBH_Min <- qaqc(data, query, query_message, values_check)
```

-   Filter to exclude DD trees. Check that DBH is not excessively large. Check that DBH values for standing trees exist.

```{r}
# Perform Rosner's Test for outliers using all values
data_temp <- Trees_all %>%
  filter(!is.na(DBH))
data_outlier <- rosnerTest(data_temp$DBH)
outliers <- data_outlier$all.stats$Value[data_outlier$all.stats$Outlier]
max_DBH <- ifelse(length(outliers) == 0, max(data_temp$DBH), min(outliers) - 1)

# Set parameters 
data <- Trees_data %>% 
  filter(!CrwnCl %in% c("DD", "BBD", "CUS"))
query <- "Trees DBH Max" 
query_message <- paste0("Tag ", data$TagNo, ". ", "DBH")
values_data <- data$DBH
values_valid <- max_DBH
values_check <- values_data <= values_valid

# Identify errors
errors_Trees_DBH_Max <- qaqc(data, query, query_message, values_check)
```

#### Combine DBH Errors

```{r}
# Combine
errors_Trees_DBH <- unique(rbind(errors_Trees_DBH_Min, errors_Trees_DBH_Max))

remove(errors_Trees_DBH_Min, errors_Trees_DBH_Max)
```

### Trees DD

[Problem]{.underline}: DBH, height, and or crown base height entered for trees that are no longer standing.

[Procedure:]{.underline}

-   Filter for CrwnCl = BBD, CUS, or DD. Check that DBH, Ht, [and]{.underline} LiCrBHt are blank

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(CrwnCl %in% c("DD", "BBD", "CUS"))
data$DBH[is.na(data$DBH)] <- ""
query <- "Trees DD DBH" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Crown Class = DD. DBH")
values_data <- data$DBH
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_DD_DBH <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(CrwnCl %in% c("DD", "BBD", "CUS"))
data$Ht[is.na(data$Ht)] <- ""
query <- "Trees DD Ht" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Crown Class = DD. Height")
values_data <- data$Ht
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_DD_Ht <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(CrwnCl %in% c("DD", "BBD", "CUS"))
data$LiCrBHt[is.na(data$LiCrBHt)] <- ""
query <- "Trees DD CBH" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Crown Class = DD. Live Crown Base Height")
values_data <- data$LiCrBHt
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_DD_LiCrBHt <- qaqc(data, query, query_message, values_check)
```

#### Combine DD Errors

```{r}
# Combine
errors_Trees_DD <- unique(rbind(errors_Trees_DD_DBH, errors_Trees_DD_Ht, errors_Trees_DD_LiCrBHt))

remove(errors_Trees_DD_DBH, errors_Trees_DD_Ht, errors_Trees_DD_LiCrBHt)
```

### Trees Duplicates

[Problem]{.underline}: Duplicate tree tag numbers entered for a plot visit.

[Procedure:]{.underline}

-   Check for duplicate tree tags across Date, MacroPlot.Name, SubFrac [and]{.underline} TagNo.

```{r}
# Check for tag duplication and store results in new column
data_temp <- Trees_data %>% 
  get_dupes(Date, MacroPlot.Name, SubFrac, TagNo)

# Set parameters 
data <- merge(Trees_data, data_temp, all.x = T)
data$dupe_count[is.na(data$dupe_count)] <- 0
query <- "Trees Duplicates" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Duplicates")
values_data <- data$dupe_count
values_valid <- 0
values_check <- values_data == values_valid

# Identify errors
errors_Trees_Duplicate <- qaqc(data, query, query_message, values_check)
```

### Trees Frac

[Problem:]{.underline} Incorrect subplot fraction entered for trees.

[Note:]{.underline} In some cases pole SubFrac will differ from 0.5 if a smaller area was sampled in a particular year. Check yearly procedures to confirm correct values.

[Procedure:]{.underline}

-   If DBH = overstory, then check that SubFrac = 1

```{r}
# Set parameters 
data <- Trees_data %>% 
   filter(DBH > 15.0)
query <- "Trees Frac Overstory" 
query_message <- paste0("Tag ", data$TagNo, ". ", "DBH = overstory. Subplot Fraction")
values_data <- data$SubFrac
values_valid <- 1
values_check <- values_data == values_valid

# Identify errors
errors_Trees_Frac_Overstory <- qaqc(data, query, query_message, values_check)
```

-   If DBH = pole, then check that SubFrac = 0.5

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(DBH <= 15.0)
query <- "Trees Frac Pole" 
query_message <- paste0("DBH = pole. Subplot Fraction")
values_data <- data$SubFrac
values_valid <- 0.5
values_check <- values_data == values_valid

# Identify errors
errors_Trees_Frac_Pole <- qaqc(data, query, query_message, values_check)
```

#### Combine Frac Errors

```{r}
# Combine
errors_Trees_Frac <- unique(rbind(errors_Trees_Frac_Overstory, errors_Trees_Frac_Pole))

remove(errors_Trees_Frac_Overstory, errors_Trees_Frac_Pole)
```

### Trees Height

#### Trees Height Standing

[Problem:]{.underline} Height not recorded. Standing trees should have height values.

[Procedure:]{.underline}

-   Filter for standing trees. Check that Ht exists.

```{r}
# Set parameters 
data <- Trees_data %>% 
   filter(!CrwnCl %in% c("DD", "BBD", "CUS"))
data$Ht[is.na(data$Ht)] <- ""
query <- "Trees Height Standing" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Standing tree. Height")
values_data <- data$Ht
values_valid <- ""
values_check <- values_data != values_valid

# Identify errors
errors_Trees_Height_Standing <- qaqc(data, query, query_message, values_check)
```

#### Trees Height Overstory

[Problem:]{.underline} Incorrect values entered for height. Overstory trees should have reasonable height values.

[Procedure:]{.underline}

-   Filter for Overstory trees. Check that Ht is not excessively large.

```{r}
# Perform Rosner's Test for outliers using all values
data_temp <- Trees_all %>%
  filter(!is.na(Ht), DBH > 15.0)
data_outlier <- rosnerTest(data_temp$Ht)
outliers <- data_outlier$all.stats$Value[data_outlier$all.stats$Outlier]
max_Ht <- ifelse(length(outliers) == 0, max(data_temp$Ht), min(outliers) - 1)

# Set parameters 
data <- Trees_data %>% 
   filter(DBH > 15.0)
query <- "Trees Height Overstory" 
query_message <- paste0("Tag ", data$TagNo, ". ", "DBH = overstory. Height")
values_data <- data$Ht
values_valid <- max_Ht
values_check <- values_data <= values_valid

# Identify errors
errors_Trees_Height_Overstory <- qaqc(data, query, query_message, values_check)
```

#### Trees Height Pole

[Problem:]{.underline} Incorrect values entered for height. Pole trees should correspond to an established height class.

[Procedure:]{.underline}

-   Filter for Pole trees. Check that Ht = 0.6, 1, 2, 3, 4, 5, 6, 7, 8, 9, or 10 or blank.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(DBH <= 15.0)
query <- "Trees Height Pole" 
query_message <- paste0("Tag ", data$TagNo, ". ", "DBH = pole. Height")
values_data <- data$Ht
values_valid <- c("", 0.6, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_Height_Pole <- qaqc(data, query, query_message, values_check)
```

#### Combine Height Errors

```{r}
# Combine
errors_Trees_Height <- unique(rbind(errors_Trees_Height_Standing, errors_Trees_Height_Overstory, errors_Trees_Height_Pole))

remove(errors_Trees_Height_Standing, errors_Trees_Height_Overstory, errors_Trees_Height_Pole)
```

### Trees Life Form

[Problem:]{.underline} Species entered that are not trees.

[Procedure:]{.underline}

-   Check that Preferred Lifeform = Tree

```{r}
# Set parameters 
data <- Trees_data
query <- "Trees Life Form" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Life Form")
values_data <- data$Preferred_Lifeform
values_valid <- "Tree"
values_check <- values_data == values_valid

# Identify errors
errors_Trees_LifeForm <- qaqc(data, query, query_message, values_check)
```

### Trees Non-Post Severity

[Problem]{.underline}: Scorch height, scorch %, and/or char height are entered in sample events other than immediate post for FMH plots. These values should only be collected at immediate post reads.

[Procedure:]{.underline}

-   Filter to exclude all POST reads (01Post, 02Post, etc). Check that CharHt = blank, ScorchHt = blank, [and]{.underline} CrScPct = blank (i.e. no records should appear in the query)

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(!Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"))
data$CharHt[is.na(data$CharHt)] <- ""
query <- "Trees Non-Post Severity CharHt" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Non-post read. Char Height")
values_data <- data$CharHt
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_NonPost_Severity_CharHt <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(!Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"))
data$ScorchHt[is.na(data$ScorchHt)] <- ""
query <- "Trees Non-Post Severity ScorchHt" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Non-post read. Scorch Height")
values_data <- data$ScorchHt
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_NonPost_Severity_ScorchHt <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(!Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"))
data$CrScPct[is.na(data$CrScPct)] <- ""
query <- "Trees Non-Post Severity CrScPct" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Non-post read. Crown Scorch Percent")
values_data <- data$CrScPct
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_NonPost_Severity_CrScPct <- qaqc(data, query, query_message, values_check)
```

#### Combine Non-Post Severity Errors

```{r}
# Combine
errors_Trees_NonPost_Severity <- unique(rbind(errors_Trees_NonPost_Severity_CharHt, errors_Trees_NonPost_Severity_ScorchHt, errors_Trees_NonPost_Severity_CrScPct))

remove(errors_Trees_NonPost_Severity_CharHt, errors_Trees_NonPost_Severity_ScorchHt, errors_Trees_NonPost_Severity_CrScPct)
```

### Trees Post

#### Trees Post Char

[Problem]{.underline}: Char height missing or excessively large for standing trees during the immediate post-fire sample events in FMH plots. Char height entered for trees that are no longer standing.

[Note]{.underline}: Post-burn measurements were not collected on pole-sized trees 1990-2002.

[Procedure:]{.underline}

-   DD trees. Filter to include all POST reads (01Post, 02Post, etc). Filter for DD crown classes (DD, BBD, and CUS). Check that CharHt = blank.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"),
         CrwnCl %in% c("DD", "BBD", "CUS"))
data$CharHt[is.na(data$CharHt)] <- ""
query <- "Trees Post Char DD" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Crown Class = DD. Char Height")
values_data <- data$CharHt
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_Post_Char_DD <- qaqc(data, query, query_message, values_check)
```

-   Standing trees. Filter to include all POST reads (01Post, 02Post, etc). Filter to exclude DD crown classes (DD, BBD, and CUS). Check that CharHt is [not]{.underline} blank.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"),
         !CrwnCl %in% c("DD", "BBD", "CUS"))
data$CharHt[is.na(data$CharHt)] <- ""
query <- "Trees Post Char Standing" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Standing tree. Char Height")
values_data <- data$CharHt
values_valid <- ""
values_check <- values_data != values_valid

# Identify errors
errors_Trees_Post_Char_Standing <- qaqc(data, query, query_message, values_check)
```

-   Max char. Filter to include all POST reads (01Post, 02Post, etc). Check that CharHt is not excessively high.

```{r}
# Perform Rosner's Test for outliers using all values
data_temp <- Trees_all %>%
  filter(Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"),
         !is.na(CharHt))
data_outlier <- rosnerTest(data_temp$CharHt)
outliers <- data_outlier$all.stats$Value[data_outlier$all.stats$Outlier]
max_Char <- ifelse(length(outliers) == 0, max(data_temp$CharHt), min(outliers) - 1)

# Set parameters 
data <- Trees_data %>% 
  filter(Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"),
         !is.na(CharHt))
query <- "Trees Post Char Max" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Char Height")
values_data <- data$CharHt
values_valid <- max_Char
values_check <- values_data <= values_valid

# Identify errors
errors_Trees_Post_Char_Max <- qaqc(data, query, query_message, values_check)
```

#### Trees Post ScorchHt

[Problem]{.underline}: Scorch height missing or excessively large for live trees during the immediate post-fire sample events in FMH plots. Scorch height entered for trees that are no longer standing or are broken.

[Note]{.underline}: Post-burn measurements were not collected on pole-sized trees 1990-2002.

[Procedure:]{.underline}

-   DD Trees. Filter to include all POST reads (01Post, 02Post, etc). Filter for DD crown classes (DD, BBD, and CUS). Check that ScorchHt = blank.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"),
         CrwnCl %in% c("DD", "BBD", "CUS"))
data$ScorchHt[is.na(data$ScorchHt)] <- ""
query <- "Trees Post Scorch DD" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Crown Class = DD. Scorch Height")
values_data <- data$ScorchHt
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_Post_Scorch_DD <- qaqc(data, query, query_message, values_check)
```

-   Standing Dead Trees. Filter to include all POST reads (01Post, 02Post, etc). Filter to exclude DD crown classes (DD, BBD, and CUS). Filter for Status = D. Check that ScorchHt = blank.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"),
         !CrwnCl %in% c("DD", "BBD", "CUS"),
         Status == "D")
data$ScorchHt[is.na(data$ScorchHt)] <- ""
query <- "Trees Post Scorch Dead"
query_message <- paste0("Tag ", data$TagNo, ". ", "Status = D. Standing tree. Scorch Height")
values_data <- data$ScorchHt
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_Post_Scorch_Dead <- qaqc(data, query, query_message, values_check)
```

-   Live trees. Filter to include all POST reads (01Post, 02Post, etc). Filter for Status = L. Check that ScorchHt is not blank or excessively high.

```{r}
# Perform Rosner's Test for outliers using all values
data_temp <- Trees_all %>%
  filter(Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"),
         !is.na(ScorchHt))
data_outlier <- rosnerTest(data_temp$ScorchHt)
outliers <- data_outlier$all.stats$Value[data_outlier$all.stats$Outlier]
max_Scorch <- ifelse(length(outliers) == 0, max(data_temp$ScorchHt), min(outliers) - 1)

# Set parameters 
data <- Trees_data %>% 
  filter(Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"),
         Status == "L")
query <- "Trees Post Scorch Live" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Status = L. Scorch Height")
values_data <- data$ScorchHt
values_valid <- max_Scorch
values_check <- values_data <= values_valid

# Identify errors
errors_Trees_Post_Scorch_Live <- qaqc(data, query, query_message, values_check)
```

#### Trees Post Scorch%

[Problem]{.underline}: Scorch % missing or excessively large for live trees during the immediate post-fire sample events in FMH plots. Scorch % entered incorrectly for trees that are no longer standing.

[Note]{.underline}: Post-burn measurements were not collected on pole-sized trees 1990-2002.

[Procedure:]{.underline}

-   DD Trees. Filter to include all POST reads (01Post, 02Post, etc). Filter for DD crown classes (DD, BBD, and CUS). Check that CrScPct = blank.

-   Check that all rows with CrwnCl = DD have CrScPct = blank or 100. (RS?)

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"),
         CrwnCl %in% c("DD", "BBD", "CUS"))
data$CrScPct[is.na(data$CrScPct)] <- ""
query <- "Trees Post Scorch% DD" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Crown Class = DD. Scorch Percent")
values_data <- data$CrScPct
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_Post_ScorchPct_DD <- qaqc(data, query, query_message, values_check)
```

-   Standing Dead Trees. Filter to include all POST reads (01Post, 02Post, etc). Filter to exclude DD crown classes (DD, BBD, and CUS). Filter for Status = D. Check that CrScPct = blank or 100.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"),
         !CrwnCl %in% c("DD", "BBD", "CUS"),
         Status == "D")
data$CrScPct[is.na(data$CrScPct)] <- ""
query <- "Trees Post Scorch% Dead"
query_message <- paste0("Tag ", data$TagNo, ". ", "Status = D. Standing tree. Scorch Percent")
values_data <- data$CrScPct
values_valid <- c("", 100)
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_Post_ScorchPct_Dead <- qaqc(data, query, query_message, values_check)
```

-   Live trees. Filter to include all POST reads (01Post, 02Post, etc). Filter for Status = L. Check that CrScPct is not blank. Check that CrScPct ≤ 100.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"),
         Status == "L")
query <- "Trees Post Scorch% Live" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Status = L. Scorch Percent")
values_data <- data$CrScPct
values_valid <- max(100)
values_check <- values_data <= values_valid

# Identify errors
errors_Trees_Post_ScorchPct_Live <- qaqc(data, query, query_message, values_check)
```

#### Combine Post Severity Errors

```{r}
# Combine
errors_Trees_Post_Severity <- unique(rbind(errors_Trees_Post_Char_DD, errors_Trees_Post_Char_Standing, errors_Trees_Post_Char_Max, errors_Trees_Post_Scorch_DD, errors_Trees_Post_Scorch_Dead, errors_Trees_Post_Scorch_Live, errors_Trees_Post_ScorchPct_DD, errors_Trees_Post_ScorchPct_Dead, errors_Trees_Post_ScorchPct_Live))

remove(errors_Trees_Post_Char_DD, errors_Trees_Post_Char_Standing, errors_Trees_Post_Char_Max, errors_Trees_Post_Scorch_DD, errors_Trees_Post_Scorch_Dead, errors_Trees_Post_Scorch_Live, errors_Trees_Post_ScorchPct_DD, errors_Trees_Post_ScorchPct_Dead, errors_Trees_Post_ScorchPct_Live)
```

### Trees Status

[Problem:]{.underline} Species do not have an entry for status or have an entry that we don’t use.

[Procedure:]{.underline}

-   Sort by Status. Check that Status = L or D

```{r}
# Set parameters 
data <- Trees_data
query <- "Trees Status" 
query_message <- paste0("Tag ", data$TagNo, ". ", "Status")
values_data <- data$Status
values_valid <- c("L", "D")
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_Status <- qaqc(data, query, query_message, values_check)
```

### Trees UV

[Problem]{.underline}: X present in the UV columns signifying respective data (DBH, Ht, or LiCrBHT) weren’t assessed.

[Procedure]{.underline}:

-   Check that UV(1, 2, [or]{.underline} 3) is blank for all plots

```{r}
# Set parameters 
data <- Trees_data
data$UV1[is.na(data$UV1)] <- ""
query <- "Trees UV1" 
query_message <- paste0("Tag ", data$TagNo, ". ", "User Variable 1")
values_data <- data$UV1
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_UV1 <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data
data$UV2[is.na(data$UV2)] <- ""
query <- "Trees UV2" 
query_message <- paste0("Tag ", data$TagNo, ". ", "User Variable 2")
values_data <- data$UV2
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_UV2 <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data
data$UV3[is.na(data$UV3)] <- ""
query <- "Trees UV3" 
query_message <- paste0("Tag ", data$TagNo, ". ", "User Variable 3")
values_data <- data$UV3
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_UV3 <- qaqc(data, query, query_message, values_check)
```

#### Combine UV Errors

```{r}
# Combine
errors_Trees_UV <- unique(rbind(errors_Trees_UV1, errors_Trees_UV2, errors_Trees_UV3))

remove(errors_Trees_UV1, errors_Trees_UV2, errors_Trees_UV3)
```

## Trees Errors

```{r}
# errors_Trees <- unique(rbind(errors_Trees_CrwnCl, errors_Trees_DamCd, errors_Trees_DBH, errors_Trees_DD, errors_Trees_CBH, errors_Trees_Duplicate, errors_Trees_Frac, errors_Trees_Header, errors_Trees_Height, errors_Trees_NonPost_Severity, errors_Trees_Post_Severity, errors_Trees_Status, errors_Trees_LifeForm, errors_Trees_UV))

errors_Trees <- rbind(errors_Trees_CrwnCl, errors_blank, errors_Trees_DamCd, errors_blank, errors_Trees_DBH, errors_blank, errors_Trees_DD, errors_blank, errors_Trees_CBH, errors_blank, errors_Trees_Duplicate, errors_blank, errors_Trees_Frac, errors_blank, errors_Trees_Header, errors_blank, errors_Trees_Height, errors_blank, errors_Trees_NonPost_Severity, errors_blank, errors_Trees_Post_Severity, errors_blank, errors_Trees_Status, errors_blank, errors_Trees_LifeForm, errors_blank, errors_Trees_UV)

remove(errors_Trees_Header, errors_Trees_CrwnCl, errors_Trees_DamCd, errors_Trees_DBH, errors_Trees_CBH, errors_Trees_DD, errors_Trees_Duplicate, errors_Trees_Frac, errors_Trees_Height, errors_Trees_NonPost_Severity, errors_Trees_Post_Severity, errors_Trees_Status, errors_Trees_LifeForm, errors_Trees_UV)
```

```{r}
# Table of results for quick check
#kable(errors_Trees, "pipe")

# Save as CSV or XLSX
path_errors
#write.csv(errors_Trees, paste0(path_errors, "errors_GRCA_FMH_Trees.csv"), quote=FALSE, row.names = FALSE, na = "")
#write_xlsx(errors_Trees, paste0(path_errors, "errors_GRCA_FMH_Trees.xlsx"))
```

# CROSS PROTOCOL CHECKS

## 1000HR VS FINE SLOPE

This code conducts quality control checks on slope data in the surface fuels coarse woody debris (1000-hr) and fine woody debris datasets. It also checks for overall consistency in slope and azimuth values across years.

```{r}
# # Set parameters 
# data_Fuels1000_slope <- Fuels1000_data %>% 
#   group_by(MacroPlot.Name, Date, Transect, Slope) %>% 
#   summarise("MacroPlot.Name" = unique(MacroPlot.Name), 
#             "Transect" = unique(Transect)) %>% 
#   mutate(Slope_1000 = Slope) %>% 
#   arrange(MacroPlot.Name, Transect) %>% 
#   select(!Slope)
# 
# data_FuelsFine_slope <- FuelsFine_data %>% 
#   group_by(MacroPlot.Name, Date, Transect, Comment, Slope) %>% 
#   summarise("MacroPlot.Name" = unique(MacroPlot.Name), 
#             "Transect" = unique(Transect)) %>% 
#   mutate(Slope_fine = Slope) %>% 
#   arrange(MacroPlot.Name, Transect) %>% 
#   select(!Slope)
# 
# errors_Fuels_slope <- merge(data_Fuels1000_slope, data_FuelsFine_slope, all.y = T) %>% 
#   filter(Slope_fine != Slope_1000) %>% 
#   mutate(SavedQuery = "Fuel Slope",
#          Error = paste("FWD slope", Slope_fine, "!=", "CWD slope", Slope_1000)) %>% 
#   reframe("SavedQuery" = unique(query),
#           "MacroPlot.Name" = unique(MacroPlot.Name), 
#           "Date" = unique(Date), 
#           "Error" = unique(Error),
#           "Comment" = unique(Comment))
# 
# # If no errors, add "No Errors" to existing data frame 
# if (nrow(errors_Fuels_slope) == 0) {   
#     errors_Fuels_slope <- errors_blank_NoError %>%     
#     mutate(SavedQuery = query) 
# } else {   
#     errors_Fuels_slope <- errors_Fuels_slope
# }
```

# CROSS YEAR CHECKS

## Trees Dead to Alive

## Trees Smaller DBH

## Trees Shorter Height

## Trees Shorter CBH

## Trees Changing Tag Numbers

# ALL ERRORS

```{r}
# Save to master error list 
# errors_all <- unique(rbind(errors_Fuels1000, errors_FuelsDuffLitt, errors_FuelsFine, errors_Trees))

errors_all <- rbind(errors_Fuels1000, errors_blank, errors_FuelsDuffLitt, errors_blank, errors_FuelsFine, errors_blank, errors_PostBurn, errors_blank, errors_Trees)

# errors_CrossProtocol <- unique(rbind(errors_Fuels_slope))
# 
# errors_CrossYear <- unique(rbind())
```

```{r}
# Table of results for quick check
#kable(errors_all, "pipe")

# Create conditional formatting rule
data_temp <- as.data.frame(grepl("PBSev", errors_all$SavedQuery, fixed = TRUE) | grepl(" Post", errors_all$SavedQuery, fixed = TRUE))
colnames(data_temp) <- c("Post")
errors_all_cf <- cbind(errors_all, data_temp) %>% 
  mutate(Post = ifelse(SavedQuery == "", NA, Post))

# Add conditional formating to ouput
errors_all <- condformat(errors_all_cf[1:(nrow(errors_all_cf)), 1:8]) %>%
  rule_fill_discrete(columns = 1:7,
                     expression = (errors_all_cf$Post == "FALSE"),
                     colours = c("TRUE"="#C6E0B4",
                                 "FALSE"="#FFC000"),
                     na.value = "#FFFFFF") %>%
  select(!c("Post"))

# Save as CSV or XLSX 
path_errors
# write.csv(errors_all, paste0(path_errors, "errors_GRCA_FMH.csv"), quote=FALSE, row.names = FALSE, na = "") 
#write_xlsx(errors_all, paste0(path_errors, "errors_GRCA_FMH.xlsx"))
# condformat2excel(errors_all, paste0(path_errors, "errors_GRCA_FMH_formatted.xlsx"), overwrite_wb = TRUE)
condformat2excel(errors_all, paste0(path_errors, "errors_GRCA_FMH_formatted.xlsx"), sheet_name = path_errors_name, overwrite_wb = TRUE)
```
