---
title: "Grand Canyon QAQC for I&M Data"
author: "Alexandra Lalor"
date: "2024-01-22"
output:
  html_document:
    theme: readable
    highlight: 
    toc: yes
    toc_depth: 3
    toc_float:
      smooth_scroll: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Install packages (if needed)
# install.packages("tidyverse")
# install.packages("Rtools")
```

# BEFORE STARTING

### Load Packages

```{r}
# Load packages
library(tidyverse)
library(writexl)
library(dplyr)
library(plyr)
library(knitr)
# EnvStats is needed for the rosnerTest() function
library(EnvStats)
```

### Adjust File Paths

Make sure to update file paths to be specific for your data.

```{r}
# Identify working directory (specifically user name)
getwd()

# Load in data. Change file path based on user name!

# Mixed Conifer
path_data_M <- "X:/FFI Data Management/Exports from FFI/GRCA_I&M/M/2025-01-23/"
# Pinyon Juniper
path_data_P <- "X:/FFI Data Management/Exports from FFI/GRCA_I&M/P/2025-01-23/"
# Errors
path_errors <- "C:/Users/alalor.NPS/OneDrive - DOI/Desktop/R/GRCA/output/errors/2024/"
```

### Load Data

```{r}
# Blank data frame for no errors
errors_blank <- data.frame("SavedQuery" = "","MacroPlot.Name" = "","Date" = "","Error" = "No Error", "Comment" = "")
```

Mixed Conifer

```{r}
# Load in data
Fuels1000_all_M <- read.csv(paste0(path_data_M, "Surface Fuels - 1000Hr_XPT.csv"))
FuelsDuffLitt_all_M <- read.csv(paste0(path_data_M, "Surface Fuels - Duff_Litter_XPT.csv"))
FuelsFine_all_M <- read.csv(paste0(path_data_M, "Surface Fuels - Fine_XPT.csv"))
Trees_all_M <- read.csv(paste0(path_data_M, "Trees - Individuals (metric)_XPT.csv"))
PostBurn_all_M <- read.csv(paste0(path_data_M, "Post Burn Severity_XPT.csv"))
```

Pinyon Juniper

```{r}
# Load in data
Fuels1000_all_P <- read.csv(paste0(path_data_P, "Surface Fuels - 1000Hr_XPT.csv"))
FuelsDuffLitt_all_P <- read.csv(paste0(path_data_P, "Surface Fuels - Duff_Litter_XPT.csv"))
FuelsFine_all_P <- read.csv(paste0(path_data_P, "Surface Fuels - Fine_XPT.csv"))
Trees_all_P <- read.csv(paste0(path_data_P, "Trees - Individuals (metric)_XPT.csv"))
PostBurn_all_P <- read.csv(paste0(path_data_P, "Post Burn Severity_XPT.csv"))
```

### Merge Data

```{r}
# Merge M and P
Fuels1000_all <- rbind(Fuels1000_all_M, Fuels1000_all_P)
FuelsDuffLitt_all <- rbind(FuelsDuffLitt_all_M, FuelsDuffLitt_all_P)
FuelsFine_all <- rbind(FuelsFine_all_M, FuelsFine_all_P)
Trees_all <- rbind(Trees_all_M, Trees_all_P)
PostBurn_all <- rbind(PostBurn_all_M, PostBurn_all_P)
```

```{r}
# Remove extra data
remove(Fuels1000_all_M, Fuels1000_all_P, FuelsDuffLitt_all_M, FuelsDuffLitt_all_P, FuelsFine_all_M, FuelsFine_all_P, Trees_all_M, Trees_all_P, PostBurn_all_M, PostBurn_all_P)
```

### Check Data

Fuels1000

```{r}
# Clean up data frame
Fuels1000_all <- Fuels1000_all %>%
  # Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>%
  select(!c("Month", "Day", "Time")) %>%
  # Add Monitoring Type
  separate(MacroPlot.Name, sep = "_", into = c("Park", "Plot"), remove = FALSE) %>%
  separate(Plot, sep = "(?=[[:digit:]])", into = c("MonitoringType", "PlotNumber"), remove = FALSE) %>%
  select(!c("Park", "PlotNumber")) %>% 
  mutate(MonitoringType = ifelse(MonitoringType == "M", "Mixed Conifer", "PIED"))

# Ensure blanks in Visited column are NA
Fuels1000_all$Visited[Fuels1000_all$Visited==""] <- NA
  
# Filter for only data, by year
Fuels1000_data <- Fuels1000_all %>%
  filter(is.na(Visited), Year %in% c(2021, 2022, 2023))

# Filter for only headers, by year
Fuels1000_header <- Fuels1000_all %>%
  filter(!is.na(Visited), Year %in% c(2021, 2022, 2023))
```

FuelsDuffLitt

```{r}
# Clean up data frame
FuelsDuffLitt_all <- FuelsDuffLitt_all %>%
  # Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>% 
  select(!c("Month", "Day", "Time")) %>% 
  # Add Monitoring Type
  separate(MacroPlot.Name, sep = "_", into = c("Park", "Plot"), remove = FALSE) %>%
  separate(Plot, sep = "(?=[[:digit:]])", into = c("MonitoringType", "PlotNumber"), remove = FALSE) %>%
  select(!c("Park", "PlotNumber")) %>% 
  mutate(MonitoringType = ifelse(MonitoringType == "M", "Mixed Conifer", "PIED"))

# Ensure blanks in Visited column are NA
FuelsDuffLitt_all$Visited[FuelsDuffLitt_all$Visited==""] <- NA
  
# Filter for only data, by year
FuelsDuffLitt_data <- FuelsDuffLitt_all %>%
  filter(is.na(Visited), Year %in% c(2021, 2022, 2023))

# Filter for only headers, by year
FuelsDuffLitt_header <- FuelsDuffLitt_all %>%
  filter(!is.na(Visited), Year %in% c(2021, 2022, 2023))
```

FuelsFine

```{r}
# Clean up data frame
FuelsFine_all <- FuelsFine_all %>%
  # Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>% 
  select(!c("Month", "Day", "Time")) %>% 
  # Add Monitoring Type
  separate(MacroPlot.Name, sep = "_", into = c("Park", "Plot"), remove = FALSE) %>%
  separate(Plot, sep = "(?=[[:digit:]])", into = c("MonitoringType", "PlotNumber"), remove = FALSE) %>%
  select(!c("Park", "PlotNumber")) %>% 
  mutate(MonitoringType = ifelse(MonitoringType == "M", "Mixed Conifer", "PIED"))

# Ensure blanks in Visited column are NA
FuelsFine_all$Visited[FuelsFine_all$Visited==""] <- NA
  
# Filter for only data, by year
FuelsFine_data <- FuelsFine_all %>%
  filter(is.na(Visited), Year %in% c(2021, 2022, 2023))

# Filter for only headers, by year
FuelsFine_header <- FuelsFine_all %>%
  filter(!is.na(Visited), Year %in% c(2021, 2022, 2023))
```

Trees

```{r}
# Clean up data frame
Trees_all <- Trees_all %>%
  # Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>% 
  select(!c("Month", "Day", "Time")) %>% 
  # Add Monitoring Type
  separate(MacroPlot.Name, sep = "_", into = c("Park", "Plot"), remove = FALSE) %>%
  separate(Plot, sep = "(?=[[:digit:]])", into = c("MonitoringType", "PlotNumber"), remove = FALSE) %>%
  select(!c("Park", "PlotNumber")) %>% 
  mutate(MonitoringType = ifelse(MonitoringType == "M", "Mixed Conifer", "PIED"))

# Ensure blanks in Visited column are NA
Trees_all$Visited[Trees_all$Visited==""] <- NA
  
# Filter for only data, by year
Trees_data <- Trees_all %>%
  filter(is.na(Visited), Year %in% c(2021, 2022, 2023))

# Filter for only headers, by year
Trees_header <- Trees_all %>%
  filter(!is.na(Visited), Year %in% c(2021, 2022, 2023))
```

PostBurn

```{r}
# Clean up data frame
PostBurn_all <- PostBurn_all %>%
  # Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>% 
  select(!c("Month", "Day", "Time")) %>% 
  # Add Monitoring Type
  separate(MacroPlot.Name, sep = "_", into = c("Park", "Plot"), remove = FALSE) %>%
  separate(Plot, sep = "(?=[[:digit:]])", into = c("MonitoringType", "PlotNumber"), remove = FALSE) %>%
  select(!c("Park", "PlotNumber")) %>% 
  mutate(MonitoringType = ifelse(MonitoringType == "M", "Mixed Conifer", "PIED"))

# Ensure blanks in Visited column are NA
PostBurn_all$Visited[PostBurn_all$Visited==""] <- NA
  
# Filter for only data, by year
PostBurn_data <- PostBurn_all %>%
  filter(is.na(Visited), Year %in% c(2021, 2022, 2023))

# Filter for only headers, by year
PostBurn_header <- PostBurn_all %>%
  filter(!is.na(Visited), Year %in% c(2021, 2022, 2023))
```

### Create Function

```{r}
# QAQC function
qaqc <- function(data, query, query_message, values_check) {
  # Identify errors 
  errors <- data %>%   
    mutate("SavedQuery" = query, 
           "Error" = paste(query_message, "=", values_data)) %>%   
    # Data is equal exactly to a value
    filter(!(values_check %>% replace_na(FALSE))) %>%   
    select("SavedQuery", "MacroPlot.Name", "Date", "Error", "Comment")
  
  # If no errors, add "No Errors" to existing data frame 
  if (nrow(errors) == 0) {   
    errors <- errors_blank %>%     
      mutate(SavedQuery = query) 
  } else {   
    errors <- errors
  }
}
```

# PROTOCOL - SURFACE FUELS

## 1000HR CWD

This code conducts quality control checks on surface fuels data within the surface fuels coarse woody debris (1000-hr) data set.

It checks for: header info (number of transects, length of transects), decay class values, fuel constants, duplicate values, transect numbers, reasonable slope numbers, reasonable diameter values.

### Fuel 1000 Header

[Problem:]{.underline} Incorrect header information entered for I&M plots.

[Procedure:]{.underline}

-   Check that NumTran = 3

```{r}
# Set parameters 
data <- Fuels1000_header
query <- "Fuel 1000 Header NumTran"
query_message <- "Number of Transects" 
values_data <- data$NumTran
values_valid <- 3
values_check <- values_data == values_valid

# Identify errors
errors_Fuels1000_NumTran <- qaqc(data, query, query_message, values_check)
```

-   Check that TranLen = 50

```{r}
# Set parameters 
data <- Fuels1000_header
query <- "Fuel 1000 Header TranLen" 
query_message <- "Transect Length" 
values_data <- data$TranLen 
values_valid <- 50  
values_check <- values_data == values_valid

# Identify errors
errors_Fuels1000_TranLen <- qaqc(data, query, query_message, values_check)
```

### Fuel 1000 DecayClass

[Problem:]{.underline} Logs have an entry for decay class that we don’t use.

[Procedure:]{.underline}

-   Check that DecayCl = 3 or 4

```{r}
# Set parameters
data <- Fuels1000_data
query <- "Fuel 1000 DecayClass"
query_message <- "Decay Class"
values_data <- data$DecayCl
values_valid <- c(3,4)
values_check <- values_data %in% values_valid

# Identify errors
errors_Fuels1000_DecayCl <- qaqc(data, query, query_message, values_check)
```

### Fuel 1000 FuelConst

[Problem:]{.underline} Incorrect coarse woody debris fuel constant entered. Fuel constant name should correspond to the Monitoring Type.

[Procedure:]{.underline}

-   Check that CWDFuConSt = MonitoringType

```{r}
# Set parameters
data <- Fuels1000_data
query <- "Fuel 1000 FuelConst"
query_message <- "Fuel Constant"
values_data <- data$CWDFuConSt
values_valid <- data$MonitoringType
values_check <- values_data == values_valid

# Identify errors
errors_Fuels1000_FuelConst <- qaqc(data, query, query_message, values_check)
```

### Fuel 1000 LogNo

[Problem:]{.underline} "Duplicate" data exists due to multiple 1000 hr fuels with the same transect / diameter / decay class. Log Numbers are missing (which would differentiate 1000 hr fuels).

[Procedure:]{.underline}

-   Check for duplication (TRUE/FALSE) and store results in new column

-   Check for TRUE values

```{r}
# Remove Index. Check for duplication (TRUE/FALSE) and store results in new column
data_temp <- Fuels1000_data %>%
  select(!"Index")

# Set parameters 
data <- data_temp %>%
  mutate("Duplicate" = duplicated(data_temp))
query <- "Fuel 1000 LogNo" 
query_message <- "Duplicate Logs" 
values_data <- data$Duplicate
values_valid <- FALSE
values_check <- values_data == values_valid

# Identify errors
errors_Fuels1000_LogNo <- qaqc(data, query, query_message, values_check)
```

### Fuel 1000 Values

[Problem:]{.underline}  Values for transect, slope, and diameter are missing or unreasonable.

[Procedure:]{.underline}

-   Check that Transect = 1, 2, or 3

```{r}
# Set parameters
data <- Fuels1000_data
query <- "Fuel 1000 Values TransNum"
query_message <- "Transect Number"
values_data <- data$Transect
values_valid <- c(1, 2, 3)
values_check <- values_data %in% values_valid

# Identify errors
errors_Fuels1000_TransNum <- qaqc(data, query, query_message, values_check)
```

-   Check that Slope values are reasonable.

```{r}
# Set parameters
data <- Fuels1000_data
query <- "Fuel 1000 Values Slope"
query_message <- "Slope"
values_data <- data$Slope
values_valid <- seq(0, 90, by = 1)
values_check <- values_data %in% values_valid

# Identify errors
errors_Fuels1000_Slope <- qaqc(data, query, query_message, values_check)
```

-   Check that diameter values are reasonable, are measured at 0.5 increments, and ≥ 3.0.

```{r}
# Perform Rosner's Test for outliers using all values
data_temp <- Fuels1000_all %>%
  filter(!is.na(Dia))
data_outlier <- rosnerTest(data_temp$Dia)
outliers <- data_outlier$all.stats$Value[data_outlier$all.stats$Outlier]
max_Dia <- ifelse(length(outliers) == 0, max(data_temp$Dia), min(outliers) - 1)

# Set parameters
data <- Fuels1000_data
query <- "Fuel 1000 Values Dia"
query_message <- "Diameter"
values_data <- data$Dia
values_valid <- seq(3, max_Dia, by = 0.5)
values_check <- values_data %in% values_valid

# Identify errors
errors_Fuels1000_Dia <- qaqc(data, query, query_message, values_check)
```

### Fuel 1000 Errors

```{r}
# Save to master error list
errors_Fuels1000 <- unique(rbind(errors_Fuels1000_DecayCl, errors_Fuels1000_FuelConst, errors_Fuels1000_NumTran, errors_Fuels1000_TranLen, errors_Fuels1000_LogNo, errors_Fuels1000_TransNum, errors_Fuels1000_Slope, errors_Fuels1000_Dia))

remove(errors_Fuels1000_DecayCl, errors_Fuels1000_FuelConst, errors_Fuels1000_NumTran, errors_Fuels1000_TranLen, errors_Fuels1000_LogNo, errors_Fuels1000_TransNum, errors_Fuels1000_Slope, errors_Fuels1000_Dia)
```

```{r}
# Table of results for quick check
kable(errors_Fuels1000, "pipe")

# Save as CSV or XLSX
#path_errors
#write.csv(errors_Fuels1000, paste0(path_errors, "errors_GRCA_I&M_Fuels1000.csv"), quote=FALSE, row.names = FALSE, na = "")
#write_xlsx(errors_Fuels1000, paste0(path_errors, "errors_GRCA_I&M_Fuels1000.xlsx"))
```

## DUFF LITTER

This code conducts quality control checks on surface fuels data within the surface fuels duff litter data set.

It checks for: number of transects, fuel constants, number of points sampled, reasonable litter values, reasonable duff values.

### Fuel DL Header

[Problem:]{.underline} Incorrect header information entered for I&M plots.

[Procedure:]{.underline}

-   Check that NumTran = 3

```{r}
# Set parameters  
data <- FuelsDuffLitt_header
query <- "Fuel DL Header NumTran" 
query_message <- "Number of Transects"  
values_data <- data$NumTran  
values_valid <- 3    
values_check <- values_data == values_valid

# Identify errors
errors_FuelsDuffLitt_NumTran <- qaqc(data, query, query_message, values_check)
```

### Fuel DL FuelConst

[Problem:]{.underline} Incorrect duff & litter fuel constant entered. Fuel constant name should correspond to the Monitoring Type.

[Procedure:]{.underline}

-   Check that DLFuConSt = Monitoring Type

```{r}
# Set parameters 
data <- FuelsDuffLitt_data
query <- "Fuel DL FuelConst" 
query_message <- "Fuel Constant" 
values_data <- data$DLFuConSt 
values_valid <- data$MonitoringType  
values_check <- values_data == values_valid

# Identify errors
errors_FuelsDuffLitt_FuelConst <- qaqc(data, query, query_message, values_check)
```

### Fuel DL Hits

[Problem:]{.underline} Extra or missing data points in I&M plots. There should be 3 transects with 10 sample points per transect. 

[Procedure:]{.underline}

-   Group by Macroplot. Calculate number of hits per plot. Check that sum_hits = 30

```{r}
# Filter NA values. Add number of hits and store in new column
data_temp1 <- FuelsDuffLitt_data %>%
  filter(!is.na(LittDep)) %>%
  filter(!is.na(DuffDep)) %>% 
  group_by(MacroPlot.Name) %>% 
  dplyr::summarise("sum_hits" = n())

data_temp2 <- FuelsDuffLitt_header %>% 
  mutate(MacroPlot.Name = unique(MacroPlot.Name)) %>% 
  select(MacroPlot.Name, Date)

data_temp <- merge(data_temp1, data_temp2, by = "MacroPlot.Name", all = TRUE)
remove(data_temp1, data_temp2)

# Set parameters 
data <- merge(FuelsDuffLitt_data, data_temp, by = c("MacroPlot.Name", "Date"), all = T)
query <- "Fuel DL Hits" 
query_message <- "Number of Hits" 
values_data <- data$sum_hits
values_valid <- 30
values_check <- values_data == values_valid

# Identify errors
errors_FuelsDuffLitt_Hits <- qaqc(data, query, query_message, values_check)
```

### Fuel DL Values

[Problem:]{.underline}  Values for litter and duff depth are missing or unreasonable.

[Procedure:]{.underline}

-   Sort by LittDep. Check that litter depth values are reasonable.

```{r}
# Perform Rosner's Test for outliers using all values
data_temp <- FuelsDuffLitt_all %>%
  filter(!is.na(LittDep))
data_outlier <- rosnerTest(data_temp$LittDep)
outliers <- data_outlier$all.stats$Value[data_outlier$all.stats$Outlier]
max_Litt <- ifelse(length(outliers) == 0, max(data_temp$LittDep), min(outliers) - 0.1)

# Set parameters 
data <- FuelsDuffLitt_data
query <- "Fuel DL Values LittDep" 
query_message <- "Litter Depth Outlier" 
values_data <- data$LittDep
values_valid <- max_Litt
values_check <- values_data <= values_valid

# Identify errors
errors_FuelsDuffLitt_LittDep <- qaqc(data, query, query_message, values_check)
```

-   Sort by DuffDep. Check that duff depth values are reasonable.

```{r}
# Perform Rosner's Test for outliers using all values
data_temp <- FuelsDuffLitt_all %>%
  filter(!is.na(DuffDep))
data_outlier <- rosnerTest(data_temp$DuffDep)
outliers <- data_outlier$all.stats$Value[data_outlier$all.stats$Outlier]
max_Duff <- ifelse(length(outliers) == 0, max(data_temp$DuffDep), min(outliers) - 0.1)

# Set parameters 
data <- FuelsDuffLitt_data
query <- "Fuel DL Values DuffDep" 
query_message <- "Duff Depth Outlier" 
values_data <- data$DuffDep
values_valid <- max_Duff
values_check <- values_data <= values_valid

# Identify errors
errors_FuelsDuffLitt_DuffDep <- qaqc(data, query, query_message, values_check)
```

### Fuel DL Errors

```{r}
# Save to master error list 
errors_FuelsDuffLitt <- unique(rbind(errors_FuelsDuffLitt_FuelConst, errors_FuelsDuffLitt_NumTran, errors_FuelsDuffLitt_Hits, errors_FuelsDuffLitt_LittDep, errors_FuelsDuffLitt_DuffDep))

remove(errors_FuelsDuffLitt_FuelConst, errors_FuelsDuffLitt_NumTran, errors_FuelsDuffLitt_Hits, errors_FuelsDuffLitt_LittDep, errors_FuelsDuffLitt_DuffDep)
```

```{r}
# Table of results for quick check
kable(errors_FuelsDuffLitt, "pipe")

# Save as CSV or XLSX
#path_errors
#write.csv(errors_FuelsDuffLitt, paste0(path_errors, "errors_GRCA_I&M_FuelsDuffLitt.csv"), quote=FALSE, row.names = FALSE, na = "")
#write_xlsx(errors_FuelsDuffLitt, paste0(path_errors, "errors_GRCA_I&M_FuelsDuffLitt.xlsx"))
```

## FINE FUELS

This code conducts quality control checks on surface fuels data within the surface fuels fine (FWD) data set.

It checks for: header info (number of transects, length of transects for One/Ten/Hundred Hour), fuel constants, transect numbers, reasonable azimuth numbers, reasonable slope numbers, reasonable One/Ten/Hundred Hour count values.

### Fuel Fine Header

[Problem:]{.underline} Incorrect header information entered for I&M plots.

[Procedure:]{.underline}

-   Check that NumTran = 3

```{r}
# Set parameters  
data <- FuelsFine_header
query <- "Fuel Fine Header NumTran" 
query_message <- "Number of Transects"  
values_data <- data$NumTran  
values_valid <- 3    
values_check <- values_data == values_valid

# Identify errors
errors_FuelsFine_NumTran <- qaqc(data, query, query_message, values_check)
```

-   Check that OneHrTranLen = 6

```{r}
# Set parameters  
data <- FuelsFine_header
query <- "Fuel Fine Header OneHrTranLen" 
query_message <- "One Hour Transect Length"  
values_data <- data$OneHrTranLen
values_valid <- 6
values_check <- values_data == values_valid

# Identify errors
errors_FuelsFine_OneHrTranLen <- qaqc(data, query, query_message, values_check)
```

-   Check that TenHrTranLen = 6

```{r}
# Set parameters  
data <- FuelsFine_header
query <- "Fuel Fine Header TenHrTranLen" 
query_message <- "Ten Hour Transect Length"  
values_data <- data$TenHrTranLen 
values_valid <- 6
values_check <- values_data == values_valid

# Identify errors
errors_FuelsFine_TenHrTranLen <- qaqc(data, query, query_message, values_check)
```

-   Check that HunHrTranLen = 12

```{r}
# Set parameters  
data <- FuelsFine_header
query <- "Fuel Fine Header HunHrTranLen" 
query_message <- "Hundrend Hour Transect Length"  
values_data <- data$HunHrTranLen 
values_valid <- 12
values_check <- values_data == values_valid

# Identify errors
errors_FuelsFine_HunHrTranLen <- qaqc(data, query, query_message, values_check)
```

### Fuel Fine FuelConst

[Problem:]{.underline} Incorrect fine woody debris fuel constant entered. Fuel constant name should correspond to the Monitoring Type.

[Procedure:]{.underline}

-   Check that FWDFuConSt = Monitoring Type

```{r}
# Set parameters 
data <- FuelsFine_data
query <- "Fuel Fine FuelConst" 
query_message <- "Fuel Constant" 
values_data <- data$FWDFuConSt 
values_valid <- data$MonitoringType  
values_check <- values_data == values_valid

# Identify errors
errors_FuelsFine_FuelConst <- qaqc(data, query, query_message, values_check)
```

### Fuel Fine Hits

[Problem:]{.underline} Extra or missing data points in I&M plots. There should be 3 transects with 1 sample point per transect. 

[Procedure:]{.underline}

-   Group by Macroplot. Calculate number of hits per plot. Check that sum_hits = 3

```{r}
# Filter NA values. Add number of hits and store in new column
data_temp1 <- FuelsFine_data %>%
  group_by(MacroPlot.Name) %>% 
  dplyr::summarize("sum_hits" = n())

data_temp2 <- FuelsFine_header %>% 
  mutate(MacroPlot.Name = unique(MacroPlot.Name)) %>% 
  select(MacroPlot.Name, Date)

data_temp <- merge(data_temp1, data_temp2, by = "MacroPlot.Name", all = TRUE)
remove(data_temp1, data_temp2)

# Set parameters 
data <- merge(FuelsFine_data, data_temp, by = c("MacroPlot.Name", "Date"), all = T)
query <- "Fuel Fine Hits" 
query_message <- "Number of Hits" 
values_data <- data$sum_hits
values_valid <- 3
values_check <- values_data == values_valid

# Identify errors
errors_FuelsFine_Hits <- qaqc(data, query, query_message, values_check)
```

### Fuel Fine Values

[Problem:]{.underline}  Values for transect, azimuth, slope, and counts for One, Ten, Hundred hour fuels are missing or unreasonable.

[Procedure:]{.underline}

-   Check that Transect = 1, 2, or 3

```{r}
# Set parameters
data <- FuelsFine_data
query <- "Fuel Fine Values TransNum"
query_message <- "Transect Number"
values_data <- data$Transect
values_valid <- c(1, 2, 3)
values_check <- values_data %in% values_valid

# Identify errors
errors_FuelsFine_TransNum <- qaqc(data, query, query_message, values_check)
```

-   Check that Azimuth values are reasonable.

```{r}
# Set parameters
data <- FuelsFine_data
query <- "Fuel Fine Values Azimuth"
query_message <- "Azimuth"
values_data <- data$Azimuth
values_valid <- seq(0, 359, by = 1)
values_check <- values_data %in% values_valid

# Identify errors
errors_FuelsFine_Azimuth <- qaqc(data, query, query_message, values_check)
```

-   Check that Slope values are reasonable.

```{r}
# Set parameters
data <- FuelsFine_data
query <- "Fuel Fine Values Slope"
query_message <- "Slope"
values_data <- data$Slope
values_valid <- seq(0, 90, by = 1)
values_check <- values_data %in% values_valid

# Identify errors
errors_FuelsFine_Slope <- qaqc(data, query, query_message, values_check)
```

-   Check that One Hour values are reasonable.

```{r}
# Perform Rosner's Test for outliers using all values
data_temp <- FuelsFine_all %>%
  filter(!is.na(OneHr))
data_outlier <- rosnerTest(data_temp$OneHr)
outliers <- data_outlier$all.stats$Value[data_outlier$all.stats$Outlier]
max_One <- ifelse(length(outliers) == 0, max(data_temp$OneHr), min(outliers) - 1)

# Set parameters 
data <- FuelsFine_data
query <- "Fuel Fine Values OneHr" 
query_message <- "One Hour Count Outlier" 
values_data <- data$OneHr
values_valid <- seq(0, max_One, by = 1)
values_check <- values_data %in% values_valid

# Identify errors
errors_FuelsFine_OneHr <- qaqc(data, query, query_message, values_check)
```

-   Check that Ten Hour values are reasonable.

```{r}
# Perform Rosner's Test for outliers using all values
data_temp <- FuelsFine_all %>%
  filter(!is.na(TenHr))
data_outlier <- rosnerTest(data_temp$TenHr)
outliers <- data_outlier$all.stats$Value[data_outlier$all.stats$Outlier]
max_Ten <- ifelse(length(outliers) == 0, max(data_temp$TenHr), min(outliers) - 1)

# Set parameters 
data <- FuelsFine_data
query <- "Fuel Fine Values TenHr" 
query_message <- "Ten Hour Count Outlier" 
values_data <- data$TenHr
values_valid <- seq(0, max_Ten, by = 1)
values_check <- values_data %in% values_valid

# Identify errors
errors_FuelsFine_TenHr <- qaqc(data, query, query_message, values_check)
```

-   Check that Hundred Hour values are reasonable.

```{r}
# Perform Rosner's Test for outliers using all values
data_temp <- FuelsFine_all %>%
  filter(!is.na(HunHr))
data_outlier <- rosnerTest(data_temp$HunHr)
outliers <- data_outlier$all.stats$Value[data_outlier$all.stats$Outlier]
max_Hun <- ifelse(length(outliers) == 0, max(data_temp$HunHr), min(outliers) - 1)

# Set parameters 
data <- FuelsFine_data
query <- "Fuel Fine Values HunHr" 
query_message <- "Hundred Hour Count Outlier" 
values_data <- data$HunHr
values_valid <- seq(0, max_Hun, by = 1)
values_check <- values_data %in% values_valid

# Identify errors
errors_FuelsFine_HunHr <- qaqc(data, query, query_message, values_check)
```

### Fuel Fine Errors

```{r}
# Save to master error list 
errors_FuelsFine <- unique(rbind(errors_FuelsFine_FuelConst, errors_FuelsFine_NumTran, errors_FuelsFine_OneHrTranLen, errors_FuelsFine_TenHrTranLen, errors_FuelsFine_HunHrTranLen, errors_FuelsFine_Hits, errors_FuelsFine_TransNum, errors_FuelsFine_Azimuth, errors_FuelsFine_Slope, errors_FuelsFine_OneHr, errors_FuelsFine_TenHr, errors_FuelsFine_HunHr))

remove(errors_FuelsFine_FuelConst, errors_FuelsFine_NumTran, errors_FuelsFine_OneHrTranLen,errors_FuelsFine_TenHrTranLen, errors_FuelsFine_HunHrTranLen, errors_FuelsFine_Hits, errors_FuelsFine_TransNum, errors_FuelsFine_Azimuth, errors_FuelsFine_Slope, errors_FuelsFine_OneHr, errors_FuelsFine_TenHr, errors_FuelsFine_HunHr)
```

```{r}
# Table of results for quick check
kable(errors_FuelsFine, "pipe")

# Save as CSV or XLSX 
#path_errors
#write.csv(errors_FuelsFine, paste0(path_errors, "errors_GRCA_I&M_FuelsFine.csv"), quote=FALSE, row.names = FALSE, na = "") 
#write_xlsx(errors_FuelsFine, paste0(path_errors, "errors_GRCA_I&M_FuelsFine.xlsx"))
```

# PROTOCOL - POST BURN

This code conducts quality control checks on post burn data within the post burn severity data set.

It checks for: header info (number of transects, length of transects, number of points per transect), substrate and vegetation values.

### PBSev Header

[Problem:]{.underline} Incorrect header information entered for I&M plots.

[Procedure:]{.underline}

-   Check that NumTran = 3

```{r}
# Set parameters
data <- PostBurn_header
query <- "PBSev Header NumTran"
query_message <- "Number of Transects"
values_data <- data$NumTran
values_valid <- 3
values_check <- values_data == values_valid

# Identify errors
errors_PostBurn_NumTran <- qaqc(data, query, query_message, values_check)
```

-   Check that TranLen = 50

```{r}
# Set parameters
data <- PostBurn_header
query <- "PBSev Header TranLen"
query_message <- "Transect Length"
values_data <- data$TranLen
values_valid <- 50
values_check <- values_data == values_valid

# Identify errors
errors_PostBurn_TranLen <- qaqc(data, query, query_message, values_check)
```

-   Check that NumPtsTran = 10

```{r}
# Set parameters
data <- PostBurn_header
query <- "PBSev Header NumPtsTran"
query_message <- "Number of Points"
values_data <- data$NumPtsTran
values_valid <- 10
values_check <- values_data == values_valid

# Identify errors
errors_PostBurn_NumPtsTran <- qaqc(data, query, query_message, values_check)
```

### PBSev Hits

[Problem:]{.underline} Extra or missing data points in I&M plots. There should be 3 transects with 10 sample points per transect.

[Procedure:]{.underline}

-   Group by Macroplot. Calculate number of hits per plot. Check that sum_hits = 30

```{r}
# Filter NA values. Add number of hits and store in new column
data_temp <- PostBurn_data %>%
  filter(!is.na(Point)) %>%
  group_by(MacroPlot.Name) %>%
  dplyr::summarize("sum_hits" = n())

# Set parameters
data <- merge(PostBurn_data, data_temp, all.x = T)
query <- "PBSev Hits"
query_message <- "Number of Hits"
values_data <- data$sum_hits
values_valid <- 30
values_check <- values_data == values_valid

# Identify errors
errors_PostBurn_Hits <- qaqc(data, query, query_message, values_check)
```

### PBSev Substrate

[Problem:]{.underline} Incorrect values entered for substrate severity.

[Procedure:]{.underline}

-   Check that Sub = 0, 1, 2, 3, 4, or 5

```{r}
# Set parameters
data <- PostBurn_data
query <- "PBSev Substrate"
query_message <- "Substrate Severity"
values_data <- data$Sub
values_valid <- c(0, 1, 2, 3, 4, 5)
values_check <- values_data %in% values_valid

# Identify errors
errors_PostBurn_Sub <- qaqc(data, query, query_message, values_check)
```

### PBSev Veg

[Problem:]{.underline} Incorrect values entered for vegetation severity.

[Procedure:]{.underline}

-   Check that Veg = 0, 1, 2, 3, 4, or 5

```{r}
# Set parameters
data <- PostBurn_data
query <- "PBSev Veg"
query_message <- "Vegetation Severity"
values_data <- data$Veg
values_valid <- c(0, 1, 2, 3, 4, 5)
values_check <- values_data %in% values_valid

# Identify errors
errors_PostBurn_Veg <- qaqc(data, query, query_message, values_check)
```

### PBSev Errors

```{r}
# Save to master error list
errors_PostBurn <- unique(rbind(errors_PostBurn_NumTran, errors_PostBurn_TranLen, errors_PostBurn_NumPtsTran, errors_PostBurn_Hits, errors_PostBurn_Sub, errors_PostBurn_Veg))

remove(errors_PostBurn_NumTran, errors_PostBurn_TranLen, errors_PostBurn_NumPtsTran, errors_PostBurn_Hits, errors_PostBurn_Sub, errors_PostBurn_Veg)
```

```{r}
# Table of results for quick check
kable(errors_PostBurn, "pipe")

# Save as CSV or XLSX
#path_errors
#write.csv(errors_PostBurn, paste0(path_errors, "errors_GRCA_I&M_PostBurn.csv"), quote=FALSE, row.names = FALSE, na = "")
#write_xlsx(errors_PostBurn, paste0(path_errors, "errors_GRCA_I&M_PostBurn.xlsx"))
```

# PROTOCOL - TREES

### Trees Header

[Problem:]{.underline} Incorrect header information entered for I&M plots.

[Procedure:]{.underline}

-   Check that MacroPlotSize = 0.1

```{r}
# Set parameters  
data <- Trees_header
query <- "Trees Header MacroPlotSize" 
query_message <- "Macro Plot Size"  
values_data <- data$MacroPlotSize  
values_valid <- 0.1    
values_check <- values_data == values_valid

# Identify errors
errors_Trees_MacroPlotSize <- qaqc(data, query, query_message, values_check)
```

-   Check that SnagPlotSize = 0.1

```{r}
# Set parameters  
data <- Trees_header
query <- "Trees Header SnagPlotSize" 
query_message <- "Snag Plot Size"  
values_data <- data$SnagPlotSize
values_valid <- 0.1    
values_check <- values_data == values_valid

# Identify errors
errors_Trees_SnagPlotSize <- qaqc(data, query, query_message, values_check)
```

-   Check that BrkPntDia = 15 or 10

```{r}
# Set parameters  
data <- Trees_header
query <- "Trees Header BrkPntDia" 
query_message <- "Break Point Diameter"  
values_data <- data$BrkPntDia
values_valid <- c(15, 10)
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_BrkPntDia <- qaqc(data, query, query_message, values_check)
```

#### Combine Header Errors

```{r}
# Save to master error list 
errors_Trees_Header <- unique(rbind(errors_Trees_MacroPlotSize, errors_Trees_SnagPlotSize, errors_Trees_BrkPntDia))

remove(errors_Trees_MacroPlotSize, errors_Trees_SnagPlotSize, errors_Trees_BrkPntDia)
```

### Trees CBH

#### Trees CBH Live

[Problem]{.underline}: Live trees do not have live crown base height entered.

[Procedure:]{.underline}

-   Sort by Status. If Status = L, then check that LiCrBHt is [not]{.underline} blank

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Status == "L")
data$LiCrBHt[is.na(data$LiCrBHt)] <- ""
query <- "Trees CBH Live" 
query_message <- "Live tree. Live Crown Base Height"  
values_data <- data$LiCrBHt
values_valid <- ""
values_check <- values_data != values_valid

# Identify errors
errors_Trees_CBH_Live <- qaqc(data, query, query_message, values_check)
```

#### Trees CBH Dead

[Problem]{.underline}: Dead trees have live crown base height entered.

[Procedure:]{.underline}

-   Sort by Status. If Status = D, then check that LiCrBHt = blank

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Status == "D")
data$LiCrBHt[is.na(data$LiCrBHt)] <- ""
query <- "Trees CBH Dead" 
query_message <- "Dead tree. Live Crown Base Height"  
values_data <- data$LiCrBHt
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_CBH_Dead <- qaqc(data, query, query_message, values_check)
```

#### Trees CBH vs Ht

[Problem]{.underline}: Check that Live Crown Base Height is shorter than total tree Height

[Procedure:]{.underline}

-   Check that LiCrBHt \< Ht

```{r}
# Set parameters
data <- Trees_data %>% 
  filter(!is.na(LiCrBHt), Status == "L")
query <- "Trees CBH vs Ht" 
query_message <- "Height shorter than CBH. CBH"  
values_data <- data$LiCrBHt
values_valid <- data$Ht
values_check <- values_data < values_valid

# Identify errors
errors_Trees_CBH_Ht <- qaqc(data, query, query_message, values_check)
```

#### Combine CBH Errors

```{r}
# Combine
errors_Trees_CBH <- unique(rbind(errors_Trees_CBH_Live, errors_Trees_CBH_Dead, errors_Trees_CBH_Ht))

remove(errors_Trees_CBH_Live, errors_Trees_CBH_Dead, errors_Trees_CBH_Ht)
```

### Trees Crown Class

[Problem]{.underline}: Incompatible values entered for tree crown class.

[Procedure:]{.underline}

-   Filter by Status = D and DBH = blank. Check that CrwnCl = BBD, CUS, or DD

```{r}
# Set parameters  
data <- Trees_data %>% 
  filter(is.na(DBH), Status == "D")
query <- "Trees Crown Class Dead" 
query_message <- "Crown Class"  
values_data <- data$CrwnCl
values_valid <- c("DD", "BBD", "CUS")
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_CrwnCl <- qaqc(data, query, query_message, values_check)
```

### Trees Damage

#### Trees Damage Live

[Problem]{.underline}: Tree damage codes are entered incorrectly.

[Procedure:]{.underline}

-   Sort by DamCd 1. Check that DamCd 1 corresponds to one of the choices on the I&M-1 overstory data sheets.

```{r}
# Make sure damage codes are blank vs NA
Trees_data$DamCd1[is.na(Trees_data$DamCd1)] <- ""
Trees_data$DamCd2[is.na(Trees_data$DamCd2)] <- ""
Trees_data$DamCd3[is.na(Trees_data$DamCd3)] <- ""
Trees_data$DamCd4[is.na(Trees_data$DamCd4)] <- ""
Trees_data$DamCd5[is.na(Trees_data$DamCd5)] <- ""
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Status == "L")
query <- "Trees Damage Live" 
query_message <- "Damage Code 1"  
values_data <- data$DamCd1
values_valid <- c("", "ABGR", "BIRD", "BLIG", "BROK", "BROM", "BURL", "CONK", "CROK", "DTOP", "EPIC", "EPIP", "FIRE", "FORK", "FRST", "GALL", "HOLW", "INSE", "LEAN", "LICH", "LIGT", "MAMM", "MISL", "MOSS", "OZON", "ROOT", "ROTT", "SPAR", "SPRT", "TWIN", "UMAN", "WOND")
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_L_1 <- qaqc(data, query, query_message, values_check)
```

-   Sort by DamCd 2. Check that DamCd 2 corresponds to one of the choices on the I&M-1 overstory data sheets.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Status == "L")
query <- "Trees Damage Live" 
query_message <- "Damage Code 2"  
values_data <- data$DamCd2
values_valid <- c("", "ABGR", "BIRD", "BLIG", "BROK", "BROM", "BURL", "CONK", "CROK", "DTOP", "EPIC", "EPIP", "FIRE", "FORK", "FRST", "GALL", "HOLW", "INSE", "LEAN", "LICH", "LIGT", "MAMM", "MISL", "MOSS", "OZON", "ROOT", "ROTT", "SPAR", "SPRT", "TWIN", "UMAN", "WOND")
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_L_2 <- qaqc(data, query, query_message, values_check)
```

-   Sort by DamCd 3. Check that DamCd 3 corresponds to one of the choices on the I&M-1 overstory data sheets.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Status == "L")
query <- "Trees Damage Live" 
query_message <- "Damage Code 3"  
values_data <- data$DamCd3
values_valid <- c("", "ABGR", "BIRD", "BLIG", "BROK", "BROM", "BURL", "CONK", "CROK", "DTOP", "EPIC", "EPIP", "FIRE", "FORK", "FRST", "GALL", "HOLW", "INSE", "LEAN", "LICH", "LIGT", "MAMM", "MISL", "MOSS", "OZON", "ROOT", "ROTT", "SPAR", "SPRT", "TWIN", "UMAN", "WOND")
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_L_3 <- qaqc(data, query, query_message, values_check)
```

-   Sort by DamCd 4. Check that DamCd 4 corresponds to one of the choices on the I&M-1 overstory data sheets.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Status == "L")
query <- "Trees Damage Live" 
query_message <- "Damage Code 4"  
values_data <- data$DamCd4
values_valid <- c("", "ABGR", "BIRD", "BLIG", "BROK", "BROM", "BURL", "CONK", "CROK", "DTOP", "EPIC", "EPIP", "FIRE", "FORK", "FRST", "GALL", "HOLW", "INSE", "LEAN", "LICH", "LIGT", "MAMM", "MISL", "MOSS", "OZON", "ROOT", "ROTT", "SPAR", "SPRT", "TWIN", "UMAN", "WOND")
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_L_4 <- qaqc(data, query, query_message, values_check)
```

-   Sort by DamCd 5. Check that DamCd 5 corresponds to one of the choices on the I&M-1 overstory data sheets.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Status == "L")
query <- "Trees Damage Live" 
query_message <- "Damage Code 5"  
values_data <- data$DamCd5
values_valid <- c("", "ABGR", "BIRD", "BLIG", "BROK", "BROM", "BURL", "CONK", "CROK", "DTOP", "EPIC", "EPIP", "FIRE", "FORK", "FRST", "GALL", "HOLW", "INSE", "LEAN", "LICH", "LIGT", "MAMM", "MISL", "MOSS", "OZON", "ROOT", "ROTT", "SPAR", "SPRT", "TWIN", "UMAN", "WOND")
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_L_5 <- qaqc(data, query, query_message, values_check)
```

#### Trees Damage Dead

[Problem]{.underline}: Damage codes entered for trees that are dead.

[Procedure:]{.underline}

-   Sort by Status. If Status = D, check that DamCd1, DamCd2, DamCd3, DamCd4, [and]{.underline} DamCd5 are blank.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Status == "D")
query <- "Trees Damage Dead" 
query_message <- "Damage Code 1"  
values_data <- data$DamCd1
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_D_1 <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Status == "D")
query <- "Trees Damage Dead" 
query_message <- "Damage Code 2"  
values_data <- data$DamCd2
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_D_2 <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Status == "D")
query <- "Trees Damage Dead" 
query_message <- "Damage Code 3"  
values_data <- data$DamCd3
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_D_3 <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Status == "D")
query <- "Trees Damage Dead" 
query_message <- "Damage Code 4"  
values_data <- data$DamCd4
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_D_4 <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Status == "D")
query <- "Trees Damage Dead" 
query_message <- "Damage Code 5"  
values_data <- data$DamCd5
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_D_5 <- qaqc(data, query, query_message, values_check)
```

#### Trees Damage Blank

[Problem]{.underline}: Tree damage codes are entered incorrectly. Blank damages exist before entered damages

[Procedure:]{.underline}

-   Filter for DamCd1 = blank. Check that DamCd2, DamCd3, DamCd4, [and]{.underline} DamCd5 are blank.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(DamCd1 == "")
query <- "Trees Damage Blank" 
query_message <- "Damage Code 1 = Blank, Damage Code 2"  
values_data <- data$DamCd2
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_1_Blank_2 <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(DamCd1 == "")
query <- "Trees Damage Blank" 
query_message <- "Damage Code 1 = Blank, Damage Code 3"  
values_data <- data$DamCd3
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_1_Blank_3 <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(DamCd1 == "")
query <- "Trees Damage Blank" 
query_message <- "Damage Code 1 = Blank, Damage Code 4"  
values_data <- data$DamCd4
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_1_Blank_4 <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(DamCd1 == "")
query <- "Trees Damage Blank" 
query_message <- "Damage Code 1 = Blank, Damage Code 5"  
values_data <- data$DamCd5
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_1_Blank_5 <- qaqc(data, query, query_message, values_check)
```

-   Filter for DamCd2 = blank. Check that DamCd3, DamCd4, [and]{.underline} DamCd5 are blank.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(DamCd2 == "")
query <- "Trees Damage Blank" 
query_message <- "Damage Code 2 = Blank, Damage Code 3"  
values_data <- data$DamCd3
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_2_Blank_3 <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(DamCd2 == "")
query <- "Trees Damage Blank" 
query_message <- "Damage Code 2 = Blank, Damage Code 4"  
values_data <- data$DamCd4
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_2_Blank_4 <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(DamCd2 == "")
query <- "Trees Damage Blank" 
query_message <- "Damage Code 2 = Blank, Damage Code 5"  
values_data <- data$DamCd5
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_2_Blank_5 <- qaqc(data, query, query_message, values_check)
```

-   Filter for DamCd3 = blank. Check that DamCd4 [and]{.underline} DamCd5 are blank.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(DamCd3 == "")
query <- "Trees Damage Blank" 
query_message <- "Damage Code 3 = Blank, Damage Code 4"  
values_data <- data$DamCd4
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_3_Blank_4 <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(DamCd3 == "")
query <- "Trees Damage Blank" 
query_message <- "Damage Code 3 = Blank, Damage Code 5"  
values_data <- data$DamCd5
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_3_Blank_5 <- qaqc(data, query, query_message, values_check)
```

-   Filter for DamCd4 = blank. Check that DamCd5 is blank.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(DamCd4 == "")
query <- "Trees Damage Blank" 
query_message <- "Damage Code 4 = Blank, Damage Code 5"  
values_data <- data$DamCd5
values_valid <- ""
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_DamCd_4_Blank_5 <- qaqc(data, query, query_message, values_check)
```

#### Combine Damage Errors

```{r}
# Damage Errors Live Trees
errors_Trees_DamCd_L <- unique(rbind(errors_Trees_DamCd_L_1, errors_Trees_DamCd_L_2, errors_Trees_DamCd_L_3, errors_Trees_DamCd_L_4, errors_Trees_DamCd_L_5))

remove(errors_Trees_DamCd_L_1, errors_Trees_DamCd_L_2, errors_Trees_DamCd_L_3, errors_Trees_DamCd_L_4, errors_Trees_DamCd_L_5)

# Damage Errors Dead Trees
errors_Trees_DamCd_D <- unique(rbind(errors_Trees_DamCd_D_1, errors_Trees_DamCd_D_2, errors_Trees_DamCd_D_3, errors_Trees_DamCd_D_4, errors_Trees_DamCd_D_5))

remove(errors_Trees_DamCd_D_1, errors_Trees_DamCd_D_2, errors_Trees_DamCd_D_3, errors_Trees_DamCd_D_4, errors_Trees_DamCd_D_5)

# Damage Errors Blank Entries
errors_Trees_DamCd_Blank <- unique(rbind(errors_Trees_DamCd_1_Blank_2, errors_Trees_DamCd_1_Blank_3, errors_Trees_DamCd_1_Blank_4, errors_Trees_DamCd_1_Blank_5, errors_Trees_DamCd_2_Blank_3, errors_Trees_DamCd_2_Blank_4, errors_Trees_DamCd_2_Blank_5, errors_Trees_DamCd_3_Blank_4, errors_Trees_DamCd_3_Blank_5, errors_Trees_DamCd_4_Blank_5))

remove(errors_Trees_DamCd_1_Blank_2, errors_Trees_DamCd_1_Blank_3, errors_Trees_DamCd_1_Blank_4, errors_Trees_DamCd_1_Blank_5, errors_Trees_DamCd_2_Blank_3, errors_Trees_DamCd_2_Blank_4, errors_Trees_DamCd_2_Blank_5, errors_Trees_DamCd_3_Blank_4, errors_Trees_DamCd_3_Blank_5, errors_Trees_DamCd_4_Blank_5)

# Damage Errors Combined
errors_Trees_DamCd <- unique(rbind(errors_Trees_DamCd_L, errors_Trees_DamCd_D, errors_Trees_DamCd_Blank))

remove(errors_Trees_DamCd_L, errors_Trees_DamCd_D, errors_Trees_DamCd_Blank)
                                  
```

### Trees DBH

[Problem]{.underline}: DBH values are smaller than sampling parameters or excessively large. DBH values for standing trees do not exist.

[Procedure:]{.underline}

-   Filter to exclude DD trees. Check that DBH ≥ 15

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(!CrwnCl %in% c("DD", "BBD", "CUS"))
query <- "Trees DBH Min" 
query_message <- "DBH"  
values_data <- data$DBH
values_valid <- min(15)
values_check <- values_data >= values_valid

# Identify errors
errors_Trees_DBH_Min <- qaqc(data, query, query_message, values_check)
```

-   Filter to exclude DD trees. Check that DBH is not excessively large. Check that DBH values for standing trees exist.

```{r}
# Perform Rosner's Test for outliers using all values
data_temp <- Trees_all %>%
  filter(!is.na(DBH))
data_outlier <- rosnerTest(data_temp$DBH)
outliers <- data_outlier$all.stats$Value[data_outlier$all.stats$Outlier]
max_DBH <- ifelse(length(outliers) == 0, max(data_temp$DBH), min(outliers) - 1)

# Set parameters 
data <- Trees_data %>% 
  filter(!CrwnCl %in% c("DD", "BBD", "CUS"))
query <- "Trees DBH Max" 
query_message <- "DBH Outlier"  
values_data <- data$DBH
values_valid <- max_DBH
values_check <- values_data <= values_valid

# Identify errors
errors_Trees_DBH_Max <- qaqc(data, query, query_message, values_check)
```

#### Combine DBH Errors

```{r}
# Combine
errors_Trees_DBH <- unique(rbind(errors_Trees_DBH_Min, errors_Trees_DBH_Max))

remove(errors_Trees_DBH_Min, errors_Trees_DBH_Max)
```

### Trees DD

[Problem]{.underline}: DBH, height, and or crown base height entered for trees that are no longer standing.

[Procedure:]{.underline}

-   Filter for CrwnCl = BBD, CUS, or DD. Check that DBH, Ht, [and]{.underline} LiCrBHt are blank

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(CrwnCl %in% c("DD", "BBD", "CUS"))
data$DBH[is.na(data$DBH)] <- ""
query <- "Trees DD DBH" 
query_message <- "DBH"  
values_data <- data$DBH
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_DD_DBH <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(CrwnCl %in% c("DD", "BBD", "CUS"))
data$Ht[is.na(data$Ht)] <- ""
query <- "Trees DD Ht" 
query_message <- "Height"  
values_data <- data$Ht
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_DD_Ht <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(CrwnCl %in% c("DD", "BBD", "CUS"))
data$LiCrBHt[is.na(data$LiCrBHt)] <- ""
query <- "Trees DD CBH" 
query_message <- "Live Crown Base Height"  
values_data <- data$LiCrBHt
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_DD_LiCrBHt <- qaqc(data, query, query_message, values_check)
```

#### Combine DD Errors

```{r}
# Combine
errors_Trees_DD <- unique(rbind(errors_Trees_DD_DBH, errors_Trees_DD_Ht, errors_Trees_DD_LiCrBHt))

remove(errors_Trees_DD_DBH, errors_Trees_DD_Ht, errors_Trees_DD_LiCrBHt)
```

### Trees Duplicates

[Problem]{.underline}: Duplicate tree tag numbers entered for a plot visit.

[Procedure:]{.underline}

-   Sort by RowCount. Check that RowCount = 1 for all entries.

```{r}
# Check for tag duplication (TRUE/FALSE) and store results in new column
data_temp <- Trees_data %>%
  group_by(MacroPlot.Name, TagNo) %>% 
  filter(!TagNo == 0) %>% 
  mutate("Duplicate" = duplicated(TagNo))

# Set parameters 
data <- merge(Trees_data, data_temp, all.x = T)
data$Duplicate[is.na(data$Duplicate)] <- ""
query <- "Trees Duplicates" 
query_message <- "Duplicate Tag"  
values_data <- data$Duplicate
values_valid <- c(FALSE, "")
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_Duplicate <- qaqc(data, query, query_message, values_check)
```

### Trees Frac

[Problem:]{.underline} Incorrect subplot fraction entered for trees.

[Procedure:]{.underline}

-   If DBH = overstory, then check that SubFrac = 1

```{r}
# Set parameters 
data <- Trees_data %>% 
   filter(DBH >= 15.0)
query <- "Trees Frac Overstory" 
query_message <- "Subplot Fraction"  
values_data <- data$SubFrac
values_valid <- 1
values_check <- values_data == values_valid

# Identify errors
errors_Trees_Frac <- qaqc(data, query, query_message, values_check)
```

### Trees Height

#### Trees Height Standing

[Problem:]{.underline} Height not recorded. Standing trees should have height values.

[Procedure:]{.underline}

-   Filter for standing trees. Check that Ht exists.

```{r}
# Set parameters 
data <- Trees_data %>% 
   filter(!CrwnCl %in% c("DD", "BBD", "CUS"))
data$Ht[is.na(data$Ht)] <- ""
query <- "Trees Height Standing" 
query_message <- "Standing tree. Height"  
values_data <- data$Ht
values_valid <- ""
values_check <- values_data != values_valid

# Identify errors
errors_Trees_Height_Standing <- qaqc(data, query, query_message, values_check)
```

#### Trees Height Overstory

[Problem:]{.underline} Incorrect values entered for height. Overstory trees should have reasonable height values.

[Procedure:]{.underline}

-   Filter for Overstory trees. Check that Ht is not excessively large.

```{r}
# Perform Rosner's Test for outliers using all values
data_temp <- Trees_all %>%
  filter(!is.na(Ht), DBH >= 15.0)
data_outlier <- rosnerTest(data_temp$Ht)
outliers <- data_outlier$all.stats$Value[data_outlier$all.stats$Outlier]
max_Ht <- ifelse(length(outliers) == 0, max(data_temp$Ht), min(outliers) - 1)

# Set parameters 
data <- Trees_data %>% 
   filter(!is.na(Ht), DBH >= 15.0)
query <- "Trees Height Overstory" 
query_message <- "Height Outlier"  
values_data <- data$Ht
values_valid <- max_Ht
values_check <- values_data <= values_valid

# Identify errors
errors_Trees_Height_Overstory <- qaqc(data, query, query_message, values_check)
```

#### Combine Height Errors

```{r}
# Combine
errors_Trees_Height <- unique(rbind(errors_Trees_Height_Standing, errors_Trees_Height_Overstory))

remove(errors_Trees_Height_Standing, errors_Trees_Height_Overstory)
```

### Trees Life Form

[Problem:]{.underline} Species entered that are not trees.

[Procedure:]{.underline}

-   Check that Preferred Lifeform = Tree

```{r}
# Set parameters 
data <- Trees_data
query <- "Trees Life Form" 
query_message <- "Life Form"  
values_data <- data$Preferred_Lifeform
values_valid <- "Tree"
values_check <- values_data == values_valid

# Identify errors
errors_Trees_LifeForm <- qaqc(data, query, query_message, values_check)
```

### Trees Non-Post Severity

[Problem]{.underline}: Scorch height, scorch %, and/or char height are entered in sample events other than immediate post for I&M plots. These values should only be collected at immediate post reads.

[Procedure:]{.underline}

-   Filter to exclude all POST reads (01Post, 02Post, etc). Check that CharHt = blank, ScorchHt = blank, [and]{.underline} CrScPct = blank (i.e. no records should appear in the query)

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(!Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"))
data$CharHt[is.na(data$CharHt)] <- ""
query <- "Trees Non-Post Severity CharHt" 
query_message <- "Char Height"  
values_data <- data$CharHt
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_NonPost_Severity_CharHt <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(!Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"))
data$ScorchHt[is.na(data$ScorchHt)] <- ""
query <- "Trees Non-Post Severity ScorchHt" 
query_message <- "Scorch Height"  
values_data <- data$ScorchHt
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_NonPost_Severity_ScorchHt <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(!Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"))
data$CrScPct[is.na(data$CrScPct)] <- ""
query <- "Trees Non-Post Severity CrScPct" 
query_message <- "Crown Scorch Percent"  
values_data <- data$CrScPct
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_NonPost_Severity_CrScPct <- qaqc(data, query, query_message, values_check)
```

#### Combine Non-Post Errors

```{r}
# Combine
errors_Trees_NonPost_Severity <- unique(rbind(errors_Trees_NonPost_Severity_CharHt, errors_Trees_NonPost_Severity_ScorchHt, errors_Trees_NonPost_Severity_CrScPct))

remove(errors_Trees_NonPost_Severity_CharHt, errors_Trees_NonPost_Severity_ScorchHt, errors_Trees_NonPost_Severity_CrScPct)
```

### Trees Post

#### Trees Post Char

[Problem]{.underline}: Char height missing or excessively large for standing trees during the immediate post-fire sample events in I&M plots. Char height entered for trees that are no longer standing.

[Procedure:]{.underline}

-   DD Trees. Filter to include all POST reads (01Post, 02Post, etc). Filter for DD crown classes (DD, BBD, and CUS). Check that CharHt = blank.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"),
         CrwnCl %in% c("DD", "BBD", "CUS"))
data$CharHt[is.na(data$CharHt)] <- ""
query <- "Trees Post Char DD" 
query_message <- "Char Height for DD Tree"  
values_data <- data$CharHt
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_Post_Char_DD <- qaqc(data, query, query_message, values_check)
```

-   Standing Dead Trees. Filter to include all POST reads (01Post, 02Post, etc). Filter to exclude DD crown classes (DD, BBD, and CUS). Filter for Status = D. Check that CharHt is [not]{.underline} blank.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"),
         !CrwnCl %in% c("DD", "BBD", "CUS"),
         Status == "D",
         # Blank Char Ht can have tag 0
         !TagNo == 0)
data$CharHt[is.na(data$CharHt)] <- ""
query <- "Trees Post Char Dead" 
query_message <- "Char Height for Standing Dead"  
values_data <- data$CharHt
values_valid <- ""
values_check <- values_data != values_valid

# Identify errors
errors_Trees_Post_Char_Dead <- qaqc(data, query, query_message, values_check)
```

-   Live trees. Filter to include all POST reads (01Post, 02Post, etc). Filter for Status = L. Check that CharHt is not blank or excessively high.

```{r}
# Perform Rosner's Test for outliers using all values
data_temp <- Trees_all %>%
  filter(Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"),
         !is.na(CharHt))
data_outlier <- rosnerTest(data_temp$CharHt)
outliers <- data_outlier$all.stats$Value[data_outlier$all.stats$Outlier]
max_Char <- ifelse(length(outliers) == 0, max(data_temp$CharHt), min(outliers) - 1)

# Set parameters 
data <- Trees_data %>% 
  filter(Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"),
         Status == "L")
query <- "Trees Post Char Live" 
query_message <- "Char Height Outlier"  
values_data <- data$CharHt
values_valid <- max_Char
values_check <- values_data <= values_valid

# Identify errors
errors_Trees_Post_Char_Live <- qaqc(data, query, query_message, values_check)
```

#### Trees Post ScorchHt

[Problem]{.underline}: Scorch height missing or excessively large for live trees during the immediate post-fire sample events in I&M plots. Scorch height entered for trees that are no longer standing or are broken.

[Procedure:]{.underline}

-   DD Trees. Filter to include all POST reads (01Post, 02Post, etc). Filter for DD crown classes (DD, BBD, and CUS). Check that ScorchHt = blank.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"),
         CrwnCl %in% c("DD", "BBD", "CUS"))
data$ScorchHt[is.na(data$ScorchHt)] <- ""
query <- "Trees Post Scorch DD" 
query_message <- "Scorch Height for DD Tree"  
values_data <- data$ScorchHt
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_Post_Scorch_DD <- qaqc(data, query, query_message, values_check)
```

-   Standing Dead Trees. Filter to include all POST reads (01Post, 02Post, etc). Filter to exclude DD crown classes (DD, BBD, and CUS). Filter for Status = D. Check that ScorchHt = blank.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"),
         !CrwnCl %in% c("DD", "BBD", "CUS"),
         Status == "D")
data$ScorchHt[is.na(data$ScorchHt)] <- ""
query <- "Trees Post Scorch Dead"
query_message <- "Scorch Height for Standing Dead"  
values_data <- data$ScorchHt
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_Post_Scorch_Dead <- qaqc(data, query, query_message, values_check)
```

-   Live trees. Filter to include all POST reads (01Post, 02Post, etc). Filter for Status = L. Check that ScorchHt is not blank or excessively high.

```{r}
# Perform Rosner's Test for outliers using all values
data_temp <- Trees_all %>%
  filter(Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"),
         !is.na(ScorchHt))
data_outlier <- rosnerTest(data_temp$ScorchHt)
outliers <- data_outlier$all.stats$Value[data_outlier$all.stats$Outlier]
max_Scorch <- ifelse(length(outliers) == 0, max(data_temp$ScorchHt), min(outliers) - 1)

# Set parameters 
data <- Trees_data %>% 
  filter(Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"),
         Status == "L")
query <- "Trees Post Scorch Live" 
query_message <- "Scorch Height Outlier"  
values_data <- data$ScorchHt
values_valid <- max_Scorch
values_check <- values_data <= values_valid

# Identify errors
errors_Trees_Post_Scorch_Live <- qaqc(data, query, query_message, values_check)
```

#### Trees Post Scorch%

[Problem]{.underline}: Scorch % missing or excessively large for live trees during the immediate post-fire sample events in I&M plots. Scorch % entered incorrectly for trees that are no longer standing.

[Procedure:]{.underline}

-   DD Trees. Filter to include all POST reads (01Post, 02Post, etc). Filter for DD crown classes (DD, BBD, and CUS). Check that CrScPct = blank.

-   Check that all rows with CrwnCl = DD have CrScPct = blank or 100. (RS?)

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"),
         CrwnCl %in% c("DD", "BBD", "CUS"))
data$CrScPct[is.na(data$CrScPct)] <- ""
query <- "Trees Post Scorch% DD" 
query_message <- "Scorch Percent for DD Tree"  
values_data <- data$CrScPct
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_Post_ScorchPct_DD <- qaqc(data, query, query_message, values_check)
```

-   Standing Dead Trees. Filter to include all POST reads (01Post, 02Post, etc). Filter to exclude DD crown classes (DD, BBD, and CUS). Filter for Status = D. Check that CrScPct = blank or 100.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"),
         !CrwnCl %in% c("DD", "BBD", "CUS"),
         Status == "D")
data$CrScPct[is.na(data$CrScPct)] <- ""
query <- "Trees Post Scorch% Dead"
query_message <- "Scorch Percent for Standing Dead"  
values_data <- data$CrScPct
values_valid <- c("", 100)
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_Post_ScorchPct_Dead <- qaqc(data, query, query_message, values_check)
```

-   Live trees. Filter to include all POST reads (01Post, 02Post, etc). Filter for Status = L. Check that CrScPct is not blank. Check that CrScPct ≤ 100.

```{r}
# Set parameters 
data <- Trees_data %>% 
  filter(Monitoring.Status %in% c("01Post", "02Post", "03Post", "04Post"),
         Status == "L")
query <- "Trees Post Scorch% Live" 
query_message <- "Scorch Percent"  
values_data <- data$CrScPct
values_valid <- max(100)
values_check <- values_data <= values_valid

# Identify errors
errors_Trees_Post_ScorchPct_Live <- qaqc(data, query, query_message, values_check)
```

#### Combine Post Severity Errors

```{r}
# Combine
errors_Trees_Post_Severity <- unique(rbind(errors_Trees_Post_Char_DD, errors_Trees_Post_Char_Dead, errors_Trees_Post_Char_Live, errors_Trees_Post_Scorch_DD, errors_Trees_Post_Scorch_Dead, errors_Trees_Post_Scorch_Live, errors_Trees_Post_ScorchPct_DD, errors_Trees_Post_ScorchPct_Dead, errors_Trees_Post_ScorchPct_Live))

remove(errors_Trees_Post_Char_DD, errors_Trees_Post_Char_Dead, errors_Trees_Post_Char_Live, errors_Trees_Post_Scorch_DD, errors_Trees_Post_Scorch_Dead, errors_Trees_Post_Scorch_Live, errors_Trees_Post_ScorchPct_DD, errors_Trees_Post_ScorchPct_Dead, errors_Trees_Post_ScorchPct_Live)
```

### Trees Status

[Problem:]{.underline} Species do not have an entry for status or have an entry that we don’t use.

[Procedure:]{.underline}

-   Sort by Status. Check that Status = L or D

```{r}
# Set parameters 
data <- Trees_data
query <- "Trees Status" 
query_message <- "Status"  
values_data <- data$Status
values_valid <- c("L", "D")
values_check <- values_data %in% values_valid

# Identify errors
errors_Trees_Status <- qaqc(data, query, query_message, values_check)
```

### Trees UV

[Problem]{.underline}: X present in the UV columns signifying respective data (DBH, Ht, or LiCrBHT) weren’t assessed.

[Procedure]{.underline}:

-   Check that UV(1, 2, [or]{.underline} 3) is blank for all plots

```{r}
# Set parameters 
data <- Trees_data
data$UV1[is.na(data$UV1)] <- ""
query <- "Trees UV1" 
query_message <- "User Variable 1"  
values_data <- data$UV1
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_UV1 <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data
data$UV2[is.na(data$UV2)] <- ""
query <- "Trees UV2" 
query_message <- "User Variable 2"  
values_data <- data$UV2
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_UV2 <- qaqc(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- Trees_data
data$UV3[is.na(data$UV3)] <- ""
query <- "Trees UV3" 
query_message <- "User Variable 3"  
values_data <- data$UV3
values_valid <- ""
values_check <- values_data == values_valid

# Identify errors
errors_Trees_UV3 <- qaqc(data, query, query_message, values_check)
```

#### Combine UV Errors

```{r}
# Combine
errors_Trees_UV <- unique(rbind(errors_Trees_UV1, errors_Trees_UV2, errors_Trees_UV3))

remove(errors_Trees_UV1, errors_Trees_UV2, errors_Trees_UV3)
```

## Trees Errors

```{r}
errors_Trees <- unique(rbind(errors_Trees_CrwnCl, errors_Trees_DamCd, errors_Trees_DBH, errors_Trees_DD, errors_Trees_CBH, errors_Trees_Duplicate, errors_Trees_Frac, errors_Trees_Header, errors_Trees_Height, errors_Trees_NonPost_Severity, errors_Trees_Post_Severity, errors_Trees_Status, errors_Trees_LifeForm, errors_Trees_UV))

remove(errors_Trees_Header, errors_Trees_CrwnCl, errors_Trees_DamCd, errors_Trees_DBH, errors_Trees_CBH, errors_Trees_DD, errors_Trees_Duplicate, errors_Trees_Frac, errors_Trees_Height, errors_Trees_NonPost_Severity, errors_Trees_Post_Severity, errors_Trees_Status, errors_Trees_LifeForm, errors_Trees_UV)
```

```{r}
# Table of results for quick check
kable(errors_Trees, "pipe")

# Save as CSV or XLSX
#path_errors
#write.csv(errors_Trees, paste0(path_errors, "errors_GRCA_I&M_Trees.csv"), quote=FALSE, row.names = FALSE, na = "")
#write_xlsx(errors_Trees, paste0(path_errors, "errors_GRCA_I&M_Trees.xlsx"))
```

# ALL ERRORS

```{r}
# Save to master error list 
errors_all <- unique(rbind(errors_Fuels1000, errors_FuelsDuffLitt, errors_FuelsFine, errors_PostBurn, errors_Trees))
```

```{r}
# Table of results for quick check
#kable(errors_all, "pipe")

# Save as CSV or XLSX 
path_errors
#write.csv(errors_all, paste0(path_errors, "errors_GRCA_I&M.csv"), quote=FALSE, row.names = FALSE, na = "") 
#write_xlsx(errors_all, paste0(path_errors, "errors_GRCA_I&M.xlsx"))
```
