---
title: "Grand Canyon QAQC for FMH Data"
author: "Alexandra Lalor"
date: "2024-11-06"
output:
  html_document:
    theme: readable
    highlight: 
    toc: yes
    toc_depth: 3
    toc_float:
      smooth_scroll: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Install packages (if needed)
#install.packages("tidyverse")
```

# BEFORE STARTING

### Load Packages

```{r}
# Load packages
library(tidyverse)
library(writexl)
library(dplyr)
library(knitr)
# EnvStats is needed for the rosnerTest() function
library(EnvStats)
```

### Adjust File Paths

Make sure to update file paths to be specific for your data.

```{r}
# Identify working directory (specifically user name)
getwd()

# Load in data and name them based on file path
# Change file path based on user name!
path_data <- "C:/Users/alalor.NPS/OneDrive - DOI/Desktop/R/GRCA/data/qaqc/2024/"
path_errors <- "C:/Users/alalor.NPS/OneDrive - DOI/Desktop/R/GRCA/output/errors/2024/"
#path_data <- "C:/Users/allie/OneDrive/Desktop/R Projects/GRCA/data/qaqc/2024/"
#path_errors <- "C:/Users/allie/OneDrive/Desktop/R Projects/GRCA/output/errors/2024/"
```

### Load Data

```{r}
# Load in data
Fuels1000_all <- read.csv(paste0(path_data, "GRCA_FMH_2024_Surface Fuels - 1000Hr_XPT.csv"))
FuelsDuffLitt_all <- read.csv(paste0(path_data, "GRCA_FMH_2024_Surface Fuels - Duff_Litter_XPT.csv"))
FuelsFine_all <- read.csv(paste0(path_data, "GRCA_FMH_2024_Surface Fuels - Fine_XPT.csv"))
HerbsPoints_all <- read.csv(paste0(path_data, "Cover - Points (metric)_XPT.csv"))
HerbsObs_all <- read.csv(paste0(path_data, "Cover - Species Composition (metric)_XPT.csv"))
Shrubs_all <- read.csv(paste0(path_data, "Density - Belts (metric)_XPT.csv"))
Seedlings_all <- read.csv(paste0(path_data, "Density - Quadrats (metric)_XPT.csv"))
Trees_all <- read.csv(paste0(path_data, "Trees - Individuals (metric)_XPT.csv"))
#PostBurn_all <- read.csv(paste0(path_data, "Post Burn Severity_XPT.csv"))

# Blank data frame for no errors
errors_blank <- data.frame("SavedQuery" = "","MacroPlot.Name" = "","Date" = "","Error" = "No Error", "Comment" = "")
```

### Check Data

Fuels1000

```{r}
# Ensure blanks in Visited column are NA
Fuels1000_all$Visited[Fuels1000_all$Visited==""] <- NA

# Clean up data frame
Fuels1000_all <- Fuels1000_all %>%
  # Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  # Separate MacroPlotName to get Monitoring type (or export data with Monitoring Type already included)
  separate(MacroPlot.Name, sep = " ", into = c("MonitoringType", "Plot"), remove = FALSE)

# Filter for only data
Fuels1000_data <- Fuels1000_all %>%
  filter(is.na(Visited))

# Filter for only headers
Fuels1000_header <- Fuels1000_all %>%
  filter(!is.na(Visited))
```

FuelsDuffLitt

```{r}
# Ensure blanks in Visited column are NA 
FuelsDuffLitt_all$Visited[FuelsDuffLitt_all$Visited==""] <- NA  

# Clean up data frame 
FuelsDuffLitt_all <- FuelsDuffLitt_all %>%   
  # Separate Date column into Date and Time columns   
  separate(Date, sep = " ", into = c("Date", "Time")) %>%   
  # Separate MacroPlotName to get Monitoring type (or export data with Monitoring Type already included)   
  separate(MacroPlot.Name, sep = " ", into = c("MonitoringType", "Plot"), remove = FALSE)

# Filter for only data 
FuelsDuffLitt_data <- FuelsDuffLitt_all %>%   
  filter(is.na(Visited))  

# Filter for only headers 
FuelsDuffLitt_header <- FuelsDuffLitt_all %>%   
  filter(!is.na(Visited))
```

FuelsFine

```{r}
# Ensure blanks in Visited column are NA 
FuelsFine_all$Visited[FuelsFine_all$Visited==""] <- NA  

# Clean up data frame 
FuelsFine_all <- FuelsFine_all %>%   
  # Separate Date column into Date and Time columns   
  separate(Date, sep = " ", into = c("Date", "Time")) %>%   
  # Separate MacroPlotName to get Monitoring type (or export data with Monitoring Type already included)   
  separate(MacroPlot.Name, sep = " ", into = c("MonitoringType", "Plot"), remove = FALSE)

# Filter for only data 
FuelsFine_data <- FuelsFine_all %>%   
  filter(is.na(Visited))  

# Filter for only headers 
FuelsFine_header <- FuelsFine_all %>%   
  filter(!is.na(Visited))
```

HerbsPoints

```{r}
# Ensure blanks in Visited column are NA 
HerbsPoints_all$Visited[HerbsPoints_all$Visited==""] <- NA  

# Clean up data frame 
HerbsPoints_all <- HerbsPoints_all %>%   
  # Separate Date column into Date and Time columns   
  separate(Date, sep = " ", into = c("Date", "Time")) %>%   
  # Separate MacroPlotName to get Monitoring type (or export data with Monitoring Type already included)   
  separate(MacroPlot.Name, sep = " ", into = c("MonitoringType", "Plot"), remove = FALSE) 

# Filter for only data 
HerbsPoints_data <- HerbsPoints_all %>%   
  filter(is.na(Visited))  

# Filter for only headers 
HerbsPoints_header <- HerbsPoints_all %>%   
  filter(!is.na(Visited))
```

HerbsObs

```{r}
# Ensure blanks in Visited column are NA 
HerbsObs_all$Visited[HerbsObs_all$Visited==""] <- NA  

# Clean up data frame 
HerbsObs_all <- HerbsObs_all %>%   
  # Separate Date column into Date and Time columns   
  separate(Date, sep = " ", into = c("Date", "Time")) %>%   
  # Separate MacroPlotName to get Monitoring type (or export data with Monitoring Type already included)   
  separate(MacroPlot.Name, sep = " ", into = c("MonitoringType", "Plot"), remove = FALSE) 

# Filter for only data 
HerbsObs_data <- HerbsObs_all %>%   
  filter(is.na(Visited))  

# Filter for only headers 
HerbsObs_header <- HerbsObs_all %>%   
  filter(!is.na(Visited))
```

Shrubs

```{r}
# Ensure blanks in Visited column are NA 
Shrubs_all$Visited[Shrubs_all$Visited==""] <- NA  

# Clean up data frame 
Shrubs_all <- Shrubs_all %>%   
  # Separate Date column into Date and Time columns   
  separate(Date, sep = " ", into = c("Date", "Time")) %>%   
  # Separate MacroPlotName to get Monitoring type (or export data with Monitoring Type already included)   
  separate(MacroPlot.Name, sep = " ", into = c("MonitoringType", "Plot"), remove = FALSE)

# Filter for only data 
Shrubs_data <- Shrubs_all %>%   
  filter(is.na(Visited))  

# Filter for only headers 
Shrubs_header <- Shrubs_all %>%   
  filter(!is.na(Visited))
```

Seedlings

```{r}
# Ensure blanks in Visited column are NA 
Seedlings_all$Visited[Seedlings_all$Visited==""] <- NA  

# Clean up data frame 
Seedlings_all <- Seedlings_all %>%   
  # Separate Date column into Date and Time columns   
  separate(Date, sep = " ", into = c("Date", "Time")) %>%   
  # Separate MacroPlotName to get Monitoring type (or export data with Monitoring Type already included)   
  separate(MacroPlot.Name, sep = " ", into = c("MonitoringType", "Plot"), remove = FALSE) 

# Filter for only data 
Seedlings_data <- Seedlings_all %>%   
  filter(is.na(Visited))  

# Filter for only headers 
Seedlings_header <- Seedlings_all %>%   
  filter(!is.na(Visited))
```

Trees

```{r}
# Ensure blanks in Visited column are NA 
Trees_all$Visited[Trees_all$Visited==""] <- NA  

# Clean up data frame 
Trees_all <- Trees_all %>%   
  # Separate Date column into Date and Time columns   
  separate(Date, sep = " ", into = c("Date", "Time")) %>%   
  # Separate MacroPlotName to get Monitoring type (or export data with Monitoring Type already included)   
  separate(MacroPlot.Name, sep = " ", into = c("MonitoringType", "Plot"), remove = FALSE) 

# Filter for only data 
Trees_data <- Trees_all %>%   
  filter(is.na(Visited))  

# Filter for only headers 
Trees_header <- Trees_all %>%   
  filter(!is.na(Visited))
```

PostBurn

```{r}
# # Ensure blanks in Visited column are NA 
# PostBurn_all$Visited[PostBurn_all$Visited==""] <- NA  
# 
# # Clean up data frame 
# PostBurn_all <- PostBurn_all %>%   
#   # Separate Date column into Date and Time columns   
#   separate(Date, sep = " ", into = c("Date", "Time")) %>%   
#   # Separate MacroPlotName to get Monitoring type (or export data with Monitoring Type already included)   
#   separate(MacroPlot.Name, sep = " ", into = c("MonitoringType", "Plot"), remove = FALSE)
# 
# # Filter for only data 
# PostBurn_data <- PostBurn_all %>%   
#   filter(is.na(Visited))
# 
# # Filter for only headers 
# PostBurn_header <- PostBurn_all %>%   
#   filter(!is.na(Visited))
```

# FUNCTIONS

Create functions for QAQC

```{r}
qaqc <- function(data, query, query_message, values_check) {
  # Identify errors 
  errors <- data %>%   
    mutate("SavedQuery" = query, 
           "Error" = paste(query_message, "=", values_data)) %>%   
    # Data is equal exactly to a value
    filter(values_check) %>%   
    select("SavedQuery", "MacroPlot.Name", "Date", "Error", "Comment")
  
  # If no errors, add "No Errors" to existing data frame 
  if (nrow(errors) == 0) {   
    errors <- errors_blank %>%     
      mutate(SavedQuery = query) 
  } else {   
    errors <- errors
  }
}
```

# PROTOCOL - SURFACE FUELS {.tabs}

## 1000HR CWD

This code conducts quality control checks on surface fuels data within the surface fuels coarse woody debris (1000-hr) data set.

It checks for: header info (number of transects, length of transects), decay class values, fuel constants, duplicate values, transect numbers, reasonable slope numbers, reasonable diameter values.

### Fuel 1000 Header FMH

[Problem:]{.underline} Incorrect header information entered for FMH plots.

[Procedure:]{.underline}

-   Check that NumTran = 4

```{r}
# Set parameters 
data <- Fuels1000_header
query <- "Fuel 1000 Header FMH NumTran"
query_message <- "Number of Transects" 
values_data <- data$NumTran
values_valid <- 4
values_check <- !values_data == values_valid

# Identify errors
errors_Fuels1000_NumTran <- qaqc(data, query, query_message, values_check)
```

-   Check that North Rim TranLen = 50 (or very close). North Rim plots include PIAB, PIEN, and PIPN.

```{r}
# Set parameters 
data <- Fuels1000_header %>% 
  filter(MonitoringType %in% c("PIAB", "PIEN", "PIPN"))
query <- "Fuel 1000 Header FMH NRim TranLen" 
query_message <- "Transect Length" 
values_data <- data$TranLen 
values_valid <- 50  
values_check <- !values_data == values_valid

# Identify errors
errors_Fuels1000_TranLen_N <- qaqc(data, query, query_message, values_check)
```

-   Check that S Rim TranLen = 100 (or very close). South Rim plots include PIED and PIPO.

```{r}
# Set parameters 
data <- Fuels1000_header %>% 
  filter(MonitoringType %in% c("PIED", "PIPO"))
query <- "Fuel 1000 Header FMH SRim TranLen" 
query_message <- "Transect Length" 
values_data <- data$TranLen 
values_valid <- 100 
values_check <- !values_data == values_valid

# Identify errors
errors_Fuels1000_TranLen_S <- qaqc(data, query, query_message, values_check)
```

### Fuel 1000 DecayClass

[Problem:]{.underline} Logs have an entry for decay class that we don’t use.

[Procedure:]{.underline}

-   Check that DecayCl = 3 or 4

```{r}
# Set parameters
data <- Fuels1000_data
query <- "Fuel 1000 DecayClass"
query_message <- "Decay Class"
values_data <- data$DecayCl
values_valid <- c(3,4)
values_check <- !values_data %in% values_valid

# Identify errors
errors_Fuels1000_DecayCl <- qaqc(data, query, query_message, values_check)
```

### Fuel 1000 FuelConst

[Problem:]{.underline} Incorrect coarse woody debris fuel constant entered for PIAB or PIED or PIEN or PIPN or PIPO. Fuel constant name should correspond to the Monitoring Type.

[Procedure:]{.underline}

-   Check that CWDFuConSt = Monitoring Type

```{r}
# Set parameters
data <- Fuels1000_data
query <- "Fuel 1000 FuelConst"
query_message <- "Fuel Constant"
values_data <- data$CWDFuConSt
values_valid <- data$MonitoringType
values_check <- !values_data == values_valid

# Identify errors
errors_Fuels1000_FuelConst <- qaqc(data, query, query_message, values_check)
```

### Fuel 1000 LogNo

[Problem:]{.underline} "Duplicate" data exists due to multiple 1000 hr fuels with the same transect / diameter / decay class. Log Numbers are missing (which would differentiate 1000 hr fuels).

[Procedure:]{.underline}

-   Check for duplication (TRUE/FALSE) and store results in new column

-   Check for TRUE values

```{r}
# Remove Index. Check for duplication (TRUE/FALSE) and store results in new column
data_temp <- Fuels1000_data %>%
  select(!"Index")

# Set parameters 
data <- data_temp %>%
  mutate("Duplicate" = duplicated(data_temp))
query <- "Fuel 1000 LogNo" 
query_message <- "Duplicate Logs" 
values_data <- data$Duplicate
values_valid <- FALSE
values_check <- !values_data == values_valid

# Identify errors
errors_Fuels1000_LogNo <- qaqc(data, query, query_message, values_check)
```

### Fuel 1000 Values

[Problem:]{.underline}  Values for transect, slope, and diameter are missing or unreasonable.

[Procedure:]{.underline}

-   Check that Transect = 1, 2, 3, or 4

```{r}
# Set parameters
data <- Fuels1000_data
query <- "Fuel 1000 Values TransNum"
query_message <- "Transect Number"
values_data <- data$Transect
values_valid <- c(1, 2, 3, 4)
values_check <- !values_data %in% values_valid

# Identify errors
errors_Fuels1000_TransNum <- qaqc(data, query, query_message, values_check)
```

-   Check that Slope values are reasonable.

```{r}
# Set parameters
data <- Fuels1000_data
query <- "Fuel 1000 Values Slope"
query_message <- "Slope"
values_data <- data$Slope
values_valid <- seq(0, 90, by = 1)
values_check <- !values_data %in% values_valid

# Identify errors
errors_Fuels1000_Slope <- qaqc(data, query, query_message, values_check)
```

-   Check that diameter values are reasonable, are measured at 0.5 increments, and ≥ 3.0.

```{r}
# Perform Rosner's Test for outliers.
# This test is most appropriate when the sample size is large (n > 20). 
# Valid values were found by using the Rosner test on the full suite of data.
# Refer to the script 2_QAQC_GRCA_FMH_Rosner.Rmd

# Set parameters
data <- Fuels1000_data
query <- "Fuel 1000 Values Dia"
query_message <- "Diameter"
values_data <- data$Dia
values_valid <- seq(3, 35, by = 0.5)
values_check <- !values_data %in% values_valid

# Identify errors
errors_Fuels1000_Dia <- qaqc(data, query, query_message, values_check)
```

### Fuel 1000 Errors

```{r}
# Save to master error list
errors_Fuels1000 <- rbind(errors_Fuels1000_DecayCl, errors_Fuels1000_FuelConst, 
                          errors_Fuels1000_NumTran, errors_Fuels1000_TranLen_N, 
                          errors_Fuels1000_TranLen_S, errors_Fuels1000_LogNo, 
                          errors_Fuels1000_TransNum, errors_Fuels1000_Slope, 
                          errors_Fuels1000_Dia)

remove(errors_Fuels1000_DecayCl, errors_Fuels1000_FuelConst, errors_Fuels1000_NumTran, 
       errors_Fuels1000_TranLen_N, errors_Fuels1000_TranLen_S, errors_Fuels1000_LogNo,
       errors_Fuels1000_TransNum, errors_Fuels1000_Slope, errors_Fuels1000_Dia)

# Table of results for quick check
kable(errors_Fuels1000, "pipe")

# Save as CSV or XLSX
path_errors
#write.csv(errors, paste0(path_errors, "errors_Fuels1000.csv"), quote=FALSE, row.names = FALSE, na = "")
#write_xlsx(errors_Fuels1000, paste0(path_errors, "errors_Fuels1000.xlsx"))
```

## DUFF LITTER

This code conducts quality control checks on surface fuels data within the surface fuels duff litter data set.

It checks for: number of transects, fuel constants, number of points sampled, reasonable litter values, reasonable duff values.

### Fuel DL Header FMH

[Problem:]{.underline} Incorrect header information entered for FMH plots.

[Procedure:]{.underline}

-   Check that NumTran = 4

```{r}
# Set parameters  
data <- FuelsDuffLitt_header
query <- "Fuel DL Header FMH NumTran" 
query_message <- "Number of Transects"  
values_data <- data$NumTran  
values_valid <- 4    
values_check <- !values_data == values_valid

# Identify errors
errors_FuelsDuffLitt_NumTran <- qaqc(data, query, query_message, values_check)
```

### Fuel DL FuelConst

[Problem:]{.underline} Incorrect duff & litter fuel constant entered for PIAB or PIED or PIEN or PIPN or PIPO. Fuel constant name should correspond to the Monitoring Type.

[Procedure:]{.underline}

-   Check that DLFuConSt = Monitoring Type

```{r}
# Set parameters 
data <- FuelsDuffLitt_data
query <- "Fuel DL FuelConst" 
query_message <- "Fuel Constant" 
values_data <- data$DLFuConSt 
values_valid <- data$MonitoringType  
values_check <- !values_data == values_valid

# Identify errors
errors_FuelsDuffLitt_FuelConst <- qaqc(data, query, query_message, values_check)

errors_FuelsDuffLitt_FuelConst <- errors_FuelsDuffLitt_FuelConst %>% 
  reframe("SavedQuery" = unique(query),
          "MacroPlot.Name" = unique(MacroPlot.Name), 
          "Date" = unique(Date),
          "Error" = unique(Error),
          "Comment" = unique(Comment)) 
```

### Fuel DL Hits FMH

[Problem:]{.underline} Extra or missing data points in FMH plots. There should be 4 transects with 10 sample points per transect. 

[Procedure:]{.underline}

-   Group by Macroplot. Calculate number of hits per plot. Check that sum_hits = 40

```{r}
# Filter NA values. Add number of hits and store in new column
data_temp <- FuelsDuffLitt_data %>%
  filter(!is.na(SampLoc)) %>% 
  group_by(MacroPlot.Name) %>% 
  summarize("sum_hits" = n())

# Set parameters 
data <- merge(FuelsDuffLitt_data, data_temp, all.x = T)
query <- "Fuel DL Hits" 
query_message <- "Number of Hits" 
values_data <- data$sum_hits
values_valid <- 40
values_check <- !values_data == values_valid

# Identify errors
errors_FuelsDuffLitt_Hits <- qaqc(data, query, query_message, values_check)

errors_FuelsDuffLitt_Hits <- errors_FuelsDuffLitt_Hits %>% 
  reframe("SavedQuery" = unique(query),
          "MacroPlot.Name" = unique(MacroPlot.Name), 
          "Date" = unique(Date),
          "Error" = unique(Error),
          "Comment" = unique(Comment)) 
```

### Fuel DL Values

[Problem:]{.underline}  Values for litter and duff depth are missing or unreasonable.

[Procedure:]{.underline}

-   Sort by LittDep. Check that litter depth values are reasonable.

```{r}
# Perform Rosner's Test for outliers.
# This test is most appropriate when the sample size is large (n > 20). 
# Valid values were found by using the Rosner test on the full suite of data.
# Refer to the script 2_QAQC_GRCA_FMH_Rosner.Rmd

# Set parameters 
data <- FuelsDuffLitt_data
query <- "Fuel DL Values LittDep" 
query_message <- "Litter Depth" 
values_data <- data$LittDep
values_valid <- max(7.6)
values_check <- (!values_data < values_valid) %>% replace_na(TRUE)

# Identify errors
errors_FuelsDuffLitt_LittDep <- qaqc(data, query, query_message, values_check)
```

-   Sort by DuffDep. Check that duff depth values are reasonable.

```{r}
# Perform Rosner's Test for outliers.
# This test is most appropriate when the sample size is large (n > 20). 
# Valid values were found by using the Rosner test on the full suite of data.
# Refer to the script 2_QAQC_GRCA_FMH_Rosner.Rmd

# Set parameters 
data <- FuelsDuffLitt_data
query <- "Fuel DL Values DuffDep" 
query_message <- "Duff Depth" 
values_data <- data$DuffDep
values_valid <- max(13)
values_check <- (!values_data < values_valid) %>% replace_na(TRUE)

# Identify errors
errors_FuelsDuffLitt_DuffDep <- qaqc(data, query, query_message, values_check)
```

### Fuel DL Errors

```{r}
# Save to master error list 
errors_FuelsDuffLitt <- rbind(errors_FuelsDuffLitt_FuelConst, errors_FuelsDuffLitt_NumTran,
                              errors_FuelsDuffLitt_Hits, errors_FuelsDuffLitt_LittDep,
                              errors_FuelsDuffLitt_DuffDep)

remove(errors_FuelsDuffLitt_FuelConst, errors_FuelsDuffLitt_NumTran,
       errors_FuelsDuffLitt_Hits, errors_FuelsDuffLitt_LittDep,
       errors_FuelsDuffLitt_DuffDep)

# Table of results for quick check
kable(errors_FuelsDuffLitt, "pipe")

# Save as CSV or XLSX
path_errors
#write.csv(errors, paste0(path_errors, "errors_Fuels1000.csv"), quote=FALSE, row.names = FALSE, na = "")
#write_xlsx(errors_Fuels1000, paste0(path_errors, "errors_Fuels1000.xlsx"))
```

## FINE FUELS

This code conducts quality control checks on surface fuels data within the surface fuels fine (FWD) data set.

It checks for: header info (number of transects, length of transects for One/Ten/Hundred Hour), fuel constants, transect numbers, reasonable azimuth numbers, reasonable slope numbers, reasonable One/Ten/Hundred Hour count values.

### Fuel Fine Header FMH

[Problem:]{.underline} Incorrect header information entered for FMH plots.

[Procedure:]{.underline}

-   Check that NumTran = 4

```{r}
# Set parameters  
data <- FuelsFine_header
query <- "Fuel Fine Header FMH NumTran" 
query_message <- "Number of Transects"  
values_data <- data$NumTran  
values_valid <- 4    
values_check <- !values_data == values_valid

# Identify errors
errors_FuelsFine_NumTran <- qaqc(data, query, query_message, values_check)
```

-   Check that OneHrTranLen = 6

```{r}
# Set parameters  
data <- FuelsFine_header
query <- "Fuel Fine Header FMH OneHrTranLen" 
query_message <- "One Hour Transect Length"  
values_data <- data$OneHrTranLen
values_valid <- 6
values_check <- !values_data == values_valid

# Identify errors
errors_FuelsFine_OneHrTranLen <- qaqc(data, query, query_message, values_check)
```

-   Check that TenHrTranLen = 6

```{r}
# Set parameters  
data <- FuelsFine_header
query <- "Fuel Fine Header FMH TenHrTranLen" 
query_message <- "Ten Hour Transect Length"  
values_data <- data$TenHrTranLen 
values_valid <- 6
values_check <- !values_data == values_valid

# Identify errors
errors_FuelsFine_TenHrTranLen <- qaqc(data, query, query_message, values_check)
```

-   Check that HunHrTranLen = 12

```{r}
# Set parameters  
data <- FuelsFine_header
query <- "Fuel Fine Header FMH HunHrTranLen" 
query_message <- "Hundrend Hour Transect Length"  
values_data <- data$HunHrTranLen 
values_valid <- 12
values_check <- !values_data == values_valid

# Identify errors
errors_FuelsFine_HunHrTranLen <- qaqc(data, query, query_message, values_check)
```

### Fuel Fine FuelConst

[Problem:]{.underline} Incorrect fine fuel constant entered for PIAB or PIED or PIEN or PIPN or PIPO. Fuel constant name should correspond to the Monitoring Type.

[Procedure:]{.underline}

-   Check that FWDFuConSt = Monitoring Type

```{r}
# Set parameters 
data <- FuelsFine_data
query <- "Fuel Fine FuelConst" 
query_message <- "Fuel Constant" 
values_data <- data$FWDFuConSt 
values_valid <- data$MonitoringType  
values_check <- !values_data == values_valid

# Identify errors
errors_FuelsFine_FuelConst <- qaqc(data, query, query_message, values_check)
```

### Fuel Fine Values

[Problem:]{.underline}  Values for transect, azimuth, slope, and counts for One, Ten, Hundred hour fuels are missing or unreasonable.

[Procedure:]{.underline}

-   Check that Transect = 1, 2, 3, or 4

```{r}
# Set parameters
data <- FuelsFine_data
query <- "Fuel Fine Values TransNum"
query_message <- "Transect Number"
values_data <- data$Transect
values_valid <- c(1, 2, 3, 4)
values_check <- !values_data %in% values_valid

# Identify errors
errors_FuelsFine_TransNum <- qaqc(data, query, query_message, values_check)
```

-   Check that Azimuth values are reasonable.

```{r}
# Set parameters
data <- FuelsFine_data
query <- "Fuel Fine Values Azimuth"
query_message <- "Azimuth"
values_data <- data$Azimuth
values_valid <- seq(0, 360, by = 1)
values_check <- !values_data %in% values_valid

# Identify errors
errors_FuelsFine_Azimuth <- qaqc(data, query, query_message, values_check)
```

-   Check that Slope values are reasonable.

```{r}
# Set parameters
data <- FuelsFine_data
query <- "Fuel Fine Values Slope"
query_message <- "Slope"
values_data <- data$Slope
values_valid <- seq(0, 90, by = 1)
values_check <- !values_data %in% values_valid

# Identify errors
errors_FuelsFine_Slope <- qaqc(data, query, query_message, values_check)
```

-   Check that One Hour values are reasonable.

```{r}
# Perform Rosner's Test for outliers.
# This test is most appropriate when the sample size is large (n > 20). 
# Valid values were found by using the Rosner test on the full suite of data.
# Refer to the script 2_QAQC_GRCA_FMH_Rosner.Rmd

# Set parameters 
data <- FuelsFine_data
query <- "Fuel Fine Values OneHr" 
query_message <- "One Hour Count" 
values_data <- data$OneHr
values_valid <- seq(0, 64, by = 1)
values_check <- !values_data %in% values_valid

# Identify errors
errors_FuelsFine_OneHr <- qaqc(data, query, query_message, values_check)
```

-   Check that Ten Hour values are reasonable.

```{r}
# Perform Rosner's Test for outliers.
# This test is most appropriate when the sample size is large (n > 20). 
# Valid values were found by using the Rosner test on the full suite of data.
# Refer to the script 2_QAQC_GRCA_FMH_Rosner.Rmd

# Set parameters 
data <- FuelsFine_data
query <- "Fuel Fine Values TenHr" 
query_message <- "Ten Hour Count" 
values_data <- data$TenHr
values_valid <- seq(0, 26, by = 1)
values_check <- !values_data %in% values_valid

# Identify errors
errors_FuelsFine_TenHr <- qaqc(data, query, query_message, values_check)
```

-   Check that Hundred Hour values are reasonable.

```{r}
# Perform Rosner's Test for outliers.
# This test is most appropriate when the sample size is large (n > 20). 
# Valid values were found by using the Rosner test on the full suite of data.
# Refer to the script 2_QAQC_GRCA_FMH_Rosner.Rmd

# Set parameters 
data <- FuelsFine_data
query <- "Fuel Fine Values HunHr" 
query_message <- "Hundred Hour Count" 
values_data <- data$HunHr
values_valid <- seq(0, 15, by = 1)
values_check <- !values_data %in% values_valid

# Identify errors
errors_FuelsFine_HunHr <- qaqc(data, query, query_message, values_check)
```

### Fuel Fine Errors

```{r}
# Save to master error list 
errors_FuelsFine <- rbind(errors_FuelsFine_FuelConst, errors_FuelsFine_NumTran,
                          errors_FuelsFine_OneHrTranLen, errors_FuelsFine_TenHrTranLen,
                          errors_FuelsFine_HunHrTranLen, errors_FuelsFine_TransNum,
                          errors_FuelsFine_Azimuth, errors_FuelsFine_Slope,
                          errors_FuelsFine_OneHr, errors_FuelsFine_TenHr,
                          errors_FuelsFine_HunHr)

remove(errors_FuelsFine_FuelConst, errors_FuelsFine_NumTran, errors_FuelsFine_OneHrTranLen,
       errors_FuelsFine_TenHrTranLen, errors_FuelsFine_HunHrTranLen,
       errors_FuelsFine_TransNum, errors_FuelsFine_Azimuth, errors_FuelsFine_Slope,
       errors_FuelsFine_OneHr, errors_FuelsFine_TenHr, errors_FuelsFine_HunHr)

# Table of results for quick check
kable(errors_FuelsFine, "pipe")

# Save as CSV or XLSX 
path_errors
#write.csv(errors, paste0(path_errors, "errors_FuelsFine.csv"), quote=FALSE, row.names = FALSE, na = "") 
#write_xlsx(errors_FuelsFine, paste0(path_errors, "errors_FuelsFine.xlsx"))
```

## 1000HR VS FINE SLOPE

This code conducts quality control checks on slope data in the surface fuels coarse woody debris (1000-hr) and fine woody debris datasets. It also checks for overall consistency in slope and azimuth values across years.

```{r}
# Set parameters 
data_Fuels1000_slope <- Fuels1000_data %>% 
  group_by(MacroPlot.Name, Date, Transect, Slope) %>% 
  summarise("MacroPlot.Name" = unique(MacroPlot.Name), 
            "Transect" = unique(Transect)) %>% 
  mutate(Slope_1000 = Slope) %>% 
  arrange(MacroPlot.Name, Transect) %>% 
  select(!Slope)

data_FuelsFine_slope <- FuelsFine_data %>% 
  group_by(MacroPlot.Name, Date, Transect, Comment, Slope) %>% 
  summarise("MacroPlot.Name" = unique(MacroPlot.Name), 
            "Transect" = unique(Transect)) %>% 
  mutate(Slope_fine = Slope) %>% 
  arrange(MacroPlot.Name, Transect) %>% 
  select(!Slope)

errors_Fuels_slope <- merge(data_Fuels1000_slope, data_FuelsFine_slope, all.y = T) %>% 
  filter(Slope_fine != Slope_1000) %>% 
  mutate(SavedQuery = "Fuel Slope FMH",
         Error = paste("FWD slope", Slope_fine, "!=", "CWD slope", Slope_1000)) %>% 
  reframe("SavedQuery" = unique(query),
          "MacroPlot.Name" = unique(MacroPlot.Name), 
          "Date" = unique(Date), 
          "Error" = unique(Error),
          "Comment" = unique(Comment))

# If no errors, add "No Errors" to existing data frame 
if (nrow(errors_Fuels_slope) == 0) {   
    errors_Fuels_slope <- errors_blank %>%     
    mutate(SavedQuery = query) 
} else {   
    errors_Fuels_slope <- errors_Fuels_slope
}
```

```{r}
# transect_slope_azimuth_qc <- function(fuel1000, fine) {
#   # Print a message indicating that plots are being compared between 1000Hr fuels and fine fuels
#   cat("Plots are the same in 1000Hr fuels and fine fuels...\n")
# 
#   # Check if all unique MacroPlot names in 'fuel1000' are the same as in 'fine'
#   if (all(unique(fuel1000$MacroPlot.Name) %in% unique(fine$MacroPlot.Name))) {
#     cat("TRUE\n")
#     cat("\n")
#   } else {
#     # Print an error message and update flags if there are differing plot names
#     flags<- c(flags, "There is a plot that differs between 1000hr fuels and fine fuels - check this before running function \n")
#     return(cat("FALSE, there is a plot that differs between 1000hr fuels and fine fuels - check this before running function \n"))
#   }
# 
#   # Get unique MacroPlot names from 'fuel1000'
#   plots <- unique(fuel1000$MacroPlot.Name)
# 
#   # Initialize empty vectors to store error flags for slope and azimuth
#   no_errors_s <- c()
#   no_errors_a <- c()
# 
#   # Loop over each unique MacroPlot name
#   for (plot in 1:length(plots)) {
#     macroplot_1000 <- fuel1000[which(fuel1000$MacroPlot.Name == plots[plot]), ]
#     macroplot_fine <- fine[which(fine$MacroPlot.Name == plots[plot]), ]
# 
#     # Determine the set of transects to compare based on their availability in 'fuel1000' and 'fine'
#     if (length(unique(macroplot_1000$Transect)) > length(unique(macroplot_fine$Transect))) {
#       transects <- na.omit(unique(macroplot_1000$Transect))
#     } else if (length(unique(macroplot_1000$Transect)) < length(unique(macroplot_fine$Transect))) {
#       transects <- na.omit(unique(macroplot_fine$Transect))
#     } else {
#       transects <- na.omit(unique(macroplot_fine$Transect))
#     }
# 
#     # Loop over each unique transect within the current MacroPlot
#     for (transect in 1:length(transects)) {
#       transect_1000 <- macroplot_1000[which(macroplot_1000$Transect == transects[transect]), ]
#       transect_fine <- macroplot_fine[which(macroplot_fine$Transect == transects[transect]), ]
# 
#       # Compare slope values within 1000Hr fuels
#       if (length(unique(transect_1000$Slope)) == 1 | length(unique(transect_1000$Slope)) == 0) {
#         no_errors_s <- c(no_errors_s, 0)
#       } else {
#         no_errors_s <- c(no_errors_s, 1)
#         cat(paste(c("Error in macroplot", plots[plot], "transect", transects[transect], "1000hr fuels: multiple different values for slopes:",unique(transect_1000$Slope),  "\n"), collapse = " "))
#         cat("\n")
#         flags<- c(flags, paste(c("Error in macroplot", plots[plot], "transect", transects[transect], "1000hr fuels: multiple different values for slopes:",unique(transect_1000$Slope)), collapse = " "))
#       }
# 
#       # Compare slope values within fine fuels
#       if (length(unique(transect_fine$Slope)) == 1 | length(unique(transect_fine$Slope)) == 0) {
#         no_errors_s <- c(no_errors_s, 0)
#       } else {
#         no_errors_s <- c(no_errors_s, 1)
#         cat(paste(c("Error in macroplot", plots[plot], "transect", transects[transect], "fine fuels: multiple different values for slopes:", unique(transect_fine$Slope), "\n"), collapse = " "))
#         cat("\n")
#         flags<- c(flags, paste(c("Error in macroplot", plots[plot], "transect", transects[transect], "fine fuels: multiple different values for slopes:", unique(transect_fine$Slope)), collapse = " "))
#       }
# 
#       # Compare MacroPlot azimuth values within 1000Hr fuels
#       if (length(unique(transect_1000$MacroPlot.Azimuth)) == 1 | length(unique(transect_1000$MacroPlot.Azimuth)) == 0) {
#         no_errors_a <- c(no_errors_a, 0)
#       } else {
#         no_errors_a <- c(no_errors_a, 1)
#         cat(paste(c("Error in macroplot", plots[plot], "transect", transects[transect], "1000hr fuels: multiple different values for MacroPlot.Azimuths:", unique(transect_1000$MacroPlot.Azimuth), "\n"), collapse = " "))
#         cat("\n")
#         flags<- c(flags, paste(c("Error in macroplot", plots[plot], "transect", transects[transect], "1000hr fuels: multiple different values for MacroPlot.Azimuths:", unique(transect_1000$MacroPlot.Azimuth)), collapse = " "))
#       }
# 
#       # Compare MacroPlot azimuth values within fine fuels
#       if (length(unique(transect_fine$MacroPlot.Azimuth)) == 1 | length(unique(transect_fine$MacroPlot.Azimuth)) == 0) {
#         no_errors_a <- c(no_errors_a, 0)
#       } else {
#         no_errors_a <- c(no_errors_a, 1)
#         cat(paste(c("Error in macroplot", plots[plot], "transect", transects[transect], "fine fuels: multiple different values for MacroPlot.Azimuths:", unique(transect_fine$MacroPlot.Azimuth),"\n"), collapse = " "))
#         cat("\n")
#         flags<- c(flags, paste(c("Error in macroplot", plots[plot], "transect", transects[transect], "fine fuels: multiple different values for MacroPlot.Azimuths:", unique(transect_fine$MacroPlot.Azimuth)), collapse = " "))
#       }
# 
#       # Compare slope values between fine and 1000Hr fuels
#       if (length(unique(transect_fine$Slope)) == 1 & length(unique(transect_1000$Slope)) == 1) {
#         if (unique(transect_1000$Slope) == unique(transect_fine$Slope)) {
#           no_errors_s <- c(no_errors_s, 0)
#         } else {
#           no_errors_s <- c(no_errors_s, 1)
#           cat(paste(c("Error in macroplot", plots[plot], "transect", transects[transect], ": different values for fine (",unique(transect_fine$Slope),") and 1000hr (",unique(transect_1000$Slope),") slopes\n"), collapse = " "))
#           cat("\n")
#           flags<- c(flags, paste(c("Error in macroplot", plots[plot], "transect", transects[transect], ": different values for fine and 1000hr slopes"), collapse = " "))
#         }
#       } else if (length(unique(transect_fine$Slope)) == 0 | length(unique(transect_1000$Slope)) == 0) {
#         # Cannot compare slope values between fine and 1000Hr fuels
#         # This is acceptable
#       } else {
#         # Error - More than 1 value for slopes in either fine or 1000Hr fuels, cannot compare values
#         # This should have already been flagged
#       }
# 
#       # Compare MacroPlot azimuth values between fine and 1000Hr fuels
#       if (length(unique(transect_fine$MacroPlot.Azimuth)) == 1 & length(unique(transect_1000$MacroPlot.Azimuth)) == 1) {
#         if (unique(transect_1000$MacroPlot.Azimuth) == unique(transect_fine$MacroPlot.Azimuth)) {
#           no_errors_a <- c(no_errors_a, 0)
#         } else {
#           no_errors_a <- c(no_errors_a, 1)
#           cat(paste(c("Error in macroplot", plots[plot], "transect", transects[transect], ": different values for fine (",unique(transect_fine$MacroPlot.Azimuth),")and 1000hr (",unique(transect_1000$MacroPlot.Azimuth),") MacroPlot.Azimuths\n"), collapse = " "))
#           cat("\n")
#           flags<- c(flags, paste(c("Error in macroplot", plots[plot], "transect", transects[transect], ": different values for fine and 1000hr MacroPlot.Azimuths"), collapse = " "))
#         }
#       } else if (length(unique(transect_fine$MacroPlot.Azimuth)) == 0 | length(unique(transect_1000$MacroPlot.Azimuth)) == 0) {
#         # Cannot compare MacroPlot azimuth values between fine and 1000Hr fuels
#         # This is acceptable
#       } else {
#         # Error - More than 1 value for MacroPlot.Azimuths in either fine or 1000Hr fuels, cannot compare values
#         # This should have already been flagged
#       }
#     }
#   }
# 
#   # Check if all transects have consistent slope values
#   if (length(unique(no_errors_s)) == 1) {
#     if (unique(no_errors_s) == 0) {
#       cat("Slope for a transect does not change across years or between fine and 1000hr fuel data\n")
#       cat("\n")
#     }
#   }
# 
#   # Check if all transects have consistent azimuth values
#   if (length(unique(no_errors_a)) == 1) {
#     if (unique(no_errors_a) == 0) {
#       cat("Azimuth for a transect does not change across years or between fine and 1000hr fuel data\n")
#       cat("\n")
#     }
#   }
# 
#   return(flags)
# } # end of function

```

# PROTOCOL - POST BURN

# PROTOCOL - TREES

## Tree CBH and Tree CBH Range

## Tree Crown Class

## Tree Damage

## Tree DBH

## Tree DD

## Tree Area Multiplier

## Tree Sample Area

## Tree Height

## Tree Severity

## Tree Status and Life Form

## Tree Duplicates

## Tree Dead to Alive and Changing DBH

# ALL ERRORS

```{r}
# Save to master error list 
errors_all <- rbind(errors_Fuels1000, errors_FuelsDuffLitt, 
                    errors_FuelsFine, errors_Fuels_slope)

# Table of results for quick check
kable(errors_all, "pipe")

# Save as CSV or XLSX 
path_errors
#write.csv(errors, paste0(path_errors, "errors_all.csv"), quote=FALSE, row.names = FALSE, na = "") 
#write_xlsx(errors_all, paste0(path_errors, "errors_all.xlsx"))
```
