---
title: "Grand Canyon QAQC for FMH Data"
author: "Alexandra Lalor"
date: "2024-11-06"
output:
  html_document:
    theme: readable
    highlight: 
    toc: yes
    toc_depth: 3
    toc_float:
      smooth_scroll: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Install packages (if needed)
#install.packages("tidyverse")
```

## R Markdown

press ctrl+alt+i to create a code block

# BEFORE STARTING

Make sure to update file paths to be specific for your data.

```{r}
# Load packages
library(tidyverse)
library(writexl)
library(dplyr)
library(knitr)
library(kableExtra)

# Identify working directory (specifically user name)
getwd()

# Load in data and name them based on file path
# Change file path based on user name!
#data_path <- "C:/Users/alalor.NPS/Desktop/R/GRCA/data/data_qaqc/2024/ORIGINALS/"
data_path <- "C:/Users/alalor.NPS/Desktop/R/GRCA/data/data_qaqc/2024/"
errors_path <- "C:/Users/alalor.NPS/Desktop/R/GRCA/data/data_errors/2024/"
```

# PROTOCOL - SURFACE FUELS

## SUBPROTOCOL - 1000HR CWD

This code conducts quality control checks on surface fuels data within the surface fuels coarse woody debris (1000-hr) data set. It checks for: header info (number of transects, length of transects), decay class values, fuel constants, duplicate values, transect numbers, reasonable slope numbers, reasonable diameter values.

\*IF ADJUSTING FOR DIFFERENT PROGRAM – you must edit code to include values specific to your program. You can edit the code in the "Set Parameters" section of each query code block.

#### Load and Check Data

```{r}
data_path
# Load in data
Fuels1000_all <- read.csv(paste0(data_path, "Surface Fuels - 1000Hr_XPT.csv"))
```

```{r}
# Ensure blanks in Visited column are NA
Fuels1000_all$Visited[Fuels1000_all$Visited==""] <- NA

# Clean up data frame
Fuels1000_all <- Fuels1000_all %>%
  # Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  # Separate MacroPlotName to get Monitoring type (or export data with Monitoring Type already included)
  separate(MacroPlot.Name, sep = " ", into = c("MonitoringType", "Plot"), remove = FALSE) %>% 
  # Add columns relevant to error checking
  mutate("SavedQuery" = "", "Error" = "", "Fixed" = "", "Explanation" = "", "Person" = "")

# Filter for only data
Fuels1000_data <- Fuels1000_all %>%
  filter(is.na(Visited))

# Filter for only headers
Fuels1000_header <- Fuels1000_all %>%
  filter(!is.na(Visited))

# Blank data frame for no errors
no_errors <- data.frame("SavedQuery" = "","MacroPlot.Name" = "","Date" = "","Error" = "No Error","Fixed" = "","Explanation" = "","Person" = "")
```

### Fuel 1000 header FMH

[Problem:]{.underline} Incorrect header information entered for FMH plots.

[Procedure:]{.underline}

-   Check that NumTran = 4

```{r}
# Set parameters 
saved_query <- "Fuel 1000 header FMH NumTran" 
error_message <- "Number of Transects" 
column_data <- Fuels1000_header$NumTran 
valid_values <- 4  

# Identify errors 
errors_NumTran <- Fuels1000_header %>%   
  mutate("SavedQuery" = saved_query, "Error" = paste(error_message, "=", column_data)) %>%   
  filter(!column_data == valid_values) %>%   
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Fixed","Explanation","Person"))  

# If no errors, add "No Errors" to existing data frame 
if (nrow(errors_NumTran) == 0) {   
  errors_NumTran <- no_errors %>%     
    mutate(SavedQuery = saved_query) 
} else {   
    errors_NumTran <- errors_NumTran }

# Table of results for quick check
table <- errors_NumTran %>% 
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error"))
kable(table) %>% 
  kable_material_dark()
```

-   Check that N Rim TranLen = 50 (or very close). North Rim plots include PIAB, PIEN, and PIPN.

```{r}
# Set parameters 
saved_query <- "Fuel 1000 header FMH TranLen NRim" 
error_message <- "Transect Length" 
column_data <- Fuels1000_header$TranLen 
valid_values <- 50  


# Identify errors 
errors_TranLen_N <- Fuels1000_header %>%   
  mutate("SavedQuery" = saved_query, "Error" = paste(error_message, "=", column_data))  %>%    
  filter(!column_data == valid_values, 
         MonitoringType %in% c("PIAB", "PIEN", "PIPN")) %>%    
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Fixed","Explanation","Person"))

# If no errors, add "No Errors" to existing data frame 
if (nrow(errors_TranLen_N) == 0) {   
  errors_TranLen_N <- no_errors %>%     
    mutate(SavedQuery = saved_query) 
} else {   
    errors_TranLen_N <- errors_TranLen_N }

# Table of results for quick check
table <- errors_TranLen_N %>% 
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error"))
kable(table) %>% 
  kable_material_dark()
```

-   Check that S Rim TranLen = 100 (or very close). South Rim plots include PIED and PIPO.

```{r}
# Set parameters 
saved_query <- "Fuel 1000 header FMH TranLen SRim" 
error_message <- "Transect Length" 
column_data <- Fuels1000_header$TranLen 
valid_values <- 100  

# Identify errors 
errors_TranLen_S <- Fuels1000_header %>%   
  mutate("SavedQuery" = saved_query, "Error" = paste(error_message, "=", column_data)) %>%    
  filter(!column_data == valid_values,           
         MonitoringType %in% c("PIED", "PIPO")) %>%    
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Fixed","Explanation","Person"))  

# If no errors, add "No Errors" to existing data frame 
if (nrow(errors_TranLen_S) == 0) {   
  errors_TranLen_S <- no_errors %>%     
    mutate(SavedQuery = saved_query) 
} else {   
    errors_TranLen_S <- errors_TranLen_S }

# Table of results for quick check
table <- errors_TranLen_S %>% 
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error"))
kable(table) %>% 
  kable_material_dark()
```

### Fuel 1000 DecayClass

[Problem:]{.underline} Logs have an entry for decay class that we don’t use.

[Procedure:]{.underline} Check that DecayCl = 3 or 4

```{r}
# Set parameters
saved_query <- "Fuel 1000 DecayClass"
error_message <- "Decay Class"
column_data <- Fuels1000_data$DecayCl
valid_values <- c(3,4)

# Identify errors
errors_DecayCl <- Fuels1000_data %>%
  mutate("SavedQuery" = saved_query, "Error" = paste(error_message, "=", column_data)) %>%
  filter(!column_data %in% valid_values) %>%
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Fixed","Explanation","Person"))

# If no errors, add "No Errors" to existing data frame
if (nrow(errors_DecayCl) == 0) {
  errors_DecayCl <- no_errors %>%
    mutate(SavedQuery = saved_query)
} else {
  errors_DecayCl <- errors_DecayCl
}

# Table of results for quick check
table <- errors_DecayCl %>% 
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error"))
kable(table) %>% 
  kable_material_dark()
```

### Fuel 1000 FuelConst

[Problem:]{.underline} Incorrect coarse woody debris fuel constant entered for PIAB or PIED or PIEN or PIPN or PIPO. Fuel constant name should correspond to the Monitoring Type

[Procedure:]{.underline} Check that CWDFuConSt = Monitoring Type

```{r}
# Set parameters
saved_query <- "Fuel 1000 FuelConst"
error_message <- "Fuel Constant"
column_data <- Fuels1000_data$CWDFuConSt
valid_values <- Fuels1000_data$MonitoringType

# Identify errors
errors_FuelConst <- Fuels1000_data %>%
  mutate("SavedQuery" = saved_query, "Error" = paste(error_message, "=", column_data)) %>%
  filter(!column_data == valid_values) %>%
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Fixed","Explanation","Person"))

# If no errors, add "No Errors" to existing data frame
if (nrow(errors_FuelConst) == 0) {
  errors_FuelConst <- no_errors %>%
    mutate(SavedQuery = saved_query)
} else {
  errors_FuelConst <- errors_FuelConst
}

# Table of results for quick check
table <- errors_FuelConst %>% 
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error"))
kable(table) %>% 
  kable_material_dark()
```

### Fuel 1000 LogNo

[Problem:]{.underline} "Duplicate" data exists due to multiple 1000 hr fuels with the same transect / diameter / decay class. Log Numbers are missing (which would differentiate 1000 hr fuels).

[Procedure:]{.underline}

-   Check for duplication (TRUE/FALSE) and store results in UV1

-   Check for TRUE values in UV1

```{r}
# Store duplication results in Duplicate column
Fuels1000_data_duplicate <- Fuels1000_data %>% 
  mutate("Duplicate" = duplicated(Fuels1000_data))

# Set parameters 
saved_query <- "Fuel 1000 LogNo" 
error_message <- "Duplicate Logs" 
column_data <- Fuels1000_data_duplicate$Duplicate
valid_values <- FALSE

# Identify errors 
errors_LogNo <- Fuels1000_data_duplicate %>%   
  mutate("SavedQuery" = saved_query, "Error" = paste(error_message, "=", column_data)) %>%    
  filter(!column_data == valid_values) %>%    
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Fixed","Explanation","Person"))

# If no errors, add "No Errors" to existing data frame 
if (nrow(errors_LogNo) == 0) {   
  errors_LogNo <- no_errors %>%     
    mutate(SavedQuery = saved_query) 
} else {   
    errors_LogNo <- errors_LogNo }

# Table of results for quick check
table <- errors_LogNo %>% 
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error"))
kable(table) %>% 
  kable_material_dark()
```

### Fuel 1000 Values

[Problem:]{.underline}  Values for transect, slope, and diameter are missing or unreasonable.

[Procedure:]{.underline}

-   Check that Transect = 1, 2, 3, or 4

```{r}
# Set parameters
saved_query <- "Fuel 1000 Values TransNum"
error_message <- "Transect Number"
column_data <- Fuels1000_data$Transect
valid_values <- c(1, 2, 3, 4)

# Identify errors
errors_TransNum <- Fuels1000_data %>%
  mutate("SavedQuery" = saved_query, "Error" = paste(error_message, "=", column_data)) %>% 
  filter(!column_data %in% valid_values) %>% 
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Fixed","Explanation","Person"))

# If no errors, add "No Errors" to existing data frame
if (nrow(errors_TransNum) == 0) {
  errors_TransNum <- no_errors %>%
    mutate(SavedQuery = saved_query)
} else {
  errors_TransNum <- errors_TransNum
}

# Table of results for quick check
table <- errors_TransNum %>% 
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error"))
kable(table) %>% 
  kable_material_dark()
```

-   Check that Slope values are reasonable.

```{r}
# Set parameters
saved_query <- "Fuel 1000 Values Slope"
error_message <- "Slope"
column_data <- Fuels1000_data$Slope
valid_values <- seq(0, 90, by = 1)

# Identify errors
errors_Slope <- Fuels1000_data %>%
  mutate("SavedQuery" = saved_query, "Error" = paste(error_message, "=", column_data)) %>% 
  filter(!column_data %in% valid_values) %>% 
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Fixed","Explanation","Person"))

# If no errors, add "No Errors" to existing data frame
if (nrow(errors_Slope) == 0) {
  errors_Slope <- no_errors %>%
    mutate(SavedQuery = saved_query)
} else {
  errors_Slope <- errors_Slope
}

# Table of results for quick check
table <- errors_Slope %>% 
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error"))
kable(table) %>% 
  kable_material_dark()
```

-   Check that diameter values are reasonable, are measured at 0.5 increments, and ≥ 3.0.

```{r}
# Set parameters
saved_query <- "Fuel 1000 Values Dia"
error_message <- "Diameter"
column_data <- Fuels1000_data$Dia
valid_values <- seq(3, 40, by = 0.5)

# Identify errors
errors_Dia <- Fuels1000_data %>%
  mutate("SavedQuery" = saved_query, "Error" = paste(error_message, "=", column_data)) %>% 
  filter(!column_data %in% valid_values) %>% 
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Fixed","Explanation","Person"))

# If no errors, add "No Errors" to existing data frame
if (nrow(errors_Dia) == 0) {
  errors_Dia <- no_errors %>%
    mutate(SavedQuery = saved_query)
} else {
  errors_Dia <- errors_Dia
}

# Table of results for quick check
table <- errors_Dia %>% 
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error"))
kable(table) %>% 
  kable_material_dark()
```

### Outlier Tests?

```{r}
### Function: Outlier Tests

  # # Function to perform outlier test using Rosner test on a specified column of data
  # # and log any outliers detected along with their details
  # cat("\n")
  # test_for_outliers <- function(data, column_name) {
  #   # Display a message indicating the test being performed
  #   cat("Are there any outlier values in", column_name, "?\n")
  #
  #   # Extract non-missing and non-NA/NaN/Inf values from the specified column
  #   column_values <- na.omit(data[[column_name]])
  #
  #   # Check if there are any valid values for testing
  #   if (length(column_values) == 0) {
  #     # Print a message if no valid values are available for testing
  #     cat("No valid values to test\n")
  #     return()  # Exit the function if there are no valid values
  #   }
  #
  #   # Perform Rosner test on the valid column values
  #   test <- rosnerTest(column_values)
  #   test <- test$all.stats
  #   outliers <- test[which(test$Outlier == TRUE), 4]
  #
  #   # Check if any outliers were detected
  #   if (length(outliers) == 0) {
  #     # Print a message if no outliers were detected
  #     cat("No\n")
  #   } else {
  #     # Print information about the detected outliers and their details
  #     cat(paste("Yes, the outlier values according to a Rosner test are", outliers, ". They are in events",
  #               data[which(data[[column_name]] %in% outliers),"MacroPlot.Name"],data[which(data[[column_name]] %in% outliers),"Monitoring.Status"], "of the data table. For reference, the max, min, and mean of",
  #               column_name, "are", max(column_values), min(column_values), mean(column_values), "respectively",
  #               collapse = "\n"), "\n")
  #
  #     # Add error messages about the detected outliers to the flags vector
  #     flags<- c(flags, paste("The", column_name, "has outlier values according to a Rosner test, which are", outliers,
  #                            ". They are in events", data[which(data[[column_name]] %in% outliers),"MacroPlot.Name"],data[which(data[[column_name]] %in% outliers),"Monitoring.Status"],
  #                            "of the data table. The max, min, and mean of", column_name, "'s are", max(column_values),
  #                            min(column_values), mean(column_values), "respectively", collapse = "\n"), "\n")
  #   }
  # }

### Outlier Test
  # # Perform outlier tests for the "Slope" and "Dia" columns in the fuel1000 dataset
  # test_for_outliers(fuel1000, "Slope")
  # cat("\n")
  # test_for_outliers(fuel1000, "Dia")

```

#### Save Errors

```{r}
# Save to master error list
errors_Fuels1000 <- rbind(errors_DecayCl, errors_FuelConst, errors_NumTran, 
                          errors_TranLen_N, errors_TranLen_S, errors_LogNo, 
                          errors_TransNum, errors_Slope, errors_Dia)
```

```{r}
# Save as CSV or XLSX
errors_path
#write.csv(errors, paste0(errors_path, "errors_Fuels1000.csv"), quote=FALSE, row.names = FALSE, na = "")
write_xlsx(errors_Fuels1000, paste0(errors_path, "errors_Fuels1000.xlsx"))
```
