---
title: "QAQC GRCA MetaData"
author: "Alexandra Lalor"
output:
  html_document:
    theme: readable
    highlight: 
    toc: yes
    toc_depth: 3
    toc_float:
      smooth_scroll: yes
    code_download: true
  pdf_document:
    toc: yes
    toc_depth: '3'
---

# Before Starting

### Setup

```{r setup, include=FALSE}
## Markdown Options
knitr::opts_chunk$set(echo = FALSE, error = FALSE, warnings = FALSE, message = FALSE)
```

```{r install packages, include=FALSE}
## Install packages (if needed)

# install.packages("tidyverse")
# install.packages("Rtools")
# install.packages("janitor")
# install.packages("condformat")
# install.packages("knitr")
# install.packages("here")
```

```{r load packages, include=FALSE}
## Load Packages

# tidyverse and dplyr have lots of useful functions for data cleaning
library(tidyverse)
library(dplyr)
library(plyr)
# janitor has functions to find duplicates
library(janitor)
# writexl is used to create excel output files
library(writexl)
# condformat is used to format excel output files
library(condformat)
# knitr is used to create output files from R Markdown
library(knitr)
# "here" helps you easily find and reference your working directory
library(here)
```

### Create Function

```{r}
# Blank data frame
errors_blank_meta <- data.frame("SavedQuery" = "", "Macroplot" = "", "SampleEventDate" = "", "Error" = "", "Fixed" = "", "Explanation" = "", "Queryers" = "") %>% 
  mutate(SampleEventDate = as.Date(SampleEventDate))

# QAQC function
qaqc_meta <- function(data, query, query_message, values_check) {
  
  # Check if there is data to QAQC
  if(nrow(data) == 0) {
  # If there is no data, default to "No Error" data frame
  errors <- errors_blank_meta %>%
    mutate(SavedQuery = query,
           Error = "No Error")
  # If there is data, perform error check
} else {
  
  # Use relevant dataset to look for errors
  errors <- data %>%  
    # Add columns relevant to error checking
           # Populate the "SavedQuery" column with the query name
    mutate("SavedQuery" = query, 
           # Populate the "Error" column with the query message and relevant data
           "Error" = paste(query_message, "=", values_data),
           # Create a blank "Fixed" column
           "Fixed" = "",
           # Create a blank "Explanation" column
           "Explanation" = "",
           # Create a blank "Queryers" column
           "Queryers" = "") %>%   
    # Filter for data which is "false" (data does not match valid conditions)
    # NA values are treated as "false" (missing data is not valid)
    filter(!(values_check %>% replace_na(FALSE))) %>%   
    # Select relevant columns to view errors
    select("SavedQuery", "Macroplot", "SampleEventDate", "Error", "Fixed", "Explanation", "Queryers")
  
  # Next, check if there are duplicate errors
  errors_temp <- errors %>% 
    get_dupes(SavedQuery, Macroplot, SampleEventDate, Error)
  # Merge duplicate errors
  errors <- unique(merge(errors, errors_temp, all = T))
  # If duplicate errors exist, add number of duplicates to error message
  errors <- errors %>% 
    mutate(Error = ifelse(is.na(dupe_count), Error, paste0("(x", dupe_count, ") ", Error))) %>% 
    select(!"dupe_count")
}
  
  # Check if there are no errors
  if (nrow(errors) == 0) {  
    # If there are no errors, default to "No Error" data frame
    errors <- errors_blank_meta %>%
      mutate(SavedQuery = query,
             Error = "No Error")
    # If there are errors, keep error log you just created
  } else {   
    errors <- errors
  }
}
```

### Adjust File Paths

```{r import data, include=FALSE}
# Identify working directory.
here()

# Load in data.
path_data <- "X:/FFI Data Management/Exports from FFI/metadata/"
path_dataclean <- paste0(here(), "/output/data_clean/")
path_errors <- paste0(here(), "/output/errors/")
path_errors_name <- "Metadata QAQC Errors"
```

### Load Data

```{r load data, include=FALSE}
## Load Data
SampleEventReport_all <- read.csv(paste0(path_dataclean, "all_SampleEventReport.csv"))
MetadataReport_all <- read.csv(paste0(path_dataclean, "all_MetadataReport.csv"))
all_MonStatus <- read.csv(paste0(path_dataclean, "all_MonStatus.csv"))
all_DisturbanceHistory <- read.csv(paste0(path_dataclean, "all_DisturbanceHistory.csv"))
```

# Metadata Report

## Disturbance History

[Problem:]{.underline} Disturbance info is missing or written incorrectly

[Procedure:]{.underline}

-   If Entry \> 0, Check that UV2-8 fields are not blank and and written correctly.

```{r}
# Set parameters 
data <- MetadataReport_all %>%
  filter(Entry == "1")
query <- "Disturbance History"
query_message <- paste0("Entry = ", data$Entry, ". ", "UV2 = ", data$UV2, ". ", "Disturbance Desc") 
values_data <- data$Disturbance_Desc
values_valid <- c("RX", "WF-Supp", "WFU", "WF")
values_check <- values_data %in% values_valid

# Identify errors
errors_Disturbance01 <- qaqc_meta(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- MetadataReport_all %>%
  filter(Entry == "2")
query <- "Disturbance History"
query_message <- paste0("Entry = ", data$Entry, ". ", "UV3 = ", data$UV3, ". ", "Disturbance Desc") 
values_data <- data$Disturbance_Desc
values_valid <- c("RX", "WF-Supp", "WFU", "WF")
values_check <- values_data %in% values_valid

# Identify errors
errors_Disturbance02 <- qaqc_meta(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- MetadataReport_all %>%
  filter(Entry == "3")
query <- "Disturbance History"
query_message <- paste0("Entry = ", data$Entry, ". ", "UV4 = ", data$UV4, ". ", "Disturbance Desc") 
values_data <- data$Disturbance_Desc
values_valid <- c("RX", "WF-Supp", "WFU", "WF")
values_check <- values_data %in% values_valid

# Identify errors
errors_Disturbance03 <- qaqc_meta(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- MetadataReport_all %>%
  filter(Entry == "4")
query <- "Disturbance History"
query_message <- paste0("Entry = ", data$Entry, ". ", "UV5 = ", data$UV5, ". ", "Disturbance Desc") 
values_data <- data$Disturbance_Desc
values_valid <- c("RX", "WF-Supp", "WFU", "WF")
values_check <- values_data %in% values_valid

# Identify errors
errors_Disturbance04 <- qaqc_meta(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- MetadataReport_all %>%
  filter(Entry == "5")
query <- "Disturbance History"
query_message <- paste0("Entry = ", data$Entry, ". ", "UV6 = ", data$UV6, ". ", "Disturbance Desc") 
values_data <- data$Disturbance_Desc
values_valid <- c("RX", "WF-Supp", "WFU", "WF")
values_check <- values_data %in% values_valid

# Identify errors
errors_Disturbance05 <- qaqc_meta(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- MetadataReport_all %>%
  filter(Entry == "6")
query <- "Disturbance History"
query_message <- paste0("Entry = ", data$Entry, ". ", "UV7 = ", data$UV7, ". ", "Disturbance Desc") 
values_data <- data$Disturbance_Desc
values_valid <- c("RX", "WF-Supp", "WFU", "WF")
values_check <- values_data %in% values_valid

# Identify errors
errors_Disturbance06 <- qaqc_meta(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- MetadataReport_all %>%
  filter(Entry == "7")
query <- "Disturbance History"
query_message <- paste0("Entry = ", data$Entry, ". ", "UV8 = ", data$UV8, ". ", "Disturbance Desc") 
values_data <- data$Disturbance_Desc
values_valid <- c("RX", "WF-Supp", "WFU", "WF")
values_check <- values_data %in% values_valid

# Identify errors
errors_Disturbance07 <- qaqc_meta(data, query, query_message, values_check)
```

#### Combine Disturbance Errors

```{r}
# Combine
errors_Disturbance <- unique(rbind(errors_Disturbance01, errors_Disturbance02, errors_Disturbance03, errors_Disturbance04, errors_Disturbance05, errors_Disturbance06, errors_Disturbance07))

remove(errors_Disturbance01, errors_Disturbance02, errors_Disturbance03, errors_Disturbance04, errors_Disturbance05, errors_Disturbance06, errors_Disturbance07)

if(nrow(errors_Disturbance) > 1) {
  errors_Disturbance <- errors_Disturbance %>%
    filter(Error != "No Error")
} else {
  errors_Disturbance <- errors_Disturbance}
```

## Monitoring Status

[Problem:]{.underline} Metadata Report does not have a Monitoring Status assigned to a plot. Monitoring Status does not match between Metadata Report and Sample Event Report. Monitoring Status does not match standard values.

[Procedure:]{.underline}

-   Check that MonStatus is not blank

```{r}
# Set parameters 
data <- MetadataReport_all
query <- "MonStatus"
query_message <- "Monitoring Status" 
values_data <- data$MonStatus
values_valid <- ""
values_check <- values_data != values_valid

# Identify errors
errors_MonStatus_Missing <- qaqc_meta(data, query, query_message, values_check)
```

-   Compare unique MonStatus between both the Sample Event Report and Metadata Report. Isolate differences between reports.

```{r}
# Compare MonStatus between Sample Event Report and Metadata Report, isolate differences
data_temp1 <- SampleEventReport_all %>% 
  select(AdministrationUnit_Name, Macroplot, SampleEventDate, MonStatus) %>% 
  unique()

data_temp2 <- MetadataReport_all %>% 
  select(AdministrationUnit_Name, Macroplot, SampleEventDate, MonStatus) %>% 
  filter(!is.na(SampleEventDate)) %>% 
  unique()

data_temp <- setdiff(data_temp1, data_temp2) %>% 
  mutate(MonStatusError = MonStatus) %>% 
  filter(MonStatus != "") %>% 
  select(!MonStatus)
remove(data_temp1, data_temp2)

# Set parameters 
data <- merge(MetadataReport_all, data_temp, by = c("AdministrationUnit_Name", "Macroplot", "SampleEventDate"), all = T)
data$MonStatusError[is.na(data$MonStatusError)] <- ""
data <- data %>% 
  filter(MonStatusError != "")
query <- "MonStatus"
query_message <- paste0("MetadataReport MonStatus = ", data$MonStatus, ". ", "SampleEventReport MonStatus") 
values_data <- data$MonStatusError
values_valid <- data$MonStatus
values_check <- values_data == values_valid

# Identify errors
errors_MonStatus_SampleEvent <- qaqc_meta(data, query, query_message, values_check)
```

-   Check that MonStatus matches expected values

```{r}
# Set parameters 
data <- MetadataReport_all
data$MonStatus[is.na(data$MonStatus)] <- ""
query <- "MonStatus"
query_message <- "Monitoring Status" 
values_data <- data$MonStatus
values_valid <- c("", "Measurement_1", "Measurement_2", "Measurement_3", "Measurement_4",
                       "Pre", "PR01", "PR02", "PR03", "PR04", "PR05",
                       "02Pre", "02PR01", "02PR02", 
                       "01Post",
                       "01Year1", "01Year2", "01Year5", "01Year10", "01Year20",
                       "02Post",
                       "02Year1", "02Year2", "02Year5", "02Year10", "02Year20",
                       "03Post",
                       "03Year1", "03Year2", "03Year5", "03Year10", "03Year20",
                       "04Post",
                       "04Year1", "04Year2", "04Year5", "04Year10", "04Year20")
values_check <- values_data %in% values_valid

# Identify errors
errors_MonStatus_OffCycle <- qaqc_meta(data, query, query_message, values_check)
```

#### Combine MonStatus Errors

```{r}
# Combine
errors_MonStatus <- unique(rbind(errors_MonStatus_Missing, errors_MonStatus_SampleEvent, errors_MonStatus_OffCycle))

remove(errors_MonStatus_Missing, errors_MonStatus_SampleEvent, errors_MonStatus_OffCycle)

if(nrow(errors_MonStatus) > 1) {
  errors_MonStatus <- errors_MonStatus %>%
    filter(Error != "No Error")
} else {
  errors_MonStatus <- errors_MonStatus}
```

## Project Unit

[Problem:]{.underline} Project Unit info is missing

[Procedure:]{.underline}

-   Check that UV1 field is not blank

```{r}
# Set parameters 
data <- MetadataReport_all
data$UV1[is.na(data$UV1)] <- ""
query <- "Project Unit (UV1)"
query_message <- paste0("UV1") 
values_data <- data$UV1
values_valid <- ""
values_check <- values_data != values_valid

# Identify errors
errors_ProjectUnit <- qaqc_meta(data, query, query_message, values_check)
```

## Protocols

[Problem:]{.underline} Incorrect protocols are used

[Procedure:]{.underline}

-   Check that protocols match expected values

```{r}
# Set parameters 
data <- MetadataReport_all %>% 
  filter(AdministrationUnit_Name == "WACA",
         PlotType_Name == "FMH")
query <- "Protocols"
query_message <- paste0("Park = ", data$Park, ". ", "Protocol") 
values_data <- data$Protocols
values_valid <- c("Surface Fuels",
                  "Post Burn Severity",
                  "Trees (metric)",
                  #"TreeCover2(metric)",
                  "Density - Belts (metric)",
                  "Cover - Species Composition (metric)",
                  "Cover - Points (metric)")
values_check <- values_data %in% values_valid

# Identify errors
errors_Protocols_WACA <- qaqc_meta(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- MetadataReport_all %>% 
  filter(AdministrationUnit_Name == "GRCA",
         PlotType_Name == "FMH")
query <- "Protocols"
query_message <- paste0("Park = ", data$Park, ". ", "Protocol") 
values_data <- data$Protocols
values_valid <- c("Surface Fuels",
                  "Post Burn Severity",
                  "Trees - Individuals (metric)",
                  "Density - Quadrats (metric)",
                  "Density - Belts (metric)",
                  "Cover - Species Composition (metric)",
                  "Cover - Points (metric)")
values_check <- values_data %in% values_valid

# Identify errors
errors_Protocols_FMH <- qaqc_meta(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- MetadataReport_all %>% 
  filter(AdministrationUnit_Name == "GRCA",
         PlotType_Name == "RAP")
query <- "Protocols"
query_message <- paste0("Park = ", data$Park, ". ", "Protocol") 
values_data <- data$Protocols
values_valid <- c("Surface Fuels",
                  "Post Burn Severity",
                  "Trees - Individuals (metric)",
                  "Density - Quadrats (metric)",
                  "Cover - Species Composition (metric)")
values_check <- values_data %in% values_valid

# Identify errors
errors_Protocols_RAP <- qaqc_meta(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- MetadataReport_all %>% 
  filter(AdministrationUnit_Name == "GRCA",
         PlotType_Name == "Meadow")
query <- "Protocols"
query_message <- paste0("Park = ", data$Park, ". ", "Protocol") 
values_data <- data$Protocols
values_valid <- c("Surface Fuels",
                  "Post Burn Severity",
                  "Trees - Individuals (metric)",
                  "Density - Quadrats (metric)",
                  "Density - Belts (metric)",
                  "Cover - Species Composition (metric)",
                  "Cover - Points (metric)")
values_check <- values_data %in% values_valid

# Identify errors
errors_Protocols_Meadow <- qaqc_meta(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- MetadataReport_all %>% 
  filter(AdministrationUnit_Name == "SCP I&M",
         PlotType_Name == "I&M")
query <- "Protocols"
query_message <- paste0("Park = ", data$Park, ". ", "Protocol") 
values_data <- data$Protocols
values_valid <- c("Surface Fuels",
                  "Post Burn Severity",
                  "Trees - Individuals (metric)",
                  "Trees - Saplings (Diameter Class) (metric)",
                  "Density - Quadrats (metric)",
                  "Cover/Frequency (metric)")
values_check <- values_data %in% values_valid

# Identify errors
errors_Protocols_IM <- qaqc_meta(data, query, query_message, values_check)
```

#### Combine Protocols Errors

```{r}
# Combine
errors_Protocols <- unique(rbind(errors_Protocols_WACA, errors_Protocols_FMH, errors_Protocols_Meadow, errors_Protocols_RAP, errors_Protocols_IM))

remove(errors_Protocols_WACA, errors_Protocols_FMH, errors_Protocols_Meadow, errors_Protocols_RAP, errors_Protocols_IM)

if(nrow(errors_Protocols) > 1) {
  errors_Protocols <- errors_Protocols %>%
    filter(Error != "No Error")
} else {
  errors_Protocols <- errors_Protocols}
```

## Time Since

[Problem:]{.underline} Time since disturbance does not match when comparing Sample Event Date to Disturbance Date.

[Procedure:]{.underline}

-   Check that TimeSince matches expected year

```{r}
# Set parameters 
data <- MetadataReport_all %>% 
  filter(TimeSince == "0")
query <- "Time Since Disturbance"
query_message <- paste0("MonStatus = ", data$MonStatus, ". ", "Fire year") 
values_data <- data$SampleEventDate_Year
values_valid <- data$DisturbanceDate_Year
values_check <- (values_data - values_valid <= 1) %>% replace_na(TRUE)

# Identify errors
errors_TimeSince_Post <- qaqc_meta(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- MetadataReport_all %>% 
  filter(TimeSince == "1")
query <- "Time Since Disturbance"
query_message <- paste0("MonStatus = ", data$MonStatus, ". ", "Fire year") 
values_data <- data$SampleEventDate_Year
values_valid <- data$DisturbanceDate_Year
values_check <- (values_data - values_valid == 1) %>% replace_na(TRUE)

# Identify errors
errors_TimeSince_Year1 <- qaqc_meta(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- MetadataReport_all %>% 
  filter(TimeSince == "2")
query <- "Time Since Disturbance"
query_message <- paste0("MonStatus = ", data$MonStatus, ". ", "Fire year") 
values_data <- data$SampleEventDate_Year
values_valid <- data$DisturbanceDate_Year
values_check <- (values_data - values_valid == 2) %>% replace_na(TRUE)

# Identify errors
errors_TimeSince_Year2 <- qaqc_meta(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- MetadataReport_all %>% 
  filter(TimeSince == "5")
query <- "Time Since Disturbance"
query_message <- paste0("MonStatus = ", data$MonStatus, ". ", "Fire year") 
values_data <- data$SampleEventDate_Year
values_valid <- data$DisturbanceDate_Year
values_check <- (values_data - values_valid == 5) %>% replace_na(TRUE)

# Identify errors
errors_TimeSince_Year5 <- qaqc_meta(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- MetadataReport_all %>% 
  filter(TimeSince == "10")
query <- "Time Since Disturbance"
query_message <- paste0("MonStatus = ", data$MonStatus, ". ", "Fire year") 
values_data <- data$SampleEventDate_Year
values_valid <- data$DisturbanceDate_Year
values_check <- (values_data - values_valid == 10) %>% replace_na(TRUE)

# Identify errors
errors_TimeSince_Year10 <- qaqc_meta(data, query, query_message, values_check)
```

```{r}
# Set parameters 
data <- MetadataReport_all %>% 
  filter(TimeSince == "20")
query <- "Time Since Disturbance"
query_message <- paste0("MonStatus = ", data$MonStatus, ". ", "Fire year") 
values_data <- data$SampleEventDate_Year
values_valid <- data$DisturbanceDate_Year
values_check <- (values_data - values_valid == 20) %>% replace_na(TRUE)

# Identify errors
errors_TimeSince_Year20 <- qaqc_meta(data, query, query_message, values_check)
```

#### Combine TimeSince Errors

```{r}
# Combine
errors_TimeSince <- unique(rbind(errors_TimeSince_Post, errors_TimeSince_Year1, errors_TimeSince_Year2, errors_TimeSince_Year5, errors_TimeSince_Year10, errors_TimeSince_Year20))

remove(errors_TimeSince_Post, errors_TimeSince_Year1, errors_TimeSince_Year2, errors_TimeSince_Year5, errors_TimeSince_Year10, errors_TimeSince_Year20)

if(nrow(errors_TimeSince) > 1) {
  errors_TimeSince <- errors_TimeSince %>%
    filter(Error != "No Error")
} else {
  errors_TimeSince <- errors_TimeSince}
```

## Visited

[Problem:]{.underline} Visited = N

[Procedure:]{.underline}

-   Check that Visited = Y

```{r}
# Set parameters 
data <- MetadataReport_all
query <- "Visited"
query_message <- paste0("MonStatus = ", data$MonStatus, ". ", "Protocol = ", data$Protocols, ". ", "Visited") 
values_data <- data$Visited
values_valid <- "Y"
values_check <- values_data == values_valid

# Identify errors
errors_Visited <- qaqc_meta(data, query, query_message, values_check)
```

# ALL ERRORS

```{r}
errors_all <- rbind(errors_Disturbance, errors_blank_meta, errors_MonStatus, errors_blank_meta, errors_ProjectUnit, errors_blank_meta, errors_TimeSince, errors_blank_meta, errors_Visited)
```

```{r}
# Create conditional formatting rules
errors_all_cf <- errors_all %>% 
  mutate(Color = ifelse(SavedQuery == "", NA, "Color"))

# Add conditional formating to ouput
errors_all <- condformat(errors_all_cf[1:(nrow(errors_all_cf)), 1:8]) %>%
  rule_fill_discrete(columns = 1:7,
                     expression = (errors_all_cf$Color),
                     colours = c("Color"="#C6E0B4"),
                     na.value = "#FFFFFF") %>%
  select(!c("Color"))
```

```{r}
# Table of results for quick check
#kable(errors_all, "pipe")

# Save as CSV or XLSX 
path_errors
# write.csv(errors_all, paste0(path_errors, "errors_GRCA_FMH.csv"), quote=FALSE, row.names = FALSE, na = "") 
#write_xlsx(errors_all, paste0(path_errors, "errors_GRCA_FMH.xlsx"))
# condformat2excel(errors_all, paste0(path_errors, "errors_GRCA_FMH_formatted.xlsx"), overwrite_wb = TRUE)
#condformat2excel(errors_all, paste0(path_errors, "errors_metadata_formatted.xlsx"), sheet_name = path_errors_name, overwrite_wb = TRUE)
```
