---
title: "3_GRCA_FMH"
author: "Alexandra Lalor"
output:
  html_document:
    theme: readable
    highlight: 
    toc: yes
    toc_depth: 3
    toc_float:
      smooth_scroll: yes
    code_download: true
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

# BEFORE STARTING

### Install Packages (if needed)

```{r}
# install.packages("here")
# install.packages("tidyverse")
# install.packages("Rtools")
# install.packages("janitor")
# install.packages("condformat")
# install.packages("knitr")
```

### Load Packages

```{r}
library(here)
# tidyverse and dplyr have lots of useful functions for data cleaning
library(tidyverse)
library(dplyr)
#library(plyr)
# janitor has functions to find duplicates
library(janitor)
# EnvStats is needed for the rosnerTest() function
library(EnvStats)
# writexl is used to create excel output files
library(writexl)
# condformat is used to format excel output files
library(condformat)
# knitr is used to create output files from R Markdown
library(knitr)
```

### Adjust File Paths

Make sure to update file paths to be specific for your data.

```{r}
# Identify working directory (specifically user name)
here()

# Load in data. Change file path based on user name!
path_data <- "X:/FFI Data Management/Exports from FFI/GRCA_FMH/2025-01-23/"
path_errors <- paste0(here(), "/output/errors/")

path_data <- "D:/Allie/_FireFX/FFI Data Management/Exports from FFI/GRCA_FMH/2025-01-23/"
#path_errors <- "C:/Users/allie/OneDrive/Desktop/R Projects/GRCA/output/errors/"
```

### Create Function

```{r}
# Blank data frame
errors_blank_CrossYear <- data.frame("SavedQuery" = "", "MacroPlot.Name" = "", "Date" = "", "Monitoring.Status" = "", "TagNo" = "", "Status" = "", "DBH" = "", "Ht" = "", "CrwnCl" = "", "LiCrBHt" = "", "Species.Symbol" = "", "Comment" = "", "Error" = "", "Fixed" = "", "Explanation" = "", "Queryers" = "") %>% 
  mutate(Date = as.Date(Date))

# QAQC function
qaqc_CrossYear <- function(data, query, query_message, values_check) {
  
  # Check if there is data to QAQC
  if(nrow(data) == 0) {
  # If there is no data, default to "No Error" data frame
  errors <- errors_blank_CrossYear %>%
    mutate(SavedQuery = query,
           Error = "No Error")
  # If there is data, perform error check
} else {
  
  # Use relevant dataset to look for errors
  errors <- data %>%  
    # Add columns relevant to error checking
           # Populate the "SavedQuery" column with the query name
    mutate("SavedQuery" = query, 
           # Populate the "Error" column with the query message and relevant data
           "Error" = paste(query_message, "=", values_data),
           # Create a blank "Fixed" column
           "Fixed" = "",
           # Create a blank "Explanation" column
           "Explanation" = "",
           # Create a blank "Queryers" column
           "Queryers" = "") %>%   
    # Filter for data which is "false" (data does not match valid conditions)
    # NA values are treated as "false" (missing data is not valid)
    filter(!(values_check %>% replace_na(FALSE))) %>%   
    # Select relevant columns to view errors
    select("SavedQuery", "MacroPlot.Name", "Date", "Monitoring.Status", "TagNo", "Status", "DBH", "Ht", "CrwnCl", "LiCrBHt", "Species.Symbol", "Comment", "Error", "Fixed", "Explanation", "Queryers")
  
  # Next, check if there are duplicate errors
  errors_temp <- errors %>% 
    get_dupes(SavedQuery,MacroPlot.Name, Date, Error)
  # Merge duplicate errors
  errors <- unique(merge(errors, errors_temp, all = T))
  # If duplicate errors exist, add number of duplicates to error message
  errors <- errors %>% 
    mutate(Error = ifelse(is.na(dupe_count), Error, paste0("(x", dupe_count, ") ", Error))) %>% 
    select(!"dupe_count")
}
  
  # Check if there are no errors
  if (nrow(errors) == 0) {  
    # If there are no errors, default to "No Error" data frame
    errors <- errors_blank_CrossYear %>%
      mutate(SavedQuery = query,
             Error = "No Error")
    # If there are errors, keep error log you just created
  } else {   
    errors <- errors
  }
}
```

### Load Data

```{r}
# # Load in data
# Fuels1000_all <- read.csv(paste0(path_data, "Surface Fuels - 1000Hr_XPT.csv"), quote = "")
# FuelsFine_all <- read.csv(paste0(path_data, "Surface Fuels - Fine_XPT.csv"), quote = "")
Trees_all <- read.csv(paste0(path_data, "Trees - Individuals (metric)_XPT.csv"), quote = "")
```

### Clean Data

```{r}
# #####
# # Fuels1000
# #####
# # Reformat date column
# Fuels1000_all <- Fuels1000_all %>%
#   # Separate Date column
#   separate(Date, sep = " ", into = c("Date", "Time")) %>%
#   separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>% 
#   # Reformat Date column
#   mutate(Date = as.Date(Date, format = "%m/%d/%Y")) %>% 
#   # Remove Month, Day, and Time columns
#   select(!c("Month", "Day", "Time"))
# 
# #####
# # FuelsFine
# #####
# # Reformat date column
# FuelsFine_all <- FuelsFine_all %>%
#   # Separate Date column
#   separate(Date, sep = " ", into = c("Date", "Time")) %>%
#   separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>% 
#   # Reformat Date column
#   mutate(Date = as.Date(Date, format = "%m/%d/%Y")) %>% 
#   # Remove Month, Day, and Time columns
#   select(!c("Month", "Day", "Time"))

#####
# Trees
#####
# Reformat date column
Trees_all <- Trees_all %>%
  # Separate Date column
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>% 
  # Reformat Date column
  mutate(Date = as.Date(Date, format = "%m/%d/%Y")) %>% 
  # Remove Month, Day, and Time columns
  select(!c("Month", "Day", "Time"))
```

### Filter Data

```{r}
# #####
# # Fuels1000
# #####
# # Ensure blanks in Visited column are NA
# Fuels1000_all$Visited[Fuels1000_all$Visited==""] <- NA
# # Fuels1000 data
# Fuels1000_data <- Fuels1000_all %>%
#   filter(is.na(Visited))
# 
# 
# #####
# # FuelsFine
# #####
# # Ensure blanks in Visited column are NA 
# FuelsFine_all$Visited[FuelsFine_all$Visited==""] <- NA  
# # FuelsFine data
# FuelsFine_data <- FuelsFine_all %>%   
#   filter(is.na(Visited))  


#####
# Trees
#####
# Ensure blanks in Visited column are NA 
Trees_all$Visited[Trees_all$Visited==""] <- NA
# Trees data
Trees_data_CrossYear <- Trees_all %>%   
  filter(is.na(Visited))  
```

# PROTOCOL - SURFACE FUELS

## 1000HR (CWD) & FINE FUELS (FWD)

This code conducts cross-year quality control checks on 2 surface fuels datasets

-   coarse woody debris (CWD) surface fuels data within the "Surface Fuels - 1000Hr" data set

-   fine woody debris (FWD) surface fuels data within the "Surface Fuels - Fine" data set.

It checks for errors in: Fuel Slope consistency across years, Fuel Azimuth consistency across years

### Fuel Slope

### Fuel Azimuth

# PROTOCOL - TREES

This code conducts quality control checks on tree data within the "Trees - Individuals (metric)" data set.

It checks for errors in:

### Add Columns

#### CBH Columns

-   MaxCBH = What is the tallest recorded CBH of each individual **live** tree? (Dead trees do not have CBH)

-   MaxCBHDate = What date was the tallest CBH recorded?

-   MaxCBHDiff = What is the size difference between the tallest recorded CBH, and the sample event CBH?

```{r}
data_temp_CBH <- Trees_data_CrossYear %>% 
  filter(Status == "L",
         !is.na(LiCrBHt)) %>% 
  group_by(TagNo, MacroPlot.Name, SubFrac) %>% 
  mutate(MaxCBH = max(LiCrBHt),
         MaxCBHDiff = MaxCBH - LiCrBHt) %>% 
  ungroup() %>% 
  arrange(TagNo, MacroPlot.Name, Date, SubFrac)

data_temp_MaxCBHDate <- data_temp_CBH %>% 
  filter(LiCrBHt == MaxCBH) %>% 
  group_by(TagNo, MacroPlot.Name, SubFrac) %>% 
  mutate(MaxCBHDate = max(Date)) %>% 
  select(TagNo, MacroPlot.Name, SubFrac, MaxCBHDate) %>% 
  unique()

data_temp_merge_CBH <- merge(data_temp_CBH, data_temp_MaxCBHDate, all = T)
remove(data_temp_CBH, data_temp_MaxCBHDate)
```

#### DBH Columns

-   MinDBH = What is the smallest recorded DBH of each individual **dead** tree?

-   MinDBHDate = What date was the smallest DBH recorded?

-   MinDBHDiff = What is the size difference between the smallest recorded DBH, and the sample event DBH?

-   MaxDBH = What is the largest recorded DBH of each individual **live** tree?

-   MaxDBHDate = What date was the largest DBH recorded?

-   MaxDBHDiff = What is the size difference between the largest recorded DBH, and the sample event DBH?

```{r}
data_temp_DBH <- Trees_data_CrossYear %>% 
  filter(Status == "D",
         !is.na(DBH)) %>% 
  group_by(TagNo, MacroPlot.Name, SubFrac) %>% 
  mutate(MinDBH = min(DBH),
         MinDBHDiff = DBH - MinDBH) %>% 
  ungroup() %>% 
  arrange(TagNo, MacroPlot.Name, Date, SubFrac)

data_temp_MinDBHDate <- data_temp_DBH %>% 
  filter(DBH == MinDBH) %>% 
  group_by(TagNo, MacroPlot.Name, SubFrac) %>% 
  mutate(MinDBHDate = min(Date)) %>% 
  select(TagNo, MacroPlot.Name, SubFrac, MinDBHDate) %>% 
  unique()

data_temp_merge_DeadDBH <- merge(data_temp_DBH, data_temp_MinDBHDate, all = T)

############################################################################################

data_temp_DBH <- Trees_data_CrossYear %>% 
  filter(Status == "L",
         !is.na(DBH)) %>% 
  group_by(TagNo, MacroPlot.Name, SubFrac) %>% 
  mutate(MaxDBH = max(DBH),
         MaxDBHDiff = MaxDBH - DBH) %>% 
  ungroup() %>% 
  arrange(TagNo, MacroPlot.Name, Date, SubFrac)

data_temp_MaxDBHDate <- data_temp_DBH %>% 
  filter(DBH == MaxDBH) %>% 
  group_by(TagNo, MacroPlot.Name, SubFrac) %>% 
  mutate(MaxDBHDate = max(Date)) %>% 
  select(TagNo, MacroPlot.Name, SubFrac, MaxDBHDate) %>% 
  unique()

data_temp_merge_LiveDBH <- merge(data_temp_DBH, data_temp_MaxDBHDate, all = T)

############################################################################################

data_temp_merge_DBH <- merge(data_temp_merge_DeadDBH, data_temp_merge_LiveDBH, all = T)
remove(data_temp_merge_DeadDBH, data_temp_merge_LiveDBH, data_temp_DBH, data_temp_MinDBHDate, data_temp_MaxDBHDate)
```

#### Height Columns

-   MinHeight= What is the shortest recorded Height of each individual **dead** tree?

-   MinHeightDate = What date was the shortest Height recorded?

-   MinHeightDiff = What is the size difference between the shortest recorded Height, and the sample event Height?

-   MaxHeight = What is the tallest recorded Height of each individual **live** tree?

-   MaxHeightDate = What date was the tallest Height recorded?

-   MaxHeightDiff = What is the size difference between the tallest recorded Height, and the sample event Height?

```{r}
data_temp_Height <- Trees_data_CrossYear %>% 
  filter(Status == "D",
         !is.na(Ht)) %>% 
  group_by(TagNo, MacroPlot.Name, SubFrac) %>% 
  mutate(MinHeight = min(Ht),
         MinHeightDiff = Ht - MinHeight) %>% 
  ungroup() %>% 
  arrange(TagNo, MacroPlot.Name, Date, SubFrac)

data_temp_MinHeightDate <- data_temp_Height %>% 
  filter(Ht == MinHeight) %>% 
  group_by(TagNo, MacroPlot.Name, SubFrac) %>% 
  mutate(MinHeightDate = min(Date)) %>% 
  select(TagNo, MacroPlot.Name, SubFrac, MinHeightDate) %>% 
  unique()

data_temp_merge_DeadHeight <- merge(data_temp_Height, data_temp_MinHeightDate, all = T)

############################################################################################

data_temp_Height <- Trees_data_CrossYear %>% 
  filter(Status == "L",
         !is.na(Ht)) %>% 
  group_by(TagNo, MacroPlot.Name, SubFrac) %>% 
  mutate(MaxHeight = max(Ht),
         MaxHeightDiff = MaxHeight - Ht) %>% 
  ungroup() %>% 
  arrange(TagNo, MacroPlot.Name, Date, SubFrac)

data_temp_MaxHeightDate <- data_temp_Height %>% 
  filter(Ht == MaxHeight) %>% 
  group_by(TagNo, MacroPlot.Name, SubFrac) %>% 
  mutate(MaxHeightDate = max(Date)) %>% 
  select(TagNo, MacroPlot.Name, SubFrac, MaxHeightDate) %>% 
  unique()

data_temp_merge_LiveHeight <- merge(data_temp_Height, data_temp_MaxHeightDate, all = T)

############################################################################################

data_temp_merge_Height <- merge(data_temp_merge_DeadHeight, data_temp_merge_LiveHeight, all = T)
remove(data_temp_merge_DeadHeight, data_temp_merge_LiveHeight, data_temp_Height, data_temp_MinHeightDate, data_temp_MaxHeightDate)
```

#### Species Columns

-   Species_Count = How many unique species have been recorded for each individual tree? Count \> 1 means that the tree has changed species.

```{r}
Trees_data_CrossYear_Species <- Trees_data_CrossYear %>% 
  group_by(TagNo, MacroPlot.Name, SubFrac) %>% 
  mutate(Species_Count = n_distinct(unique(Species.Symbol))) %>% 
  ungroup() %>% 
  arrange(TagNo, MacroPlot.Name, Date, SubFrac)

data_temp_Species <- Trees_data_CrossYear_Species %>% 
  filter(Species_Count > 1) %>% 
  group_by(TagNo, MacroPlot.Name, SubFrac) %>% 
  mutate(Species_Change = ifelse(duplicated(Species.Symbol), "N", "Y")) %>% 
  filter(Species_Change == "Y") %>% 
  unique()

data_temp_merge_Species <- data_temp_Species
remove(Trees_data_CrossYear_Species, data_temp_Species)
```

#### Status Columns

-   Status_Count = How many unique statuses have been recorded for each individual tree? Count of 1 means that the tree has been alive or dead for the entirety of data collection. Count of 2 means that the tree has been both alive and dead.

-   MinDeadDate = What date was the tree first recorded as Dead?

-   MinLiveDate = What date was the tree first recorded as Live?

```{r}
Trees_data_CrossYear_Status <- Trees_data_CrossYear %>% 
  group_by(TagNo, MacroPlot.Name, SubFrac) %>% 
  mutate(Status_Count = n_distinct(unique(Status))) %>% 
  ungroup() %>% 
  arrange(TagNo, MacroPlot.Name, Date, SubFrac)

# add min death date to data_temp
data_temp_Status_AllDead <- Trees_data_CrossYear_Status %>% 
  filter(Status_Count == 1 & Status == "D")
data_temp_Status_SomeDead <- Trees_data_CrossYear_Status %>% 
  filter(Status_Count > 1)
data_temp_Status_Dead <- rbind(data_temp_Status_AllDead, data_temp_Status_SomeDead)

data_temp_MinDeadDate <- data_temp_Status_Dead %>% 
  filter(Status == "D") %>% 
  group_by(TagNo, MacroPlot.Name, SubFrac) %>% 
  mutate(MinDeadDate = min(Date)) %>% 
  select(TagNo, MacroPlot.Name, SubFrac, MinDeadDate) %>% 
  unique()

data_temp_merge_Status_Dead <- merge(data_temp_Status_Dead, data_temp_MinDeadDate, all = T)

# add min live date to data_temp
data_temp_Status_AllLive <- Trees_data_CrossYear_Status %>% 
  filter(Status_Count == 1 & Status == "L")
data_temp_Status_SomeLive <- Trees_data_CrossYear_Status %>% 
  filter(Status_Count > 1)
data_temp_Status_Live <- rbind(data_temp_Status_AllLive, data_temp_Status_SomeLive)

data_temp_MinLiveDate <- data_temp_Status_Live %>% 
  filter(Status == "L") %>% 
  group_by(TagNo, MacroPlot.Name, SubFrac) %>% 
  mutate(MinLiveDate = min(Date)) %>% 
  select(TagNo, MacroPlot.Name, SubFrac, MinLiveDate) %>% 
  unique()

data_temp_merge_Status_Live <- merge(data_temp_Status_Live, data_temp_MinLiveDate, all = T)

# merge
data_temp_merge_Status <- merge(data_temp_merge_Status_Dead, data_temp_merge_Status_Live, all = T)

remove(Trees_data_CrossYear_Status, data_temp_Status_AllDead, data_temp_Status_SomeDead, data_temp_Status_Dead, data_temp_MinDeadDate, data_temp_merge_Status_Dead, data_temp_Status_AllLive, data_temp_Status_SomeLive, data_temp_Status_Live, data_temp_MinLiveDate, data_temp_merge_Status_Live)
```

### Trees CBH

#### CBH Decrease

[Problem:]{.underline} Live trees have Crown Base Height (CBH) which decreases over time.

[Procedure:]{.underline}

-   Filter for Status = L. Filter to exclude a certain threshold where the difference in CBH is not a concern. Remaining entries should include all CBHs that are smaller than the maximum CBH.

-   We are concerned with live trees which decrease in CBH over time. Ensure that each sample event date is *earlier* than the date with the tallest recorded CBH (**MaxCBHDate**).

```{r}
# Set parameters 
data <- merge(Trees_data_CrossYear, data_temp_merge_CBH, all = T) %>% 
  filter(Status == "L", !is.na(LiCrBHt),
         # set threshold of concern
         MaxCBHDiff >= 5)
query <- "Trees Live CBH Decrease" 
query_message <- paste0("LiCrBHt (CBH) = ", data$LiCrBHt, ". ", "Max CBH = ", data$MaxCBH, " on date")
values_data <- data$MaxCBHDate
values_valid <- data$Date
values_check <- values_data >= values_valid

# Identify errors
errors_Trees_CrossYear_CBH <- qaqc_CrossYear(data, query, query_message, values_check) %>% 
  arrange(MacroPlot.Name, TagNo, Date)
```

### Trees DBH

#### Live DBH Decrease

[Problem:]{.underline} Live trees have DBH which decreases over time.

[Procedure:]{.underline}

-   Filter for Status = L. Filter to exclude a certain threshold where the difference in DBH is not a concern. Remaining entries should include all DBHs that are smaller than the maximum DBH.

-   We are concerned with live trees which decrease in DBH while they are still alive. Ensure that each sample event date is *earlier* than the date with the largest recorded DBH (**MaxDBHDate**).

```{r}
# Set parameters 
data <- merge(Trees_data_CrossYear, data_temp_merge_DBH, all = T) %>% 
  filter(Status == "L", !is.na(DBH),
         # set threshold of concern
         MaxDBHDiff >= 5)
query <- "Trees Live DBH Decrease" 
query_message <- paste0("Status = ", data$Status, ". ", "DBH = ", data$DBH, ". ", "Max DBH = ", data$MaxDBH, " on date")
values_data <- data$MaxDBHDate
values_valid <- data$Date
values_check <- values_data >= values_valid

# Identify errors
errors_Trees_CrossYear_DBH_Live <- qaqc_CrossYear(data, query, query_message, values_check) %>% 
  arrange(MacroPlot.Name, TagNo, Date)
```

#### Dead DBH Increase

[Problem:]{.underline} Dead trees have DBH which increases over time.

[Procedure:]{.underline}

-   Filter for Status = D. Filter to exclude a certain threshold where the difference in DBH is not a concern. Remaining entries should include all DBHs that are larger than the minimum DBH.

-   We are concerned with dead trees which increase DBH post-death. Ensure that each sample event date is *earlier* than the date with the smallest recorded DBH (**MinDBHDate**).

```{r}
# Set parameters 
data <- merge(Trees_data_CrossYear, data_temp_merge_DBH, all = T) %>% 
  filter(Status == "D", !is.na(DBH),
         # set threshold of concern
         MinDBHDiff >= 5)
query <- "Trees Dead DBH Increase" 
query_message <- paste0("Status = ", data$Status, ". ", "DBH = ", data$DBH, ". ", "Min DBH = ", data$MinDBH, " on date")
values_data <- data$MinDBHDate
values_valid <- data$Date
values_check <- values_data >= values_valid

# Identify errors
errors_Trees_CrossYear_DBH_Dead <- qaqc_CrossYear(data, query, query_message, values_check) %>% 
  arrange(MacroPlot.Name, TagNo, Date)
```

### Trees Height

#### Live Height Decrease

[Problem:]{.underline} Live trees have Height which decreases over time.

[Procedure:]{.underline}

-   Filter for Status = L. Filter to exclude a certain threshold where the difference in Height is not a concern. Remaining entries should include all Heights that are shorter than the maximum Height.

-   We are concerned with live trees which decrease in Height over time. Ensure that each sample event date is *earlier* than the date with the tallest recorded Height (**MaxHeightDate**).

```{r}
# Set parameters 
data <- merge(Trees_data_CrossYear, data_temp_merge_Height, all = T) %>% 
  filter(Status == "L", !is.na(Ht),
         # set threshold of concern
         MaxHeightDiff >= 5)
query <- "Trees Live Height Decrease" 
query_message <- paste0("Status = ", data$Status, ". ", "Height = ", data$Ht, ". ", "Max Height = ", data$MaxHeight, " on date")
values_data <- data$MaxHeightDate
values_valid <- data$Date
values_check <- values_data >= values_valid

# Identify errors
errors_Trees_CrossYear_Height_Live <- qaqc_CrossYear(data, query, query_message, values_check) %>% 
  arrange(MacroPlot.Name, TagNo, Date)
```

#### Dead Height Increase

[Problem:]{.underline} Dead trees have Height which increases over time.

[Procedure:]{.underline}

-   Filter for Status = D. Filter to exclude a certain threshold where the difference in Height is not a concern. Remaining entries should include all Heights that are taller than the minimum Height.

-   We are concerned with dead trees which increase in Height over time. Ensure that each sample event date is *earlier* than the date with the shortest recorded Height (**MinHeightDate**).

```{r}
# Set parameters 
data <- merge(Trees_data_CrossYear, data_temp_merge_Height, all = T) %>% 
  filter(Status == "D", !is.na(Ht),
         # set threshold of concern
         MinHeightDiff >= 5)
query <- "Trees Dead Height Increase" 
query_message <- paste0("Status = ", data$Status, ". ", "Height = ", data$Ht, ". ", "Min Height = ", data$MinHeight, " on date")
values_data <- data$MinHeightDate
values_valid <- data$Date
values_check <- values_data >= values_valid

# Identify errors
errors_Trees_CrossYear_Height_Dead <- qaqc_CrossYear(data, query, query_message, values_check) %>% 
  arrange(MacroPlot.Name, TagNo, Date)
```

### Trees Species

#### Species Change

[Problem:]{.underline} Trees change species over time.

[Procedure:]{.underline}

-   Isolate rows which have more than 1 entry for Species.Symbol for each individual tree, and the earliest date when each species was recorded.

```{r}
# Set parameters 
data <- merge(Trees_data_CrossYear, data_temp_merge_Species, all = T)
query <- "Trees Species Change" 
query_message <- paste0("Species = ", data$Species.Symbol, ". ", "Species Change")
values_data <- data$Species_Change
values_valid <- "Y"
values_check <- (values_data != values_valid) %>%  replace_na(TRUE)

# Identify errors
errors_Trees_CrossYear_Species <- qaqc_CrossYear(data, query, query_message, values_check) %>% 
  arrange(MacroPlot.Name, TagNo, Date)
```

### Trees Status

#### Dead to Live

[Problem:]{.underline} Trees change status from D to L over time.

[Procedure:]{.underline}

-   Filter for Status = L. Check that the sample event date is not later than the death date.

```{r}
# Set parameters 
data <- merge(Trees_data_CrossYear, data_temp_merge_Status, all = T) %>% 
  filter(Status == "L")
query <- "Trees Status Dead to Live" 
query_message <- paste0("Status = ", data$Status, ". ", "Death date")
values_data <- data$MinDeadDate
values_valid <- data$Date
values_check <- (values_data >= values_valid) %>%  replace_na(TRUE)

# Identify errors
errors_Trees_CrossYear_Status <- qaqc_CrossYear(data, query, query_message, values_check) %>% 
  arrange(MacroPlot.Name, TagNo, Date)
```

# ALL ERRORS

```{r}
errors_Trees_CrossYear <- rbind(errors_Trees_CrossYear_CBH, errors_blank_CrossYear, errors_Trees_CrossYear_DBH_Live, errors_blank_CrossYear, errors_Trees_CrossYear_DBH_Dead, errors_blank_CrossYear, errors_Trees_CrossYear_Height_Live, errors_blank_CrossYear, errors_Trees_CrossYear_Height_Dead, errors_blank_CrossYear, errors_Trees_CrossYear_Species, errors_blank_CrossYear, errors_Trees_CrossYear_Status)

remove(errors_Trees_CrossYear_CBH, errors_Trees_CrossYear_DBH_Live, errors_Trees_CrossYear_DBH_Dead, errors_Trees_CrossYear_Height_Live, errors_Trees_CrossYear_Height_Dead, errors_Trees_CrossYear_Species,  errors_Trees_CrossYear_Status)
```

```{r}
# Table of results for quick check
#kable(errors_all, "pipe")

# Save as CSV or XLSX 
path_errors

write_xlsx(errors_Trees_CrossYear, paste0(path_errors, "errors_CrossYear.xlsx"))
```

# Data Viz - Trees

```{r}
graph_Trees_data <- Trees_data_CrossYear %>% 
  mutate(Status = ifelse(CrwnCl == "DD", "DD", Status)) %>% 
  mutate(DBH = (ifelse(is.na(DBH), 0 , DBH)),
         LiCrBHt = ifelse(is.na(LiCrBHt), 0, LiCrBHt),
         Ht = ifelse(is.na(Ht), 0, Ht)) %>% 
  mutate(SubFrac = factor(SubFrac),
         TagNo = factor(TagNo),
         Status = factor(Status, levels = c("L", "D", "DD")),
         CrwnCl = factor(CrwnCl, levels = c("X", "",
                                            "SC", "I", "C", "D",
                                            "RS", "LBS", "CS", "BAD", 
                                            "BBD", "CUS", "DD")))
```

### Isolate Problematic Tree Tags

```{r}
#unique(Trees_data_CrossYear$MacroPlot.Name)
target_plot <- "PIAB 15"
```

```{r}
badtags_CBH <- errors_Trees_CrossYear %>% 
  filter(grepl("CBH", errors_Trees_CrossYear$SavedQuery)) %>% 
  filter(MacroPlot.Name == target_plot)
badtags_CBH <- unique(badtags_CBH$TagNo)

badtags_DBH <- errors_Trees_CrossYear %>% 
  filter(grepl("DBH", errors_Trees_CrossYear$SavedQuery)) %>% 
  filter(MacroPlot.Name == target_plot)
badtags_DBH <- unique(badtags_DBH$TagNo)

badtags_Height <- errors_Trees_CrossYear %>% 
  filter(grepl("Height", errors_Trees_CrossYear$SavedQuery)) %>% 
  filter(MacroPlot.Name == target_plot)
badtags_Height <- unique(badtags_Height$TagNo)

badtags_Species <- errors_Trees_CrossYear %>% 
  filter(grepl("Species", errors_Trees_CrossYear$SavedQuery)) %>% 
  filter(MacroPlot.Name == target_plot)
badtags_Species <- unique(badtags_Species$TagNo)

badtags_Status <- errors_Trees_CrossYear %>% 
  filter(grepl("Status", errors_Trees_CrossYear$SavedQuery)) %>% 
  filter(MacroPlot.Name == target_plot)
badtags_Status <- unique(badtags_Status$TagNo)
```

## Graphs: [`r target_plot`]{style="color:red"}

### Status

```{r}
graph_Trees_data %>% 
  filter(MacroPlot.Name == target_plot,
         TagNo %in% c(badtags_Status)) %>% 
  arrange(Date) %>% 
  ggplot(aes(x = Year, y = TagNo)) +
  geom_line(aes(group = TagNo, color = Status), size = 1) +
  geom_point(aes(fill = Status, size = DBH), shape = 21) +
  scale_fill_manual(values = c("L" = "forestgreen", "D" = "darkorange", "DD" = "red")) +
  scale_color_manual(values=c("L" = "forestgreen", "D" = "darkorange", "DD" = "red")) +
  labs(title = target_plot,
       caption = "0 values represent missing data") +
  theme_light() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(size = 20, face="bold", hjust = 0.5),
        plot.subtitle = element_text(size = 12, face="italic", hjust = 0.5),
        axis.title = element_text(size = 12, face="bold"),
        axis.title.y = element_text(margin = unit(c(0,4,0,0), "mm")),
        axis.title.x = element_text(margin = unit(c(4,0,0,0), "mm")),
        axis.text.y = element_text(size = 5),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1))
```

### Species

```{r}
graph_Trees_data %>% 
  filter(MacroPlot.Name == target_plot,
         TagNo %in% c(badtags_Species)) %>% 
  arrange(Date) %>% 
  ggplot(aes(x = Year, y = TagNo)) +
  geom_line(aes(group = TagNo, color = Species.Symbol), size = 1) +
  geom_point(aes(fill = Status), shape = 21, size = 3) +
  scale_fill_manual(values = c("L" = "forestgreen", "D" = "darkorange", "DD" = "red")) +
  labs(title = target_plot,
       caption = "0 values represent missing data") +
  theme_light() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(size = 20, face="bold", hjust = 0.5),
        plot.subtitle = element_text(size = 12, face="italic", hjust = 0.5),
        axis.title = element_text(size = 12, face="bold"),
        axis.title.y = element_text(margin = unit(c(0,4,0,0), "mm")),
        axis.title.x = element_text(margin = unit(c(4,0,0,0), "mm")),
        axis.text.y = element_text(size = 5),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1))
```

### DBH

Live trees should only increase DBH. Dead trees should only decrease DBH.

```{r}
graph_Trees_data %>% 
  filter(MacroPlot.Name == target_plot,
         TagNo %in% c(badtags_DBH),
         Status != "DD") %>% 
  arrange(Date) %>% 
  ggplot(aes(x = Year, y = DBH)) +
  geom_line(aes(group = TagNo, color = TagNo), size = 1) +
  geom_point(aes(fill = Status), shape = 21, size = 3) +
  scale_fill_manual(values=c("L" = "forestgreen", "D" = "darkorange")) +
  labs(title = target_plot,
       caption = "0 values represent missing data") +
  facet_wrap(~Status) +
  theme_light() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(size = 20, face="bold", hjust = 0.5),
        plot.subtitle = element_text(size = 12, face="italic", hjust = 0.5),
        axis.title = element_text(size = 12, face="bold"),
        axis.title.y = element_text(margin = unit(c(0,4,0,0), "mm")),
        axis.title.x = element_text(margin = unit(c(4,0,0,0), "mm")),
        axis.text.y = element_text(size = 5),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1))
```

### Height

Live trees should only increase in height. Dead trees should only decrease in height.

```{r}
graph_Trees_data %>% 
  filter(MacroPlot.Name == target_plot,
         TagNo %in% c(badtags_Height),
         Status != "DD") %>% 
  arrange(Date) %>% 
  ggplot(aes(x = Year, y = Ht)) +
  geom_line(aes(group = TagNo, color = TagNo), size = 1) +
  geom_point(aes(fill = Status), shape = 21, size = 3) +
  scale_fill_manual(values=c("L" = "forestgreen", "D" = "darkorange")) +
  labs(title = target_plot,
       caption = "0 values represent missing data") +
  facet_wrap(~Status) +
  theme_light() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(size = 20, face="bold", hjust = 0.5),
        plot.subtitle = element_text(size = 12, face="italic", hjust = 0.5),
        axis.title = element_text(size = 12, face="bold"),
        axis.title.y = element_text(margin = unit(c(0,4,0,0), "mm")),
        axis.title.x = element_text(margin = unit(c(4,0,0,0), "mm")),
        axis.text.y = element_text(size = 5),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1))
```

### CBH

Live trees should only increase crown base height

```{r}
graph_Trees_data %>% 
  filter(MacroPlot.Name == target_plot,
         TagNo %in% c(badtags_CBH),
         Status == "L") %>% 
  arrange(Date) %>% 
  ggplot(aes(x = Year, y = LiCrBHt)) +
  geom_line(aes(group = TagNo, color = TagNo), size = 1) +
  geom_point(aes(fill = Status), shape = 21, size = 3) +
  scale_fill_manual(values=c("L" = "forestgreen")) +
  labs(title = target_plot,
       caption = "0 values represent missing data") +
  theme_light() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(size = 20, face="bold", hjust = 0.5),
        plot.subtitle = element_text(size = 12, face="italic", hjust = 0.5),
        axis.title = element_text(size = 12, face="bold"),
        axis.title.y = element_text(margin = unit(c(0,4,0,0), "mm")),
        axis.title.x = element_text(margin = unit(c(4,0,0,0), "mm")),
        axis.text.y = element_text(size = 5),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1))
```
