---
title: "FFI Database Summary"
subtitle: "Grand Canyon"
output: 
  html_document:
    toc: true
    toc-float: true
    code_download: true
date: "2025-03-17"
---

# Purpose

This report summarizes plot work by the Fire Ecology Program. The included tables and enclosed code allow a user to quickly view the overall status of program, park, project, or protocol activities in an FFI admin unit. The code summarizes records in the Macroplot Report and Sample Event Report that can be downloaded via the Utilities menu on the Project Management tab on the FFI Remote App Server.

## Before Starting

### Setup

```{r setup, include=FALSE}
## Markdown Options
knitr::opts_chunk$set(echo = FALSE, error = FALSE, warning = FALSE, message = FALSE)
```

```{r install packages, include=FALSE}
## Install packages (if needed)

# install.packages("tidyverse")
# install.packages("kableExtra")
# install.packages("ggplot2")
```

```{r load packages, include=FALSE}
## Load Packages

# "here" helps you easily find and reference your working directory
library(here)
# "tidyverse" has lots of useful functions for data cleaning
library(tidyverse)
# "kableExtra" is used to produce formatted tables
library(kableExtra)
library(ggplot2)
```

### Adjust File Paths

```{r import data, include=FALSE}
# Identify working directory
here()

# Load in data.
path_data <- "X:/FFI Data Management/Exports from FFI/GRCA_FMH/metadata/"
path_output <- paste0(here(), "/output/")
path_data <- "D:/Allie/_FireFX/FFI Data Management/Exports from FFI/metadata/"
```

### Load Data

```{r load data, include=FALSE}
## Load Data

# GRCA
GRCA_MacroplotReport <- read.csv(paste0(path_data, "GRCA_MacroplotReport.csv"))
GRCA_SampleEventReport <- read.csv(paste0(path_data, "GRCA_SampleEventReport.csv"))
#GRCA_ProjectUnitReport <- read.csv(paste0(path_data, "GRCA_ProjectUnitAssignmentReport.csv"))

# I&M
IM_MacroplotReport <- read.csv(paste0(path_data, "I&M_MacroplotReport.csv"))
IM_SampleEventReport <- read.csv(paste0(path_data, "I&M_SampleEventReport.csv"))
#IM_ProjectUnitReport <- read.csv(paste0(path_data, "I&M_ProjectUnitAssignmentReport.csv"))

# WACA
WACA_MacroplotReport <- read.csv(paste0(path_data, "WACA_MacroplotReport.csv"))
WACA_SampleEventReport <- read.csv(paste0(path_data, "WACA_SampleEventReport.csv"))
#WACA_ProjectUnitReport <- read.csv(paste0(path_data, "WACA_ProjectUnitAssignmentReport.csv"))
```

### Clean Data

-   remove trailing/leading whitespace from character columns

-   reformat date columns

-   make column names match between reports

#### Sample Event Report

```{r}
GRCA_SampleEventReport <- GRCA_SampleEventReport %>% 
  
  # remove trailing/leading whitespace
  mutate(across(where(is.character), str_trim)) %>%
  
  # separate Date columns into Date and Time columns
  separate(SampleEvent_Date, sep = " ", into = c("SampleEventDate", "SampleEventTime")) %>%
  # reformate Date columns
  mutate(SampleEventDate = as.Date(SampleEventDate, "%m/%d/%Y")) %>% 
  # separate Date columns into Year, Month, and Day columns
  separate(SampleEventDate, sep = "-", into = c("SampleEventDate_Year", "SampleEventDate_Month", "SampleEventDate_Day"), remove = FALSE) %>% 
  # remove unnecessary Date columns
  select(!c(SampleEventTime, SampleEventDate_Month, SampleEventDate_Day)) %>% 
  
  # make column names match those in MacroplotReport
  mutate(Macroplot = MacroPlot_Name,
         LegacyMonitoringStatus = SampleEvent_LegacyMonitoringStatus,
         MonStatus = MonitoringStatus_Name) %>% 
  select(!c(MacroPlot_Name, SampleEvent_LegacyMonitoringStatus, MonitoringStatus_Name))

##############################################

IM_SampleEventReport <- IM_SampleEventReport %>% 
  
  # remove trailing/leading whitespace
  mutate(across(where(is.character), str_trim)) %>%
  
  # separate Date columns into Date and Time columns
  separate(SampleEvent_Date, sep = " ", into = c("SampleEventDate", "SampleEventTime")) %>%
  # reformate Date columns
  mutate(SampleEventDate = as.Date(SampleEventDate, "%m/%d/%Y")) %>% 
  # separate Date columns into Year, Month, and Day columns
  separate(SampleEventDate, sep = "-", into = c("SampleEventDate_Year", "SampleEventDate_Month", "SampleEventDate_Day"), remove = FALSE) %>% 
  # remove unnecessary Date columns
  select(!c(SampleEventTime, SampleEventDate_Month, SampleEventDate_Day)) %>% 

  # make column names match those in MacroplotReport
  mutate(Macroplot = MacroPlot_Name,
         LegacyMonitoringStatus = SampleEvent_LegacyMonitoringStatus,
         MonStatus = MonitoringStatus_Name) %>% 
  select(!c(MacroPlot_Name, SampleEvent_LegacyMonitoringStatus, MonitoringStatus_Name)) %>% 
  
  # edit Macroplot column
  separate(Macroplot, sep = "_", into = c("Park", "Macroplot"), remove = F) %>% 
  select(!Park)

##############################################

WACA_SampleEventReport <- WACA_SampleEventReport %>% 
  
  # remove trailing/leading whitespace
  mutate(across(where(is.character), str_trim)) %>%
  
  # separate Date columns into Date and Time columns
  separate(SampleEvent_Date, sep = " ", into = c("SampleEventDate", "SampleEventTime")) %>%
  # reformate Date columns
  mutate(SampleEventDate = as.Date(SampleEventDate, "%m/%d/%Y")) %>% 
  # separate Date columns into Year, Month, and Day columns
  separate(SampleEventDate, sep = "-", into = c("SampleEventDate_Year", "SampleEventDate_Month", "SampleEventDate_Day"), remove = FALSE) %>% 
  # remove unnecessary Date columns
  select(!c(SampleEventTime, SampleEventDate_Month, SampleEventDate_Day)) %>% 

  # make column names match those in MacroplotReport
  mutate(Macroplot = MacroPlot_Name,
         LegacyMonitoringStatus = SampleEvent_LegacyMonitoringStatus,
         MonStatus = MonitoringStatus_Name) %>% 
  select(!c(MacroPlot_Name, SampleEvent_LegacyMonitoringStatus, MonitoringStatus_Name)) %>% 
  
  # edit Macroplot and ProjectUnit_Name columns
  mutate(Macroplot = ifelse(Macroplot == "B:FPIED1D02:01", "PIED 01", 
                            ifelse(Macroplot == "B:FPIED1D02:02", "PIED 02", Macroplot))) %>%
  mutate(ProjectUnit_Name = ifelse(ProjectUnit_Name == "FPIED1D02", "PIED", ProjectUnit_Name))
```

#### Macroplot Report

```{r}
GRCA_MacroplotReport <- GRCA_MacroplotReport %>% 
  
  # remove trailing/leading whitespace
  mutate(across(where(is.character), str_trim)) %>%
  
  # separate Date columns into Date and Time columns
  separate(SampleEventDate, sep = " ", into = c("SampleEventDate", "SampleEventTime")) %>% 
  separate(DateIn, sep = " ", into = c("DateIn", "DateInTime")) %>% 
  # reformate Date columns
  mutate(SampleEventDate = as.Date(SampleEventDate, "%m/%d/%Y"),
         DateIn = as.Date(DateIn, "%m/%d/%Y")) %>% 
  # separate Date columns into Year, Month, and Day columns
  separate(SampleEventDate, sep = "-", into = c("SampleEventDate_Year", "SampleEventDate_Month", "SampleEventDate_Day"), remove = FALSE) %>% 
  separate(DateIn, sep = "-", into = c("DateIn_Year", "DateIn_Month", "DateIn_Day"), remove = FALSE) %>% 
  # remove unnecessary Date columns
  select(!c(SampleEventTime, SampleEventDate_Month, SampleEventDate_Day, 
            DateInTime, DateIn_Month, DateIn_Day))

##############################################

IM_MacroplotReport <- IM_MacroplotReport %>% 
  
  # remove trailing/leading whitespace
  mutate(across(where(is.character), str_trim)) %>%
  
  # separate Date columns into Date and Time columns
  separate(SampleEventDate, sep = " ", into = c("SampleEventDate", "SampleEventTime")) %>% 
  separate(DateIn, sep = " ", into = c("DateIn", "DateInTime")) %>% 
  # reformate Date columns
  mutate(SampleEventDate = as.Date(SampleEventDate, "%m/%d/%Y"),
         DateIn = as.Date(DateIn, "%m/%d/%Y")) %>% 
  # separate Date columns into Year, Month, and Day columns
  separate(SampleEventDate, sep = "-", into = c("SampleEventDate_Year", "SampleEventDate_Month", "SampleEventDate_Day"), remove = FALSE) %>% 
  separate(DateIn, sep = "-", into = c("DateIn_Year", "DateIn_Month", "DateIn_Day"), remove = FALSE) %>% 
  # remove unnecessary Date columns
  select(!c(SampleEventTime, SampleEventDate_Month, SampleEventDate_Day, 
            DateInTime, DateIn_Month, DateIn_Day)) %>% 
  
  # edit Macroplot column
  separate(Macroplot, sep = "_", into = c("Park", "Macroplot"), remove = F)

##############################################

WACA_MacroplotReport <- WACA_MacroplotReport %>% 
  
  # remove trailing/leading whitespace
  mutate(across(where(is.character), str_trim)) %>%
  
  # separate Date columns into Date and Time columns
  separate(SampleEventDate, sep = " ", into = c("SampleEventDate", "SampleEventTime")) %>% 
  separate(DateIn, sep = " ", into = c("DateIn", "DateInTime")) %>% 
  # reformate Date columns
  mutate(SampleEventDate = as.Date(SampleEventDate, "%m/%d/%Y"),
         DateIn = as.Date(DateIn, "%m/%d/%Y")) %>% 
  # separate Date columns into Year, Month, and Day columns
  separate(SampleEventDate, sep = "-", into = c("SampleEventDate_Year", "SampleEventDate_Month", "SampleEventDate_Day"), remove = FALSE) %>% 
  separate(DateIn, sep = "-", into = c("DateIn_Year", "DateIn_Month", "DateIn_Day"), remove = FALSE) %>% 
  # remove unnecessary Date columns
  select(!c(SampleEventTime, SampleEventDate_Month, SampleEventDate_Day, 
            DateInTime, DateIn_Month, DateIn_Day)) %>% 
  
  # edit Metadata Macroplot columns
  mutate(MetaData = ifelse(Macroplot == "B:FPIED1D02:01" & Purpose == "PIED", Macroplot,
                           ifelse(Macroplot == "B:FPIED1D02:02" & Purpose == "PIED", Macroplot, MetaData))) %>% 
  mutate(Macroplot = ifelse(Macroplot == "B:FPIED1D02:01", "PIED 01", 
                            ifelse(Macroplot == "B:FPIED1D02:02", "PIED 02", Macroplot)))
```

### Filter Data

#### Sample Event Report

The Sample Event Report contains multiple project units, and each plot can be assigned to any project unit. To reduce duplicates, select only the core project units.

```{r}
GRCA_SampleEventReport_merge <- GRCA_SampleEventReport %>% 
  filter(!ProjectUnit_Name %in% c("z2014 + 2017_Slopes RX", "z2017_Tipover East RX", "z2017_Slopes RX", "zHiSevMSO", "zSRIMRx2024Spring_Latest")) %>% 
  filter(!ProjectUnit_Name %in% c("ALL_FMH", "ALL_RAP_NRim", "ALL_RAP_SRim")) %>% 
  select(!c(Multi_PU, SampleEventDate_Year, LegacyMonitoringStatus))

##############################################

IM_SampleEventReport_merge <- IM_SampleEventReport %>% 
  filter(!ProjectUnit_Name %in% c("z2014 + 2017_Slopes RX")) %>% 
  select(!c(Multi_PU, SampleEventDate_Year, LegacyMonitoringStatus))

##############################################

WACA_SampleEventReport_merge <- WACA_SampleEventReport %>% 
  select(!c(Multi_PU, SampleEventDate_Year, LegacyMonitoringStatus))
```

#### Macroplot Report

```{r}
GRCA_MacroplotReport_MetaData <- GRCA_MacroplotReport %>% 
  filter(is.na(SampleEventDate)) %>% 
  select(!c(SampleEventDate, SampleEventDate_Year, SampleEventTeam, SampleEventComment, MonStatusOrd, LegacyMonStatus, MonStatus))

GRCA_MacroplotReport_SampleEvents <- GRCA_MacroplotReport %>% 
  filter(!is.na(SampleEventDate)) %>% 
  select(c(Macroplot, SampleEventDate, SampleEventDate_Year, SampleEventTeam, SampleEventComment, MonStatusOrd, LegacyMonStatus, MonStatus))

##############################################

IM_MacroplotReport_MetaData <- IM_MacroplotReport %>% 
  filter(is.na(SampleEventDate)) %>% 
  select(!c(SampleEventDate, SampleEventDate_Year, SampleEventTeam, SampleEventComment, MonStatusOrd, LegacyMonStatus, MonStatus))

IM_MacroplotReport_SampleEvents <- IM_MacroplotReport %>% 
  filter(!is.na(SampleEventDate)) %>% 
  select(c(Macroplot, SampleEventDate, SampleEventDate_Year, SampleEventTeam, SampleEventComment, MonStatusOrd, LegacyMonStatus, MonStatus))

##############################################

WACA_MacroplotReport_MetaData <- WACA_MacroplotReport %>% 
  filter(is.na(SampleEventDate)) %>% 
  select(!c(SampleEventDate, SampleEventDate_Year, SampleEventTeam, SampleEventComment, MonStatusOrd, LegacyMonStatus, MonStatus))

WACA_MacroplotReport_SampleEvents <- WACA_MacroplotReport %>% 
  filter(!is.na(SampleEventDate)) %>% 
  select(c(Macroplot, SampleEventDate, SampleEventDate_Year, SampleEventTeam, SampleEventComment, MonStatusOrd, LegacyMonStatus, MonStatus))
```

### Add Columns

-   Disturbance_Type (burn, thin, etc...)

-   PlotType_Name (FMH, RAP, etc...)

-   Park (GRCA, WACA)

-   Park_Unit (North, South)

-   Burn_Unit

-   Macroplot_Code

-   Macroplot_Num

```{r}
GRCA_MacroplotReport_MetaData <- GRCA_MacroplotReport_MetaData %>% 
  
  # PlotType_Name
  mutate(PlotType_Name = ifelse(Purpose %in% c("PIED","PIPO","PIPN","PIAB","PIEN"), "FMH", 
                                ifelse (Purpose %in% c("GRIN","GRED"), "Meadow", "RAP"))) %>%
  
  # Park
  mutate(Park = "GRCA") %>% 
  
  # Macroplot_Code and Macroplot_Number
  separate(Macroplot, sep = " ", into = c("Macroplot_Code", "Macroplot_Num"), remove = F) %>%
  separate(Macroplot, sep = "(?<=[A-Z])(?=[0-9])", into = c("Macroplot_Code1", "Macroplot_Num1"), remove = F) %>% 
  mutate(Macroplot_Code = ifelse(PlotType_Name == "RAP", Macroplot_Code1, Macroplot_Code)) %>% 
  mutate(Macroplot_Num = ifelse(PlotType_Name == "RAP", Macroplot_Num1, Macroplot_Num)) %>% 
  select(!c(Macroplot_Code1, Macroplot_Num1)) %>% 
  
  # Park_Unit
  mutate(Park_Unit = ifelse(Purpose %in% c("PIED", "PIPO", "Moqui", "Picnic", "Quarry", "Tusayan Pueblo (Thinning)"), "South", "North")) %>% 
  
  # Burn_Unit
  separate(UV1, sep = " Project Unit", into = c("Burn_Unit"), remove = F) %>% 
  
  # Disturbance_Type
  mutate(Disturbance_Type = ifelse(Purpose %in% c("Tusayan Pueblo (Thinning)"), "thin", "burn")) %>% 
  
  # reorder columns
  relocate(PlotType_Name, .before = Macroplot) %>% 
  relocate(Park, .before = PlotType_Name) %>% 
  relocate(Type, .after = PlotType_Name) %>% 
  relocate(Park_Unit, .after = Type) %>% 
  relocate(Purpose, .after = Park_Unit) %>%
  relocate(Burn_Unit, .after = Purpose) %>% 
  relocate(MetaData, .after = Burn_Unit)

##############################################

IM_MacroplotReport_MetaData <- IM_MacroplotReport_MetaData %>% 
  
  # PlotType_Name
  mutate(PlotType_Name = "I&M") %>% 

  # Macroplot_Code and Macroplot_Number
   separate(Macroplot, sep = "(?<=[A-Z])(?=[0-9])", into = c("Macroplot_Code", "Macroplot_Num"), remove = F) %>% 
  
  # Purpose
  mutate(Purpose = ifelse(Macroplot_Code == "M", "Mixed Conifer", "Pinyon Juniper")) %>% 

  # Park_Unit
  mutate(Park_Unit = ifelse(Macroplot_Code == "M", "North", "South")) %>% 
  
  # Burn_Unit
  separate(UV1, sep = " Project Unit", into = c("Burn_Unit"), remove = F) %>% 
  
  # Disturbance_Type
  mutate(Disturbance_Type = "burn") %>% 
  
  # reorder columns
  relocate(PlotType_Name, .before = Macroplot) %>% 
  relocate(Park, .before = PlotType_Name) %>% 
  relocate(Type, .after = PlotType_Name) %>% 
  relocate(Park_Unit, .after = Type) %>% 
  relocate(Purpose, .after = Park_Unit) %>%
  relocate(Burn_Unit, .after = Purpose) %>% 
  relocate(MetaData, .after = Burn_Unit)

##############################################

WACA_MacroplotReport_MetaData <- WACA_MacroplotReport_MetaData %>% 

  # PlotType_Name
  mutate(PlotType_Name = "FMH") %>% 
  
  # Park
  mutate(Park = "WACA") %>% 

  # Macroplot_Code and Macroplot_Number
  separate(Macroplot, sep = " ", into = c("Macroplot_Code", "Macroplot_Num"), remove = F) %>% 
  # Park_Unit
  mutate(Park_Unit = ifelse(Macroplot %in% c("PIPO 06", "PIPO 07", "PIPO 08", "PIPO 09", "PIPO 13", "PIPO 14"), "South", "North")) %>% 

  # Burn_Unit
  separate(UV1, sep = " Project Unit", into = c("Burn_Unit"), remove = F) %>% 
  
  # Disturbance_Type
  mutate(Disturbance_Type = "burn") %>% 
  
  # reorder columns
  relocate(PlotType_Name, .before = Macroplot) %>% 
  relocate(Park, .before = PlotType_Name) %>% 
  relocate(Type, .after = PlotType_Name) %>% 
  relocate(Park_Unit, .after = Type) %>% 
  relocate(Purpose, .after = Park_Unit) %>%
  relocate(Burn_Unit, .after = Purpose) %>% 
  relocate(MetaData, .after = Burn_Unit)
```

## Summary Tables

### Monitoring Status

Create columns for Entry (number of disturbances), Status (pre or post disturbance), and Time Since disturbance (i.e., Year 1) via break down of the Monitoring Status field

```{r}
# Isolate different Monitoring Statuses from each park
GRCA_MonStatus <- GRCA_MacroplotReport_SampleEvents %>% 
  select(MonStatusOrd, MonStatus) %>% 
  mutate(MonStatusOrd = NA)

IM_MonStatus <- IM_MacroplotReport_SampleEvents %>% 
  select(MonStatusOrd, MonStatus) %>% 
  mutate(MonStatusOrd = NA)

WACA_MonStatus <- WACA_MacroplotReport_SampleEvents %>% 
  select(MonStatusOrd, MonStatus) %>% 
  mutate(MonStatusOrd = NA)

# Combine Monitoring Statuses, keep only unique values
all_MonStatus <- rbind(GRCA_MonStatus, IM_MonStatus, WACA_MonStatus) %>% 
  unique() %>% 
  filter(MonStatus != "")

# Create column with consistent naming
all_MonStatus <- all_MonStatus %>% 
  mutate(MonStatus_pre = ifelse(grepl("Measurement", MonStatus), "00Measurement0", 
                                ifelse(grepl("Pre|PR", MonStatus), "00Pre0", NA))) %>% 
  mutate(MonStatus_post = ifelse(grepl("01Post|01 Post", MonStatus), "01Post0",
                                 ifelse(grepl("02Post", MonStatus), "02Post0",
                                        ifelse(grepl("03Post", MonStatus), "03Post0",
                                               ifelse(grepl("04Post", MonStatus), "04Post0", NA))))) %>% 
  mutate(MonStatus_year = ifelse(grepl("01 yr01", MonStatus), "01Year1",
                                ifelse(grepl("01 yr02", MonStatus), "01Year2", NA))) %>% 
  mutate(MonStatus_new = ifelse(!is.na(MonStatus_pre), MonStatus_pre, 
                                ifelse(!is.na(MonStatus_post), MonStatus_post,
                                       ifelse(!is.na(MonStatus_year), MonStatus_year, MonStatus)))) %>% 
  select(!c(MonStatus_pre, MonStatus_post, MonStatus_year))

# Extract info 
all_MonStatus <- all_MonStatus %>% 
  unique() %>% 
  # extract info based on monitoring status names
  extract(MonStatus_new, into = c("Entry","Status","TimeSince"),
          regex = "([0-9]+)([A-Z][a-z]+)([0-9]+)",
          remove = FALSE) %>%
  # fix up classifications
  mutate(Status = ifelse(Status == "Year", "Post", Status)) %>%
  mutate(Status = ifelse(Status == "Measurement", "Pre", Status)) %>% 
  mutate(TimeSince = ifelse(Status == "Pre", NA, TimeSince)) %>% 
  # make columns numeric
  mutate(Entry = as.numeric(Entry),
         TimeSince = as.numeric(TimeSince)) %>% 
  # create factor order
  mutate(MonStatus = factor(MonStatus, levels = c("Measurement_1", "Measurement_2",
                                                  "Measurement_3", "Measurement_4",
                                                  "PR01", "PR02", "PR03", "PR04", "PR05",
                                                  "Pre", "00 PRE",
                                                  "01Post", "01PostB", "01 Post",
                                                  "01Year1", "01 yr01",
                                                  "01Year2", "01 yr02",
                                                  "01Year3" ,
                                                  "01Year5", "01Year10", "01Year20",
                                                  "02PR01", "02PR02", "02Pre",
                                                  "02Post", "02PostB",
                                                  "02Year1", "02Year2",
                                                  "02Year5", "02Year10", "02Year20",
                                                  "03Post",
                                                  "03Year1", "03Year2",
                                                  "03Year5", "03Year10",
                                                  "04Post",
                                                  "04Year1", "04Year2",
                                                  "04Year5"))) %>%
  arrange(MonStatus) %>%
  mutate(MonStatusOrd = row_number()) %>% 
  select(!c(MonStatus_new))
```

### Disturbance History

Create columns for Entry (number of disturbances), Disturbance_Name (i.e., fire name or project name), Disturbance_Desc (i.e., RX, WF, Supp), DisturbanceDate, and Severity (i.e., low, mod, high).

```{r}
# Ensure blanks in UV columns are NA
GRCA_MacroplotReport_MetaData$UV2[GRCA_MacroplotReport_MetaData$UV2==""] <- NA
GRCA_MacroplotReport_MetaData$UV3[GRCA_MacroplotReport_MetaData$UV3==""] <- NA
GRCA_MacroplotReport_MetaData$UV4[GRCA_MacroplotReport_MetaData$UV4==""] <- NA
GRCA_MacroplotReport_MetaData$UV5[GRCA_MacroplotReport_MetaData$UV5==""] <- NA
#GRCA_MacroplotReport_MetaData$UV6[GRCA_MacroplotReport_MetaData$UV6==""] <- NA
#GRCA_MacroplotReport_MetaData$UV7[GRCA_MacroplotReport_MetaData$UV7==""] <- NA
#GRCA_MacroplotReport_MetaData$UV8[GRCA_MacroplotReport_MetaData$UV8==""] <- NA

IM_MacroplotReport_MetaData$UV2[IM_MacroplotReport_MetaData$UV2==""] <- NA

WACA_MacroplotReport_MetaData$UV2[WACA_MacroplotReport_MetaData$UV2==""] <- NA
```

#### 01 Disturbance

```{r}
# add add relevant columns
GRCA_Disturbance01 <- GRCA_MacroplotReport_MetaData %>% 
  select(c(Park, PlotType_Name, Macroplot, Disturbance_Type, UV2)) %>% 
  filter(!is.na(UV2)) %>% 
  separate(UV2, sep = ": ", into = c("Entry", "Disturbance_Info"), remove = FALSE) %>% 
  separate(Disturbance_Info, sep = ",", into = c("Disturbance_Name", "Disturbance_Desc", "DisturbanceDate", "Severity"), remove = FALSE) %>% 
  mutate(across(where(is.character), str_trim)) %>% 
  mutate(Entry = 1)

# specific edits
GRCA_Disturbance01 <- GRCA_Disturbance01 %>% 
  mutate(Disturbance_Name = ifelse(Macroplot == "SC03", "Tipover East", Disturbance_Name),
         Disturbance_Desc = ifelse(Macroplot == "SC03", "Rx", Disturbance_Desc),
         DisturbanceDate = ifelse(Macroplot == "SC03", "12/22/17", DisturbanceDate),
         Severity = ifelse(Macroplot == "SC03", "(POST read '18)", Severity)) %>% 
  mutate(Disturbance_Name = ifelse(Macroplot == "SC16", "Tipover East", Disturbance_Name),
         Disturbance_Desc = ifelse(Macroplot == "SC16", "Rx", Disturbance_Desc),
         DisturbanceDate = ifelse(Macroplot == "SC16", "12/22/17", DisturbanceDate),
         Severity = ifelse(Macroplot == "SC16", "(POST read '18)", Severity)) %>% 
  mutate(Disturbance_Name = ifelse(Macroplot == "PIAB 06", "NW 1,3,5", Disturbance_Name),
         Disturbance_Desc = ifelse(Macroplot == "PIAB 06", "Rx", Disturbance_Desc),
         DisturbanceDate = ifelse(Macroplot == "PIAB 06", "11/9/2007", DisturbanceDate),
         Severity = ifelse(Macroplot == "PIAB 06", "Low Severity", Severity)) %>% 
  mutate(Disturbance_Name = ifelse(Macroplot == "PIAB 24", "NW 1,3,5", Disturbance_Name),
         Disturbance_Desc = ifelse(Macroplot == "PIAB 24", "Rx", Disturbance_Desc),
         DisturbanceDate = ifelse(Macroplot == "PIAB 24", "11/10/2007", DisturbanceDate),
         Severity = ifelse(Macroplot == "PIAB 24", "Moderate/Low Severity", Severity)) %>% 
  mutate(Disturbance_Desc = ifelse(Disturbance_Desc == "Rx", "RX", Disturbance_Desc))

# edit date column
GRCA_Disturbance01 <- GRCA_Disturbance01 %>% 
  mutate(DisturbanceDate = parse_date_time(DisturbanceDate, orders = c("mdy", "my", "y"))) %>% 
  separate(DisturbanceDate, sep = "-", into = c("DisturbanceDate_Year", "Month", "Day"), remove = F) %>% 
  select(!c(Month, Day, UV2, Disturbance_Info)) %>% 
  relocate(DisturbanceDate_Year, .after = DisturbanceDate) %>% 
  relocate(Disturbance_Type, .before = Disturbance_Name)

##############################################

IM_Disturbance01 <- IM_MacroplotReport_MetaData %>% 
  # filter for relevant columns
  select(c(Park, PlotType_Name, Macroplot, Disturbance_Type, UV2)) %>% 
  filter(!is.na(UV2)) %>% 
  # add relevant columns
  separate(UV2, sep = ": ", into = c("Entry", "Disturbance_Info"), remove = FALSE) %>% 
  separate(Disturbance_Info, sep = ",", into = c("Disturbance_Name", "Disturbance_Desc", "DisturbanceDate", "Severity"), remove = FALSE) %>% 
  # trim white space
  mutate(across(where(is.character), str_trim)) %>% 
  # create Entry column
  mutate(Entry = 1) %>% 
  # edit Date column
  mutate(DisturbanceDate = parse_date_time(DisturbanceDate, orders = c("mdy", "my", "y"))) %>% 
  separate(DisturbanceDate, sep = "-", into = c("DisturbanceDate_Year", "Month", "Day"), remove = F) %>% 
  select(!c(Month, Day, UV2, Disturbance_Info)) %>%
  # reorder columns
  relocate(DisturbanceDate_Year, .after = DisturbanceDate) %>% 
  relocate(Disturbance_Type, .before = Disturbance_Name)

##############################################

WACA_Disturbance01 <- WACA_MacroplotReport_MetaData %>% 
  # filter for relevant columns
  select(c(Park, PlotType_Name, Macroplot, Disturbance_Type, UV2)) %>% 
  filter(!is.na(UV2)) %>% 
  # add relevant columns
  separate(UV2, sep = ": ", into = c("Entry", "Disturbance_Info"), remove = FALSE) %>% 
  separate(Disturbance_Info, sep = ",", into = c("Disturbance_Name", "Disturbance_Desc", "DisturbanceDate", "Severity"), remove = FALSE) %>% 
  # trim white space
  mutate(across(where(is.character), str_trim)) %>% 
  # create Entry column
  mutate(Entry = 1) %>% 
  # edit Date column
  mutate(DisturbanceDate = parse_date_time(DisturbanceDate, orders = c("mdy", "my", "y"))) %>% 
  separate(DisturbanceDate, sep = "-", into = c("DisturbanceDate_Year", "Month", "Day"), remove = F) %>% 
  select(!c(Month, Day, UV2, Disturbance_Info)) %>%
  # reorder columns
  relocate(DisturbanceDate_Year, .after = DisturbanceDate) %>% 
  relocate(Disturbance_Type, .before = Disturbance_Name)
```

#### 02 Disturbance

```{r}
# add add relevant columns
GRCA_Disturbance02 <- GRCA_MacroplotReport_MetaData %>% 
  select(c(Park, PlotType_Name, Macroplot, Disturbance_Type, UV3)) %>% 
  filter(!is.na(UV3)) %>% 
  separate(UV3, sep = ": ", into = c("Entry", "Disturbance_Info"), remove = FALSE) %>% 
  separate(Disturbance_Info, sep = ",", into = c("Disturbance_Name", "Disturbance_Desc", "DisturbanceDate", "Severity"), remove = FALSE) %>% 
  mutate(across(where(is.character), str_trim)) %>% 
  mutate(Entry = 2) %>% 
  select(!c("Disturbance_Info"))

# specific edits
GRCA_Disturbance02 <-GRCA_Disturbance02 %>%
  mutate(Disturbance_Name = ifelse(Macroplot == "PIAB 07", "NW 1,3,5", Disturbance_Name),
         Disturbance_Desc = ifelse(Macroplot == "PIAB 07", "Rx", Disturbance_Desc),
         DisturbanceDate = ifelse(Macroplot == "PIAB 07", "11/10/2007", DisturbanceDate),
         Severity = ifelse(Macroplot == "PIAB 07", "Low Severity", Severity)) %>%
  mutate(Disturbance_Name = ifelse(Macroplot == "PIAB 25", "NW 1,3,5", Disturbance_Name),
         Disturbance_Desc = ifelse(Macroplot == "PIAB 25", "Rx", Disturbance_Desc),
         DisturbanceDate = ifelse(Macroplot == "PIAB 25", "11/10/2007", DisturbanceDate),
         Severity = ifelse(Macroplot == "PIAB 25", "Moderate/Low Severity", Severity)) %>%
  mutate(Disturbance_Name = ifelse(Macroplot == "PIPN 01", "NW 1,3,5", Disturbance_Name),
         Disturbance_Desc = ifelse(Macroplot == "PIPN 01", "Rx", Disturbance_Desc),
         DisturbanceDate = ifelse(Macroplot == "PIPN 01", "11/10/2007", DisturbanceDate),
         Severity = ifelse(Macroplot == "PIPN 01", "Moderate/Low Severity", Severity)) %>%
  mutate(Disturbance_Name = ifelse(Macroplot == "PIPN 02", "NW 1,3,5", Disturbance_Name),
         Disturbance_Desc = ifelse(Macroplot == "PIPN 02", "Rx", Disturbance_Desc),
         DisturbanceDate = ifelse(Macroplot == "PIPN 02", "11/10/2007", DisturbanceDate),
         Severity = ifelse(Macroplot == "PIPN 02", "Moderate/Low Severity", Severity)) %>%
  mutate(Disturbance_Name = ifelse(Macroplot == "PIPO 24", "Watson 2,3,4", Disturbance_Name),
         Disturbance_Desc = ifelse(Macroplot == "PIPO 24", "Rx", Disturbance_Desc),
         DisturbanceDate = ifelse(Macroplot == "PIPO 24", "10/14/2011", DisturbanceDate),
         Severity = ifelse(Macroplot == "PIPO 24", "Low Severity", Severity)) %>% 
  mutate(Disturbance_Desc = ifelse(Disturbance_Desc == "Rx", "RX", Disturbance_Desc)) %>% 
  select(!c("UV3"))

# edit date column
GRCA_Disturbance02 <- GRCA_Disturbance02 %>% 
  mutate(DisturbanceDate = parse_date_time(DisturbanceDate, orders = c("mdy", "my", "y"))) %>% 
  separate(DisturbanceDate, sep = "-", into = c("DisturbanceDate_Year", "Month", "Day"), remove = F) %>% 
  select(!c("Month", "Day")) %>% 
  relocate(DisturbanceDate_Year, .after = DisturbanceDate) %>% 
  relocate(Disturbance_Type, .before = Disturbance_Name)
```

#### 03 Disturbance

```{r}
# add add relevant columns
GRCA_Disturbance03 <- GRCA_MacroplotReport_MetaData %>% 
  select(c(Park, PlotType_Name, Macroplot, Disturbance_Type, UV4)) %>% 
  filter(!is.na(UV4)) %>% 
  separate(UV4, sep = ": ", into = c("Entry", "Disturbance_Info"), remove = FALSE) %>% 
  separate(Disturbance_Info, sep = ",", into = c("Disturbance_Name", "Disturbance_Desc", "DisturbanceDate", "Severity"), remove = FALSE) %>% 
  mutate(across(where(is.character), str_trim)) %>% 
  mutate(Entry = 3) %>% 
  select(!c("Disturbance_Info"))

# specific edits
GRCA_Disturbance03 <- GRCA_Disturbance03 %>%
  mutate(Disturbance_Name = ifelse(Macroplot == "PIPO 17", "Watson 2,3,4", Disturbance_Name),
         Disturbance_Desc = ifelse(Macroplot == "PIPO 17", "Rx", Disturbance_Desc),
         DisturbanceDate = ifelse(Macroplot == "PIPO 17", "10/14/2011", DisturbanceDate),
         Severity = ifelse(Macroplot == "PIPO 17", "Low Severity", Severity)) %>%
  mutate(Disturbance_Name = ifelse(Macroplot == "PIPO 20", "Watson 2,3,4", Disturbance_Name),
         Disturbance_Desc = ifelse(Macroplot == "PIPO 20", "Rx", Disturbance_Desc),
         DisturbanceDate = ifelse(Macroplot == "PIPO 20", "10/14/2011", DisturbanceDate),
         Severity = ifelse(Macroplot == "PIPO 20", "Low Severity", Severity)) %>%
  mutate(Disturbance_Name = ifelse(Macroplot == "PIPO 21", "Watson 2,3,4", Disturbance_Name),
         Disturbance_Desc = ifelse(Macroplot == "PIPO 21", "Rx", Disturbance_Desc),
         DisturbanceDate = ifelse(Macroplot == "PIPO 21", "10/14/2011", DisturbanceDate),
         Severity = ifelse(Macroplot == "PIPO 21", "Low Severity", Severity)) %>%
  mutate(Disturbance_Name = ifelse(Macroplot == "PIPO 22", "Watson 2,3,4", Disturbance_Name),
         Disturbance_Desc = ifelse(Macroplot == "PIPO 22", "Rx", Disturbance_Desc),
         DisturbanceDate = ifelse(Macroplot == "PIPO 22", "10/14/2011", DisturbanceDate),
         Severity = ifelse(Macroplot == "PIPO 22", "Low Severity", Severity)) %>%
  mutate(Disturbance_Desc = ifelse(Disturbance_Desc == "Rx", "RX", Disturbance_Desc)) %>% 
  select(!c("UV4"))

# edit date column
GRCA_Disturbance03 <- GRCA_Disturbance03 %>% 
  mutate(DisturbanceDate = parse_date_time(DisturbanceDate, orders = c("mdy", "my", "y"))) %>% 
  separate(DisturbanceDate, sep = "-", into = c("DisturbanceDate_Year", "Month", "Day"), remove = F) %>% 
  select(!c("Month", "Day")) %>% 
  relocate(DisturbanceDate_Year, .after = DisturbanceDate) %>% 
  relocate(Disturbance_Type, .before = Disturbance_Name)
```

#### 04 Disturbance

```{r}
GRCA_Disturbance04 <- GRCA_MacroplotReport_MetaData %>% 
  select(c(Park, PlotType_Name, Macroplot, Disturbance_Type, UV5)) %>% 
  filter(!is.na(UV5)) %>% 
  # add add relevant columns
  separate(UV5, sep = ": ", into = c("Entry", "Disturbance_Info"), remove = FALSE) %>% 
  separate(Disturbance_Info, sep = ",", into = c("Disturbance_Name", "Disturbance_Desc", "DisturbanceDate", "Severity"), remove = FALSE) %>% 
  mutate(across(where(is.character), str_trim)) %>% 
  mutate(Entry = 4) %>% 
  select(!c("Disturbance_Info", "UV5")) %>% 
  # edit date column
  mutate(DisturbanceDate = parse_date_time(DisturbanceDate, orders = c("mdy", "my", "y"))) %>% 
  separate(DisturbanceDate, sep = "-", into = c("DisturbanceDate_Year", "Month", "Day"), remove = F) %>% 
  select(!c("Month", "Day")) %>% 
  relocate(DisturbanceDate_Year, .after = DisturbanceDate) %>% 
  relocate(Disturbance_Type, .before = Disturbance_Name)
```

#### All Disturbances

```{r}
GRCA_DisturbanceHistory <- rbind(GRCA_Disturbance01, GRCA_Disturbance02, GRCA_Disturbance03, GRCA_Disturbance04)

##############################################

IM_DisturbanceHistory <- rbind(IM_Disturbance01)

##############################################

WACA_DisturbanceHistory <- rbind(WACA_Disturbance01)
```

```{r}
# merge all disturbances
all_DisturbanceHistory <- rbind(GRCA_DisturbanceHistory, IM_DisturbanceHistory, WACA_DisturbanceHistory)

# reorganize severity column for consistency
all_DisturbanceHistory <- all_DisturbanceHistory %>% 
  mutate(Disturbance_Severity = ifelse(Severity %in% c("Low Severity"), "Low", 
                      ifelse(Severity %in% c("Moderate/Low Severity", "Moderate/Low Sev.", "Mod/Low Severity"), "Low/Mod", 
                             ifelse(Severity %in% c("Moderate Severity"), "Mod", 
                                    ifelse(Severity %in% c("Moderate/High Severity", "Mod-High Severity"), "Mod/High", 
                                           ifelse(Severity %in% c("High Severity"), "High", 
                                                  ifelse(Severity %in% c("Unburned"), "Unburned", NA))))))) %>% 
  select(!Severity)
```

## Merge

```{r}
# merge metadata and sample events
GRCA_MacroplotReport_all <- merge(GRCA_MacroplotReport_MetaData, GRCA_MacroplotReport_SampleEvents, by = c("Macroplot")) %>% 
  select(!MonStatusOrd)

# merge Macroplot Report and Sample Event Report
GRCA_MetadataReport <- merge(GRCA_MacroplotReport_all, GRCA_SampleEventReport_merge, by = c("Macroplot", "SampleEventDate", "MonStatus")) %>% 
  relocate(AdministrationUnit_Name, .before = Park) %>% 
  relocate(ProjectUnit_Name, .before = Purpose) %>% 
  relocate(Macroplot, .after = MetaData) %>% 
  relocate(SampleEventDate, .before = SampleEventDate_Year) %>% 
  relocate(MonStatus, .after = LegacyMonStatus)

##############################################

# merge metadata and sample events
IM_MacroplotReport_all <- merge(IM_MacroplotReport_MetaData, IM_MacroplotReport_SampleEvents, by = c("Macroplot")) %>% 
  select(!MonStatusOrd)

# merge Macroplot Report and Sample Event Report
IM_MetadataReport <- merge(IM_MacroplotReport_all, IM_SampleEventReport_merge, by = c("Macroplot", "SampleEventDate", "MonStatus")) %>% 
  relocate(AdministrationUnit_Name, .before = Park) %>% 
  relocate(ProjectUnit_Name, .before = Purpose) %>% 
  relocate(Macroplot, .after = MetaData) %>% 
  relocate(SampleEventDate, .before = SampleEventDate_Year) %>% 
  relocate(MonStatus, .after = LegacyMonStatus)

##############################################

# merge metadata and sample events
WACA_MacroplotReport_all <- merge(WACA_MacroplotReport_MetaData, WACA_MacroplotReport_SampleEvents, by = c("Macroplot")) %>% 
  select(!MonStatusOrd)

# merge Macroplot Report and Sample Event Report
WACA_MetadataReport <- merge(WACA_MacroplotReport_all, WACA_SampleEventReport_merge, by = c("Macroplot", "SampleEventDate", "MonStatus")) %>% 
  relocate(AdministrationUnit_Name, .before = Park) %>% 
  relocate(ProjectUnit_Name, .before = Purpose) %>% 
  relocate(Macroplot, .after = MetaData) %>% 
  relocate(SampleEventDate, .before = SampleEventDate_Year) %>% 
  relocate(MonStatus, .after = LegacyMonStatus)
```

```{r}
# merge all admin units
all_MetadataReport <- rbind(GRCA_MetadataReport, IM_MetadataReport, WACA_MetadataReport)

# merge monitoring status summary
all_MetadataReport <- merge(all_MetadataReport, all_MonStatus, by = c("MonStatus"), all = T) %>% 
  relocate(MonStatus, .before = MonStatusOrd) %>% 
  relocate(LegacyMonStatus, .before = MonStatus)

# merge disturbance history summary
all_MetadataReport <- merge(all_MetadataReport, all_DisturbanceHistory, by = c("Park", "PlotType_Name", "Macroplot", "Entry", "Disturbance_Type"), all = T) %>% 
  relocate(AdministrationUnit_Name, .before = Park) %>% 
  relocate(Macroplot, .after = MetaData) %>% 
  relocate(Entry, .after = MonStatus) %>% 
  relocate(Disturbance_Type, .before = Disturbance_Name)
```

# Load Metadata

```{r}
#str(all_MetadataReport)

# reclassify data as correct type
all_MetadataReport <- all_MetadataReport %>% 
  mutate(AdministrationUnit_Name = factor(AdministrationUnit_Name)) %>% 
  mutate(Park = factor(Park)) %>% 
  mutate(PlotType_Name = factor(PlotType_Name)) %>% 
  mutate(Type = factor(Type)) %>% 
  mutate(Park_Unit = factor(Park_Unit)) %>% 
  mutate(ProjectUnit_Name = factor(ProjectUnit_Name)) %>% 
  mutate(Purpose = factor(Purpose)) %>% 
  mutate(Burn_Unit = factor(Burn_Unit)) %>% 
  mutate(MetaData = factor(MetaData)) %>% 
  mutate(Macroplot = factor(Macroplot)) %>% 
  mutate(Macroplot_Code = factor(Macroplot_Code)) %>% 
  mutate(Macroplot_Num = factor(Macroplot_Num)) %>% 
  #mutate(DateIn_Year = as.numeric()) %>%
  mutate(SampleEventDate_Year = as.numeric(SampleEventDate_Year)) %>% 
  mutate(Protocols = factor(Protocols)) %>%
  mutate(Visited = factor(Visited)) %>%
  mutate(LegacyMonStatus = factor(LegacyMonStatus)) %>%
  mutate(MonStatus = factor(MonStatus)) %>%
  mutate(Entry = factor(Entry)) %>% 
  mutate(Status = factor(Status)) %>% 
  #mutate(Disturbance_Type = factor()) %>% 
  #mutate(Disturbance_Name = factor()) %>% 
  #mutate(Disturbance_Desc = factor()) %>% 
  #mutate(Disturbance_Year = as.numeric()) %>% 
  mutate(Disturbance_Severity = factor(Disturbance_Severity, levels = c("Unburned", "Low", "Low/Mod", "Mod", "Mod/High", "High")))
```

# Tables

```{r manipulate data, include=FALSE}
# check occurrence of unique values in newly created columns, make corrections as needed
all_MetadataReport$PlotType_Name %>% table(exclude = NULL)
all_MetadataReport$Status %>% table(exclude = NULL)
all_MetadataReport$Disturbance_Type %>% table(exclude = NULL)
all_MetadataReport$TimeSince %>% table(exclude = NULL)
```

### Overall Program Summary

```{r program summary}
## Summarise projects and events by park
program_summ <- all_MetadataReport %>%
  summarise(`Parks` = n_distinct(AdministrationUnit_Name),
            `Projects` = n_distinct(ProjectUnit_Name),
            `Number of Plots` = n_distinct(Macroplot),
            `Monitoring Statuses` = n_distinct(MonStatus),
            `Visit Dates` = n_distinct(SampleEventDate),
            `Protocols Used` = n_distinct(Protocols))

program_summ %>% 
  kbl(align = "lcccccc") %>%
  kable_styling(bootstrap_options = c("striped","condensed"), 
                full_width = F, position = "left", fixed_thead = T,
                row_label_position = "c") %>% 
  row_spec(0, font_size = 14, bold = TRUE) %>% 
  row_spec(nrow(program_summ), font_size = 14, bold = TRUE)
```

### Filtered Summaries

```{r set parameters, include=F}
### Select Data to Summarize
# use commas to build lists of acceptable values for filters

# designate target park(s)
unique(all_MetadataReport$Park)
target_park <- c("GRCA")

# designate target project(s)
# see all project names
unique(all_MetadataReport$PlotType_Name)
target_project <- c("FMH")

# designate target year or years
unique(all_MetadataReport$SampleEventDate_Year)
target_year <- c(2024)
```

The code underlying this report has been manipulated to display summaries of the following selected park(s), project(s), and year(s).\
\
Selected park(s): [`r target_park`]{style="color:red"}\
Selected project(s): [`r target_project`]{style="color:red"}\
Selected year(s): [`r target_year`]{style="color:red"}\

#### Park Summary: [`r target_park`]{style="color:red"}

The table below summarizes current and past projects in `r target_park`.

```{r park summ}
## Summarise projects by park
# uses 'target park' variable designated above
target_park_summ <- all_MetadataReport %>%
  filter(Park == target_park) %>% 
  group_by(`Project Name` = ProjectUnit_Name) %>%
  summarise(`Plot Type` = unique(PlotType_Name),
            `First Visit` = min(as.numeric(SampleEventDate_Year)),
            `Latest Visit` = max(as.numeric(SampleEventDate_Year)),
            `Number of Plots` = n_distinct(Macroplot),
            `Monitoring Statuses` = n_distinct(MonStatus),
            `Protocols Used` = n_distinct(Protocols)) %>% 
  arrange(`Project Name`) %>% 
  ungroup() %>% 
  mutate_all(as.character)

# Calculate totals
target_park_summ_totals <- target_park_summ %>%
  summarise(`Number of Plots` = sum(as.numeric(`Number of Plots`))) %>% 
  mutate(`Project Name` = "",
         `Plot Type` = "",
         `First Visit` = "",
         `Latest Visit` = "Total",
         `Monitoring Statuses` = "",
         `Protocols Used` = "") %>% 
  mutate_all(as.character)

# Bind the totals row to the original data frame
target_park_summ_w_totals <- bind_rows(target_park_summ, target_park_summ_totals)
  
target_park_summ_w_totals %>% 
  kbl(align = "lccccc") %>%
  kable_styling(bootstrap_options = c("striped","condensed"), 
                full_width = F, position = "left", fixed_thead = T) %>% 
  row_spec(nrow(target_park_summ_w_totals), font_size = 14, bold = TRUE)
```

#### Project Summary: [`r target_project`]{style="color:red"}

The table below summarizes sample events recorded within the `r target_project` project.

```{r project summ}
## Summarise project
# uses 'target project' variable designated above
target_project_summ <- all_MetadataReport %>%
  filter(Park == target_park,
         PlotType_Name == target_project) %>%
  group_by(`Monitoring Status` = MonStatus,
           `Order` = MonStatusOrd) %>%
  summarise(`Pre/Post` = unique(Status),
            `First Visit` = min(SampleEventDate_Year),
            `Latest Visit` = max(SampleEventDate_Year),
            `Number of Plots Visited` = n_distinct(Macroplot),
            `Protocols Used` = n_distinct(Protocols)) %>% 
  arrange(`Order`) %>% 
  select(!`Order`)

target_project_summ %>% 
  kbl(align = "lccccc") %>%
  kable_styling(bootstrap_options = c("striped","condensed"), 
                full_width = F, position = "left", fixed_thead = T) %>% 
  row_spec(0, font_size = 14, bold = TRUE)
```

#### Plot Activity Summary: [`r target_year`]{style="color:red"}

The table below summarizes all projects with plots sampled within the following year(s): `r target_year`.

```{r year summ}
## Summarise yearly plot activity
# uses 'target year' variable designated above
target_year_summ <- all_MetadataReport %>%
  filter(SampleEventDate_Year %in% target_year) %>% 
  group_by(`Park` = Park,
           `Project Name` = ProjectUnit_Name) %>%
  summarise(`Plot Type` = unique(PlotType_Name),
            `Start Date` = as.character(min(SampleEventDate)),
            `End Date` = as.character(max(SampleEventDate)),
            `Number of Plots` = n_distinct(Macroplot),
            `Protocols Used` = n_distinct(Protocols)) %>% 
  arrange(`Project Name`) %>% 
  ungroup()

# Calculate totals
year_summ_totals <- target_year_summ %>%
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE))) %>%
  mutate(`Park` = "",
         `Project Name` = "",
         `Plot Type` = "",
         `Start Date` = "",
         `End Date` = "Total")

# Bind the totals row to the original data frame
target_year_summ_w_totals <- bind_rows(target_year_summ, year_summ_totals)

target_year_summ_w_totals %>% 
  kbl(align = "lcccccc") %>%
  kable_styling(bootstrap_options = c("striped","condensed"), 
                full_width = F, position = "left", fixed_thead = T) %>% 
  row_spec(0, font_size = 14, bold = TRUE) %>% 
  row_spec(nrow(target_year_summ_w_totals), font_size = 14, bold = TRUE)
```

### Protocol Use Summary

The table below summarizes the frequency of use of protocols occurring in the FFI database, as utilized by the NPS Alaska Fire Ecology program. Protocols are arranged in descending order by number of plots such that the most frequently used protocols are listed first.

```{r protocol summ}
## Summarise protocol use by status
protocol_use_summ <- all_MetadataReport %>%
  group_by(`Protocol` = Protocols) %>%
  summarise(`First Use` = min(as.numeric(SampleEventDate_Year)),
            `Latest Use` = max(as.numeric(SampleEventDate_Year)),
            `Number of Plots` = n_distinct(Macroplot),
            `Number of Projects` = n_distinct(ProjectUnit_Name),
            `Number of Monitoring Statuses` = n_distinct(MonStatus),
            `Number of Parks` = n_distinct(AdministrationUnit_Name)) %>% 
  arrange(desc(`Number of Plots`))

protocol_use_summ %>% 
  kbl(align = "lcccccc") %>%
  kable_styling(bootstrap_options = c("striped","condensed"), 
                full_width = F, position = "left", fixed_thead = T) %>% 
  row_spec(0, font_size = 14, bold = TRUE)
```

# Graphs

```{r}
graph_MetadataReport <- all_MetadataReport %>% 
  filter(!is.na(AdministrationUnit_Name)) %>% 
  # remove unnecessary columns
  select(!c(Latitude, Longitude, UTM_X, UTM_Y, UTM_Zone, Datum, PDOP, Precision, )) %>% 
  select(!c(Elevation, ElevationUnits, Azimuth, Aspect, SlopeHill, SlopeTransect)) %>% 
  select(!c(LocatedBy, FutureVisit, DateOut, StartPoint, Directions, Comments, SampleEventTeam, SampleEventComment, LegacyMonStatus)) %>% 
  select(!c(UV1, UV2, UV3, UV4, UV5, UV6, UV7, UV8))
```

## WACA

### All FMH

```{r}
graph_MetadataReport %>% 
  filter(Park == "WACA",
         PlotType_Name == "FMH") %>% 
  ggplot(aes(x = SampleEventDate_Year, y = Macroplot_Num, color = Entry)) +
  geom_point() +
  geom_line(aes(group = Macroplot)) +
  facet_grid(~Purpose) +
  labs(title = "WACA Sample Event History",
       subtitle = "Fire Monitoring Handbook (FMH) Plots",
       y = "Plot Number",
       x = "Year",
       color = "Entry") +
  scale_color_manual(values=c('forestgreen', "yellow", 'darkorange', "red", "darkred")) +
  theme_light() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(size = 20, face="bold", hjust = 0.5),
        plot.subtitle = element_text(size = 12, face="italic", hjust = 0.5),
        axis.title = element_text(size = 12, face="bold"),
        axis.title.y = element_text(margin = unit(c(0,4,0,0), "mm")),
        axis.title.x = element_text(margin = unit(c(4,0,0,0), "mm")),
        axis.text.y = element_text(size = 5),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1))
```

## GRCA

### All I&M

```{r}
graph_MetadataReport %>% 
  filter(Park == "GRCA",
         PlotType_Name == "I&M") %>% 
  ggplot(aes(x = SampleEventDate_Year, y = Macroplot_Num, color = Entry)) +
  geom_point() +
  geom_line(aes(group = Macroplot)) +
  facet_wrap(~Macroplot_Code) +
  labs(title = "GRCA Sample Event History",
       subtitle = "Inventory and Monitoring (I&M) Plots",
       y = "Plot Number",
       x = "Year",
       color = "Entry") +
  scale_color_manual(values=c('forestgreen', "yellow", 'darkorange', "red", "darkred")) +
  theme_light() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(size = 20, face="bold", hjust = 0.5),
        plot.subtitle = element_text(size = 12, face="italic", hjust = 0.5),
        axis.title = element_text(size = 12, face="bold"),
        axis.title.y = element_text(margin = unit(c(0,4,0,0), "mm")),
        axis.title.x = element_text(margin = unit(c(4,0,0,0), "mm")),
        axis.text.y = element_blank(),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1))
```

### All RAP

```{r}
graph_MetadataReport %>% 
  filter(Park == "GRCA",
         PlotType_Name == "RAP") %>% 
  ggplot(aes(x = SampleEventDate_Year, y = Macroplot_Num, color = Entry)) +
  geom_point() +
  geom_line(aes(group = Macroplot)) +
  facet_wrap(~Macroplot_Code) +
  labs(title = "GRCA Sample Event History",
       subtitle = "Rapid Assessment Protocol (RAP) Plots",
       y = "Plot Number",
       x = "Year",
       color = "Entry") +
  scale_color_manual(values=c('forestgreen', "yellow", 'darkorange', "red", "darkred")) +
  theme_light() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(size = 20, face="bold", hjust = 0.5),
        plot.subtitle = element_text(size = 12, face="italic", hjust = 0.5),
        axis.title = element_text(size = 12, face="bold"),
        axis.title.y = element_text(margin = unit(c(0,4,0,0), "mm")),
        axis.title.x = element_text(margin = unit(c(4,0,0,0), "mm")),
        axis.text.y = element_blank(),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1))
```

### All FMH

```{r}
graph_MetadataReport %>% 
  filter(Park == "GRCA",
         PlotType_Name == "FMH") %>% 
  filter(!Macroplot %in% c("SHOS 01", "SHOS 02")) %>% 
  ggplot(aes(x = SampleEventDate_Year, y = Macroplot_Num, color = Entry)) +
  geom_point() +
  geom_line(aes(group = Macroplot)) +
  facet_grid(~Purpose) +
  labs(title = "GRCA Sample Event History",
       subtitle = "Fire Monitoring Handbook (FMH) Plots",
       y = "Plot Number",
       x = "Year",
       color = "Entry") +
  scale_color_manual(values=c('forestgreen', "yellow", 'darkorange', "red", "darkred")) +
  theme_light() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(size = 20, face="bold", hjust = 0.5),
        plot.subtitle = element_text(size = 12, face="italic", hjust = 0.5),
        axis.title = element_text(size = 12, face="bold"),
        axis.title.y = element_text(margin = unit(c(0,4,0,0), "mm")),
        axis.title.x = element_text(margin = unit(c(4,0,0,0), "mm")),
        axis.text.y = element_text(size = 5),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1))
```

#### Summarize

```{r}
# Summarize to one point per (Year, Macroplot, Visited)

summary_points_PIED <- graph_MetadataReport %>%
  filter(Park == "GRCA", Purpose == "PIED") %>%
  group_by(SampleEventDate_Year, Macroplot, Visited) %>%
  summarise(n_protocols = n(), .groups = "drop")  # optional: keep count of protocols

summary_points_PIPO <- graph_MetadataReport %>%
  filter(Park == "GRCA", Purpose == "PIPO") %>%
  group_by(SampleEventDate_Year, Macroplot, Visited) %>%
  summarise(n_protocols = n(), .groups = "drop")

summary_points_PIPN <- graph_MetadataReport %>%
  filter(Park == "GRCA", Purpose == "PIPN") %>%
  group_by(SampleEventDate_Year, Macroplot, Visited) %>%
  summarise(n_protocols = n(), .groups = "drop")

summary_points_PIAB <- graph_MetadataReport %>%
  filter(Park == "GRCA", Purpose == "PIAB") %>%
  group_by(SampleEventDate_Year, Macroplot, Visited) %>%
  summarise(n_protocols = n(), .groups = "drop")

summary_points_PIEN <- graph_MetadataReport %>%
  filter(Park == "GRCA", Purpose == "PIEN") %>%
  group_by(SampleEventDate_Year, Macroplot, Visited) %>%
  summarise(n_protocols = n(), .groups = "drop")
```

#### PIED

```{r}
graph_MetadataReport %>% 
  filter(Park == "GRCA",
         Purpose %in% c("PIED")) %>% 
  group_by(SampleEventDate_Year, Macroplot, Visited) %>% 
  ggplot(aes(x = SampleEventDate_Year, y = Macroplot)) +
  # Line plot
  geom_line(aes(group = Macroplot, color = Entry)) +
  # Plot green (Visited = Y) points first (background)
  geom_jitter(data = summary_points_PIED %>% filter(Visited == "Y"),
              aes(x = SampleEventDate_Year, y = Macroplot, fill = Visited),
              shape = 21, size = 2, width = 0.1, height = 0.1) +
  # Plot red (Visited = N) points second (foreground)
  geom_jitter(data = summary_points_PIED %>% filter(Visited == "N"),
              aes(x = SampleEventDate_Year, y = Macroplot, fill = Visited),
              shape = 21, size = 2, width = 0.1, height = 0.1) +
  #geom_jitter(aes(group = Visited, fill = Visited), shape = 21) +
  labs(title = "GRCA Sample Event History",
       subtitle = "Pinyon Juniper Ecosystem (PIED)",
       y = "Macroplot",
       x = "Year",
       color = "Entry") +
  scale_fill_manual(values = c("Y" = "forestgreen", "N" = "red")) +
  scale_color_manual(values=c('forestgreen', "yellow", 'darkorange', "red", "darkred")) +
  theme_light() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(size = 20, face="bold", hjust = 0.5),
        plot.subtitle = element_text(size = 12, face="italic", hjust = 0.5),
        axis.title = element_text(size = 12, face="bold"),
        axis.title.y = element_text(margin = unit(c(0,4,0,0), "mm")),
        axis.title.x = element_text(margin = unit(c(4,0,0,0), "mm")),
        axis.text.y = element_text(size = 5),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1))
```

#### PIPO

```{r}
graph_MetadataReport %>% 
  filter(Park == "GRCA",
         Purpose %in% c("PIPO")) %>% 
  group_by(SampleEventDate_Year, Macroplot, Visited) %>% 
  ggplot(aes(x = SampleEventDate_Year, y = Macroplot)) +
  # Line plot
  geom_line(aes(group = Macroplot, color = Entry)) +
  # Plot green (Visited = Y) points first (background)
  geom_jitter(data = summary_points_PIPO %>% filter(Visited == "Y"),
              aes(x = SampleEventDate_Year, y = Macroplot, fill = Visited),
              shape = 21, size = 2, width = 0.1, height = 0.1) +
  # Plot red (Visited = N) points second (foreground)
  geom_jitter(data = summary_points_PIPO %>% filter(Visited == "N"),
              aes(x = SampleEventDate_Year, y = Macroplot, fill = Visited),
              shape = 21, size = 2, width = 0.1, height = 0.1) +
  #geom_jitter(aes(group = Visited, fill = Visited), shape = 21) +
  labs(title = "GRCA Sample Event History",
       subtitle = "South Rim Ponderosa Pine Ecosystem (PIPO)",
       y = "Macroplot",
       x = "Year",
       color = "Entry",
       fill = "Visited") +
  scale_fill_manual(values = c("Y" = "forestgreen", "N" = "red")) +
  scale_color_manual(values=c('forestgreen', "yellow", 'darkorange', "red", "darkred")) +
  theme_light() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(size = 20, face="bold", hjust = 0.5),
        plot.subtitle = element_text(size = 12, face="italic", hjust = 0.5),
        axis.title = element_text(size = 12, face="bold"),
        axis.title.y = element_text(margin = unit(c(0,4,0,0), "mm")),
        axis.title.x = element_text(margin = unit(c(4,0,0,0), "mm")),
        axis.text.y = element_text(size = 5),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1))
```

#### PIPN

```{r}
graph_MetadataReport %>% 
  filter(Park == "GRCA",
         Purpose %in% c("PIPN")) %>% 
  group_by(SampleEventDate_Year, Macroplot, Visited) %>% 
  ggplot(aes(x = SampleEventDate_Year, y = Macroplot)) +
  # Line plot
  geom_line(aes(group = Macroplot, color = Entry)) +
  # Plot green (Visited = Y) points first (background)
  geom_jitter(data = summary_points_PIPN %>% filter(Visited == "Y"),
              aes(x = SampleEventDate_Year, y = Macroplot, fill = Visited),
              shape = 21, size = 2, width = 0.1, height = 0.1) +
  # Plot red (Visited = N) points second (foreground)
  geom_jitter(data = summary_points_PIPN %>% filter(Visited == "N"),
              aes(x = SampleEventDate_Year, y = Macroplot, fill = Visited),
              shape = 21, size = 2, width = 0.1, height = 0.1) +
  #geom_jitter(aes(group = Visited, fill = Visited), shape = 21) +
  labs(title = "GRCA Sample Event History",
       subtitle = "North Rim Ponderosa Pine Ecosystem (PIPN)",
       y = "Macroplot",
       x = "Year",
       color = "Entry") +
  scale_fill_manual(values = c("Y" = "forestgreen", "N" = "red")) +
  scale_color_manual(values=c('forestgreen', "yellow", 'darkorange', "red", "darkred")) +
  theme_light() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(size = 20, face="bold", hjust = 0.5),
        plot.subtitle = element_text(size = 12, face="italic", hjust = 0.5),
        axis.title = element_text(size = 12, face="bold"),
        axis.title.y = element_text(margin = unit(c(0,4,0,0), "mm")),
        axis.title.x = element_text(margin = unit(c(4,0,0,0), "mm")),
        axis.text.y = element_text(size = 5),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1))
```

#### PIAB

```{r}
graph_MetadataReport %>% 
  filter(Park == "GRCA",
         Purpose %in% c("PIAB")) %>% 
  group_by(SampleEventDate_Year, Macroplot, Visited) %>% 
  ggplot(aes(x = SampleEventDate_Year, y = Macroplot)) +
  # Line plot
  geom_line(aes(group = Macroplot, color = Entry)) +
  # Plot green (Visited = Y) points first (background)
  geom_jitter(data = summary_points_PIAB %>% filter(Visited == "Y"),
              aes(x = SampleEventDate_Year, y = Macroplot, fill = Visited),
              shape = 21, size = 2, width = 0.1, height = 0.1) +
  # Plot red (Visited = N) points second (foreground)
  geom_jitter(data = summary_points_PIAB %>% filter(Visited == "N"),
              aes(x = SampleEventDate_Year, y = Macroplot, fill = Visited),
              shape = 21, size = 2, width = 0.1, height = 0.1) +
  #geom_jitter(aes(group = Visited, fill = Visited), shape = 21) +
  labs(title = "GRCA Sample Event History",
       subtitle = "Pine-Fir Ecosystem (PIAB)",
       y = "Macroplot",
       x = "Year",
       color = "Entry") +
  scale_fill_manual(values = c("Y" = "forestgreen", "N" = "red")) +
  scale_color_manual(values=c('forestgreen', "yellow", 'darkorange', "red", "darkred")) +
  theme_light() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(size = 20, face="bold", hjust = 0.5),
        plot.subtitle = element_text(size = 12, face="italic", hjust = 0.5),
        axis.title = element_text(size = 12, face="bold"),
        axis.title.y = element_text(margin = unit(c(0,4,0,0), "mm")),
        axis.title.x = element_text(margin = unit(c(4,0,0,0), "mm")),
        axis.text.y = element_text(size = 5),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1))
```

#### PIEN

```{r}
graph_MetadataReport %>% 
  filter(Park == "GRCA",
         Purpose %in% c("PIEN")) %>% 
  group_by(SampleEventDate_Year, Macroplot, Visited) %>% 
  ggplot(aes(x = SampleEventDate_Year, y = Macroplot)) +
  # Line plot
  geom_line(aes(group = Macroplot, color = Entry)) +
  # Plot green (Visited = Y) points first (background)
  geom_jitter(data = summary_points_PIEN %>% filter(Visited == "Y"),
              aes(x = SampleEventDate_Year, y = Macroplot, fill = Visited),
              shape = 21, size = 2, width = 0.1, height = 0.1) +
  # Plot red (Visited = N) points second (foreground)
  geom_jitter(data = summary_points_PIEN %>% filter(Visited == "N"),
              aes(x = SampleEventDate_Year, y = Macroplot, fill = Visited),
              shape = 21, size = 2, width = 0.1, height = 0.1) +
  #geom_jitter(aes(group = Visited, fill = Visited), shape = 21) +
  labs(title = "GRCA Sample Event History",
       subtitle = "Spruce-Fir Ecosystem (PIEN)",
       y = "Macroplot",
       x = "Year",
       color = "Entry") +
  scale_fill_manual(values = c("Y" = "forestgreen", "N" = "red")) +
  scale_color_manual(values=c('forestgreen', "yellow", 'darkorange', "red", "darkred")) +
  theme_light() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(size = 20, face="bold", hjust = 0.5),
        plot.subtitle = element_text(size = 12, face="italic", hjust = 0.5),
        axis.title = element_text(size = 12, face="bold"),
        axis.title.y = element_text(margin = unit(c(0,4,0,0), "mm")),
        axis.title.x = element_text(margin = unit(c(4,0,0,0), "mm")),
        axis.text.y = element_text(size = 5),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1))
```

### Fires

```{r}
graph_fires_MetadataReport <- graph_MetadataReport %>% 
  filter(Park == "GRCA",
         PlotType_Name == "FMH") %>% 
  filter(!Macroplot %in% c("SHOS 01", "SHOS 02")) %>% 
  filter(Entry != "0") %>% 
  group_by(Burn_Unit, DisturbanceDate, DisturbanceDate_Year, SampleEventDate_Year, Park_Unit, MonStatus, Disturbance_Name, Disturbance_Desc, Disturbance_Severity) %>% 
  filter(DisturbanceDate_Year == SampleEventDate_Year) %>% 
  summarise(n_macroplot = n())
```

```{r}
graph_fires_MetadataReport %>% 
  filter(Park_Unit == "North") %>% 
  ggplot(aes(x = DisturbanceDate, y = Burn_Unit, label = Disturbance_Desc)) +
  geom_line(aes(group = Burn_Unit)) +
  #geom_point(aes(size  = n_macroplot), shape = "\U1F525") +
  geom_point(aes(fill = Disturbance_Severity, size  = n_macroplot), shape = 21) +
  #geom_text(size = 4) +
  #geom_label_repel(aes(size = 4)) +
  labs(title = "GRCA North Rim Fire History",
       y = "Burn Unit",
       x = "Year",
       fill = "Severity",
       size = "Number of Plots") +
  scale_fill_manual(values=c('yellowgreen', "yellow", 'orange', "red", "darkred","grey")) +
  theme_light() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(size = 20, face="bold", hjust = 0.5),
        axis.title = element_text(size = 12, face="bold"),
        axis.title.y = element_text(margin = unit(c(0,4,0,0), "mm")),
        axis.title.x = element_text(margin = unit(c(4,0,0,0), "mm")),
        axis.text.y = element_text(size = 8),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1))
```

```{r}
graph_fires_MetadataReport %>% 
  filter(Park_Unit == "South") %>% 
  ggplot(aes(x = DisturbanceDate, y = Burn_Unit, label = Disturbance_Desc)) +
  geom_line(aes(group = Burn_Unit)) +
  #geom_point(aes(size  = n_macroplot), shape = "\U1F525") +
  geom_point(aes(fill = Disturbance_Severity, size  = n_macroplot), shape = 21) +
  #geom_text(size = 4) +
  #geom_label_repel(aes(size = 4)) +
  labs(title = "GRCA South Rim Fire History",
       y = "Burn Unit",
       x = "Year",
       fill = "Severity",
       size = "Number of Plots") +
  scale_fill_manual(values=c('yellowgreen', "yellow", 'orange', "red", "darkred","grey")) +
  theme_light() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(size = 20, face="bold", hjust = 0.5),
        axis.title = element_text(size = 12, face="bold"),
        axis.title.y = element_text(margin = unit(c(0,4,0,0), "mm")),
        axis.title.x = element_text(margin = unit(c(4,0,0,0), "mm")),
        axis.text.y = element_text(size = 8),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1))
```
