---
title: "FFI Database Summary"
subtitle: "Grand Canyon"
output: 
  html_document:
    toc: true
    toc-float: true
    code_download: true
date: "2025-03-17"
---

## Purpose

This report summarizes plot work by the Fire Ecology Program. The included tables and enclosed code allow a user to quickly view the overall status of program, park, project, or protocol activities in an FFI admin unit. The code summarizes records in the Macroplot Report and Sample Event Report that can be downloaded via the Utilities menu on the Project Management tab on the FFI Remote App Server.

### Setup

```{r setup, include=FALSE}
## Markdown Options
knitr::opts_chunk$set(echo = FALSE, error = FALSE, warnings = FALSE, message = FALSE)
```

```{r install packages, include=FALSE}
## Install packages (if needed)

# install.packages("tidyverse")
# install.packages("kableExtra")
```

```{r load packages, include=FALSE}
## Load Packages

# "here" helps you easily find and reference your working directory
library(here)
# "tidyverse" has lots of useful functions for data cleaning
library(tidyverse)
# "kableExtra" is used to produce formatted tables
library(kableExtra)
# library(stringr)
# library(DT)
# library(fs)
```

### Adjust File Paths

```{r import data, include=FALSE}
# Identify working directory
here()

# Load in data.
path_data <- "X:/FFI Data Management/Exports from FFI/GRCA_FMH/metadata/"
path_output <- paste0(here(), "/output/")
path_data <- "D:/Allie/_FireFX/FFI Data Management/Exports from FFI/GRCA_FMH/"
```

### Load Data

```{r load data, include=FALSE}
## Load Data

MacroplotReport <- read.csv(paste0(path_data, "metadata/MacroplotReport.csv"))
SampleEventReport <- read.csv(paste0(path_data, "metadata/SampleEventReport.csv"))
#ProjectUnitReport <- read.csv(paste0(path_data, "metadata/ProjectUnitAssignmentReport.csv"))
```

### Clean Data

#### Macroplot Report

```{r}
MacroplotReport <- MacroplotReport %>% 
  # separate Date columns into Date and Time columns
  separate(SampleEventDate, sep = " ", into = c("SampleEventDate", "SampleEventTime")) %>% 
  separate(DateIn, sep = " ", into = c("DateIn", "DateInTime")) %>% 
  # reformate Date columns
  mutate(SampleEventDate = as.Date(SampleEventDate, "%m/%d/%Y"),
         DateIn = as.Date(DateIn, "%m/%d/%Y")) %>% 
  # separate Date columns into Year, Month, and Day columns
  separate(SampleEventDate, sep = "-", into = c("SampleEventDate_Year", "SampleEventDate_Month", "SampleEventDate_Day"), remove = FALSE) %>% 
  separate(DateIn, sep = "-", into = c("DateIn_Year", "DateIn_Month", "DateIn_Day"), remove = FALSE) %>% 
  # remove unnecessary columns
  select(!c(SampleEventTime, SampleEventDate_Month, SampleEventDate_Day, 
            DateInTime, DateIn_Month, DateIn_Day)) %>% 
  # add Disturbance_Type column
  mutate(Disturbance_Type = ifelse(Purpose %in% c("Tusayan Pueblo (Thinning)"), "thin", "burn"))
```

#### Sample Event Report

```{r}
SampleEventReport <- SampleEventReport %>% 
  # remove trailing/leading whitespace
  mutate(across(where(is.character), str_trim)) %>%
  # separate Date columns into Date and Time columns
  separate(SampleEvent_Date, sep = " ", into = c("SampleEventDate", "SampleEventTime")) %>%
  # reformate Date columns
  mutate(SampleEventDate = as.Date(SampleEventDate, "%m/%d/%Y")) %>% 
  # separate Date columns into Year, Month, and Day columns
  separate(SampleEventDate, sep = "-", into = c("SampleEventDate_Year", "SampleEventDate_Month", "SampleEventDate_Day"), remove = FALSE) %>% 
  # remove unnecessary columns
  select(!c(SampleEventTime, SampleEventDate_Month, SampleEventDate_Day)) %>%
  # make column names match those in MacroplotReport
  mutate(Macroplot = MacroPlot_Name,
         LegacyMonitoringStatus = SampleEvent_LegacyMonitoringStatus,
         MonStatus = MonitoringStatus_Name) %>% 
  # add PlotType_Name column
  mutate(PlotType_Name = ifelse(grepl("RAP", ProjectUnit_Name), "RAP", 
                                 ifelse(grepl("GR", ProjectUnit_Name), "Meadow", "FMH"))) %>%
  # reorganize columns
  relocate(Macroplot, .after = ProjectUnit_Name) %>% 
  relocate(LegacyMonitoringStatus, .after = SampleEventDate_Year) %>% 
  relocate(MonStatus, .after = LegacyMonitoringStatus) %>%
  relocate(PlotType_Name, .after = AdministrationUnit_Name) %>% 
  select(!c(MacroPlot_Name, SampleEvent_LegacyMonitoringStatus, MonitoringStatus_Name))
```

### Filter Data

```{r}
MacroplotReport_MetaData <- MacroplotReport %>% 
  filter(is.na(SampleEventDate)) %>% 
  select(!c(SampleEventDate, SampleEventDate_Year, SampleEventTeam, SampleEventComment, MonStatusOrd, LegacyMonStatus, MonStatus))

MacroplotReport_SampleEvents <- MacroplotReport %>% 
  filter(!is.na(SampleEventDate)) %>% 
  select(c(Macroplot, SampleEventDate, SampleEventDate_Year, SampleEventTeam, SampleEventComment, MonStatusOrd, LegacyMonStatus, MonStatus))
```

## Summary Tables

### Monitoring Status

Create columns for Entry (number of disturbances), Status (pre or post disturbance), and Time Since disturbance (i.e., Year 1) via break down of the Monitoring Status field

```{r}
sum_MonStatus <- MacroplotReport_SampleEvents %>% 
  select(MonStatus) %>% 
  unique() %>% 
  filter(MonStatus != "") %>% 
  # extract info based on monitoring status names
  extract(MonStatus, into = c("Entry","Status","TimeSince"),
          regex = "([0-9]+)([A-Z][a-z]+)([0-9]+)", 
          remove = FALSE) %>% 
  # fix up classification of "Year" reads
  mutate(Status = ifelse(Status == "Year", "Post", Status)) %>% 
  # fix up classifications of "Pre" reads
  mutate(Entry = ifelse(grepl("Pre", MonStatus), 0, Entry),
         Status = ifelse(grepl("Pre", MonStatus), "Pre", Status)) %>% 
  # fix up classifications of "Post" reads
  mutate(Entry = ifelse(grepl("Post", MonStatus), substr(MonStatus, start = 1, stop = 2), Entry),
         Status = ifelse(grepl("Post", MonStatus), "Post", Status),
         TimeSince = ifelse(grepl("Post", MonStatus), 0, TimeSince)) %>% 
  # fix up classifications of "PR" reads
  mutate(Entry = ifelse(grepl("PR", MonStatus), "0", Entry),
         Status = ifelse(grepl("PR", MonStatus), "Pre", Status)) %>%
  # make columns numeric
  mutate(Entry = as.numeric(Entry),
         TimeSince = as.numeric(TimeSince))
```

### Disturbance History

Create columns for Entry (number of disturbances), Disturbance_Name (i.e., fire name or project name), Disturbance_Desc (i.e., RX, WF, Supp), DisturbanceDate, and Severity (i.e., low, mod, high).

```{r}
# Ensure blanks in UV columns are NA
MacroplotReport_MetaData$UV1[MacroplotReport_MetaData$UV1==""] <- NA
MacroplotReport_MetaData$UV2[MacroplotReport_MetaData$UV2==""] <- NA
MacroplotReport_MetaData$UV3[MacroplotReport_MetaData$UV3==""] <- NA
MacroplotReport_MetaData$UV4[MacroplotReport_MetaData$UV4==""] <- NA
MacroplotReport_MetaData$UV5[MacroplotReport_MetaData$UV5==""] <- NA
MacroplotReport_MetaData$UV6[MacroplotReport_MetaData$UV6==""] <- NA
MacroplotReport_MetaData$UV7[MacroplotReport_MetaData$UV7==""] <- NA
MacroplotReport_MetaData$UV8[MacroplotReport_MetaData$UV8==""] <- NA
```

#### 01 Disturbance

```{r}
# add add relevant columns
Disturbance01 <- MacroplotReport_MetaData %>% 
  select(c("Macroplot", "Disturbance_Type", "UV2")) %>% 
  filter(!is.na(UV2)) %>% 
  separate(UV2, sep = ": ", into = c("Entry", "Disturbance_Info"), remove = FALSE) %>% 
  separate(Disturbance_Info, sep = ",", into = c("Disturbance_Name", "Disturbance_Desc", "DisturbanceDate", "Severity"), remove = FALSE) %>% 
  mutate(across(where(is.character), str_trim)) %>% 
  mutate(Entry = 1) %>% 
  select(!c("Disturbance_Info"))

# specific edits
Disturbance01 <- Disturbance01 %>% 
  mutate(Disturbance_Name = ifelse(Macroplot == "SC03", "Tipover East", Disturbance_Name),
         Disturbance_Desc = ifelse(Macroplot == "SC03", "Rx", Disturbance_Desc),
         DisturbanceDate = ifelse(Macroplot == "SC03", "12/22/17", DisturbanceDate),
         Severity = ifelse(Macroplot == "SC03", "(POST read '18)", Severity)) %>% 
  mutate(Disturbance_Name = ifelse(Macroplot == "SC16", "Tipover East", Disturbance_Name),
         Disturbance_Desc = ifelse(Macroplot == "SC16", "Rx", Disturbance_Desc),
         DisturbanceDate = ifelse(Macroplot == "SC16", "12/22/17", DisturbanceDate),
         Severity = ifelse(Macroplot == "SC16", "(POST read '18)", Severity)) %>% 
  mutate(Disturbance_Name = ifelse(Macroplot == "PIAB 06", "NW 1,3,5", Disturbance_Name),
         Disturbance_Desc = ifelse(Macroplot == "PIAB 06", "Rx", Disturbance_Desc),
         DisturbanceDate = ifelse(Macroplot == "PIAB 06", "11/9/2007", DisturbanceDate),
         Severity = ifelse(Macroplot == "PIAB 06", "Low Severity", Severity)) %>% 
  mutate(Disturbance_Name = ifelse(Macroplot == "PIAB 24", "NW 1,3,5", Disturbance_Name),
         Disturbance_Desc = ifelse(Macroplot == "PIAB 24", "Rx", Disturbance_Desc),
         DisturbanceDate = ifelse(Macroplot == "PIAB 24", "11/10/2007", DisturbanceDate),
         Severity = ifelse(Macroplot == "PIAB 24", "Moderate/Low Severity", Severity)) %>% 
  mutate(Disturbance_Desc = ifelse(Disturbance_Desc == "Rx", "RX", Disturbance_Desc)) %>% 
  select(!c("UV2"))

# edit date column
Disturbance01 <- Disturbance01 %>% 
  mutate(DisturbanceDate = parse_date_time(DisturbanceDate, orders = c("mdy", "my", "y"))) %>% 
  separate(DisturbanceDate, sep = "-", into = c("DisturbanceDate_Year", "Month", "Day"), remove = F) %>% 
  select(!c("Month", "Day")) %>% 
  relocate(DisturbanceDate_Year, .after = DisturbanceDate) %>% 
  relocate(Disturbance_Type, .before = Disturbance_Name)
```

#### 02 Disturbance

```{r}
# add add relevant columns
Disturbance02 <- MacroplotReport_MetaData %>% 
  select(c("Macroplot", "Disturbance_Type", "UV3")) %>% 
  filter(!is.na(UV3)) %>% 
  separate(UV3, sep = ": ", into = c("Entry", "Disturbance_Info"), remove = FALSE) %>% 
  separate(Disturbance_Info, sep = ",", into = c("Disturbance_Name", "Disturbance_Desc", "DisturbanceDate", "Severity"), remove = FALSE) %>% 
  mutate(across(where(is.character), str_trim)) %>% 
  mutate(Entry = 2) %>% 
  select(!c("Disturbance_Info"))

# specific edits
Disturbance02 <- Disturbance02 %>%
  mutate(Disturbance_Name = ifelse(Macroplot == "PIAB 07", "NW 1,3,5", Disturbance_Name),
         Disturbance_Desc = ifelse(Macroplot == "PIAB 07", "Rx", Disturbance_Desc),
         DisturbanceDate = ifelse(Macroplot == "PIAB 07", "11/10/2007", DisturbanceDate),
         Severity = ifelse(Macroplot == "PIAB 07", "Low Severity", Severity)) %>%
  mutate(Disturbance_Name = ifelse(Macroplot == "PIAB 25", "NW 1,3,5", Disturbance_Name),
         Disturbance_Desc = ifelse(Macroplot == "PIAB 25", "Rx", Disturbance_Desc),
         DisturbanceDate = ifelse(Macroplot == "PIAB 25", "11/10/2007", DisturbanceDate),
         Severity = ifelse(Macroplot == "PIAB 25", "Moderate/Low Severity", Severity)) %>%
  mutate(Disturbance_Name = ifelse(Macroplot == "PIPN 01", "NW 1,3,5", Disturbance_Name),
         Disturbance_Desc = ifelse(Macroplot == "PIPN 01", "Rx", Disturbance_Desc),
         DisturbanceDate = ifelse(Macroplot == "PIPN 01", "11/10/2007", DisturbanceDate),
         Severity = ifelse(Macroplot == "PIPN 01", "Moderate/Low Severity", Severity)) %>%
  mutate(Disturbance_Name = ifelse(Macroplot == "PIPN 02", "NW 1,3,5", Disturbance_Name),
         Disturbance_Desc = ifelse(Macroplot == "PIPN 02", "Rx", Disturbance_Desc),
         DisturbanceDate = ifelse(Macroplot == "PIPN 02", "11/10/2007", DisturbanceDate),
         Severity = ifelse(Macroplot == "PIPN 02", "Moderate/Low Severity", Severity)) %>%
  mutate(Disturbance_Name = ifelse(Macroplot == "PIPO 24", "Watson 2,3,4", Disturbance_Name),
         Disturbance_Desc = ifelse(Macroplot == "PIPO 24", "Rx", Disturbance_Desc),
         DisturbanceDate = ifelse(Macroplot == "PIPO 24", "10/14/2011", DisturbanceDate),
         Severity = ifelse(Macroplot == "PIPO 24", "Low Severity", Severity)) %>% 
  mutate(Disturbance_Desc = ifelse(Disturbance_Desc == "Rx", "RX", Disturbance_Desc)) %>% 
  select(!c("UV3"))

# edit date column
Disturbance02 <- Disturbance02 %>% 
  mutate(DisturbanceDate = parse_date_time(DisturbanceDate, orders = c("mdy", "my", "y"))) %>% 
  separate(DisturbanceDate, sep = "-", into = c("DisturbanceDate_Year", "Month", "Day"), remove = F) %>% 
  select(!c("Month", "Day")) %>% 
  relocate(DisturbanceDate_Year, .after = DisturbanceDate) %>% 
  relocate(Disturbance_Type, .before = Disturbance_Name)
```

#### 03 Disturbance

```{r}
# add add relevant columns
Disturbance03 <- MacroplotReport_MetaData %>% 
  select(c("Macroplot", "Disturbance_Type", "UV4")) %>% 
  filter(!is.na(UV4)) %>% 
  separate(UV4, sep = ": ", into = c("Entry", "Disturbance_Info"), remove = FALSE) %>% 
  separate(Disturbance_Info, sep = ",", into = c("Disturbance_Name", "Disturbance_Desc", "DisturbanceDate", "Severity"), remove = FALSE) %>% 
  mutate(across(where(is.character), str_trim)) %>% 
  mutate(Entry = 3) %>% 
  select(!c("Disturbance_Info"))

# specific edits
Disturbance03 <- Disturbance03 %>%
  mutate(Disturbance_Name = ifelse(Macroplot == "PIPO 17", "Watson 2,3,4", Disturbance_Name),
         Disturbance_Desc = ifelse(Macroplot == "PIPO 17", "Rx", Disturbance_Desc),
         DisturbanceDate = ifelse(Macroplot == "PIPO 17", "10/14/2011", DisturbanceDate),
         Severity = ifelse(Macroplot == "PIPO 17", "Low Severity", Severity)) %>%
  mutate(Disturbance_Name = ifelse(Macroplot == "PIPO 20", "Watson 2,3,4", Disturbance_Name),
         Disturbance_Desc = ifelse(Macroplot == "PIPO 20", "Rx", Disturbance_Desc),
         DisturbanceDate = ifelse(Macroplot == "PIPO 20", "10/14/2011", DisturbanceDate),
         Severity = ifelse(Macroplot == "PIPO 20", "Low Severity", Severity)) %>%
  mutate(Disturbance_Name = ifelse(Macroplot == "PIPO 21", "Watson 2,3,4", Disturbance_Name),
         Disturbance_Desc = ifelse(Macroplot == "PIPO 21", "Rx", Disturbance_Desc),
         DisturbanceDate = ifelse(Macroplot == "PIPO 21", "10/14/2011", DisturbanceDate),
         Severity = ifelse(Macroplot == "PIPO 21", "Low Severity", Severity)) %>%
  mutate(Disturbance_Name = ifelse(Macroplot == "PIPO 22", "Watson 2,3,4", Disturbance_Name),
         Disturbance_Desc = ifelse(Macroplot == "PIPO 22", "Rx", Disturbance_Desc),
         DisturbanceDate = ifelse(Macroplot == "PIPO 22", "10/14/2011", DisturbanceDate),
         Severity = ifelse(Macroplot == "PIPO 22", "Low Severity", Severity)) %>%
  mutate(Disturbance_Desc = ifelse(Disturbance_Desc == "Rx", "RX", Disturbance_Desc)) %>% 
  select(!c("UV4"))

# edit date column
Disturbance03 <- Disturbance03 %>% 
  mutate(DisturbanceDate = parse_date_time(DisturbanceDate, orders = c("mdy", "my", "y"))) %>% 
  separate(DisturbanceDate, sep = "-", into = c("DisturbanceDate_Year", "Month", "Day"), remove = F) %>% 
  select(!c("Month", "Day")) %>% 
  relocate(DisturbanceDate_Year, .after = DisturbanceDate) %>% 
  relocate(Disturbance_Type, .before = Disturbance_Name)
```

#### 04 Disturbance

```{r}
Disturbance04 <- MacroplotReport_MetaData %>% 
  select(c("Macroplot", "Disturbance_Type", "UV5")) %>% 
  filter(!is.na(UV5)) %>% 
  # add add relevant columns
  separate(UV5, sep = ": ", into = c("Entry", "Disturbance_Info"), remove = FALSE) %>% 
  separate(Disturbance_Info, sep = ",", into = c("Disturbance_Name", "Disturbance_Desc", "DisturbanceDate", "Severity"), remove = FALSE) %>% 
  mutate(across(where(is.character), str_trim)) %>% 
  mutate(Entry = 4) %>% 
  select(!c("Disturbance_Info", "UV5")) %>% 
  # edit date column
  mutate(DisturbanceDate = parse_date_time(DisturbanceDate, orders = c("mdy", "my", "y"))) %>% 
  separate(DisturbanceDate, sep = "-", into = c("DisturbanceDate_Year", "Month", "Day"), remove = F) %>% 
  select(!c("Month", "Day")) %>% 
  relocate(DisturbanceDate_Year, .after = DisturbanceDate) %>% 
  relocate(Disturbance_Type, .before = Disturbance_Name)
```

#### All Disturbances

```{r}
sum_DisturbanceHistory <- rbind(Disturbance01, Disturbance02, Disturbance03, Disturbance04)

remove(Disturbance01, Disturbance02, Disturbance03, Disturbance04)
```

## Macroplot Report

```{r}
# merge metadata and sample events
MacroplotReport_all <- merge(MacroplotReport_MetaData, MacroplotReport_SampleEvents, by = c("Macroplot"))

# merge monitoring status summary
MacroplotReport_all <- merge(MacroplotReport_all, sum_MonStatus, by = c("MonStatus"), all = T) %>% 
  relocate(MonStatus, .after = LegacyMonStatus)

# merge disturbance history summmary
MacroplotReport_all <- merge(MacroplotReport_all, sum_DisturbanceHistory, by = c("Macroplot", "Entry", "Disturbance_Type"), all = T) %>% 
  relocate(Entry, .after = MonStatus) %>% 
  relocate(Disturbance_Type, .before = Disturbance_Name)
```

## Sample Event Report

```{r}
# merge monitoring status summary
SampleEventReport_all <- merge(SampleEventReport, sum_MonStatus, by = c("MonStatus"), all = T) %>% 
  relocate(MonStatus, .before = Entry)

# merge disturbance history summmary
SampleEventReport_all <- merge(SampleEventReport_all, sum_DisturbanceHistory, by = c("Macroplot", "Entry"), all = T) %>% 
  relocate(Entry, .after = MonStatus)
```

## Tables

```{r manipulate data, include=FALSE}
# check occurrence of unique values in newly created columns, make corrections as needed
SampleEventReport_all$PlotType_Name %>% table(exclude = NULL)
SampleEventReport_all$Status %>% table(exclude = NULL)
SampleEventReport_all$Disturbance_Type %>% table(exclude = NULL)
SampleEventReport_all$TimeSince %>% table(exclude = NULL)
```

### Overall Program Summary

```{r program summary}
## Summarise projects and events by park
program_summ <- SampleEventReport_all %>%
  filter(!ProjectUnit_Name %in% c("z2014 + 2017_Slopes RX", "z2017_Tipover East RX", "z2017_Slopes RX", "zHiSevMSO", "zSRIMRx2024Spring_Latest")) %>% 
  filter(!ProjectUnit_Name %in% c("ALL_FMH", "ALL_RAP_NRim", "ALL_RAP_SRim")) %>% 
  filter(!ProjectUnit_Name %in% c("GRIN", "GRED")) %>% 
  summarise(`Park` = "GRCA",
            `Projects` = n_distinct(ProjectUnit_Name),
            `Number of Plots` = n_distinct(Macroplot),
            `Monitoring Statuses` = n_distinct(MonStatus),
            `Visit Dates` = n_distinct(SampleEventDate),
            `Protocols Used` = n_distinct(Protocols))

# Calculate totals
program_summ_totals <- program_summ %>%
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE))) %>%
  mutate(Park = "Total")

# Bind the totals row to the original data frame
program_summ_w_totals <- bind_rows(program_summ, program_summ_totals)

program_summ_w_totals %>% 
  kbl(align = "lcccccc") %>%
  kable_styling(bootstrap_options = c("striped","condensed"), 
                full_width = F, position = "left", fixed_thead = T,
                row_label_position = "c") %>% 
  row_spec(0, font_size = 14, bold = TRUE) %>% 
  row_spec(nrow(program_summ_w_totals), font_size = 14, bold = TRUE)
```

### Filtered Summaries

The code underlying this report has been manipulated to display summaries of the following selected park(s), project(s), and year(s).\
\
Selected park(s): [`r target_park`]{style="color:red"}\
Selected project(s): [`r target_project`]{style="color:red"}\
Selected year(s): [`r target_year`]{style="color:red"}\

```{r set parameters, include=F}
### Select Data to Summarize
# use commas to build lists of acceptable values for filters

# designate target park(s)
target_park <- c("GRCA")

# designate target project(s)
# see all project names
unique(SampleEventReport_all$ProjectUnit_Name)
target_project <- c("ALL_FMH")

# designate target year or years
target_year <- c(2024)
```

#### Park Summary: `r target_park`

The table below summarizes current and past projects in `r target_park`.

```{r park summ}
## Summarise projects by park
# uses 'target park' variable designated above
target_park_summ <- SampleEventReport_all %>%
  filter(!ProjectUnit_Name %in% c("z2014 + 2017_Slopes RX", "z2017_Tipover East RX", "z2017_Slopes RX", "zHiSevMSO", "zSRIMRx2024Spring_Latest")) %>% 
  filter(!ProjectUnit_Name %in% c("ALL_FMH", "ALL_RAP_NRim", "ALL_RAP_SRim")) %>% 
  filter(!ProjectUnit_Name %in% c("GRIN", "GRED")) %>% 
  filter(AdministrationUnit_Name == target_park) %>% 
  group_by(`Project Name` = ProjectUnit_Name) %>%
  summarise(`First Visit` = min(as.numeric(SampleEventDate_Year)),
            `Latest Visit` = max(as.numeric(SampleEventDate_Year)),
            `Number of Plots` = n_distinct(Macroplot),
            `Monitoring Statuses` = n_distinct(MonStatus),
            `Protocols Used` = n_distinct(Protocols)) %>% 
  arrange(desc(`First Visit`)) %>% 
  ungroup() %>% 
  mutate_all(as.character)

# Calculate totals
target_park_summ_totals <- target_park_summ %>%
  summarise(`Number of Plots` = sum(as.numeric(`Number of Plots`))) %>% 
  mutate(`Project Name` = "",
         `First Visit` = "",
         `Latest Visit` = "Total",
         `Monitoring Statuses` = "",
         `Protocols Used` = "") %>% 
  mutate_all(as.character)

# Bind the totals row to the original data frame
target_park_summ_w_totals <- bind_rows(target_park_summ, target_park_summ_totals)
  
target_park_summ_w_totals %>% 
  kbl(align = "lccccc") %>%
  kable_styling(bootstrap_options = c("striped","condensed"), 
                full_width = F, position = "left", fixed_thead = T) %>% 
  row_spec(nrow(target_park_summ_w_totals), font_size = 14, bold = TRUE)
```

#### Project Summary: `r target_project`

The table below summarizes sample events recorded within the `r target_project` project.

```{r project summ}
## Summarise project
# uses 'target project' variable designated above
target_project_summ <- SampleEventReport_all %>%
  filter(ProjectUnit_Name == target_project) %>% 
  group_by(`Monitoring Status` = MonStatus) %>%
  summarise(`First Visit` = min(SampleEventDate_Year),
            `Latest Visit` = max(SampleEventDate_Year),
            `Number of Plots Visited in Monitoring Status` = n_distinct(Macroplot),
            `Protocols Used during Monitoring Status` = n_distinct(Protocols)) %>% 
  arrange(`First Visit`)

target_project_summ %>% 
  kbl(align = "lcccc") %>%
  kable_styling(bootstrap_options = c("striped","condensed"), 
                full_width = F, position = "left", fixed_thead = T) %>% 
  row_spec(0, font_size = 14, bold = TRUE)
```

#### Plot Activity Summary: `r target_year`

The table below summarizes all projects with plots sampled within the following year(s): `r target_year`.

```{r year summ}
## Summarise yearly plot activity
# uses 'target year' variable designated above
target_year_summ <- SampleEventReport_all %>%
  filter(!ProjectUnit_Name %in% c("z2014 + 2017_Slopes RX", "z2017_Tipover East RX", "z2017_Slopes RX", "zHiSevMSO", "zSRIMRx2024Spring_Latest")) %>% 
  filter(!ProjectUnit_Name %in% c("ALL_FMH", "ALL_RAP_NRim", "ALL_RAP_SRim")) %>% 
  filter(!ProjectUnit_Name %in% c("GRIN", "GRED")) %>% 
  filter(SampleEventDate_Year %in% target_year) %>% 
  group_by(`Project Name` = ProjectUnit_Name) %>%
  summarise(`Start Date` = as.character(min(SampleEventDate)),
            `End Date` = as.character(max(SampleEventDate)),
            `Number of Plots` = n_distinct(Macroplot),
            `Protocols Used` = n_distinct(Protocols)) %>% 
  arrange(`Start Date`) %>% 
  ungroup()

# Calculate totals
year_summ_totals <- target_year_summ %>%
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE))) %>%
  mutate(`Project Name` = "",
         `Start Date` = "",
         `End Date` = "Total")

# Bind the totals row to the original data frame
target_year_summ_w_totals <- bind_rows(target_year_summ, year_summ_totals)

target_year_summ_w_totals %>% 
  kbl(align = "lcccc") %>%
  kable_styling(bootstrap_options = c("striped","condensed"), 
                full_width = F, position = "left", fixed_thead = T) %>% 
  row_spec(0, font_size = 14, bold = TRUE) %>% 
  row_spec(nrow(target_year_summ_w_totals), font_size = 14, bold = TRUE)
```

### Protocol Use Summary

The table below summarizes the frequency of use of protocols occurring in the FFI database, as utilized by the NPS Alaska Fire Ecology program. Protocols are arranged in descending order by number of plots such that the most frequently used protocols are listed first.

```{r protocol summ}
## Summarise protocol use by status
protocol_use_summ <- SampleEventReport_all %>%
  filter(!ProjectUnit_Name %in% c("z2014 + 2017_Slopes RX", "z2017_Tipover East RX", "z2017_Slopes RX", "zHiSevMSO", "zSRIMRx2024Spring_Latest")) %>% 
  filter(!ProjectUnit_Name %in% c("ALL_FMH", "ALL_RAP_NRim", "ALL_RAP_SRim")) %>% 
  filter(!ProjectUnit_Name %in% c("GRIN", "GRED")) %>% 
  group_by(`Protocol` = Protocols) %>%
  summarise(`First Use` = min(as.numeric(SampleEventDate_Year)),
            `Latest Use` = max(as.numeric(SampleEventDate_Year)),
            `Number of Plots` = n_distinct(Macroplot),
            `Number of Projects` = n_distinct(ProjectUnit_Name),
            `Number of Monitoring Statuses` = n_distinct(MonStatus),
            `Number of Parks` = n_distinct(AdministrationUnit_Name)) %>% 
  arrange(desc(`Number of Plots`))

protocol_use_summ %>% 
  kbl(align = "lcccccc") %>%
  kable_styling(bootstrap_options = c("striped","condensed"), 
                full_width = F, position = "left", fixed_thead = T) %>% 
  row_spec(0, font_size = 14, bold = TRUE)
```

## Graphs

### Macroplot Report

#### PIED

```{r}
MacroplotReport_all %>% 
  mutate(MonStatus = fct_reorder(factor(MonStatus), as.integer(MonStatusOrd)),
         SampleEventDate_Year = as.factor(SampleEventDate_Year)) %>% 
  filter(Purpose %in% c("PIED")) %>% 
  ggplot(aes(x = SampleEventDate_Year, y = Macroplot, color = as.factor(Entry))) +
  geom_point() +
  geom_line(aes(group = Macroplot)) +
  labs(title = "Sample Event History",
       y = "Macroplot",
       x = "Year",
       color = "Entry") +
  scale_color_manual(values=c('forestgreen', "yellow", 'darkorange', "red", "darkred")) +
  theme_light() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(size = 20, face="bold", hjust = 0.5),
        axis.title = element_text(size = 15, face="bold"),
        axis.title.y = element_text(margin = unit(c(0,4,0,0), "mm")),
        axis.title.x = element_text(margin = unit(c(4,0,0,0), "mm")),
        axis.text.y = element_text(size = 5),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1))
```

#### PIPO

```{r}
MacroplotReport_all %>% 
  mutate(MonStatus = fct_reorder(factor(MonStatus), as.integer(MonStatusOrd))) %>% 
  filter(Purpose %in% c("PIPO")) %>% 
  ggplot(aes(x = SampleEventDate_Year, y = Macroplot, color = as.factor(Entry))) +
  geom_point() +
  geom_line(aes(group = Macroplot)) +
  labs(title = "Sample Event History",
       y = "Macroplot",
       x = "Year",
       color = "Entry") +
  scale_color_manual(values=c('forestgreen', "yellow", 'darkorange', "red", "darkred")) +
  theme_light() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(size = 20, face="bold", hjust = 0.5),
        axis.title = element_text(size = 15, face="bold"),
        axis.title.y = element_text(margin = unit(c(0,4,0,0), "mm")),
        axis.title.x = element_text(margin = unit(c(4,0,0,0), "mm")),
        axis.text.y = element_text(size = 5),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1))
```

#### PIPN

```{r}
MacroplotReport_all %>% 
  mutate(MonStatus = fct_reorder(factor(MonStatus), as.integer(MonStatusOrd))) %>% 
  filter(Purpose %in% c("PIPN")) %>% 
  ggplot(aes(x = SampleEventDate_Year, y = Macroplot, color = as.factor(Entry))) +
  geom_point() +
  geom_line(aes(group = Macroplot)) +
  labs(title = "Sample Event History",
       y = "Macroplot",
       x = "Year",
       color = "Entry") +
  scale_color_manual(values=c('forestgreen', "yellow", 'darkorange', "red", "darkred")) +
  theme_light() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(size = 20, face="bold", hjust = 0.5),
        axis.title = element_text(size = 15, face="bold"),
        axis.title.y = element_text(margin = unit(c(0,4,0,0), "mm")),
        axis.title.x = element_text(margin = unit(c(4,0,0,0), "mm")),
        axis.text.y = element_text(size = 5),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1))
```

#### PIAB

```{r}
MacroplotReport_all %>% 
  mutate(MonStatus = fct_reorder(factor(MonStatus), as.integer(MonStatusOrd))) %>% 
  filter(Purpose %in% c("PIAB")) %>% 
  ggplot(aes(x = SampleEventDate_Year, y = Macroplot, color = as.factor(Entry))) +
  geom_point() +
  geom_line(aes(group = Macroplot)) +
  labs(title = "Sample Event History",
       y = "Macroplot",
       x = "Year",
       color = "Entry") +
  scale_color_manual(values=c('forestgreen', "yellow", 'darkorange', "red", "darkred")) +
  theme_light() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(size = 20, face="bold", hjust = 0.5),
        axis.title = element_text(size = 15, face="bold"),
        axis.title.y = element_text(margin = unit(c(0,4,0,0), "mm")),
        axis.title.x = element_text(margin = unit(c(4,0,0,0), "mm")),
        axis.text.y = element_text(size = 5),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1))
```

#### PIEN

```{r}
MacroplotReport_all %>% 
  mutate(MonStatus = fct_reorder(factor(MonStatus), as.integer(MonStatusOrd))) %>% 
  filter(Purpose %in% c("PIEN")) %>% 
  ggplot(aes(x = SampleEventDate_Year, y = Macroplot, color = as.factor(Entry))) +
  geom_point() +
  geom_line(aes(group = Macroplot)) +
  labs(title = "Sample Event History",
       y = "Macroplot",
       x = "Year",
       color = "Entry") +
  scale_color_manual(values=c('forestgreen', "yellow", 'darkorange', "red", "darkred")) +
  theme_light() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(size = 20, face="bold", hjust = 0.5),
        axis.title = element_text(size = 15, face="bold"),
        axis.title.y = element_text(margin = unit(c(0,4,0,0), "mm")),
        axis.title.x = element_text(margin = unit(c(4,0,0,0), "mm")),
        axis.text.y = element_text(size = 5),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1))
```
