---
title: "FFI Database Summary"
subtitle: "Grand Canyon"
output: 
  html_document:
    theme: readable
    highlight: 
    toc: yes
    toc_depth: 3
    toc_float:
      smooth_scroll: yes
    code_download: true
date: "`r Sys.Date()`"
---

# Purpose

### Setup

```{r}
## Markdown Options
knitr::opts_chunk$set(echo = FALSE, error = FALSE, warning = FALSE, message = FALSE)
```

```{r}
## Install packages (if needed)

# install.packages("tidyverse")
# install.packages("kableExtra")
# install.packages("ggplot2")
# install.packages("readxl")
```

```{r}
## Load Packages

# "here" helps you easily find and reference your working directory
library(here)
# "tidyverse" has lots of useful functions for data cleaning
library(tidyverse)
# "kableExtra" is used to produce formatted tables
library(kableExtra)
library(ggplot2)
library(readxl)
```

### Adjust File Paths

```{r}
# Adjust File Paths
path_data <- "X:/FFI Data Management/Exports from FFI/GRCA_FMH/2025-01-23/"
path_errors <- "X:/FFI Data Management/Queries/QAQC/GRCA_FMH+RAP/"
```

### Load Data

```{r}
# Load in data
Trees_all <- read.csv(paste0(path_data, "Trees - Individuals (metric)_XPT.csv"), quote = "")
errors_Trees_CrossYear <- read_excel(paste0(path_errors, "errors_GRCA_FMH_CrossYear_Trees_formatted.xlsx"))
```

### Clean Data

```{r}
# Reformat date column
Trees_all <- Trees_all %>%
  # Separate Date column
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>% 
  # Reformat Date column
  mutate(Date = as.Date(Date, format = "%m/%d/%Y")) %>% 
  # Remove Month, Day, and Time columns
  select(!c("Month", "Day", "Time"))
```

### Filter Data

```{r}
# Ensure blanks in Visited column are NA 
Trees_all$Visited[Trees_all$Visited==""] <- NA
# Trees data
Trees_data_CrossYear <- Trees_all %>%   
  filter(is.na(Visited)) 
```

# Data Viz - Trees

#### Mutate Data for Graphing

```{r}
graph_Trees_data <- Trees_data_CrossYear %>% 
  mutate(Status = ifelse(CrwnCl == "DD", "DD", Status)) %>% 
  mutate(SubFrac = factor(SubFrac),
         TagNo = factor(TagNo),
         Status = factor(Status, levels = c("L", "D", "DD")),
         CrwnCl = factor(CrwnCl, levels = c("X", "",
                                            "SC", "I", "C", "D",
                                            "RS", "LBS", "CS", "BAD", 
                                            "BBD", "CUS", "DD")))
```

#### Define Target Plot & Isolate Problematic Tree Tags

```{r}
##################################################################################
#unique(errors_Trees_CrossYear$MacroPlot.Name)
# plots_2025 <- c("PIPO 02", "PIPO 03", "PIPO 09", "PIPO 13", "PIPO 14", "PIPO 19", "PIPO 39", "PIPN 05", "PIPN 12", "PIEN 16")
#plots_2025_errors <- c("PIPO 09", "PIPO 14", "PIPO 19", "PIPO 39", "PIPN 05", "PIEN 16")
target_plot <- "PIAB 03"
##################################################################################


badtags_CBH <- errors_Trees_CrossYear %>% 
  filter(grepl("CBH", errors_Trees_CrossYear$SavedQuery)) %>% 
  filter(MacroPlot.Name == target_plot)
badtags_CBH <- unique(badtags_CBH$TagNo)

badtags_DBH <- errors_Trees_CrossYear %>% 
  filter(grepl("DBH", errors_Trees_CrossYear$SavedQuery)) %>% 
  filter(MacroPlot.Name == target_plot)
badtags_DBH <- unique(badtags_DBH$TagNo)

badtags_Height <- errors_Trees_CrossYear %>% 
  filter(grepl("Height", errors_Trees_CrossYear$SavedQuery)) %>% 
  filter(MacroPlot.Name == target_plot)
badtags_Height <- unique(badtags_Height$TagNo)

badtags_Species <- errors_Trees_CrossYear %>% 
  filter(grepl("Species", errors_Trees_CrossYear$SavedQuery)) %>% 
  filter(MacroPlot.Name == target_plot)
badtags_Species <- unique(badtags_Species$TagNo)

badtags_Status <- errors_Trees_CrossYear %>% 
  filter(grepl("Status", errors_Trees_CrossYear$SavedQuery)) %>% 
  filter(MacroPlot.Name == target_plot)
badtags_Status <- unique(badtags_Status$TagNo)
```

## [`r target_plot`]{style="color:red"}

### Status

```{r}
graph_Trees_data %>% 
  filter(MacroPlot.Name == target_plot,
         TagNo %in% c(badtags_Status)) %>% 
  arrange(Date) %>% 
  ggplot(aes(x = Year, y = TagNo)) +
  geom_line(aes(group = TagNo, color = Status), size = 1) +
  geom_point(aes(fill = Status), shape = 21, size = 3) +
  scale_fill_manual(values = c("L" = "forestgreen", "D" = "darkorange", "DD" = "red")) +
  scale_color_manual(values=c("L" = "forestgreen", "D" = "darkorange", "DD" = "red")) +
  labs(title = target_plot) +
  theme_light() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(size = 20, face="bold", hjust = 0.5),
        plot.subtitle = element_text(size = 12, face="italic", hjust = 0.5),
        axis.title = element_text(size = 12, face="bold"),
        axis.title.y = element_text(margin = unit(c(0,4,0,0), "mm")),
        axis.title.x = element_text(margin = unit(c(4,0,0,0), "mm")),
        axis.text.y = element_text(size = 5),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1))
```

### Species

```{r}
graph_Trees_data %>% 
  filter(MacroPlot.Name == target_plot,
         TagNo %in% c(badtags_Species)) %>% 
  arrange(Date) %>% 
  ggplot(aes(x = Year, y = TagNo)) +
  geom_line(aes(group = TagNo, color = Species.Symbol), size = 1) +
  geom_point(aes(fill = Status), shape = 21, size = 3) +
  scale_fill_manual(values = c("L" = "forestgreen", "D" = "darkorange", "DD" = "red")) +
  labs(title = target_plot) +
  theme_light() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(size = 20, face="bold", hjust = 0.5),
        plot.subtitle = element_text(size = 12, face="italic", hjust = 0.5),
        axis.title = element_text(size = 12, face="bold"),
        axis.title.y = element_text(margin = unit(c(0,4,0,0), "mm")),
        axis.title.x = element_text(margin = unit(c(4,0,0,0), "mm")),
        axis.text.y = element_text(size = 5),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1))
```

### DBH

Live trees should only increase DBH. Dead trees should only decrease DBH.

```{r}
graph_Trees_data %>% 
  filter(MacroPlot.Name == target_plot,
         TagNo %in% c(badtags_DBH),
         !is.na(DBH),
         Status != "DD") %>% 
  arrange(Date) %>% 
  ggplot(aes(x = Year, y = DBH)) +
  geom_line(aes(group = TagNo, color = TagNo), size = 1) +
  geom_point(aes(fill = Status), shape = 21, size = 3) +
  scale_fill_manual(values=c("L" = "forestgreen", "D" = "darkorange")) +
  labs(title = target_plot) +
  #facet_wrap(~Status) +
  theme_light() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(size = 20, face="bold", hjust = 0.5),
        plot.subtitle = element_text(size = 12, face="italic", hjust = 0.5),
        axis.title = element_text(size = 12, face="bold"),
        axis.title.y = element_text(margin = unit(c(0,4,0,0), "mm")),
        axis.title.x = element_text(margin = unit(c(4,0,0,0), "mm")),
        axis.text.y = element_text(size = 5),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1))
```

### Height

Live trees should only increase in height. Dead trees should only decrease in height.

```{r}
graph_Trees_data %>% 
  filter(MacroPlot.Name == target_plot,
         TagNo %in% c(badtags_Height),
         !is.na(Ht),
         Status != "DD") %>% 
  arrange(Date) %>% 
  ggplot(aes(x = Year, y = Ht)) +
  geom_line(aes(group = TagNo, color = TagNo), size = 1) +
  geom_point(aes(fill = Status), shape = 21, size = 3) +
  scale_fill_manual(values=c("L" = "forestgreen", "D" = "darkorange")) +
  labs(title = target_plot) +
  #facet_wrap(~Status) +
  theme_light() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(size = 20, face="bold", hjust = 0.5),
        plot.subtitle = element_text(size = 12, face="italic", hjust = 0.5),
        axis.title = element_text(size = 12, face="bold"),
        axis.title.y = element_text(margin = unit(c(0,4,0,0), "mm")),
        axis.title.x = element_text(margin = unit(c(4,0,0,0), "mm")),
        axis.text.y = element_text(size = 5),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1))
```

### CBH

Live trees should only increase crown base height

```{r}
graph_Trees_data %>% 
  filter(MacroPlot.Name == target_plot,
         TagNo %in% c(badtags_CBH),
         !is.na(LiCrBHt),
         Status == "L") %>% 
  arrange(Date) %>% 
  ggplot(aes(x = Year, y = LiCrBHt)) +
  geom_line(aes(group = TagNo, color = TagNo), size = 1) +
  geom_point(aes(fill = Status), shape = 21, size = 3) +
  scale_fill_manual(values=c("L" = "forestgreen")) +
  labs(title = target_plot) +
  theme_light() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        plot.title = element_text(size = 20, face="bold", hjust = 0.5),
        plot.subtitle = element_text(size = 12, face="italic", hjust = 0.5),
        axis.title = element_text(size = 12, face="bold"),
        axis.title.y = element_text(margin = unit(c(0,4,0,0), "mm")),
        axis.title.x = element_text(margin = unit(c(4,0,0,0), "mm")),
        axis.text.y = element_text(size = 5),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1))
```
