---
title: "Grand Canyon QAQC for FMH Data"
author: "Alexandra Lalor"
date: "2024-11-06"
output:
  html_document:
    theme: readable
    highlight: 
    toc: yes
    toc_depth: 3
    toc_float:
      smooth_scroll: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Install packages (if needed)
#install.packages("tidyverse")
```

# BEFORE STARTING

### Load Packages

```{r}
# Load packages
library(tidyverse)
#library(writexl)
library(dplyr)
library(knitr)
```

### Adjust File Paths

Make sure to update file paths to be specific for your data.

```{r}
# Identify working directory (specifically user name)
getwd()

# Load in data and name them based on file path
# Change file path based on user name!
#path_data <- "C:/Users/alalor.NPS/Desktop/R/GRCA/data/data_qaqc/2024/ORIGINALS/"
path_data <- "C:/Users/allie/OneDrive/Desktop/R Projects/GRCA/data/qaqc/2024/"
path_errors <- "C:/Users/allie/OneDrive/Desktop/R Projects/GRCA/output/errors/2024/"
```

### Load Data

```{r}
# Load in data
Fuels1000_all <- read.csv(paste0(path_data, "GRCA_FMH_2024_Surface Fuels - 1000Hr_XPT.csv"))
FuelsDuffLitt_all <- read.csv(paste0(path_data, "GRCA_FMH_2024_Surface Fuels - Duff_Litter_XPT.csv"))
FuelsFine_all <- read.csv(paste0(path_data, "Surface Fuels - Fine_XPT.csv"))
HerbsPoints_all <- read.csv(paste0(path_data, "Cover - Points (metric)_XPT.csv"))
HerbsObs_all <- read.csv(paste0(path_data, "Cover - Species Composition (metric)_XPT.csv"))
Shrubs_all <- read.csv(paste0(path_data, "Density - Belts (metric)_XPT.csv"))
Seedlings_all <- read.csv(paste0(path_data, "Density - Quadrats (metric)_XPT.csv"))
Trees_all <- read.csv(paste0(path_data, "Trees - Individuals (metric)_XPT.csv"))
#PostBurn_all <- read.csv(paste0(path_data, "Post Burn Severity_XPT.csv"))

# Blank data frame for no errors
errors_blank <- data.frame("SavedQuery" = "","MacroPlot.Name" = "","Date" = "","Error" = "No Error", "Comment" = "")
```

### Check Data

Fuels1000

```{r}
# Ensure blanks in Visited column are NA
Fuels1000_all$Visited[Fuels1000_all$Visited==""] <- NA

# Clean up data frame
Fuels1000_all <- Fuels1000_all %>%
  # Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  # Separate MacroPlotName to get Monitoring type (or export data with Monitoring Type already included)
  separate(MacroPlot.Name, sep = " ", into = c("MonitoringType", "Plot"), remove = FALSE)

# Filter for only data
Fuels1000_data <- Fuels1000_all %>%
  filter(is.na(Visited))

# Filter for only headers
Fuels1000_header <- Fuels1000_all %>%
  filter(!is.na(Visited))
```

FuelsDuffLitt

```{r}
# Ensure blanks in Visited column are NA 
FuelsDuffLitt_all$Visited[FuelsDuffLitt_all$Visited==""] <- NA  

# Clean up data frame 
FuelsDuffLitt_all <- FuelsDuffLitt_all %>%   
  # Separate Date column into Date and Time columns   
  separate(Date, sep = " ", into = c("Date", "Time")) %>%   
  # Separate MacroPlotName to get Monitoring type (or export data with Monitoring Type already included)   
  separate(MacroPlot.Name, sep = " ", into = c("MonitoringType", "Plot"), remove = FALSE)

# Filter for only data 
FuelsDuffLitt_data <- FuelsDuffLitt_all %>%   
  filter(is.na(Visited))  

# Filter for only headers 
FuelsDuffLitt_header <- FuelsDuffLitt_all %>%   
  filter(!is.na(Visited))
```

FuelsFine

```{r}
# Ensure blanks in Visited column are NA 
FuelsFine_all$Visited[FuelsFine_all$Visited==""] <- NA  

# Clean up data frame 
FuelsFine_all <- FuelsFine_all %>%   
  # Separate Date column into Date and Time columns   
  separate(Date, sep = " ", into = c("Date", "Time")) %>%   
  # Separate MacroPlotName to get Monitoring type (or export data with Monitoring Type already included)   
  separate(MacroPlot.Name, sep = " ", into = c("MonitoringType", "Plot"), remove = FALSE)

# Filter for only data 
FuelsFine_data <- FuelsFine_all %>%   
  filter(is.na(Visited))  

# Filter for only headers 
FuelsFine_header <- FuelsFine_all %>%   
  filter(!is.na(Visited))
```

HerbsPoints

```{r}
# Ensure blanks in Visited column are NA 
HerbsPoints_all$Visited[HerbsPoints_all$Visited==""] <- NA  

# Clean up data frame 
HerbsPoints_all <- HerbsPoints_all %>%   
  # Separate Date column into Date and Time columns   
  separate(Date, sep = " ", into = c("Date", "Time")) %>%   
  # Separate MacroPlotName to get Monitoring type (or export data with Monitoring Type already included)   
  separate(MacroPlot.Name, sep = " ", into = c("MonitoringType", "Plot"), remove = FALSE) 

# Filter for only data 
HerbsPoints_data <- HerbsPoints_all %>%   
  filter(is.na(Visited))  

# Filter for only headers 
HerbsPoints_header <- HerbsPoints_all %>%   
  filter(!is.na(Visited))
```

HerbsObs

```{r}
# Ensure blanks in Visited column are NA 
HerbsObs_all$Visited[HerbsObs_all$Visited==""] <- NA  

# Clean up data frame 
HerbsObs_all <- HerbsObs_all %>%   
  # Separate Date column into Date and Time columns   
  separate(Date, sep = " ", into = c("Date", "Time")) %>%   
  # Separate MacroPlotName to get Monitoring type (or export data with Monitoring Type already included)   
  separate(MacroPlot.Name, sep = " ", into = c("MonitoringType", "Plot"), remove = FALSE) 

# Filter for only data 
HerbsObs_data <- HerbsObs_all %>%   
  filter(is.na(Visited))  

# Filter for only headers 
HerbsObs_header <- HerbsObs_all %>%   
  filter(!is.na(Visited))
```

Shrubs

```{r}
# Ensure blanks in Visited column are NA 
Shrubs_all$Visited[Shrubs_all$Visited==""] <- NA  

# Clean up data frame 
Shrubs_all <- Shrubs_all %>%   
  # Separate Date column into Date and Time columns   
  separate(Date, sep = " ", into = c("Date", "Time")) %>%   
  # Separate MacroPlotName to get Monitoring type (or export data with Monitoring Type already included)   
  separate(MacroPlot.Name, sep = " ", into = c("MonitoringType", "Plot"), remove = FALSE)

# Filter for only data 
Shrubs_data <- Shrubs_all %>%   
  filter(is.na(Visited))  

# Filter for only headers 
Shrubs_header <- Shrubs_all %>%   
  filter(!is.na(Visited))
```

Seedlings

```{r}
# Ensure blanks in Visited column are NA 
Seedlings_all$Visited[Seedlings_all$Visited==""] <- NA  

# Clean up data frame 
Seedlings_all <- Seedlings_all %>%   
  # Separate Date column into Date and Time columns   
  separate(Date, sep = " ", into = c("Date", "Time")) %>%   
  # Separate MacroPlotName to get Monitoring type (or export data with Monitoring Type already included)   
  separate(MacroPlot.Name, sep = " ", into = c("MonitoringType", "Plot"), remove = FALSE) 

# Filter for only data 
Seedlings_data <- Seedlings_all %>%   
  filter(is.na(Visited))  

# Filter for only headers 
Seedlings_header <- Seedlings_all %>%   
  filter(!is.na(Visited))
```

Trees

```{r}
# Ensure blanks in Visited column are NA 
Trees_all$Visited[Trees_all$Visited==""] <- NA  

# Clean up data frame 
Trees_all <- Trees_all %>%   
  # Separate Date column into Date and Time columns   
  separate(Date, sep = " ", into = c("Date", "Time")) %>%   
  # Separate MacroPlotName to get Monitoring type (or export data with Monitoring Type already included)   
  separate(MacroPlot.Name, sep = " ", into = c("MonitoringType", "Plot"), remove = FALSE) 

# Filter for only data 
Trees_data <- Trees_all %>%   
  filter(is.na(Visited))  

# Filter for only headers 
Trees_header <- Trees_all %>%   
  filter(!is.na(Visited))
```

PostBurn

```{r}
# # Ensure blanks in Visited column are NA 
# PostBurn_all$Visited[PostBurn_all$Visited==""] <- NA  
# 
# # Clean up data frame 
# PostBurn_all <- PostBurn_all %>%   
#   # Separate Date column into Date and Time columns   
#   separate(Date, sep = " ", into = c("Date", "Time")) %>%   
#   # Separate MacroPlotName to get Monitoring type (or export data with Monitoring Type already included)   
#   separate(MacroPlot.Name, sep = " ", into = c("MonitoringType", "Plot"), remove = FALSE)
# 
# # Filter for only data 
# PostBurn_data <- PostBurn_all %>%   
#   filter(is.na(Visited))
# 
# # Filter for only headers 
# PostBurn_header <- PostBurn_all %>%   
#   filter(!is.na(Visited))
```

# PROTOCOL - SURFACE FUELS {.tabs}

## SUBPROTOCOL - 1000HR CWD

This code conducts quality control checks on surface fuels data within the surface fuels coarse woody debris (1000-hr) data set.

It checks for: header info (number of transects, length of transects), decay class values, fuel constants, duplicate values, transect numbers, reasonable slope numbers, reasonable diameter values.

### Fuel 1000 Header FMH

[Problem:]{.underline} Incorrect header information entered for FMH plots.

[Procedure:]{.underline}

-   Check that NumTran = 4

```{r}
# Set parameters 
query <- "Fuel 1000 Header FMH NumTran"
query_message <- "Number of Transects" 
values_data <- Fuels1000_header$NumTran 
values_valid <- 4  

# Identify errors 
errors_Fuels1000_NumTran <- Fuels1000_header %>%   
  mutate("SavedQuery" = query, 
         "Error" = paste(query_message, "=", values_data)) %>%   
  filter(!values_data == values_valid) %>%   
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Comment"))  

# If no errors, add "No Errors" to existing data frame 
if (nrow(errors_Fuels1000_NumTran) == 0) {   
  errors_Fuels1000_NumTran <- errors_blank %>%     
    mutate(SavedQuery = query) 
} else {   
    errors_Fuels1000_NumTran <- errors_Fuels1000_NumTran 
}
```

-   Check that North Rim TranLen = 50 (or very close). North Rim plots include PIAB, PIEN, and PIPN.

```{r}
# Set parameters 
query <- "Fuel 1000 Header FMH NRim TranLen" 
query_message <- "Transect Length" 
values_data <- Fuels1000_header$TranLen 
values_valid <- 50  


# Identify errors 
errors_Fuels1000_TranLen_N <- Fuels1000_header %>%   
  mutate("SavedQuery" = query, 
         "Error" = paste(query_message, "=", values_data))  %>%    
  filter(!values_data == values_valid, 
         MonitoringType %in% c("PIAB", "PIEN", "PIPN")) %>%    
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Comment"))

# If no errors, add "No Errors" to existing data frame 
if (nrow(errors_Fuels1000_TranLen_N) == 0) {   
  errors_Fuels1000_TranLen_N <- errors_blank %>%     
    mutate(SavedQuery = query) 
} else {   
    errors_Fuels1000_TranLen_N <- errors_Fuels1000_TranLen_N 
}
```

-   Check that S Rim TranLen = 100 (or very close). South Rim plots include PIED and PIPO.

```{r}
# Set parameters 
query <- "Fuel 1000 Header FMH SRim TranLen" 
query_message <- "Transect Length" 
values_data <- Fuels1000_header$TranLen 
values_valid <- 100  

# Identify errors 
errors_Fuels1000_TranLen_S <- Fuels1000_header %>%   
  mutate("SavedQuery" = query, 
         "Error" = paste(query_message, "=", values_data)) %>%    
  filter(!values_data == values_valid,           
         MonitoringType %in% c("PIED", "PIPO")) %>%    
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Comment"))  

# If no errors, add "No Errors" to existing data frame 
if (nrow(errors_Fuels1000_TranLen_S) == 0) {   
  errors_Fuels1000_TranLen_S <- errors_blank %>%     
    mutate(SavedQuery = query) 
} else {   
    errors_Fuels1000_TranLen_S <- errors_Fuels1000_TranLen_S 
}
```

### Fuel 1000 DecayClass

[Problem:]{.underline} Logs have an entry for decay class that we don’t use.

[Procedure:]{.underline}

-   Check that DecayCl = 3 or 4

```{r}
# Set parameters
query <- "Fuel 1000 DecayClass"
query_message <- "Decay Class"
values_data <- Fuels1000_data$DecayCl
values_valid <- c(3,4)

# Identify errors
errors_Fuels1000_DecayCl <- Fuels1000_data %>%
  mutate("SavedQuery" = query, 
         "Error" = paste(query_message, "=", values_data)) %>%
  filter(!values_data %in% values_valid) %>%
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Comment"))

# If no errors, add "No Errors" to existing data frame
if (nrow(errors_Fuels1000_DecayCl) == 0) {
  errors_Fuels1000_DecayCl <- errors_blank %>%
    mutate(SavedQuery = query)
} else {
  errors_Fuels1000_DecayCl <- errors_Fuels1000_DecayCl
}
```

### Fuel 1000 FuelConst

[Problem:]{.underline} Incorrect coarse woody debris fuel constant entered for PIAB or PIED or PIEN or PIPN or PIPO. Fuel constant name should correspond to the Monitoring Type.

[Procedure:]{.underline}

-   Check that CWDFuConSt = Monitoring Type

```{r}
# Set parameters
query <- "Fuel 1000 FuelConst"
query_message <- "Fuel Constant"
values_data <- Fuels1000_data$CWDFuConSt
values_valid <- Fuels1000_data$MonitoringType

# Identify errors
errors_Fuels1000_FuelConst <- Fuels1000_data %>%
  mutate("SavedQuery" = query, 
         "Error" = paste(query_message, "=", values_data)) %>%
  filter(!values_data == values_valid) %>%
  reframe("SavedQuery" = unique(query),
          "MacroPlot.Name" = unique(MacroPlot.Name), 
          "Date" = unique(Date), 
          "Error" = unique(Error),
          "Comment" = unique(Comment))

# If no errors, add "No Errors" to existing data frame
if (nrow(errors_Fuels1000_FuelConst) == 0) {
  errors_Fuels1000_FuelConst <- errors_blank %>%
    mutate(SavedQuery = query)
} else {
  errors_Fuels1000_FuelConst <- errors_Fuels1000_FuelConst
}
```

### Fuel 1000 LogNo

[Problem:]{.underline} "Duplicate" data exists due to multiple 1000 hr fuels with the same transect / diameter / decay class. Log Numbers are missing (which would differentiate 1000 hr fuels).

[Procedure:]{.underline}

-   Check for duplication (TRUE/FALSE) and store results in new column

-   Check for TRUE values

```{r}
# Check for duplication (TRUE/FALSE) and store results in new column
Fuels1000_data_duplicate <- Fuels1000_data %>%
  select(!"Index")
Fuels1000_data_duplicate <- Fuels1000_data_duplicate %>%
  mutate("Duplicate" = duplicated(Fuels1000_data_duplicate))

# Set parameters 
query <- "Fuel 1000 LogNo" 
query_message <- "Duplicate Logs" 
values_data <- Fuels1000_data_duplicate$Duplicate
values_valid <- FALSE

# Identify errors 
errors_Fuels1000_LogNo <- Fuels1000_data_duplicate %>%   
  mutate("SavedQuery" = query, 
         "Error" = paste(query_message, "=", values_data)) %>%    
  filter(!values_data == values_valid) %>%    
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Comment"))

# If no errors, add "No Errors" to existing data frame 
if (nrow(errors_Fuels1000_LogNo) == 0) {   
  errors_Fuels1000_LogNo <- errors_blank %>%     
    mutate(SavedQuery = query) 
} else {   
    errors_Fuels1000_LogNo <- errors_Fuels1000_LogNo 
}
```

### Fuel 1000 Values

[Problem:]{.underline}  Values for transect, slope, and diameter are missing or unreasonable.

[Procedure:]{.underline}

-   Check that Transect = 1, 2, 3, or 4

```{r}
# Set parameters
query <- "Fuel 1000 Values TransNum"
query_message <- "Transect Number"
values_data <- Fuels1000_data$Transect
values_valid <- c(1, 2, 3, 4)

# Identify errors
errors_Fuels1000_TransNum <- Fuels1000_data %>%
  mutate("SavedQuery" = query, 
         "Error" = paste(query_message, "=", values_data)) %>% 
  filter(!values_data %in% values_valid) %>% 
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Comment"))

# If no errors, add "No Errors" to existing data frame
if (nrow(errors_Fuels1000_TransNum) == 0) {
  errors_Fuels1000_TransNum <- errors_blank %>%
    mutate(SavedQuery = query)
} else {
  errors_Fuels1000_TransNum <- errors_Fuels1000_TransNum
}
```

-   Check that Slope values are reasonable.

```{r}
# Set parameters
query <- "Fuel 1000 Values Slope"
query_message <- "Slope"
values_data <- Fuels1000_data$Slope
values_valid <- seq(0, 90, by = 1)

# Identify errors
errors_Fuels1000_Slope <- Fuels1000_data %>%
  mutate("SavedQuery" = query, 
         "Error" = paste(query_message, "=", values_data)) %>% 
  filter(!values_data %in% values_valid) %>% 
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Comment"))

# If no errors, add "No Errors" to existing data frame
if (nrow(errors_Fuels1000_Slope) == 0) {
  errors_Fuels1000_Slope <- errors_blank %>%
    mutate(SavedQuery = query)
} else {
  errors_Fuels1000_Slope <- errors_Fuels1000_Slope
}
```

-   Check that diameter values are reasonable, are measured at 0.5 increments, and ≥ 3.0.

```{r}
# #remove NA values for outlier test
# Fuels1000_data_dia <- Fuels1000_data %>% 
#   filter(!is.na(Dia))
# # Perform Rosner test on the valid column values
# test <- rosnerTest(Fuels1000_data_dia$Dia)
# test <- test$all.stats
# outliers <- test[which(test$Outlier == TRUE), 4]

# Set parameters
query <- "Fuel 1000 Values Dia"
query_message <- "Diameter"
values_data <- Fuels1000_data$Dia
values_valid <- seq(3, 40, by = 0.5)

# Identify errors
errors_Fuels1000_Dia <- Fuels1000_data %>%
  mutate("SavedQuery" = query, 
         "Error" = paste(query_message, "=", values_data)) %>% 
  filter(!values_data %in% values_valid) %>% 
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Comment"))

# If no errors, add "No Errors" to existing data frame
if (nrow(errors_Fuels1000_Dia) == 0) {
  errors_Fuels1000_Dia <- errors_blank %>%
    mutate(SavedQuery = query)
} else {
  errors_Fuels1000_Dia <- errors_Fuels1000_Dia
}
```

### Fuel 1000 Errors

```{r}
# Save to master error list
errors_Fuels1000 <- rbind(errors_Fuels1000_DecayCl, errors_Fuels1000_FuelConst, errors_Fuels1000_NumTran, errors_Fuels1000_TranLen_N, errors_Fuels1000_TranLen_S, errors_Fuels1000_LogNo, errors_Fuels1000_TransNum, errors_Fuels1000_Slope, errors_Fuels1000_Dia)
```

```{r}
# Table of results for quick check
errors_Fuels1000 %>% 
  kable("pipe")
```

```{r}
# Save as CSV or XLSX
path_errors
#write.csv(errors, paste0(path_errors, "errors_Fuels1000.csv"), quote=FALSE, row.names = FALSE, na = "")
#write_xlsx(errors_Fuels1000, paste0(path_errors, "errors_Fuels1000.xlsx"))
```

## SUBPROTOCOL - DUFF LITTER

This code conducts quality control checks on surface fuels data within the surface fuels duff litter data set.

It checks for: number of transects, fuel constants, number of points sampled, reasonable litter values, reasonable duff values.

### Fuel DL Header FMH

[Problem:]{.underline} Incorrect header information entered for FMH plots.

[Procedure:]{.underline}

-   Check that NumTran = 4

```{r}
# Set parameters  
query <- "Fuel DL Header FMH NumTran" 
query_message <- "Number of Transects"  
values_data <- FuelsDuffLitt_header$NumTran  
values_valid <- 4    

# Identify errors  
errors_FuelsDuffLitt_NumTran <- FuelsDuffLitt_header %>%      
  mutate("SavedQuery" = query, 
         "Error" = paste(query_message, "=", values_data)) %>%      
  filter(!values_data == values_valid) %>%      
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Comment"))    

# If no errors, add "No Errors" to existing data frame  
if (nrow(errors_FuelsDuffLitt_NumTran) == 0) {      
  errors_FuelsDuffLitt_NumTran <- errors_blank %>%          
    mutate(SavedQuery = query)  
  } else {        
    errors_FuelsDuffLitt_NumTran <- errors_FuelsDuffLitt_NumTran 
}
```

### Fuel DL FuelConst

[Problem:]{.underline} Incorrect duff & litter fuel constant entered for PIAB or PIED or PIEN or PIPN or PIPO. Fuel constant name should correspond to the Monitoring Type.

[Procedure:]{.underline}

-   Check that DLFuConSt = Monitoring Type

```{r}
# Set parameters 
query <- "Fuel DL FuelConst" 
query_message <- "Fuel Constant" 
values_data <- FuelsDuffLitt_data$DLFuConSt 
values_valid <- FuelsDuffLitt_data$MonitoringType  

# Identify errors 
errors_FuelsDuffLitt_FuelConst <- FuelsDuffLitt_data %>%   
  mutate("SavedQuery" = query, 
         "Error" = paste(query_message, "=", values_data)) %>%   
  filter(!values_data == values_valid) %>%   
  reframe("SavedQuery" = unique(query),
          "MacroPlot.Name" = unique(MacroPlot.Name), 
          "Date" = unique(Date), 
          "Error" = unique(Error),
          "Comment" = unique(Comment))

# If no errors, add "No Errors" to existing data frame 
if (nrow(errors_FuelsDuffLitt_FuelConst) == 0) {   
  errors_FuelsDuffLitt_FuelConst <- errors_blank %>%     
    mutate(SavedQuery = query)
  } else {   
    errors_FuelsDuffLitt_FuelConst <- errors_FuelsDuffLitt_FuelConst 
}
```

### Fuel DL Hits FMH

[Problem:]{.underline} Extra or missing data points in FMH plots. There should be 4 transects with 10 sample points per transect. 

[Procedure:]{.underline}

-   Group by Macroplot. Calculate number of hits per plot. Check that sum_hits = 40

```{r}
# Filter NA values. Add number of hits and store in new column
FuelsDuffLitt_summary_hits <- FuelsDuffLitt_data %>%
  filter(!is.na(LittDep), !is.na(DuffDep)) %>% 
  group_by(MacroPlot.Name) %>% 
  summarize("sum_hits" = n())

FuelsDuffLitt_data <- merge(FuelsDuffLitt_data, FuelsDuffLitt_summary_hits, all.x = T)

# Set parameters 
query <- "Fuel DL Hits" 
query_message <- "Number of Hits" 
values_data <- FuelsDuffLitt_data$sum_hits
values_valid <- 40

# Identify errors 
errors_FuelsDuffLitt_Hits <- FuelsDuffLitt_data %>%   
  mutate("SavedQuery" = query, 
         "Error" = paste(query_message, "=", values_data)) %>%
  filter(!values_data == values_valid) %>%   
  reframe("SavedQuery" = unique(query),
          "MacroPlot.Name" = unique(MacroPlot.Name), 
          "Date" = unique(Date), 
          "Error" = unique(Error),
          "Comment" = unique(Comment))

# If no errors, add "No Errors" to existing data frame 
if (nrow(errors_FuelsDuffLitt_Hits) == 0) {   
  errors_FuelsDuffLitt_Hits <- errors_blank %>%     
    mutate(SavedQuery = query)
  } else {   
    errors_FuelsDuffLitt_Hits <- errors_FuelsDuffLitt_Hits
}
```

### Fuel DL Values

[Problem:]{.underline}  Values for litter and duff depth are missing or unreasonable.

[Procedure:]{.underline}

-   Sort by LittDep. Check that litter depth values are reasonable.

```{r}
# #remove NA values for outlier test
# FuelsDuffLitt_data_litt <- FuelsDuffLitt_data %>%
#   filter(!is.na(LittDep))
# # Perform Rosner test on the valid column values
# test <- rosnerTest(FuelsDuffLitt_data_litt$LittDep)
# outliers <- test$all.stats$Value[test$all.stats$Outlier]
# 
# # Generate a message about outlier values
# outlier_indices <- which(data %in% outliers)
# max_value <- max(na.omit(data))
# min_value <- min(na.omit(data))
# mean_value <- mean(na.omit(data))


# Set parameters 
query <- "Fuel DL Values LittDep" 
query_message <- "Litter Depth" 
values_data <- FuelsDuffLitt_data$LittDep
values_valid <- max(5)

# Identify errors 
errors_FuelsDuffLitt_LittDep <- FuelsDuffLitt_data %>%   
  mutate("SavedQuery" = query, 
         "Error" = paste(query_message, "=", values_data)) %>%    
  filter(!values_data < values_valid) %>%    
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Comment"))  

# If no errors, add "No Errors" to existing data frame 
if (nrow(errors_FuelsDuffLitt_LittDep) == 0) {   
  errors_FuelsDuffLitt_LittDep <- errors_blank %>%     
    mutate(SavedQuery = query) 
} else {   
    errors_FuelsDuffLitt_LittDep <- errors_FuelsDuffLitt_LittDep 
}
```

-   Sort by DuffDep. Check that duff depth values are reasonable.

```{r}
# #remove NA values for outlier test
# FuelsDuffLitt_data_duff <- FuelsDuffLitt_data %>%
#   filter(!is.na(DuffDep))
# # Perform Rosner test on the valid column values
# test <- rosnerTest(FuelsDuffLitt_data_duff$DuffDep)
# outliers <- test$all.stats$Value[test$all.stats$Outlier]
# 
# # Generate a message about outlier values
# outlier_indices <- which(data %in% outliers)
# max_value <- max(na.omit(data))
# min_value <- min(na.omit(data))
# mean_value <- mean(na.omit(data))


# Set parameters 
query <- "Fuel DL Values DuffDep" 
query_message <- "Duff Depth" 
values_data <- FuelsDuffLitt_data$DuffDep
values_valid <- max(10)

# Identify errors 
errors_FuelsDuffLitt_DuffDep <- FuelsDuffLitt_data %>%   
  mutate("SavedQuery" = query, 
         "Error" = paste(query_message, "=", values_data)) %>%    
  filter(!values_data < values_valid) %>%    
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Comment"))  

# If no errors, add "No Errors" to existing data frame 
if (nrow(errors_FuelsDuffLitt_DuffDep) == 0) {   
  errors_FuelsDuffLitt_DuffDep <- errors_blank %>%     
    mutate(SavedQuery = query) 
} else {   
    errors_FuelsDuffLitt_DuffDep <- errors_FuelsDuffLitt_DuffDep 
}
```

### Fuel DL Errors

```{r}
# Save to master error list 
errors_FuelsDuffLitt <- rbind(errors_FuelsDuffLitt_FuelConst, errors_FuelsDuffLitt_NumTran, errors_FuelsDuffLitt_Hits, errors_FuelsDuffLitt_LittDep, errors_FuelsDuffLitt_DuffDep)
```

```{r}
# Table of results for quick check
errors_FuelsDuffLitt %>% 
  kable("pipe")
```

```{r}
# Save as CSV or XLSX 
path_errors
#write.csv(errors, paste0(path_errors, "errors_FuelsDuffLitt.csv"), quote=FALSE, row.names = FALSE, na = "") 
#write_xlsx(errors_FuelsDuffLitt, paste0(path_errors, "errors_FuelsDuffLitt.xlsx"))
```

## SUBPROTOCOL - FINE FUELS

This code conducts quality control checks on surface fuels data within the surface fuels fine (FWD) data set.

It checks for: header info (number of transects, length of transects for One/Ten/Hundred Hour), fuel constants, transect numbers, reasonable azimuth numbers, reasonable slope numbers, reasonable One/Ten/Hundred Hour count values.

### Fuel Fine Header FMH

[Problem:]{.underline} Incorrect header information entered for FMH plots.

[Procedure:]{.underline}

-   Check that NumTran = 4

```{r}
# Set parameters  
query <- "Fuel Fine Header FMH NumTran" 
query_message <- "Number of Transects"  
values_data <- FuelsFine_header$NumTran  
values_valid <- 4    

# Identify errors  
errors_FuelsFine_NumTran <- FuelsFine_header %>%      
  mutate("SavedQuery" = query, 
         "Error" = paste(query_message, "=", values_data)) %>%      
  filter(!values_data == values_valid) %>%      
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Comment"))    

# If no errors, add "No Errors" to existing data frame  
if (nrow(errors_FuelsFine_NumTran) == 0) {      
  errors_FuelsFine_NumTran <- errors_blank %>%          
    mutate(SavedQuery = query)  
  } else {        
    errors_FuelsFine_NumTran <- errors_FuelsFine_NumTran 
}
```

-   Check that OneHrTranLen = 6

```{r}
# Set parameters  
query <- "Fuel Fine Header FMH OneHrTranLen" 
query_message <- "One Hour Transect Length"  
values_data <- FuelsFine_header$OneHrTranLen
values_valid <- 6

# Identify errors  
errors_FuelsFine_OneHrTranLen <- FuelsFine_header %>%      
  mutate("SavedQuery" = query, 
         "Error" = paste(query_message, "=", values_data)) %>%      
  filter(!values_data == values_valid) %>%      
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Comment"))    

# If no errors, add "No Errors" to existing data frame  
if (nrow(errors_FuelsFine_OneHrTranLen) == 0) {      
  errors_FuelsFine_OneHrTranLen <- errors_blank %>%          
    mutate(SavedQuery = query)  
  } else {        
    errors_FuelsFine_OneHrTranLen <- errors_FuelsFine_OneHrTranLen 
}
```

-   Check that TenHrTranLen = 6

```{r}
# Set parameters  
query <- "Fuel Fine Header FMH TenHrTranLen" 
query_message <- "Ten Hour Transect Length"  
values_data <- FuelsFine_header$TenHrTranLen 
values_valid <- 6

# Identify errors  
errors_FuelsFine_TenHrTranLen <- FuelsFine_header %>%      
  mutate("SavedQuery" = query, 
         "Error" = paste(query_message, "=", values_data)) %>%      
  filter(!values_data == values_valid) %>%      
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Comment"))    

# If no errors, add "No Errors" to existing data frame  
if (nrow(errors_FuelsFine_TenHrTranLen) == 0) {      
  errors_FuelsFine_TenHrTranLen <- errors_blank %>%          
    mutate(SavedQuery = query)  
  } else {        
    errors_FuelsFine_TenHrTranLen <- errors_FuelsFine_TenHrTranLen 
}
```

-   Check that HunHrTranLen = 12

```{r}
# Set parameters  
query <- "Fuel Fine Header FMH HunHrTranLen" 
query_message <- "Hundrend Hour Transect Length"  
values_data <- FuelsFine_header$HunHrTranLen 
values_valid <- 12

# Identify errors  
errors_FuelsFine_HunHrTranLen <- FuelsFine_header %>%      
  mutate("SavedQuery" = query, 
         "Error" = paste(query_message, "=", values_data)) %>%      
  filter(!values_data == values_valid) %>%      
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Comment"))    

# If no errors, add "No Errors" to existing data frame  
if (nrow(errors_FuelsFine_HunHrTranLen) == 0) {      
  errors_FuelsFine_HunHrTranLen <- errors_blank %>%          
    mutate(SavedQuery = query)  
  } else {        
    errors_FuelsFine_HunHrTranLen <- errors_FuelsFine_HunHrTranLen 
}
```

### Fuel Fine FuelConst

[Problem:]{.underline} Incorrect fine fuel constant entered for PIAB or PIED or PIEN or PIPN or PIPO. Fuel constant name should correspond to the Monitoring Type.

[Procedure:]{.underline}

-   Check that FWDFuConSt = Monitoring Type

```{r}
# Set parameters 
query <- "Fuel Fine FuelConst" 
query_message <- "Fuel Constant" 
values_data <- FuelsFine_data$FWDFuConSt 
values_valid <- FuelsFine_data$MonitoringType  

# Identify errors 
errors_FuelsFine_FuelConst <- FuelsFine_data %>%   
  mutate("SavedQuery" = query, 
         "Error" = paste(query_message, "=", values_data)) %>%   
  filter(!values_data == values_valid) %>%   
  reframe("SavedQuery" = unique(query),
          "MacroPlot.Name" = unique(MacroPlot.Name), 
          "Date" = unique(Date), 
          "Error" = unique(Error),
          "Comment" = unique(Comment))

# If no errors, add "No Errors" to existing data frame 
if (nrow(errors_FuelsFine_FuelConst) == 0) {   
  errors_FuelsFine_FuelConst <- errors_blank %>%     
    mutate(SavedQuery = query)
  } else {   
    errors_FuelsFine_FuelConst <- errors_FuelsFine_FuelConst 
}
```

### Fuel Fine Values

[Problem:]{.underline}  Values for transect, azimuth, slope, and counts for One, Ten, Hundred hour fuels are missing or unreasonable.

[Procedure:]{.underline}

-   Check that Transect = 1, 2, 3, or 4

```{r}
# Set parameters
query <- "Fuel Fine Values TransNum"
query_message <- "Transect Number"
values_data <- FuelsFine_data$Transect
values_valid <- c(1, 2, 3, 4)

# Identify errors
errors_FuelsFine_TransNum <- FuelsFine_data %>%
  mutate("SavedQuery" = query, 
         "Error" = paste(query_message, "=", values_data)) %>% 
  filter(!values_data %in% values_valid) %>% 
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Comment"))

# If no errors, add "No Errors" to existing data frame
if (nrow(errors_FuelsFine_TransNum) == 0) {
  errors_FuelsFine_TransNum <- errors_blank %>%
    mutate(SavedQuery = query)
} else {
  errors_FuelsFine_TransNum <- errors_FuelsFine_TransNum
}
```

-   Check that Azimuth values are reasonable.

```{r}
# Set parameters
query <- "Fuel Fine Values Azimuth"
query_message <- "Azimuth"
values_data <- FuelsFine_data$Azimuth
values_valid <- seq(0, 360, by = 1)

# Identify errors
errors_FuelsFine_Azimuth <- FuelsFine_data %>%
  mutate("SavedQuery" = query, 
         "Error" = paste(query_message, "=", values_data)) %>% 
  filter(!values_data %in% values_valid) %>% 
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Comment"))

# If no errors, add "No Errors" to existing data frame
if (nrow(errors_FuelsFine_Azimuth) == 0) {
  errors_FuelsFine_Azimuth <- errors_blank %>%
    mutate(SavedQuery = query)
} else {
  errors_FuelsFine_Azimuth <- errors_FuelsFine_Azimuth
}
```

-   Check that Slope values are reasonable.

```{r}
# Set parameters
query <- "Fuel Fine Values Slope"
query_message <- "Slope"
values_data <- FuelsFine_data$Slope
values_valid <- seq(0, 90, by = 1)

# Identify errors
errors_FuelsFine_Slope <- FuelsFine_data %>%
  mutate("SavedQuery" = query, 
         "Error" = paste(query_message, "=", values_data)) %>% 
  filter(!values_data %in% values_valid) %>% 
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Comment"))

# If no errors, add "No Errors" to existing data frame
if (nrow(errors_FuelsFine_Slope) == 0) {
  errors_FuelsFine_Slope <- errors_blank %>%
    mutate(SavedQuery = query)
} else {
  errors_FuelsFine_Slope <- errors_FuelsFine_Slope
}
```

-   Check that One Hour values are reasonable.

```{r}
# #remove NA values for outlier test
# FuelsFine_data_OneHr <- FuelsFine_data %>%
#   filter(!is.na(LittDep))
# # Perform Rosner test on the valid column values
# test <- rosnerTest(FuelsFine_data_litt$LittDep)
# outliers <- test$all.stats$Value[test$all.stats$Outlier]
# 
# # Generate a message about outlier values
# outlier_indices <- which(data %in% outliers)
# max_value <- max(na.omit(data))
# min_value <- min(na.omit(data))
# mean_value <- mean(na.omit(data))


# Set parameters 
query <- "Fuel Fine Values OneHr" 
query_message <- "One Hour Count" 
values_data <- FuelsFine_data$OneHr
values_valid <- max(50)

# Identify errors 
errors_FuelsFine_OneHr <- FuelsFine_data %>%   
  mutate("SavedQuery" = query, 
         "Error" = paste(query_message, "=", values_data)) %>%    
  filter(!values_data < values_valid) %>%    
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Comment"))  

# If no errors, add "No Errors" to existing data frame 
if (nrow(errors_FuelsFine_OneHr) == 0) {   
  errors_FuelsFine_OneHr <- errors_blank %>%     
    mutate(SavedQuery = query) 
} else {   
    errors_FuelsFine_OneHr <- errors_FuelsFine_OneHr 
}
```

-   Check that Ten Hour values are reasonable.

```{r}
# #remove NA values for outlier test
# FuelsFine_data_OneHr <- FuelsFine_data %>%
#   filter(!is.na(LittDep))
# # Perform Rosner test on the valid column values
# test <- rosnerTest(FuelsFine_data_litt$LittDep)
# outliers <- test$all.stats$Value[test$all.stats$Outlier]
# 
# # Generate a message about outlier values
# outlier_indices <- which(data %in% outliers)
# max_value <- max(na.omit(data))
# min_value <- min(na.omit(data))
# mean_value <- mean(na.omit(data))


# Set parameters 
query <- "Fuel Fine Values TenHr" 
query_message <- "Ten Hour Count" 
values_data <- FuelsFine_data$TenHr
values_valid <- max(25)

# Identify errors 
errors_FuelsFine_TenHr <- FuelsFine_data %>%   
  mutate("SavedQuery" = query, 
         "Error" = paste(query_message, "=", values_data)) %>%    
  filter(!values_data < values_valid) %>%    
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Comment"))  

# If no errors, add "No Errors" to existing data frame 
if (nrow(errors_FuelsFine_TenHr) == 0) {   
  errors_FuelsFine_TenHr <- errors_blank %>%     
    mutate(SavedQuery = query) 
} else {   
    errors_FuelsFine_TenHr <- errors_FuelsFine_TenHr 
}
```

-   Check that Hundred Hour values are reasonable.

```{r}
# #remove NA values for outlier test
# FuelsFine_data_OneHr <- FuelsFine_data %>%
#   filter(!is.na(LittDep))
# # Perform Rosner test on the valid column values
# test <- rosnerTest(FuelsFine_data_litt$LittDep)
# outliers <- test$all.stats$Value[test$all.stats$Outlier]
# 
# # Generate a message about outlier values
# outlier_indices <- which(data %in% outliers)
# max_value <- max(na.omit(data))
# min_value <- min(na.omit(data))
# mean_value <- mean(na.omit(data))


# Set parameters 
query <- "Fuel Fine Values HunHr" 
query_message <- "Hundred Hour Count" 
values_data <- FuelsFine_data$HunHr
values_valid <- max(10)

# Identify errors 
errors_FuelsFine_HunHr <- FuelsFine_data %>%   
  mutate("SavedQuery" = query, 
         "Error" = paste(query_message, "=", values_data)) %>%    
  filter(!values_data < values_valid) %>%    
  select(c("SavedQuery", "MacroPlot.Name", "Date", "Error", "Comment"))  

# If no errors, add "No Errors" to existing data frame 
if (nrow(errors_FuelsFine_HunHr) == 0) {   
  errors_FuelsFine_HunHr <- errors_blank %>%     
    mutate(SavedQuery = query) 
} else {   
    errors_FuelsFine_HunHr <- errors_FuelsFine_HunHr 
}
```

### Fuel Fine Errors

```{r}
# Save to master error list 
errors_FuelsFine <- rbind(errors_FuelsFine_FuelConst, errors_FuelsFine_NumTran, errors_FuelsFine_OneHrTranLen, errors_FuelsFine_TenHrTranLen, errors_FuelsFine_HunHrTranLen, errors_FuelsFine_TransNum, errors_FuelsFine_Azimuth, errors_FuelsFine_Slope, errors_FuelsFine_OneHr, errors_FuelsFine_TenHr, errors_FuelsFine_HunHr)
```

```{r}
# Table of results for quick check
errors_FuelsFine %>% 
  kable("pipe")
```

```{r}
# Save as CSV or XLSX 
path_errors
#write.csv(errors, paste0(path_errors, "errors_FuelsFine.csv"), quote=FALSE, row.names = FALSE, na = "") 
#write_xlsx(errors_FuelsFine, paste0(path_errors, "errors_FuelsFine.xlsx"))
```

# PROTOCOL - TREES

# ALL ERRORS

```{r}
# Save to master error list 
errors_all <- rbind(errors_Fuels1000, errors_FuelsDuffLitt, errors_FuelsFine)
```

```{r}
# Table of results for quick check
errors_all %>% 
  kable("pipe")
```

```{r}
# Save as CSV or XLSX 
path_errors
#write.csv(errors, paste0(path_errors, "errors_all.csv"), quote=FALSE, row.names = FALSE, na = "") 
#write_xlsx(errors_all, paste0(path_errors, "errors_all.xlsx"))
